### 2020-09-09 paper/main.tex
Add paper docs

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
\usepackage{url}
\def\UrlFont{\rmfamily}

% Bibliography
\usepackage{biblatex}
\addbibresource{references.bib}

% Math
\usepackage{physics}
\usepackage{siunitx}
\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

% Optimization and Algorithms
\usepackage{optidef}
\usepackage{algorithmicx}
\usepackage{algorithm,algpseudocode}

% Formatting
\usepackage{xcolor}
\usepackage{bm}  % for bold symbols 
\usepackage{booktabs}  % better tables
\usepackage{pifont}  % for x mark
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{wrapfig}
\usepackage{graphicx}

% Plotting
\usepackage{pgfplots}
\pgfplotsset{compat=1.15,
	legend style={font=\footnotesize},
}
\usepackage{tikzscale}

% Custom commands
\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{S}^3}
\newcommand{\skewmat}[1]{[#1]^\times}

\newcommand{\rmap}{\varphi}
\newcommand{\invrmap}{\varphi^{-1}}

\newcommand{\dR}{\delta \mathcal{R}}
\newcommand{\rot}{ \mathcal{R} }
\newcommand{\dq}{\delta q}
\newcommand{\q}{\textbf{q}}
\newcommand{\eq}{_\text{eq}}
\newcommand{\traj}[2][N]{#2_{0:{#1}}}
\newcommand{\pass}{{\color{green} \checkmark}}
\newcommand{\fail}{{\color{red} \ding{55}}}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ BEGIN DOCUMENT ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~%
\begin{document}
\mainmatter              % start of a contribution

\title{Planning and Control with Attitude}
\author{Brian Jackson \and Zachary Manchester}

\authorrunning{Jackson and Manchester}
\institute{Stanford University, 496 Lomita Mall, Stanford, CA, USA,\\
\email{\{bjack205,zacm\}@stanford.edu}}

\maketitle

\begin{abstract}
Planning and controlling trajectories for floating-base robotic systems that experience large attitude changes is challenging due to the nontrivial group structure of 3D rotations. 
This paper introduces an accessible and unified approach for tracking control and optimization-based planning on the space of rotations based on vector calculus and linear algebra. The methodology is used to derive an extension of the Linear-Quadratic Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR (MLQR). 
We compare MLQR to a state-of-the-art geometric tracking controller designed specifically for quadrotors and demonstrate superior tracking and robustness. 
An iterative variant of MLQR is compared to existing methods on useveral constrained trajectory optimization problems, again demonstrating improved runtime performance. %to optimize trajectories for a variety of robotic systems, we provide benchmark comparisons between several of the most common attitude representations used for motion planning and control and find that the combination of unit quaternion state variables with Rodrigues parameters to represent attitude errors provides an excellent combination of performance and algorithmic simplicity. 
\keywords{motion planning and control, quaternions, optimal control, linear quadratic regulator}
\end{abstract}

\section{Introduction}
    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional translations and rotations as part of their normal operation. 
    While representing translations is straightforward and intuitive, effectively representing the nontrivial group structure of 3D rotations has been a topic of study for many decades. Although we can intuitively deduce that rotations are three-dimensional, a globally non-singular three-parameter representation of the space of rotations does not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing rotations, we must either a) pick a three-parameter representation and deal with discontinuities, or b) pick a higher-dimensional representation and deal with constraints between the parameters. While simply representing attitude is nontrivial, generating and tracking motion plans for floating-base systems is an even more challenging problem.

    Early work on control problems involving the rotation group dates back to the 1970s, with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ \cite{Baillieul1978}. Effective attitude tracking controllers have been developed for satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude errors.

    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control. Approaches to optimal control on attitude problems include analytical methods applied to satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,Lee2008}, a combination of sampling-based planning and constrained trajectory optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators \cite{Saccon2013}, or more general theory for optimization on manifolds \cites{watterson2018trajectory}. Nearly all of these methods rely heavily on principles from differential geometry and Lie group theory; however, despite these works, many recent papers in the robotics community continue to apply traditional methods for motion planning and control with no regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,Williams2017,Geisert2016}.

    In this paper, we make a departure from previous approaches to geometric planning and control that rely heavily on ideas and notation from differential geometry, and instead use only basic mathematical tools from linear algebra and calculus that should be familiar to most roboticists. Similar to \cite{Mandic2011,Xu2016}, in Sec. \ref{sec:Quaternion_Calculus} we introduce a quaternion differential calculus, but take a significantly simpler and more general approach that unifies both planning and control, enabling straight-forward adaptation of existing algorithms to systems with quaternion states. To make this concrete, in Sec. \ref{sec:MLQR} we apply our approach to derive an extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), which is the control dual to the multiplicative extended Kalman filter (MEKF) from the state estimation literature \cite{markley2014fundamentals}. In Sec. \ref{sec:experiments} we provide several simulation results demonstrating the application of our method to tracking control and constrained trajectory optimization.
    In summary, our contributions include:
    %\vspace{-10pt}
    \begin{itemize}
        \item A unified approach to quaternion differential calculus entirely based on standard vector calculus and linear algebraic operations
        \item Derivation of multiplicative LQR, an adaptation of LQR to the control of systems with quaternion states
        \item A set of benchmark motion-planning and trajectory-tracking problems in which we compare our approach to existing methods commonly used in the robotics community to account for 3D rotations
    \end{itemize}

\section{Background}
    We begin by defining some useful conventions and notation. Attitude is defined as the rotation from the robot's body frame to a global inertial frame. We also define gradients to be row vectors, that is, for $f(x) : \R^n \to \R$, $\pdv{f}{x} \in \R^{1\times n}$.

    \subsection{Unit Quaternions} \label{sec:quaternions}
        We leverage the fact that quaternions are linear operators and that the space of quaternions $\mathbb{H}$ is isomorphic to $\R^4$ to explicitly represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively.

        Quaternion multiplication is defined as
        \begin{equation} \label{eq:quat_mult}
            \q_2 \otimes \q_1 = L(q_2) q_1 = R(q_1) q_2
        \end{equation}
        where $L(q)$ and $R(q)$ are orthonormal matrices defined as
        \begin{align}
            L(q) &:= \begin{bmatrix} q_s \;\; & -q_v^T \\ q_v \;\; & q_s I + \skewmat{q_v} \end{bmatrix} 
            \label{eq:Lmult} \\
            %= \begin{bmatrix} q & G(q) \end{bmatrix} \label{eq:Lmult} \\
            R(q) &:=\begin{bmatrix} q_s \;\; & -q_v^T \\ q_v \;\; & q_s I - \skewmat{q_v} \end{bmatrix} \label{eq:Rmult},
        \end{align}
        and $\skewmat{x}$ is the skew-symmetric matrix operator
        \begin{equation}
            \skewmat{x} := \begin{bmatrix} 0 & -x_3 & x_2 \\ x_3 & 0 & -x_1\\ -x_2 & x_1 & 0 \end{bmatrix}.
        \end{equation}

        The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal to its conjugate $q^*$, which is simply the same quaternion with a negated vector part:
        \begin{equation} \label{eq:T}
            \q^* = T q := \begin{bmatrix} 1 & \\ & -I_3 \end{bmatrix} q
        \end{equation}
        The following identities, which are easily derived from \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
        \begin{align}
            &L(Tq) = L(q)^T = L(q)^{-1} \\
            &R(Tq) = R(q)^T = R(q)^{-1} .
        \end{align}

        We will sometimes find it helpful to create a quaternion with zero scalar part from a vector $r \in \R^3$. We denote this operation as,
        \begin{equation}
            \hat{r} = H r \equiv \begin{bmatrix} 0 \\ I_3 \end{bmatrix} r.
        \end{equation}
        Unit quaternions rotate a vector through the operation $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. This can be equivalently expressed using matrix multiplication as
        \begin{align} 
            r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation} %\\ 
               %&= H^T L(q) L(Hr) T q = H^T R(Hr) R(q)^T q %\label{eq:quaternion_rotation2}
        \end{align}
        where $A(q)$ is the rotation matrix in terms of the elements of the quaternion \cite{markley2014fundamentals}.

    % \todo{Merge 3.1 and 3.2. Start with unit quaternions (maybe put some of the section 4 material here), then define RPs and MRPs with the simple equations relating them to the quaternion. Keep it high level and cite references for details.}

    % \subsection{Rodrigues Parameters}
    %     Rodrigues parameters, also known as Gibbs vectors or the Cayley transformation of quaternion, are a computationally-efficient 3-parameter represention of rotations, obtained by projecting onto the plane tangent to the quaternion identity.
    %     Their definition (up to the scaling factor) is given in Table \ref{tab:retraction_maps} as the Cayley Map. 
    %     Unlike quaternions and MRPs, they do not have a sign ambiguity.

    %  \subsection{Modified Rodrigues Parameters}
    %     Modified Rodrigues Parameters (MRPs) are the newest parameter representation and can be considered a stereographic projection of the quaternion. 
    %     They are defined (up to the scaling factor) in Table \ref{tab:retraction_maps}. 
    %     MRPs have gained a lot of attention in the aerospace community since their singularity happens at \ang{360} and are a minimal 3-parameter representation. 
    %     %\cite{markley2014fundamentals} is an excellent resource for gaining further understanding of both Rodrigues parameters and MRPs. While predominantly only seen in the aerospace community, MRPs have recently been introduced to other fields such as computer graphics \cite{Terzakis2018}.

    \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
        In the current work we will restrict our focus to rigid bodies moving freely in 3D space. That is, we consider systems with dynamics of the following form:
        \begin{equation} \label{eq:rigid_body_dynamics}
            x = \begin{bmatrix} r \\ R \\ v \\ \omega \end{bmatrix}, \quad 
            \dot{x} = \begin{bmatrix} 
                v \\ 
                \texttt{kinematics}(R,\omega) \\ 
                \frac{1}{m} F_G(x,u) \\ 
                J^{-1}(\tau_L(x,u) - \omega \times J \omega) 
            \end{bmatrix}
        \end{equation}
        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, $R \in SO(3)$ is the attitude, $v \in \R^3$ is the linear velocity, and $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the global frame, $\tau_L(x,u)$ are the moments in the local (body) frame, and $\texttt{kinematics}$ are the kinematics for the chosen attitude representation. The kinematics for unit quaternions are
        \begin{equation}
            \dot{q} = \half \q \otimes \hat{\omega} = \half L(q) H \omega.
        \end{equation}
    %     \todo{Get rid of this table. Give the kinematics for each representation in line with its description above. Skip what you can and cite.}
    %     \begin{table}
    % 		\centering
    % 		\caption{Attitude Representations}
    % 		\begin{tabular}{lll}
    % 			\toprule
    % 			\textbf{Name} & \textbf{Kinematics} & \textbf{Singularity} \\
    % 			\midrule
    % 			Rotation Matrix & $R \skewmat{\omega_L}$ & None \\
    % 			Quaternions & $\half q \hat{\omega_L}$ & None \\
    % 			Euler Angles & see \cite{michael2010grasp} & $\theta \text{ (pitch)} = \ang{90}$ \\
    % 		    Rodrigues Param & $\half(I_3 + \skewmat{g} + g g^T) \omega$ & $\ang{180}$ \\
    % 		    MRP & $\frac{1+\norm{p}^2}{4}\left(I_3 + 2\frac{(\skewmat{p})^2 + \skewmat{p}}{1+\norm{p}^2}\right)\omega_L$ & \ang{360} \\
    % 			\toprule
    % 		\end{tabular}
    % 		\label{tab:attitude_kinematics}
    % 	\end{table}

    % \subsection{Trajectory Optimization}
    % \todo{I would try to get rid of this section}
    %     Trajectory optimization is a powerful framework for controlling complicated robotic systems. The value of trajectory optimization lies primarily in its generality, allowing it to be applied to a very broad class of dynamical systems. One of the objectives of this paper is to present a clear and performant method for solving trajectory optimization problems for rigid body motions, i.e. $SE(3)$. Trajectory optimization methods solve some variant of the following problem:
    %     \begin{mini}[3]
    %     	{x_{0:N},u_{0:N-1}}{\ell_N(x_N) + \sum_{k=0}^{N-1} \ell_k(x_k,u_k,\Delta t) }{}{}
    %     	\addConstraint{x_{k+1}}{=f(x_k,u_k,\Delta t), \; k = 1,...,N\text-1}
    %     	\addConstraint{g_k(x_k,u_k)}{\leq 0,\; \forall k}
    %     	\addConstraint{h_k(x_k,u_k)}{=0,\; \forall k}
    %     	\label{opt:discrete_trajopt},
    %     \end{mini}
    %     where $x_k \in \R^n, u_k \in \R^m$ are the state and control values at time step $k$, $N$ is the total number of knot points, $\ell_k$ is the stage cost function, $\ell_N$ is the terminal cost, and $f(x,u,\Delta t)$ are the discretized dynamics with step length $\Delta t$.

    %     There are many methods for solving trajectory optimization problems. Iterative LQR (iLQR) is a common method that has gain a lot of attention in recent years, mainly for it's speed and relative simplicity \cite{li2004iterative}. In the current work, we extend our implementation as part of the ALTRO solver detailed in \cite{howell2019altro}.

\section{Quaternion Differential Calculus} \label{sec:Quaternion_Calculus}
    We now present a simple but powerful method for taking derivatives of functions involving quaternions based on the notation and linear algebraic operations outlined in Sec. \ref{sec:quaternions}.

    \begin{figure}
        \centering
        % \includegraphics[height=6cm]{figures/tangent_plane.png}
        \includegraphics[height=5cm]{figures/tangent_plane.tikz}
        \caption{When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in n-dimensional space, the tangent space $T$ is a plane living in $\R^{n-1}$, illustrated here with $n=3$. Therefore, when linearizing about a unit quaternion $q \in \Q$, the space of differential rotations lives in $\R^3$.}
        \label{fig:tangent_plane}
    \end{figure}

        Derivatives consider the effect an infinitesimal perturbation to the input has on an infinitesimal perturbation to the output. For vector spaces, the composition of the perturbation with the nominal value is simple addition and the infinitesimal perturbation lives in the same space as the original vector. For unit quaternions, however, neither of these are true; instead, they compose according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig. \ref{fig:tangent_plane}).
        %, and are related through a non-linear mapping $\varphi : \R^3 \to \Q$ )
        The fact that differential unit quaternions are three-dimensional should make intuitive sense: Rotations are inherently three-dimensional and differential rotations should live in the same space as angular velocity, i.e. $\R^3$.

        There are many possible three-parameter representations for small rotations in the literature. Many authors use the exponential map \cite{Baillieul1978,Zefran1998,Lee2008,Saccon2013,Sola2017,Fan2016,watterson2018trajectory}, while others have used the Cayley map (also known as Rodrigues parameters) \cite{Kobilarov2011,Kobilarov2014}, Modified Rodrigues Parameters (MRPs) \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}. 
        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are computationally efficient and do not inherit the sign ambiguity associated with unit quaternions. The mapping between a vector of Rodrigues parameters $\phi \in \R^3$ and a unit quaternion $q$ is known as the Cayley map:
        \begin{equation} \label{eq:cayley}
            q = \varphi(\phi) = \frac{1}{\sqrt{1 + \norm{\phi}^2}} \begin{bmatrix} 1 \\ \phi \end{bmatrix}.
        \end{equation}
        We will also make use of the inverse Cayley map:
        \begin{equation}
            \phi = \varphi^{-1}(q) = \frac{q_v}{q_s}.
        \end{equation}

    \subsection{Jacobian of Vector-Valued Functions}
        When taking derivatives with respect to quaternions, we must take into account both the composition rule and the nonlinear
        mapping between the space of unit quaternions and our chosen three-parameter error representation. 
        % Before demonstrating how this is done, let's consider the real-valued case we should already be familiar with. For some function $y = r(x) : \R^n \to \R^p$ with a differential value $\delta x \in \R^n$ applied to the input:
        % \begin{align}
        %     y + \delta y = r(x+\delta x) &\approx r(x) + \nabla r(x) \delta x,
        %     % &= \lim_{\delta x \to 0} x + \pdv{r}{x} \pdv{r + \delta x}{\delta x} = r + \pdv{r}{x} 
        % \end{align}
        % we calculate $\nabla r$ by taking the derivative with respect $\delta x$, as $\delta x$ approaches zero:
        % \begin{equation}
        %     \nabla r(x) = \lim_{\delta x \to 0} \pdv{r}{x} \pdv{(x + \delta x)}{\delta x} = \lim_{\delta x \to 0} \pdv{r}{x} I = \pdv{r}{x}
        % \end{equation}
        Let $\phi \in \R^3$ be a differential rotation applied to a function with quaternion inputs $y = h(q): \Q \to \R^p$, such that
        \begin{equation} \label{eq:vector_function}
            y + \delta y = h(L(q) \varphi(\phi)) \approx h(q) +  \nabla h(q) \phi.
        \end{equation}
        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at $\phi = 0$:
        \begin{equation} \label{eq:quat_gradient}
            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) = \pdv{h}{q} \begin{bmatrix} -q_v^T \\ sI_3 + \skewmat{q_v} \end{bmatrix}
        \end{equation}
        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which essentially becomes a ``conversion factor'' allowing us to apply results from standard vector calculus to the space of unit quaternions. This form is particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be obtained using finite difference or automatic differentiation.
        As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the same (up to a constant scaling factor) for any choice of three-parameter attitude representation.

    \subsection{Hessian of Scalar-Valued Functions}
	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by taking the Jacobian of \eqref{eq:quat_gradient} with respect to $\phi$ using the product rule, again evaluated at $\phi = 0$:
	   % \begin{equation} \label{eq:hess1}
	   %     \nabla^2 h(q) = \lim_{\phi\to0} \left(L(q) \pdv{\varphi}{q} \right)^T \pdv[2]{h}{q} \left(L(q) \pdv{\varphi}{\phi} \right)+ \pdv{\phi}  \left(G(L(q) \varphi(\phi))^T \pdv{h}{q}^T \right) .
	   % \end{equation}
	   % Expanding the second term in \eqref{eq:hess1} and simplifying results in:
	    \begin{equation} \label{eq:quat_hessian}
            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q}q,
	    \end{equation}
	    where the second term comes from the second derivative of $\varphi(\phi)$. Similar to $G(q)$, this ends up being the same (up to a scaling factor) for any choice of three-parameter attitude representation.

    \subsection{Jacobian of Quaternion-Valued Functions}
        We now consider the case of a function that maps unit quaternions to unit quaternions, $q' = f(q) : \Q \to \Q$. Here we must also consider the non-trivial effect of a differential value applied to the output, i.e.:
        \begin{equation} \label{eq:dqoutput}
            L(q') \varphi(\phi') = f(L(q)\varphi(\phi)) .
        \end{equation}
        Solving \eqref{eq:dqoutput} for $\phi'$ we find,
        \begin{equation} \label{eq:phiprime}
            \phi' = \varphi^{-1} \left( L(q')^T f(L(q)\varphi(\phi)) \right) \approx \nabla f(q) \, \phi.
        \end{equation}
        Finally, the desired Jacobian is obtained by taking the derivative of \eqref{eq:phiprime} with respect to $\phi$:
        \begin{equation} \label{eq:quat_jacobian}
            \nabla f(q) = H^T L(q')^T \pdv{f}{q} L(q) H = G(q')^T \pdv{f}{q} G(q).
        \end{equation}
        The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to I_q$, where $I_q$ is the quaternion identity. Differentiating through the inverse map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q} \to H^T$ for any three-parameter attitude representation.

    % \subsection{The Attitude Jacobian}
    %     We now derive the unit quaternion \textit{attitude Jacobian} that maps infinitesimal rotations into changes in the attitude quaternion. Consider the fairly trivial function $g(q,\phi) = L(q) \varphi(\phi)$, which simply composes a small rotation $\phi$ with a nominal attitude $q$, analagous to $x + \delta x$ in a vector space. Taking the derivative with respect to $\phi$ and evaluating it at $\phi=0$, we find the following Jacobian matrix, which we denote $G(q)$:
    %     \begin{equation} \label{eq:differential_jacobian}
    %         \pdv{f}{\phi}\Big|_{\phi=0} = L(q) \pdv{\varphi}{\phi}\Big|_{\phi=0} = L(q) H = \begin{bmatrix} -q_v^T \\ q_s I + \skewmat{q_v} \end{bmatrix} \equiv G(q)
    %     \end{equation}
    %     \todo{I still don't like this...} By differentiating through the inverse of $f$, $f^{-1}(q,q') = \varphi^{-1}(L(q')^T q)$, and evaluating at the unit quaternion identity, we
    %     can similarly find that
    %     \begin{equation}
    %         \pdv{f^{-1}}{q}\Big|_{q=I_q} = H^T L^T(q') = G(q')^T
    %     \end{equation}
    %     The attitude Jacobian essentially becomes a ``conversion factor'' allowing us to apply results from standard vector calculus to the space of unit quaternions. 
    %     As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the same (up to a constant scaling factor) for any choice of three-parameter attitude representation.

        % It can also be helpful to consider how changes in the rotation affect the differential rotation, or the \textit{inverse differential Jacobian}. 
    % Consider the simple function $\phi = g(q,q_0) = \varphi^{-1}\left(L(q_0)^T q \right) \iff \varphi^{-1}(\q_0^{-1} \otimes \q)$, similar to $\delta x = x - x_0$ in vector space.
    % Taking the derivative with respect to $q$ we get
    % \begin{equation} \label{eq:inverse_differential_Jacobian}
    %      \pdv{g}{q} = \pdv{\varphi^{-1}\left(L(q_0)^T q \right)}{q} L(q_0)^T
    % \end{equation}
    % In the limit as $\phi \to 0$, $L(q_0)^T q$ approaches the identity unit quaternion, $I_q$. Similar to above, it can be shown that for all the retraction mentioned, $\pdv*{\varphi^{-1}}{q}|_{q=I_q} = H^T$. The inverse differential Jacobian is then simply $H^T L(q_0)^T = G(q_0)^T$.
    %     \begin{table}
    % 		\centering
    % 		\caption{Retraction Maps \todo{I would merge this table into the background section on different attitude descriptions. I would also ditch the ``retraction map'' lingo.}}
    % 		\begin{tabular}{lll}
    % 			\toprule
    % 			\textbf{Name} & \textbf{Definition} & \textbf{Inverse} \\
    % 			\midrule
    % 			Exponential Map & $\exp(\phi)$ & $\log(q)$ \\
    % 			Cayley Map & $\begin{bmatrix} 1 \\ g \end{bmatrix}/\sqrt{1 + \norm{g}^2}$ & $ q_v/q_s$ \\
    % 			MRP Map & $\begin{bmatrix} 1 -  \norm{p/2}^2 \\ 2 (p/2) \end{bmatrix}/(1+\norm{p/2}^2)$ & $2 q_v/(1+q_s)$ \\
    % 		    Vector Map & $\begin{bmatrix} \sqrt{1 -  \norm{\upsilon}^2} \\ \upsilon \end{bmatrix}$	& $q_v$ \\
    % 			\toprule
    % 		\end{tabular}
    % 		\label{tab:retraction_maps}
    % 	\end{table}

% 	\subsection{Jacobians of Vector-Valued Functions}
% 	    We now consider taking derivatives of a function $h(q) : \Q \to \R^p$ that has unit quaternion inputs and scalar or vector outputs. 
% 	    Specifically, we seek the Jacobian that maps an infinitesimal rotation $\phi$ applied to the input into changes in the output $\delta x$:
% 	    \begin{equation}
% 	        x + \delta x = h(L(q)\varphi(\phi)) \approx h(q) + \nabla h(q) \, \phi.
% 	    \end{equation}
% 	    Using the results of the previous section and applying the chain rule we find,
% 	    \begin{equation} \label{eq:quat_gradient}
% 	        \nabla h(q) = \pdv{\phi}\Big|_{\phi=0} \Big( h(L(q) \varphi(\phi)) \Big) = \pdv{h}{q} L(q) H = \pdv{h}{q} G(q) ,
% 	    \end{equation}
% 	    where $\pdv*{h}{q} \in \R^{p\times4}$ is the Jacobian (or gradient if $p=1$) of $h$ with respect to $q$, treating $q$ as a normal vector. This form is particularly useful in practice since $\pdv*{h}{q}$ can be obtained using finite differences or automatic differentiation.

%     \subsection{Hessians of Scalar-Valued Functions}
% 	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by taking the Jacobian of \eqref{eq:quat_gradient}. Using the product rule, we obtain,
% 	    \begin{equation} \label{eq:hess1}
% 	        \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + \pdv{\phi}\bigg|_{\phi=0}  \left(G(L(q) \varphi(\phi))^T \pdv{h}{q}^T \right) .
% 	    \end{equation}
% 	    Expanding the second term in \eqref{eq:hess1} results in:
% 	    \begin{equation} \label{eq:quat_hessian}
% 	            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q}q .
% 	    \end{equation}

	   % We obtain the actual gradient / Jacobian and Hessian of $h$ by evaluating \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} at $\phi = 0$:
	   % \begin{equation}
	   %     \pdv{h}{\phi} = \pdv{h}{q} G(q)
	   % \end{equation}
	   % and for $p = 1$
	   % \begin{equation}
	   %     \pdv[2]{h}{\phi} = G(\q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q} q
	   % \end{equation}
	   % where the last term in the Hessian comes from the fact that, for all the differential maps mentioned, $\nabla^2 \varphi(\phi, b)|_{\phi=0} = I_3 b_1$ where $b_1$ is the first element of the input vector $b$. Since $b = L(q)^T \pdv*{h}{q}$, by inspecting \eqref{eq:Lmult} we see $b_1 = q^T \pdv*{h}{q}^T = \pdv*{h}{q} q$.

% 	\subsection{Jacobians of Quaternion-Valued Functions}

%         We now consider the case of a function that maps unit quaternions to unit quaternions, $q' = f(q)$. The key idea is to consider the effect an infinitesimal rotation applied to the input has on an infinitesimal rotation applied to the output:
%         \begin{equation} \label{eq:dqoutput}
%             L(q') \varphi(\phi') = f(L(q)\varphi(\phi)) .
%         \end{equation}
%         %This means we expect the Jacobian of $f$ to be a square $3 \times 3$ matrix. 
%         Solving \eqref{eq:dqoutput} for $\phi'$ we find,
%         \begin{equation} \label{eq:phiprime}
%             \phi' = \varphi^{-1} \left( L(q')^T f(L(q)\varphi(\phi)) \right) \approx \nabla f(q) \, \phi.
%         \end{equation}
%         Finally, the desired Jacobian is obtained by taking the derivative of \eqref{eq:phiprime} with respect to $\phi$:
%         \begin{equation} \label{eq:quat_jacobian}
%             \nabla f(q) = \pdv{\phi'}{\phi}\Big|_{\phi=0} = H^T L(q')^T \pdv{f}{q} L(q) H = G(q')^T \pdv{f}{q} G(q).
%         \end{equation}

    % \subsection{Summary of Quaternion Calculus}
    % \todo{get rid of this subsection}
    %     Inspecting \eqref{eq:quat_gradient}, \eqref{eq:quat_hessian} and \eqref{eq:quat_jacobian} we see that the differential Jacobian $G(q)$ effectively ``maps'' the gradients, Hessians, and Jacobians we would obtain by naively treating the quaternion as a vector to the correct differential values that account for the group structure of rotations, along with the one additional piece required in the Hessian \eqref{eq:quat_hessian}. This is nice for a couple reasons: first, we can use standard techniques such as finite-differencing or automatic differentiation to calculate the derivatives we need, and then simply ``convert'' them with a few extra matrix multiplications; secondly, we can easily adapt existing algorithms developed on vector spaces to handle quaternions, such as the extremely useful Linear Quadratic Regulator.

\section{Multiplicative LQR} \label{sec:MLQR}
    Leveraging the methods from the previous section, we derive multiplicative LQR (MLQR), a variant of LQR that correctly accounts for the group structure of rotations. For concreteness, we consider a system with rigid body dynamics, as presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize the system about a dynamically feasible discretized reference trajectory $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps. 
    We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
    \begin{equation} \label{eq:linearized_dynamics}
        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k 
    \end{equation}
    where \begin{equation}
        A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \quad
        B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
    \end{equation}
    and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state error Jacobian:
    \begin{equation} \label{eq:state_error}
        \delta x_k = \begin{bmatrix} 
            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
        \end{bmatrix}, \; 
        E(x) = \begin{bmatrix}
            I_3 & & & \\
            & G(q) & & \\
            & & I_3 & \\
            & & & I_3 \\
        \end{bmatrix}.
    \end{equation}

    With our linearized system, we can apply the traditional LQR Riccati recursion,
    \begin{equation}
         P_k = W_k +  A_k^T P_{k+1} A_k - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
    \end{equation}
    which minimizes the quadratic cost function
    \begin{equation}
        \half \sum_{k=0}^N \delta x_k^T Q_k \delta x_k + \delta u_k^T R_k \delta u_k
    \end{equation}
    of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From this, we get our non-linear feedback policy
    \begin{equation} \label{eq:mlqr_control}
        u = K_k \delta x_k + \bar{u}_k.
    \end{equation}
    where $\delta x_k$ is computed online using the non-linear error function \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the standard LQR formula:
    \begin{equation} \label{eq:LQR_gain}
         K_k := -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k.
    \end{equation}

    The multiplicative Linear Quadratic Regulator is summarized as follows:
    \begin{enumerate}
        \item Pick $Q_k \in \R^{12 \times 12}$ to weight the state \textit{error}, and $R_k \in \R^{m \times m}$
        \item Linearize the dynamics using the state error Jacobian \eqref{eq:state_error}
        \item Compute $K_k$ using the standard LQR Riccati recursion
        \item Online, compute the control using \eqref{eq:state_error} and \eqref{eq:mlqr_control}.
    \end{enumerate}

    % We now wish to find a sequence of controls that minimize a quadratic cost function penalizing deviations from the reference trajectory and large control values, while obeying our system dynamics. This is succinctly stated by the following optimization problem:
    % \begin{mini}[2]
    % 	{x_{0:N},u_{0:N-1}}{\half x_N^T W_N x_N + \half\sum_{k=0}^{N-1} x_k^T W_k x_k + u_k^T R_k u_k }{}{}
    % 	\addConstraint{x_{k+1}}{=A_k x_k + B_k u_k}
    % 	\addConstraint{x_0}{=x_\text{init}}
    % 	\label{opt:discrete_LQR},
    % \end{mini}
    % where we have dropped the $\delta$ leading the states and controls to reduce visual clutter.

    % We now define the \textit{value function} $V_k(x)$ which is simply the cost beginning at time step $k$, such that $V_0(x)$ is the total cost, and $V_N(x) = \half x_N^T W_N x_N$ is the terminal cost. Using the Bellman equation and the principle of optimality, we can define it recursively: %, along with the \textit{action-value function} $Q(x_k,u_k)$:
    % \begin{equation} \label{eq:value_function}
    %     V_k(x) = \min_{u_k} \half x_k^T W_k x_k + \half u_k^T R_k u_k + V_{k+1}(f(x_k,u_k))
    % \end{equation}
    % % \begin{equation} \label{eq:action_value_function}
    % %     V_k(x) = \min_{u_k} Q(x_k,u_k)
    % % \end{equation}
    % By assuming $V_k(x) = \half x_k^T P_k x_k$ for some symmetric positive-definite matrix $P$, we can solve \eqref{eq:value_function} in terms of $P_{k+1}$ using the first-order necessary condition:
    % \begin{equation}
    %     \pdv{Q}{u} = R_k u_k + B^T P_{k+1}(A_k x_k + B_k u_k) = 0
    % \end{equation}
    % Solving for $u_k$ we find the optimal feedback policy: 
    % \begin{equation} \label{eq:LQR_gain}
    %     u_k^* = -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k x_k \equiv -K_k x_k
    % \end{equation}
    % % where $Q_{uu} = \pdv*[2]{Q}{u}$ and $Q_{ux} = \pdv*{Q}{u}{x}$.

    % With an optimal feedback policy, we now substitute \eqref{eq:LQR_gain} into \eqref{eq:value_function}, and after simplifying the the result to a quadratic expression in $x_k$, obtain the recurrence relationship:
    % \begin{equation}
    %      P_k = W_k +  A_k^T P_{k+1} A_k - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k.
    % \end{equation}
    % Iterating backwards in time defines the control gains $K \in \R^{m \times 12}$ for each time step. These gains are cached in the controller and looked up online.

    % When running the controller online, we compare our current state estimate $\hat{x}$ to our reference state $\bar{x}$ and compute our state differential $\delta x$ using the nonlinear function \eqref{eq:state_error}. The new control value is then computed as

    \subsection{Constrained Iterative MLQR}
        As demonstrated above, it is very straight-forward to adapt LQR to use quaternions. We similarly use this technique to adapt nonlinear trajectory optimization algorithms. Here we present the modifications to the ALTRO solver \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an augmented Lagrangian approach to handle constraints.

        For each iteration of iLQR within ALTRO, we can calculate a quadratic approximation of the cost function using \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian}. 
        Identical to \eqref{eq:linearized_dynamics}, we use \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the feedback and feedforward gains.

        The only other modification to the algorithm is during the iLQR ``forward pass'' that simulates the system with the closed-loop MLQR controller, where we use \eqref{eq:state_error} to calculate the nonlinear feedback policy during the forward roll-outs.
        For robustness, we may also add an additional criterion to the line search that checks for singularities in the three-parameter error state which, for the Cayley map, occur at \ang{180}, at which point the step size should be reduced.

        % While the core algorithm doesn't change in practice, sometimes the approach used to solve the problem has to be modified, especially with cost functions. Since many cost function penalize distance from some desired state, there are better ways of doing this than naively subtracting two quaternions. The following section describes a few cost functions that are better-suited for this type of task.

        % \begin{algorithm}
        % \begin{algorithmic}[1]
        % \caption{iMLQR} \label{alg:iMLQR}
        % \Function{ForwardPass}{$\traj[N-1]{u}$} 
        %     \State $x,u \leftarrow $ simulate the system forward 
        %     \While{not converged}
        %         \State {\color{blue} $W,R,w,r \leftarrow$ Quadratic expansion \eqref{eq:quat_gradient}, \eqref{eq:quat_hessian}}
        %         \State {\color{blue} $A,B \leftarrow$ Linearize dynamics \eqref{eq:linearized_dynamics}}
        %         \State $K,d,\Delta V \leftarrow$ \textproc{LQR}$(W,R,w,r,A,B)$
        %         \State {\color{blue} $x,u \leftarrow$ \textproc{ForwardPass}$(\traj{x},\traj{u},K,d,\Delta V)$}
        %     \EndWhile\\
        %     \Return $X,U,J$
        % \EndFunction
        % \end{algorithmic}
        % \end{algorithm}

        % \begin{algorithm}
        % \begin{algorithmic}[1]
        % \caption{Forward Pass} \label{alg:FP}
        % \Function{ForwardPass}{$X,U,K,d,\Delta V,J$} 
        %     \State Initialize $\bar{x}_0 = x_0$, $\alpha = 1$, $J^{\small{-}} \leftarrow J$
        %     \For{k=0:1:N-1}
        %         \State {\color{blue} $\delta x \leftarrow$ Using \eqref{eq:state_error}}
        %         \State $\bar{u}_k = u_k + K_k \delta x + \alpha d_k$
        %         \State $\bar{x}_{k+1} \leftarrow f(\bar{x}_k, \bar{u}_k)$
        %     \EndFor
        %     \State $J \leftarrow$ Calculate cost using $\traj{x},\traj{u}$ 
        %     \If{$J$ {\color{blue} satisfies line search conditions}}
        %         \State $x_{0:N} \leftarrow \bar{x}_{0:N}, u_{0:N-1} \leftarrow \bar{u}_{0:N-1}$
        %     \Else
        %         \State Reduce $\alpha$ and go to line 3
        %     \EndIf\\
        %     \Return $X,U,J$
        % \EndFunction
        % \end{algorithmic}
        % \end{algorithm}

    \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
        In addition to the straight-forward modifications to the LQR algorithm itself, we need to carefully consider the types of cost functions we minimize. We frequently minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We propose two different cost functions that accomplish similar behavior. 
        For sake of clarity and space, we only consider the costs on the quaternion variables: costs on the other states and the control variables remain unaffected.

        \subsubsection{Error Quadratic}
            Rather than simple subtraction, we can use a quadratic function on the three-parameter error state \eqref{eq:state_error}:
            \begin{equation} \label{eq:error_quadratic}
                J_\text{err} = \half \phi^T Q \phi = \half \left(\varphi^{-1}(\delta q)\right)^T Q \left(\varphi^{-1}(\delta q)\right).
            \end{equation}
            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient and Hessian of \eqref{eq:error_quadratic} are 
            \begin{equation}
                \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
            \end{equation}
            \begin{equation}
                \begin{aligned}
                     \nabla^2 J_\text{err} &= 
                        G(\delta q)^T \! \left(
                        D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q) 
                      + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
                    %  &+ G(\delta q)^T \nabla D \left(\delta q, Q \phi \right) G(\delta q) \\
                \end{aligned}
            \end{equation}
            where, for the Cayley map,
            \begin{equation}
                D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
                    q_v \;\; & -\frac{1}{q_s} I_3
                \end{bmatrix}
            \end{equation}
            \begin{equation}
                \nabla D = \pdv{q}(D(q)^T Q \phi) 
                = -\frac{1}{q_s^2} \begin{bmatrix} -2 \frac{q_v}{q_s}^T Q \phi \;\;  & \phi^T Q \\
                                   Q \phi \;\; & 0 \\
                \end{bmatrix}.
            \end{equation}

            % where $\delta q \in \R^4 = L(q_g)^T q$, $\psi \in \R^3 = \varphi(\delta q)$, and $\nabla \varphi^{-1}(q) \in \R^{3 \times 4} \equiv \pdv*{\varphi^{-1}}{q}$.

            % \begin{equation}
            %     \nabla_{x} = \psi^T Q \pdv{\varphi^{-1}}{q} L(\delta q) \pdv{\varphi}{\phi}
            % \end{equation}

            % \noindent We can obtain the Hessian through the chain rule:
            % \begin{equation}
            %     \begin{aligned}
            %          \nabla^2_{xx} &= \pdv{\varphi}{\phi}^T L(\delta q)^T \left(\pdv{\varphi^{-1}}{q}\right)^T Q \left(\pdv{\varphi^{-1}}{q}\right) L(\delta q) \pdv{\varphi}{\phi} \\
            %          &+ \pdv{\varphi}{\phi}^T L(\delta q)^T 
            %          \nabla^2 \varphi^{-1} \left(\delta q, Q \psi \right) L(\delta q) \pdv{\varphi}{\phi} \\
            %          &+ \nabla^2 \varphi \left(\delta q, L(\delta q)^T \pdv{\varphi^{-1}}{q} Q \psi \right)
            %     \end{aligned}
            % \end{equation}
            % Evaluating at $\phi = 0$ we get the gradient

        \subsubsection{Geodesic Distance}
            Alternatively, we can use the geodesic distance between two quaternions \cite{Kuffner2004},
            \begin{equation} \label{eq:quat_geodesic}
                J_\text{geo} = (1-\abs{q_g^T q}) ,
            \end{equation}
            whose gradient and Hessian are,
            \begin{align}
                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
            \end{align}
            where the sign of the Hessian corresponds to the sign of $q_d^T q$.

\section{Experiments} \label{sec:experiments}
    Code code for all experiments is available on GitHub at \url{\texttt{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}.

    \subsection{Trajectory Optimization}
        In this section we present several trajectory optimization problems solved using our modified version of ALTRO. 
        The two quadrotor examples use 101 knot points for 5 second trajectories, and the airplane example uses 51 knot points for a 1.25 second trajectory. All results were run on a desktop computer with an Intel i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three experiments are given in Fig. \ref{fig:timing_chart}.

        In all examples, MLQR is implemented with a unit quaternion attitude state and Rodrigues parameters as the error state. The first example compares the two cost functions given in Sec. \ref{sec:cost_functions}, while the last two use the geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec. \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention from \cite{michael2010grasp}, and quaternions (treating them na\"ively as vectors). All examples re-normalize the quaternion in the dynamics function and use a third-order explicit Runge-Kutta integrator.

        \begin{figure}
            \centering
            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
            \caption{Timing results for the airplane barrell roll, quadrotor flip, and quadrotor zig-zag examples using iLQR with roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The timing result for the Quadrotor flip with Euler angles is omitted since it failed to converge.}
            \label{fig:timing_chart}
        \end{figure}

        \subsubsection{Quadrotor Zig-Zag}
            \begin{figure}[t]
                \centering
                \begin{subfigure}[t]{0.55\textwidth}
                    \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
                    \caption{Zig-zag trajectory. The black quadrotors indicate the reference trajectory generated by iMLQR. The yellow quadrotors track the reference trajectory using time-varying MLQR, while the red uses the HFCA controller from \cite{watterson2020control}.}
                    \label{fig:zig-zag}
                \end{subfigure}
                ~
                \begin{subfigure}[t]{0.40\textwidth}
                    \centering
                    \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
                    \caption{Snapshots of the quadrotor flip trajectory. The red-colored quadrotors represent the state near t=0 s and the green-colored quadrotors represent the state near t=5.0 s}
                    \label{fig:quad_flip}
                \end{subfigure}
                \caption{Examples applying MLQR to trajectory optimization on quadrotors}
            \end{figure}

        % \begin{figure}[t]
        %     \begin{minipage}{0.69\linewidth}
        %         \centering
        %         \includegraphics[height=7cm]{figures/zig_zag_tracking.png}
        %         \caption{Zig-zag trajectory. The black quadrotors indicate the reference trajectory generated by iMLQR. The yellow quadrotors track the reference trajectory using time-varying MLQR, while the red uses the SE(3) controller from \cite{lee2010geometric}}
        %         \label{fig:zig-zag}
        %     \end{minipage}
        %     \begin{minipage}{0.3\linewidth}
        %         \centering
        %         \begin{tabular}{cc}
        %             \toprule
        %             \textbf{Method} & \textbf{Success} \\
        %             \midrule
        %             Quat & 62\% \\
        %             RPY  & 79\% \\
        %             HFCA & 97\% \\
        %             MLQR & 98\% \\
        %             \bottomrule
        %         \end{tabular}
        %         \captionsetup{type=table}
        %         \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization}
        %         \label{tab:mlqr_success}
        %     \end{minipage}
        % \end{figure}

        We set up a fairly simple trajectory optimization problem to achieve a ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of the form 
        \begin{equation}
                  \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
                + \half \sum_{k=0} u_k^T R u_k
        \end{equation}
        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and $(0,10,0)$. We compared the two cost functions from Sec. \ref{sec:cost_functions}, using a standard quadratic penalty on the non-quaternion states. 
        \begin{align}
            \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1 \pm \bar{q}^T q)
        \end{align}
        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular velocity.
        We used the following weighting matrices: 
        $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
        $Q_w = \text{diag}(10^3 I_3,I_9)$, 
        $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and terminal cost functions, respectively.

        The geodesic cost function performed significantly better than the error quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29 iterations. 
        In addition to better convergence behavior, the geodesic cost is about half as expensive to compute, which is unsurprising when comparing the gradients and Hessians in Sec. \ref{sec:cost_functions}.

        \subsubsection{Quadrotor Flip}

	    Similar to the previous example, we use a set of waypoints to encourage the quadrotor to do a flip. The cost weights for this example were 
	    \begin{align*}
	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \times 10^{-2}, 10^{-3} I_3, 10^{-3} I_3, 10^{-2} I_3) \\
	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
	    \end{align*}
	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles of \ang{90},  \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained to stay above the floor and move to a goal state 2 meters away in the $+y$ direction. The solver is initialized with a dynamically infeasible trajectory that linearly interpolates between the initial and final states, rotating the quad around the x-axis a full \ang{360}.

	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using iMLQR within the ALTRO solver to handle constraints. The trajectory generated by iLQR with quaternions and quadratic costs was successful but had a slightly different trajectory. iLQR with Euler angles failed to solve the problem. 
        % \begin{figure}[t]
        %     \centering
        %     \includegraphics[height=6cm]{figures/quad_flip.png}
        %     \caption{Snapshots of the quadrotor flip trajectory. The red-colored quadrotors represent the state near t=0 s and the green-colored quadrotors represent the state near t=5.0 s}
        %     \label{fig:quad_flip}
        % \end{figure}

        \subsection{Airplane Barrel Roll}
        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is tasked to do a barrel roll by setting a high terminal cost for being upside-down, see Fig. \ref{fig:barrellroll}.
        The solver is initialized with level flight trim conditions. 
        For both the quadrotor flip and the airplane barrel roll, iMLQR converged faster than the pure quaternion version. Despite the extra matrix multiplications and extra Hessian term, it also was faster per iteration than its iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected, better performance by correctly leveraging the structure of the rotation group during the optimization routine.

        \begin{figure}[h]
            \centering
            \includegraphics[height=3.5cm]{figures/barrellroll.png}
            \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
            \label{fig:barrellroll}
        \end{figure}

	    \begin{figure}[h]
	        \centering
	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller. Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics. }
	        \label{fig:tracking_err}
	    \end{figure}
	    \vspace{-20pt}

	\subsection{Tracking Control}
	   % \begin{wrapfigure}{r}{0.5\textwidth}
	   %     \centering
	   %     \includegraphics{figures/mlqr_basin.tikz}
	   %     \caption{Basin of attraction}
	   %     \label{fig:mlqr_basin}
	   % \end{wrapfigure}

	    This section provides examples using MLQR for tracking and stabilizing control for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art geometric tracking controller \cite{watterson2020control} based on the Hopf Fibration.

	    \subsubsection{MLQR Stabilization}
	    To test robustness and the domain of attraction of the MLQR controller, a quadrotor is linearized about hover at the origin and initialized from states uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular velocities, and uniformly over the entire space of rotations. The success rates for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is defined as the terminal error (calculated using \eqref{eq:state_error}) being within a unit ball of radius 0.1 after 10.0 seconds.
	    To further analyze the basins of attraction, we uniformly sampled over only roll (which goes singular at \ang{90} for our Euler angle representation), and yaw. The results are shown in Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the successfully stabilized initial conditions.

	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also being more generally applicable, easier to tune, and easier to implement. It is also noteworthy that na\"ively using LQR with quaternions (incorrectly treating them as vectors) results in poorer performance than Euler angles, clearly demonstrating the value of the methodology presented in the current paper.

	    \subsubsection{MLQR Tracking}
	    Time-varying MLQR is used to track the zig-zag trajectory generated in the previous section. %The controller calculated the gains at $0.01 s$ intervals, and used zero-order-hold between intervals. 
	    To test tracking performance, zero-mean Gaussian noise was added as force and moments applied to the quadrotor.

	    The trajectories and tracking error for MLQR and HFCA are shown in Figs. \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller dramatically out-performs HFCA, we found no significant difference in performance between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion. Since the tracking controller only deals with relatively small state and attitude errors, this is unsurprising. Additionally, we found the MLQR controller ran about 20x faster than the HFCA controller.

	    \begin{figure}[h]
	        \begin{minipage}{0.60\linewidth}
    	        \centering
    	        \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
    	        \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black). All points within each contour were successfully stabilized by the controller.}
    	        \label{fig:mlqr_basin}
	        \end{minipage}
	        \hspace{0.05\linewidth}
            \begin{minipage}{0.34\linewidth}
                \centering
                \begin{tabular}{cc}
                    \toprule
                    \textbf{Method} & \textbf{Success} \\
                    \midrule
                    Quat & 61\% \\
                    RPY  & 80\% \\
                    HFCA & 97\% \\
                    MLQR & 99\% \\
                    \bottomrule
                \end{tabular}
                \captionsetup{type=table}
                \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
                \label{tab:mlqr_success}
            \end{minipage}
	    \end{figure}

\section{Conclusions} \label{sec:conclusion}
    We have presented a general, unified method for planning and control on rigid-body systems with arbitrary attitude using standard linear algebra and vector calculus. By applying these methods to LQR to correctly account for the group structure of rotations, we have matched or exceeded the performance of a state-of-the-art geometric controller designed specifically for quadrotors, while being more general, requiring less system-dependent tuning, having less computational overhead, and being easier to implement.

    We have also demonstrated a straight-forward way to adapt nonlinear trajectory optimization techniques to work with quaternion-valued states.
    For these problems, we found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be computationally efficient and provide excellent convergence in practice. We recommend the use of unit quaternions within planning and control algorithms for their simplicity, computational efficiency, and lack of singularities. We further recommend the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.

    Future areas of research include the application of these methodologies to more complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as more in-depth analysis of the convergence behavior of the algorithms proposed in the current work.

\paragraph{Acknowledgements}
This material is based upon work
supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE-1656518.
Any opinions, findings, and conclusions or recommendations
expressed in this material are those of the author(s) and
do not necessarily reflect the views of the National Science
Foundation

\printbibliography
\end{document}

### 2020-09-09 paper/iros_abstract.tex
Begin adding IEEE version

--- 2020-09-09 paper/main.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% sample root file for your "contribution" to a proceedings volume
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{svproc}
-\usepackage{url}
-\def\UrlFont{\rmfamily}
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
+\IEEEoverridecommandlockouts                              % This command is only needed if 
+                                                          % you want to use the \thanks command
+\overrideIEEEmargins                                      % Needed to meet printer requirements.
 
@@
 \usepackage{pifont}  % for x mark
-\usepackage{subcaption}
-\captionsetup{compatibility=false}
-\usepackage{wrapfig}
 \usepackage{graphicx}
@@
 
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ BEGIN DOCUMENT ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~%
+\title{\LARGE \bf
+Planning with Attitude
+}
+
+\author{Brian Jackson$^1$ and Zachary Manchester$^1$%
+    \thanks{
+        $^1$Robotics Institute, 
+        Carnegie Mellon University, 
+        5000 Forbes Ave, Pittsburgh, PA, USA
+    }
+}
+
 \begin{document}
-\mainmatter              % start of a contribution
-
-\title{Planning and Control with Attitude}
-\author{Brian Jackson \and Zachary Manchester}
-
-\authorrunning{Jackson and Manchester}
-\institute{Stanford University, 496 Lomita Mall, Stanford, CA, USA,\\
-\email{\{bjack205,zacm\}@stanford.edu}}
-
 \maketitle
 
-\begin{abstract}
-Planning and controlling trajectories for floating-base robotic systems that experience large attitude changes is challenging due to the nontrivial group structure of 3D rotations. 
-This paper introduces an accessible and unified approach for tracking control and optimization-based planning on the space of rotations based on vector calculus and linear algebra.
-The methodology is used to derive an extension of the Linear-Quadratic Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR (MLQR).
-We compare MLQR to a state-of-the-art geometric tracking controller designed specifically for quadrotors and demonstrate superior tracking and robustness. 
-An iterative variant of MLQR is compared to existing methods on useveral constrained trajectory optimization problems, again demonstrating improved runtime performance.
-%to optimize trajectories for a variety of robotic systems, we provide benchmark comparisons between several of the most common attitude representations used for motion planning and control and find that the combination of unit quaternion state variables with Rodrigues parameters to represent attitude errors provides an excellent combination of performance and algorithmic simplicity.
-\keywords{motion planning and control, quaternions, optimal control, linear quadratic regulator}
-\end{abstract}
+\section{INTRODUCTION}
+    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous 
+    underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional 
+    translations and rotations as part of their normal operation. 
+    While representing translations is straightforward and intuitive, effectively 
+    representing the nontrivial group structure of 3D rotations has been a topic of study 
+    for many decades. Although we can intuitively deduce that rotations are three-dimensional, 
+    a globally non-singular three-parameter representation of the space of rotations does 
+    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing rotations, 
+    we must either a) pick a three-parameter representation and deal with discontinuities, 
+    or b) pick a higher-dimensional representation and deal with constraints between the 
+    parameters. While simply representing attitude is nontrivial, generating and tracking 
+    motion plans for floating-base systems is an even more challenging problem.
 
-\section{Introduction}
-    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional translations and rotations as part of their normal operation.
-    While representing translations is straightforward and intuitive, effectively representing the nontrivial group structure of 3D rotations has been a topic of study for many decades.
-Although we can intuitively deduce that rotations are three-dimensional, a globally non-singular three-parameter representation of the space of rotations does not exist \cite{stuelpnagel1964parametrization}.
-As a result, when parameterizing rotations, we must either a) pick a three-parameter representation and deal with discontinuities, or b) pick a higher-dimensional representation and deal with constraints between the parameters.
-While simply representing attitude is nontrivial, generating and tracking motion plans for floating-base systems is an even more challenging problem.
+    Early work on control problems involving the rotation group dates back to the 1970s, 
+    with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ 
+    \cite{Baillieul1978}. Effective attitude tracking controllers have been developed for 
+    satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,
+    Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum 
+    \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude 
+    errors.
 
-    Early work on control problems involving the rotation group dates back to the 1970s, with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ \cite{Baillieul1978}.
-Effective attitude tracking controllers have been developed for satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude errors.
+    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, 
+    sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control. 
+    Approaches to optimal control on attitude problems include analytical methods applied to
+    satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,
+    Lee2008}, a combination of sampling-based planning and constrained trajectory 
+    optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators 
+    \cite{Saccon2013}, or more general theory for optimization on manifolds 
+    \cites{watterson2018trajectory}. 
+    Nearly all of these methods rely heavily on principles from differential geometry and 
+    Lie group theory; however, despite these works, many recent papers in the robotics 
+    community continue to apply traditional methods for motion planning and control with no 
+    regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
+    Williams2017,Geisert2016}.
 
-    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control.
-Approaches to optimal control on attitude problems include analytical methods applied to satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,Lee2008}, a combination of sampling-based planning and constrained trajectory optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators \cite{Saccon2013}, or more general theory for optimization on manifolds \cites{watterson2018trajectory}.
-Nearly all of these methods rely heavily on principles from differential geometry and Lie group theory; however, despite these works, many recent papers in the robotics community continue to apply traditional methods for motion planning and control with no regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,Williams2017,Geisert2016}.
+    In this paper, we make a departure from previous approaches to geometric planning and 
+    control that rely heavily on ideas and notation from differential geometry, 
+    and instead use only basic mathematical tools from linear algebra and calculus that 
+    should be familiar to most roboticists. 
+    Similar to \cite{Mandic2011,Xu2016}, in Sec. \ref{sec:Quaternion_Calculus} we introduce 
+    a quaternion differential calculus, but take a significantly simpler and more general 
+    approach that unifies both planning and control, enabling straight-forward adaptation of 
+    existing algorithms to systems with quaternion states. 
+    To make this concrete, in Sec. \ref{sec:MLQR} we apply our approach to derive an 
+    extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), 
+    which is the control dual to the multiplicative extended Kalman filter (MEKF) from the 
+    state estimation literature \cite{markley2014fundamentals}. 
+    In Sec. \ref{sec:experiments} we provide several simulation results demonstrating the 
+    application of our method to tracking control and constrained trajectory optimization.
+    In summary, our contributions include:
 
-    In this paper, we make a departure from previous approaches to geometric planning and control that rely heavily on ideas and notation from differential geometry, and instead use only basic mathematical tools from linear algebra and calculus that should be familiar to most roboticists.
-Similar to \cite{Mandic2011,Xu2016}, in Sec.
-\ref{sec:Quaternion_Calculus} we introduce a quaternion differential calculus, but take a significantly simpler and more general approach that unifies both planning and control, enabling straight-forward adaptation of existing algorithms to systems with quaternion states.
-To make this concrete, in Sec.
-\ref{sec:MLQR} we apply our approach to derive an extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), which is the control dual to the multiplicative extended Kalman filter (MEKF) from the state estimation literature \cite{markley2014fundamentals}.
-In Sec.
-\ref{sec:experiments} we provide several simulation results demonstrating the application of our method to tracking control and constrained trajectory optimization.
-    In summary, our contributions include:
-    %\vspace{-10pt}
     \begin{itemize}
@@
 \section{Background}
+
     We begin by defining some useful conventions and notation. 
-Attitude is defined as the rotation from the robot's body frame to a global inertial frame.
-We also define gradients to be row vectors, that is, for $f(x) : \R^n \to \R$, $\pdv{f}{x} \in \R^{1\times n}$.
+    Attitude is defined as the rotation from the robot's body frame to a global inertial 
+        frame. 
+    We also define gradients to be row vectors, that is, for 
+        $f(x) : \R^n \to \R$, $\pdv{f}{x} \in \R^{1\times n}$.
 
     \subsection{Unit Quaternions} \label{sec:quaternions}
-        We leverage the fact that quaternions are linear operators and that the space of quaternions $\mathbb{H}$ is isomorphic to $\R^4$ to explicitly represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively.
+        We leverage the fact that quaternions are linear operators and that the space of 
+        quaternions $\mathbb{H}$ is isomorphic to $\R^4$ to explicitly 
+        represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as 
+        a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and 
+        $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively.
 
@@
         \begin{align}
-            L(q) &:= \begin{bmatrix} q_s \;\; & -q_v^T \\ q_v \;\; & q_s I + \skewmat{q_v} \end{bmatrix} 
+            L(q) &:= \begin{bmatrix} 
+                q_s \;\; & -q_v^T \\ 
+                q_v \;\; & q_s I + \skewmat{q_v} 
+            \end{bmatrix} 
             \label{eq:Lmult} \\
             %= \begin{bmatrix} q & G(q) \end{bmatrix} \label{eq:Lmult} \\
-            R(q) &:=\begin{bmatrix} q_s \;\; & -q_v^T \\ q_v \;\; & q_s I - \skewmat{q_v} \end{bmatrix} \label{eq:Rmult},
+            R(q) &:=\begin{bmatrix} 
+                q_s \;\; & -q_v^T \\ 
+                q_v \;\; & q_s I - \skewmat{q_v} 
+            \end{bmatrix} \label{eq:Rmult},
         \end{align}
@@
         \begin{equation}
-            \skewmat{x} := \begin{bmatrix} 0 & -x_3 & x_2 \\ x_3 & 0 & -x_1\\ -x_2 & x_1 & 0 \end{bmatrix}.
+            \skewmat{x} := \begin{bmatrix} 
+                0 & -x_3 & x_2 \\ 
+                x_3 & 0 & -x_1\\ 
+                -x_2 & x_1 & 0 
+            \end{bmatrix}.
         \end{equation}
 
-        The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal to its conjugate $q^*$, which is simply the same quaternion with a negated vector part:
+        The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal 
+        to its conjugate $q^*$, which is simply the same quaternion with a negated vector 
+        part:
         \begin{equation} \label{eq:T}
-            \q^* = T q := \begin{bmatrix} 1 & \\ & -I_3 \end{bmatrix} q
+            \q^* = T q := \begin{bmatrix} 
+                1 & \\ 
+                & -I_3 
+            \end{bmatrix} q
         \end{equation}
-        The following identities, which are easily derived from \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
+        The following identities, which are easily derived from 
+        \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
         \begin{align}
@@
 
-        We will sometimes find it helpful to create a quaternion with zero scalar part from a vector $r \in \R^3$. We denote this operation as,
+        We will sometimes find it helpful to create a quaternion with zero scalar part from 
+        a vector $r \in \R^3$. We denote this operation as,
         \begin{equation}
@@
         \end{equation}
-        Unit quaternions rotate a vector through the operation $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. This can be equivalently expressed using matrix multiplication as
+        Unit quaternions rotate a vector through the operation 
+        $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. 
+        This can be equivalently expressed using matrix multiplication as
         \begin{align} 
-            r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation} %\\ 
-               %&= H^T L(q) L(Hr) T q = H^T R(Hr) R(q)^T q %\label{eq:quaternion_rotation2}
+            r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation}
         \end{align}
-        where $A(q)$ is the rotation matrix in terms of the elements of the quaternion \cite{markley2014fundamentals}.
-
-    % \todo{Merge 3.1 and 3.2.
-Start with unit quaternions (maybe put some of the section 4 material here), then define RPs and MRPs with the simple equations relating them to the quaternion.
-Keep it high level and cite references for details.}
-
-    % \subsection{Rodrigues Parameters}
-    %     Rodrigues parameters, also known as Gibbs vectors or the Cayley transformation of quaternion, are a computationally-efficient 3-parameter represention of rotations, obtained by projecting onto the plane tangent to the quaternion identity.
-    %     Their definition (up to the scaling factor) is given in Table \ref{tab:retraction_maps} as the Cayley Map. 
-    %     Unlike quaternions and MRPs, they do not have a sign ambiguity.
-
-    %  \subsection{Modified Rodrigues Parameters}
-    %     Modified Rodrigues Parameters (MRPs) are the newest parameter representation and can be considered a stereographic projection of the quaternion. 
-    %     They are defined (up to the scaling factor) in Table \ref{tab:retraction_maps}. 
-    %     MRPs have gained a lot of attention in the aerospace community since their singularity happens at \ang{360} and are a minimal 3-parameter representation. 
-    %     %\cite{markley2014fundamentals} is an excellent resource for gaining further understanding of both Rodrigues parameters and MRPs.
-While predominantly only seen in the aerospace community, MRPs have recently been introduced to other fields such as computer graphics \cite{Terzakis2018}.
+        where $A(q)$ is the rotation matrix in terms of the elements of the quaternion 
+        \cite{markley2014fundamentals}.
 
     \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
-        In the current work we will restrict our focus to rigid bodies moving freely in 3D space. That is, we consider systems with dynamics of the following form:
+        In the current work we will restrict our focus to rigid bodies moving freely in 3D 
+        space. That is, we consider systems with dynamics of the following form:
         \begin{equation} \label{eq:rigid_body_dynamics}
@@
         \end{equation}
-        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, $R \in SO(3)$ is the attitude, $v \in \R^3$ is the linear velocity, and $\omega \in \R^3$ is the angular velocity.
-$m \in \R$ is the mass, $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the global frame, $\tau_L(x,u)$ are the moments in the local (body) frame, and $\texttt{kinematics}$ are the kinematics for the chosen attitude representation.
-The kinematics for unit quaternions are
+        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, 
+        $R \in SO(3)$ is the attitude, $v \in \R^3$ is the linear velocity, and 
+        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, 
+        $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the 
+        global frame, $\tau_L(x,u)$ are the moments in the local (body) frame, and 
+        $\texttt{kinematics}$ are the kinematics for the chosen attitude representation. The 
+        kinematics for unit quaternions are
         \begin{equation}
@@
         \end{equation}
-    %     \todo{Get rid of this table. Give the kinematics for each representation in line with its description above. Skip what you can and cite.}
-    %     \begin{table}
-    % 		\centering
-    % 		\caption{Attitude Representations}
-    % 		\begin{tabular}{lll}
-    % 			\toprule
-    % 			\textbf{Name} & \textbf{Kinematics} & \textbf{Singularity} \\
-    % 			\midrule
-    % 			Rotation Matrix & $R \skewmat{\omega_L}$ & None \\
-    % 			Quaternions & $\half q \hat{\omega_L}$ & None \\
-    % 			Euler Angles & see \cite{michael2010grasp} & $\theta \text{ (pitch)} = \ang{90}$ \\
-    % 		    Rodrigues Param & $\half(I_3 + \skewmat{g} + g g^T) \omega$ & $\ang{180}$ \\
-    % 		    MRP & $\frac{1+\norm{p}^2}{4}\left(I_3 + 2\frac{(\skewmat{p})^2 + \skewmat{p}}{1+\norm{p}^2}\right)\omega_L$ & \ang{360} \\
-    % 			\toprule
-    % 		\end{tabular}
-    % 		\label{tab:attitude_kinematics}
-    % 	\end{table}
-
-    % \subsection{Trajectory Optimization}
-    % \todo{I would try to get rid of this section}
-    %     Trajectory optimization is a powerful framework for controlling complicated robotic systems.
-The value of trajectory optimization lies primarily in its generality, allowing it to be applied to a very broad class of dynamical systems.
-One of the objectives of this paper is to present a clear and performant method for solving trajectory optimization problems for rigid body motions, i.e.
-$SE(3)$.
-Trajectory optimization methods solve some variant of the following problem:
-    %     \begin{mini}[3]
-    %     	{x_{0:N},u_{0:N-1}}{\ell_N(x_N) + \sum_{k=0}^{N-1} \ell_k(x_k,u_k,\Delta t) }{}{}
-    %     	\addConstraint{x_{k+1}}{=f(x_k,u_k,\Delta t), \; k = 1,...,N\text-1}
-    %     	\addConstraint{g_k(x_k,u_k)}{\leq 0,\; \forall k}
-    %     	\addConstraint{h_k(x_k,u_k)}{=0,\; \forall k}
-    %     	\label{opt:discrete_trajopt},
-    %     \end{mini}
-    %     where $x_k \in \R^n, u_k \in \R^m$ are the state and control values at time step $k$, $N$ is the total number of knot points, $\ell_k$ is the stage cost function, $\ell_N$ is the terminal cost, and $f(x,u,\Delta t)$ are the discretized dynamics with step length $\Delta t$.
-
-    %     There are many methods for solving trajectory optimization problems.
-Iterative LQR (iLQR) is a common method that has gain a lot of attention in recent years, mainly for it's speed and relative simplicity \cite{li2004iterative}.
-In the current work, we extend our implementation as part of the ALTRO solver detailed in \cite{howell2019altro}.
 
 \section{Quaternion Differential Calculus} \label{sec:Quaternion_Calculus}
-    We now present a simple but powerful method for taking derivatives of functions involving quaternions based on the notation and linear algebraic operations outlined in Sec. \ref{sec:quaternions}.
+    We now present a simple but powerful method for taking derivatives of functions 
+    involving quaternions based on the notation and linear algebraic operations outlined 
+    in Sec. \ref{sec:quaternions}.
 
@@
         \centering
-        % \includegraphics[height=6cm]{figures/tangent_plane.png}
         \includegraphics[height=5cm]{figures/tangent_plane.tikz}
-        \caption{When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in n-dimensional space, the tangent space $T$ is a plane living in $\R^{n-1}$, illustrated here with $n=3$.
-Therefore, when linearizing about a unit quaternion $q \in \Q$, the space of differential rotations lives in $\R^3$.}
+        \caption{
+            When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in 
+            n-dimensional space, the tangent space $T$ is a plane living in $\R^{n-1}$, 
+            illustrated here with $n=3$. Therefore, when linearizing about a unit 
+            quaternion $q \in \Q$, the space of differential rotations lives in $\R^3$.
+        }
         \label{fig:tangent_plane}
@@
 
-        Derivatives consider the effect an infinitesimal perturbation to the input has on an infinitesimal perturbation to the output.
-For vector spaces, the composition of the perturbation with the nominal value is simple addition and the infinitesimal perturbation lives in the same space as the original vector.
-For unit quaternions, however, neither of these are true; instead, they compose according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig.
+        Derivatives consider the effect an infinitesimal perturbation to the input has on
+        an infinitesimal perturbation to the output. For vector spaces, the composition
+        of the perturbation with the nominal value is simple addition and the
+        infinitesimal perturbation lives in the same space as the original vector. For
+        unit quaternions, however, neither of these are true; instead, they compose
+        according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to
+        first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig.
         \ref{fig:tangent_plane}).
+
         %, and are related through a non-linear mapping $\varphi : \R^3 \to \Q$ )
-        The fact that differential unit quaternions are three-dimensional should make intuitive sense: Rotations are inherently three-dimensional and differential rotations should live in the same space as angular velocity, i.e.
-$\R^3$.
+        The fact that differential unit quaternions are three-dimensional should make
+        intuitive sense: Rotations are inherently three-dimensional and differential
+        rotations should live in the same space as angular velocity, i.e. $\R^3$.
 
-        There are many possible three-parameter representations for small rotations in the literature.
-Many authors use the exponential map \cite{Baillieul1978,Zefran1998,Lee2008,Saccon2013,Sola2017,Fan2016,watterson2018trajectory}, while others have used the Cayley map (also known as Rodrigues parameters) \cite{Kobilarov2011,Kobilarov2014}, Modified Rodrigues Parameters (MRPs) \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}.
-        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are computationally efficient and do not inherit the sign ambiguity associated with unit quaternions.
-The mapping between a vector of Rodrigues parameters $\phi \in \R^3$ and a unit quaternion $q$ is known as the Cayley map:
-        \begin{equation} \label{eq:cayley}
+        skinny dog runs after skinny cat
+
+        There are many possible three-parameter representations for small rotations in
+        the literature. Many authors use the exponential map \cite{Baillieul1978,
+        Zefran1998, Lee2008, Saccon2013, Sola2017, Fan2016, watterson2018trajectory},
+        while others have used the Cayley map (also known as Rodrigues parameters)
+        \cite{Kobilarov2011, Kobilarov2014}, Modified Rodrigues Parameters (MRPs)
+        \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}.
+        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are
+        computationally efficient and do not inherit the sign ambiguity associated with
+        unit quaternions. The mapping between a vector of Rodrigues parameters $\phi \in
+        \R^3$ and a unit quaternion $q$ is known as the Cayley map: \begin{equation}
+        \label{eq:cayley}
             q = \varphi(\phi) = \frac{1}{\sqrt{1 + \norm{\phi}^2}} \begin{bmatrix} 1 \\ \phi \end{bmatrix}.
@@
     \subsection{Jacobian of Vector-Valued Functions}
-        When taking derivatives with respect to quaternions, we must take into account both the composition rule and the nonlinear
-        mapping between the space of unit quaternions and our chosen three-parameter error representation. 
-        % Before demonstrating how this is done, let's consider the real-valued case we should already be familiar with.
-For some function $y = r(x) : \R^n \to \R^p$ with a differential value $\delta x \in \R^n$ applied to the input:
-        % \begin{align}
-        %     y + \delta y = r(x+\delta x) &\approx r(x) + \nabla r(x) \delta x,
-        %     % &= \lim_{\delta x \to 0} x + \pdv{r}{x} \pdv{r + \delta x}{\delta x} = r + \pdv{r}{x} 
-        % \end{align}
-        % we calculate $\nabla r$ by taking the derivative with respect $\delta x$, as $\delta x$ approaches zero:
-        % \begin{equation}
-        %     \nabla r(x) = \lim_{\delta x \to 0} \pdv{r}{x} \pdv{(x + \delta x)}{\delta x} = \lim_{\delta x \to 0} \pdv{r}{x} I = \pdv{r}{x}
-        % \end{equation}
-        Let $\phi \in \R^3$ be a differential rotation applied to a function with quaternion inputs $y = h(q): \Q \to \R^p$, such that
+        When taking derivatives with respect to quaternions, we must take into account
+        both the composition rule and the nonlinear mapping between the space of unit
+        quaternions and our chosen three-parameter error representation.
+
+        Let $\phi \in \R^3$ be a differential rotation applied to a function with
+        quaternion inputs $y = h(q): \Q \to \R^p$, such that
         \begin{equation} \label{eq:vector_function}
@@
         \end{equation}
-        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at $\phi = 0$:
+        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by
+        differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at
+        $\phi = 0$:
         \begin{equation} \label{eq:quat_gradient}
-            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) = \pdv{h}{q} \begin{bmatrix} -q_v^T \\ sI_3 + \skewmat{q_v} \end{bmatrix}
+            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) 
+                        = \pdv{h}{q} \begin{bmatrix} 
+                            -q_v^T \\ 
+                            sI_3 + \skewmat{q_v}
+                        \end{bmatrix}
         \end{equation}
-        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which essentially becomes a ``conversion factor'' allowing us to apply results from standard vector calculus to the space of unit quaternions.
-This form is particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be obtained using finite difference or automatic differentiation.
-        As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the same (up to a constant scaling factor) for any choice of three-parameter attitude representation.
+        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which
+        essentially becomes a ``conversion factor'' allowing us to apply results from
+        standard vector calculus to the space of unit quaternions. This form is
+        particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be
+        obtained using finite difference or automatic differentiation.
+        As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the
+        same (up to a constant scaling factor) for any choice of three-parameter attitude
+        representation.
 
     \subsection{Hessian of Scalar-Valued Functions}
-	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by taking the Jacobian of \eqref{eq:quat_gradient} with respect to $\phi$ using the product rule, again evaluated at $\phi = 0$:
-	   % \begin{equation} \label{eq:hess1}
-	   %     \nabla^2 h(q) = \lim_{\phi\to0} \left(L(q) \pdv{\varphi}{q} \right)^T \pdv[2]{h}{q} \left(L(q) \pdv{\varphi}{\phi} \right)+ \pdv{\phi}  \left(G(L(q) \varphi(\phi))^T \pdv{h}{q}^T \right) .
-	   % \end{equation}
-	   % Expanding the second term in \eqref{eq:hess1} and simplifying results in:
+	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by
+	    taking the Jacobian of \eqref{eq:quat_gradient} with respect to $\phi$ using the
+        product rule, again evaluated at $\phi = 0$:
+
 	    \begin{equation} \label{eq:quat_hessian}
@@
 	    where the second term comes from the second derivative of $\varphi(\phi)$.
-Similar to $G(q)$, this ends up being the same (up to a scaling factor) for any choice of three-parameter attitude representation.
+	    Similar to $G(q)$, this ends up being the same (up to a scaling factor) for any
+        choice of three-parameter attitude representation.
 
     \subsection{Jacobian of Quaternion-Valued Functions}
-        We now consider the case of a function that maps unit quaternions to unit quaternions, $q' = f(q) : \Q \to \Q$.
-Here we must also consider the non-trivial effect of a differential value applied to the output, i.e.:
+        We now consider the case of a function that maps unit quaternions to unit
+        quaternions, $q' = f(q) : \Q \to \Q$. Here we must also consider the non-trivial
+        effect of a differential value applied to the output, i.e.:
         \begin{equation} \label{eq:dqoutput}
@@
         \end{equation}
-        Finally, the desired Jacobian is obtained by taking the derivative of \eqref{eq:phiprime} with respect to $\phi$:
+        Finally, the desired Jacobian is obtained by taking the derivative of
+        \eqref{eq:phiprime} with respect to $\phi$:
         \begin{equation} \label{eq:quat_jacobian}
@@
         \end{equation}
-        The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to I_q$, where $I_q$ is the quaternion identity.
-Differentiating through the inverse map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q} \to H^T$ for any three-parameter attitude representation.
-
-    % \subsection{The Attitude Jacobian}
-    %     We now derive the unit quaternion \textit{attitude Jacobian} that maps infinitesimal rotations into changes in the attitude quaternion.
-Consider the fairly trivial function $g(q,\phi) = L(q) \varphi(\phi)$, which simply composes a small rotation $\phi$ with a nominal attitude $q$, analagous to $x + \delta x$ in a vector space.
-Taking the derivative with respect to $\phi$ and evaluating it at $\phi=0$, we find the following Jacobian matrix, which we denote $G(q)$:
-    %     \begin{equation} \label{eq:differential_jacobian}
-    %         \pdv{f}{\phi}\Big|_{\phi=0} = L(q) \pdv{\varphi}{\phi}\Big|_{\phi=0} = L(q) H = \begin{bmatrix} -q_v^T \\ q_s I + \skewmat{q_v} \end{bmatrix} \equiv G(q)
-    %     \end{equation}
-    %     \todo{I still don't like this...} By differentiating through the inverse of $f$, $f^{-1}(q,q') = \varphi^{-1}(L(q')^T q)$, and evaluating at the unit quaternion identity, we
-    %     can similarly find that
-    %     \begin{equation}
-    %         \pdv{f^{-1}}{q}\Big|_{q=I_q} = H^T L^T(q') = G(q')^T
-    %     \end{equation}
-    %     The attitude Jacobian essentially becomes a ``conversion factor'' allowing us to apply results from standard vector calculus to the space of unit quaternions. 
-    %     As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the same (up to a constant scaling factor) for any choice of three-parameter attitude representation.
-
-        % It can also be helpful to consider how changes in the rotation affect the differential rotation, or the \textit{inverse differential Jacobian}. 
-    % Consider the simple function $\phi = g(q,q_0) = \varphi^{-1}\left(L(q_0)^T q \right) \iff \varphi^{-1}(\q_0^{-1} \otimes \q)$, similar to $\delta x = x - x_0$ in vector space.
-    % Taking the derivative with respect to $q$ we get
-    % \begin{equation} \label{eq:inverse_differential_Jacobian}
-    %      \pdv{g}{q} = \pdv{\varphi^{-1}\left(L(q_0)^T q \right)}{q} L(q_0)^T
-    % \end{equation}
-    % In the limit as $\phi \to 0$, $L(q_0)^T q$ approaches the identity unit quaternion, $I_q$.
-Similar to above, it can be shown that for all the retraction mentioned, $\pdv*{\varphi^{-1}}{q}|_{q=I_q} = H^T$.
-The inverse differential Jacobian is then simply $H^T L(q_0)^T = G(q_0)^T$.
-    %     \begin{table}
-    % 		\centering
-    % 		\caption{Retraction Maps \todo{I would merge this table into the background section on different attitude descriptions. I would also ditch the ``retraction map'' lingo.}}
-    % 		\begin{tabular}{lll}
-    % 			\toprule
-    % 			\textbf{Name} & \textbf{Definition} & \textbf{Inverse} \\
-    % 			\midrule
-    % 			Exponential Map & $\exp(\phi)$ & $\log(q)$ \\
-    % 			Cayley Map & $\begin{bmatrix} 1 \\ g \end{bmatrix}/\sqrt{1 + \norm{g}^2}$ & $ q_v/q_s$ \\
-    % 			MRP Map & $\begin{bmatrix} 1 -  \norm{p/2}^2 \\ 2 (p/2) \end{bmatrix}/(1+\norm{p/2}^2)$ & $2 q_v/(1+q_s)$ \\
-    % 		    Vector Map & $\begin{bmatrix} \sqrt{1 -  \norm{\upsilon}^2} \\ \upsilon \end{bmatrix}$	& $q_v$ \\
-    % 			\toprule
-    % 		\end{tabular}
-    % 		\label{tab:retraction_maps}
-    % 	\end{table}
-
-% 	\subsection{Jacobians of Vector-Valued Functions}
-% 	    We now consider taking derivatives of a function $h(q) : \Q \to \R^p$ that has unit quaternion inputs and scalar or vector outputs. 
-% 	    Specifically, we seek the Jacobian that maps an infinitesimal rotation $\phi$ applied to the input into changes in the output $\delta x$:
-% 	    \begin{equation}
-% 	        x + \delta x = h(L(q)\varphi(\phi)) \approx h(q) + \nabla h(q) \, \phi.
-% 	    \end{equation}
-% 	    Using the results of the previous section and applying the chain rule we find,
-% 	    \begin{equation} \label{eq:quat_gradient}
-% 	        \nabla h(q) = \pdv{\phi}\Big|_{\phi=0} \Big( h(L(q) \varphi(\phi)) \Big) = \pdv{h}{q} L(q) H = \pdv{h}{q} G(q) ,
-% 	    \end{equation}
-% 	    where $\pdv*{h}{q} \in \R^{p\times4}$ is the Jacobian (or gradient if $p=1$) of $h$ with respect to $q$, treating $q$ as a normal vector.
-This form is particularly useful in practice since $\pdv*{h}{q}$ can be obtained using finite differences or automatic differentiation.
-
-%     \subsection{Hessians of Scalar-Valued Functions}
-% 	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by taking the Jacobian of \eqref{eq:quat_gradient}. Using the product rule, we obtain,
-% 	    \begin{equation} \label{eq:hess1}
-% 	        \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + \pdv{\phi}\bigg|_{\phi=0}  \left(G(L(q) \varphi(\phi))^T \pdv{h}{q}^T \right) .
-% 	    \end{equation}
-% 	    Expanding the second term in \eqref{eq:hess1} results in:
-% 	    \begin{equation} \label{eq:quat_hessian}
-% 	            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q}q .
-% 	    \end{equation}
-
-	   % We obtain the actual gradient / Jacobian and Hessian of $h$ by evaluating \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} at $\phi = 0$:
-	   % \begin{equation}
-	   %     \pdv{h}{\phi} = \pdv{h}{q} G(q)
-	   % \end{equation}
-	   % and for $p = 1$
-	   % \begin{equation}
-	   %     \pdv[2]{h}{\phi} = G(\q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q} q
-	   % \end{equation}
-	   % where the last term in the Hessian comes from the fact that, for all the differential maps mentioned, $\nabla^2 \varphi(\phi, b)|_{\phi=0} = I_3 b_1$ where $b_1$ is the first element of the input vector $b$.
-Since $b = L(q)^T \pdv*{h}{q}$, by inspecting \eqref{eq:Lmult} we see $b_1 = q^T \pdv*{h}{q}^T = \pdv*{h}{q} q$.
-
-% 	\subsection{Jacobians of Quaternion-Valued Functions}
-
-%         We now consider the case of a function that maps unit quaternions to unit quaternions, $q' = f(q)$.
-The key idea is to consider the effect an infinitesimal rotation applied to the input has on an infinitesimal rotation applied to the output:
-%         \begin{equation} \label{eq:dqoutput}
-%             L(q') \varphi(\phi') = f(L(q)\varphi(\phi)) .
-%         \end{equation}
-%         %This means we expect the Jacobian of $f$ to be a square $3 \times 3$ matrix. 
-%         Solving \eqref{eq:dqoutput} for $\phi'$ we find,
-%         \begin{equation} \label{eq:phiprime}
-%             \phi' = \varphi^{-1} \left( L(q')^T f(L(q)\varphi(\phi)) \right) \approx \nabla f(q) \, \phi.
-%         \end{equation}
-%         Finally, the desired Jacobian is obtained by taking the derivative of \eqref{eq:phiprime} with respect to $\phi$:
-%         \begin{equation} \label{eq:quat_jacobian}
-%             \nabla f(q) = \pdv{\phi'}{\phi}\Big|_{\phi=0} = H^T L(q')^T \pdv{f}{q} L(q) H = G(q')^T \pdv{f}{q} G(q).
-%         \end{equation}
-
-    % \subsection{Summary of Quaternion Calculus}
-    % \todo{get rid of this subsection}
-    %     Inspecting \eqref{eq:quat_gradient}, \eqref{eq:quat_hessian} and \eqref{eq:quat_jacobian} we see that the differential Jacobian $G(q)$ effectively ``maps'' the gradients, Hessians, and Jacobians we would obtain by naively treating the quaternion as a vector to the correct differential values that account for the group structure of rotations, along with the one additional piece required in the Hessian \eqref{eq:quat_hessian}.
-This is nice for a couple reasons: first, we can use standard techniques such as finite-differencing or automatic differentiation to calculate the derivatives we need, and then simply ``convert'' them with a few extra matrix multiplications; secondly, we can easily adapt existing algorithms developed on vector spaces to handle quaternions, such as the extremely useful Linear Quadratic Regulator.
+        The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to
+        I_q$, where $I_q$ is the quaternion identity. Differentiating through the inverse
+        map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q}
+        \to H^T$ for any three-parameter attitude representation.
 
 \section{Multiplicative LQR} \label{sec:MLQR}
-    Leveraging the methods from the previous section, we derive multiplicative LQR (MLQR), a variant of LQR that correctly accounts for the group structure of rotations.
-For concreteness, we consider a system with rigid body dynamics, as presented in Sec.
-\ref{sec:rigidbody_dynamics}, and design a controller to stabilize the system about a dynamically feasible discretized reference trajectory $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
+    Leveraging the methods from the previous section, we derive multiplicative LQR
+    (MLQR), a variant of LQR that correctly accounts for the group structure of
+    rotations. For concreteness, we consider a system with rigid body dynamics, as
+    presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize
+    the system about a dynamically feasible discretized reference trajectory
+    $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
     We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
@@
     where \begin{equation}
-        A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \quad
+        \begin{aligned}
+            A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
             B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
+        \end{aligned}
     \end{equation}
@@
     \begin{equation} \label{eq:state_error}
+        \setlength\arraycolsep{1pt}
         \delta x_k = \begin{bmatrix} 
             r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
-        \end{bmatrix}, \; 
+        \end{bmatrix}\!, \;
         E(x) = \begin{bmatrix}
@@
             & & & I_3 \\
-        \end{bmatrix}.
+        \end{bmatrix}\!.
     \end{equation}
@@
     With our linearized system, we can apply the traditional LQR Riccati recursion,
-    \begin{equation}
-         P_k = W_k +  A_k^T P_{k+1} A_k - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
-    \end{equation}
+    \begin{multline}
+         P_k = W_k +  A_k^T P_{k+1} A_k \\
+         - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
+    \end{multline}
     which minimizes the quadratic cost function
@@
     \end{equation}
-    of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls.
-From this, we get our non-linear feedback policy
+    of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on
+    the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From
+    this, we get our non-linear feedback policy
     \begin{equation} \label{eq:mlqr_control}
@@
     \end{equation}
-    where $\delta x_k$ is computed online using the non-linear error function \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the standard LQR formula:
+    where $\delta x_k$ is computed online using the non-linear error function
+    \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the
+    standard LQR formula:
     \begin{equation} \label{eq:LQR_gain}
@@
 
-    % We now wish to find a sequence of controls that minimize a quadratic cost function penalizing deviations from the reference trajectory and large control values, while obeying our system dynamics.
-This is succinctly stated by the following optimization problem:
-    % \begin{mini}[2]
-    % 	{x_{0:N},u_{0:N-1}}{\half x_N^T W_N x_N + \half\sum_{k=0}^{N-1} x_k^T W_k x_k + u_k^T R_k u_k }{}{}
-    % 	\addConstraint{x_{k+1}}{=A_k x_k + B_k u_k}
-    % 	\addConstraint{x_0}{=x_\text{init}}
-    % 	\label{opt:discrete_LQR},
-    % \end{mini}
-    % where we have dropped the $\delta$ leading the states and controls to reduce visual clutter.
+    \subsection{Constrained Iterative MLQR}
+        As demonstrated above, it is very straight-forward to adapt LQR to use
+        quaternions. We similarly use this technique to adapt nonlinear trajectory
+        optimization algorithms. Here we present the modifications to the ALTRO solver
+        \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an
+        augmented Lagrangian approach to handle constraints.
 
-    % We now define the \textit{value function} $V_k(x)$ which is simply the cost beginning at time step $k$, such that $V_0(x)$ is the total cost, and $V_N(x) = \half x_N^T W_N x_N$ is the terminal cost.
-Using the Bellman equation and the principle of optimality, we can define it recursively: %, along with the \textit{action-value function} $Q(x_k,u_k)$:
-    % \begin{equation} \label{eq:value_function}
-    %     V_k(x) = \min_{u_k} \half x_k^T W_k x_k + \half u_k^T R_k u_k + V_{k+1}(f(x_k,u_k))
-    % \end{equation}
-    % % \begin{equation} \label{eq:action_value_function}
-    % %     V_k(x) = \min_{u_k} Q(x_k,u_k)
-    % % \end{equation}
-    % By assuming $V_k(x) = \half x_k^T P_k x_k$ for some symmetric positive-definite matrix $P$, we can solve \eqref{eq:value_function} in terms of $P_{k+1}$ using the first-order necessary condition:
-    % \begin{equation}
-    %     \pdv{Q}{u} = R_k u_k + B^T P_{k+1}(A_k x_k + B_k u_k) = 0
-    % \end{equation}
-    % Solving for $u_k$ we find the optimal feedback policy: 
-    % \begin{equation} \label{eq:LQR_gain}
-    %     u_k^* = -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k x_k \equiv -K_k x_k
-    % \end{equation}
-    % % where $Q_{uu} = \pdv*[2]{Q}{u}$ and $Q_{ux} = \pdv*{Q}{u}{x}$.
+        For each iteration of iLQR within ALTRO, we can calculate a quadratic
+        approximation of the cost function using \eqref{eq:quat_gradient} and
+        \eqref{eq:quat_hessian}. Identical to \eqref{eq:linearized_dynamics}, we use
+        \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the
+        feedback and feedforward gains.
 
-    % With an optimal feedback policy, we now substitute \eqref{eq:LQR_gain} into \eqref{eq:value_function}, and after simplifying the the result to a quadratic expression in $x_k$, obtain the recurrence relationship:
-    % \begin{equation}
-    %      P_k = W_k +  A_k^T P_{k+1} A_k - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k.
-    % \end{equation}
-    % Iterating backwards in time defines the control gains $K \in \R^{m \times 12}$ for each time step. These gains are cached in the controller and looked up online.
-
-    % When running the controller online, we compare our current state estimate $\hat{x}$ to our reference state $\bar{x}$ and compute our state differential $\delta x$ using the nonlinear function \eqref{eq:state_error}.
-The new control value is then computed as
-
-    \subsection{Constrained Iterative MLQR}
-        As demonstrated above, it is very straight-forward to adapt LQR to use quaternions.
-We similarly use this technique to adapt nonlinear trajectory optimization algorithms.
-Here we present the modifications to the ALTRO solver \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an augmented Lagrangian approach to handle constraints.
-
-        For each iteration of iLQR within ALTRO, we can calculate a quadratic approximation of the cost function using \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian}. 
-        Identical to \eqref{eq:linearized_dynamics}, we use \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the feedback and feedforward gains.
-
-        The only other modification to the algorithm is during the iLQR ``forward pass'' that simulates the system with the closed-loop MLQR controller, where we use \eqref{eq:state_error} to calculate the nonlinear feedback policy during the forward roll-outs.
-        For robustness, we may also add an additional criterion to the line search that checks for singularities in the three-parameter error state which, for the Cayley map, occur at \ang{180}, at which point the step size should be reduced.
-
-        % While the core algorithm doesn't change in practice, sometimes the approach used to solve the problem has to be modified, especially with cost functions.
-Since many cost function penalize distance from some desired state, there are better ways of doing this than naively subtracting two quaternions.
-The following section describes a few cost functions that are better-suited for this type of task.
-
-        % \begin{algorithm}
-        % \begin{algorithmic}[1]
-        % \caption{iMLQR} \label{alg:iMLQR}
-        % \Function{ForwardPass}{$\traj[N-1]{u}$} 
-        %     \State $x,u \leftarrow $ simulate the system forward 
-        %     \While{not converged}
-        %         \State {\color{blue} $W,R,w,r \leftarrow$ Quadratic expansion \eqref{eq:quat_gradient}, \eqref{eq:quat_hessian}}
-        %         \State {\color{blue} $A,B \leftarrow$ Linearize dynamics \eqref{eq:linearized_dynamics}}
-        %         \State $K,d,\Delta V \leftarrow$ \textproc{LQR}$(W,R,w,r,A,B)$
-        %         \State {\color{blue} $x,u \leftarrow$ \textproc{ForwardPass}$(\traj{x},\traj{u},K,d,\Delta V)$}
-        %     \EndWhile\\
-        %     \Return $X,U,J$
-        % \EndFunction
-        % \end{algorithmic}
-        % \end{algorithm}
-
-        % \begin{algorithm}
-        % \begin{algorithmic}[1]
-        % \caption{Forward Pass} \label{alg:FP}
-        % \Function{ForwardPass}{$X,U,K,d,\Delta V,J$} 
-        %     \State Initialize $\bar{x}_0 = x_0$, $\alpha = 1$, $J^{\small{-}} \leftarrow J$
-        %     \For{k=0:1:N-1}
-        %         \State {\color{blue} $\delta x \leftarrow$ Using \eqref{eq:state_error}}
-        %         \State $\bar{u}_k = u_k + K_k \delta x + \alpha d_k$
-        %         \State $\bar{x}_{k+1} \leftarrow f(\bar{x}_k, \bar{u}_k)$
-        %     \EndFor
-        %     \State $J \leftarrow$ Calculate cost using $\traj{x},\traj{u}$ 
-        %     \If{$J$ {\color{blue} satisfies line search conditions}}
-        %         \State $x_{0:N} \leftarrow \bar{x}_{0:N}, u_{0:N-1} \leftarrow \bar{u}_{0:N-1}$
-        %     \Else
-        %         \State Reduce $\alpha$ and go to line 3
-        %     \EndIf\\
-        %     \Return $X,U,J$
-        % \EndFunction
-        % \end{algorithmic}
-        % \end{algorithm}
+        The only other modification to the algorithm is during the iLQR ``forward pass''
+        that simulates the system with the closed-loop MLQR controller, where we use
+        \eqref{eq:state_error} to calculate the nonlinear feedback policy during the
+        forward roll-outs.
+        For robustness, we may also add an additional criterion to the line search that
+        checks for singularities in the three-parameter error state which, for the Cayley
+        map, occur at \ang{180}, at which point the step size should be reduced.
 
     \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
-        In addition to the straight-forward modifications to the LQR algorithm itself, we need to carefully consider the types of cost functions we minimize.
-We frequently minimize costs that penalize distance from a goal state, e.g.
-$\half (x-x_g)^T Q (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined.
-We propose two different cost functions that accomplish similar behavior.
-        For sake of clarity and space, we only consider the costs on the quaternion variables: costs on the other states and the control variables remain unaffected.
+        In addition to the straight-forward modifications to the LQR algorithm itself, we
+        need to carefully consider the types of cost functions we minimize. We frequently
+        minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
+        (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We
+        propose two different cost functions that accomplish similar behavior.
+        For sake of clarity and space, we only consider the costs on the quaternion
+        variables: costs on the other states and the control variables remain unaffected.
 
         \subsubsection{Error Quadratic}
-            Rather than simple subtraction, we can use a quadratic function on the three-parameter error state \eqref{eq:state_error}:
+            Rather than simple subtraction, we can use a quadratic function on the
+            three-parameter error state \eqref{eq:state_error}:
             \begin{equation} \label{eq:error_quadratic}
-                J_\text{err} = \half \phi^T Q \phi = \half \left(\varphi^{-1}(\delta q)\right)^T Q \left(\varphi^{-1}(\delta q)\right).
+                J_\text{err} = \half \phi^T Q \phi 
+                = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
+                        \left(\varphi^{-1}(\delta q)\right).
             \end{equation}
-            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient and Hessian of \eqref{eq:error_quadratic} are 
+            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
+            and Hessian of \eqref{eq:error_quadratic} are
             \begin{equation}
@@
                       + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
-                    %  &+ G(\delta q)^T \nabla D \left(\delta q, Q \phi \right) G(\delta q) \\
                 \end{aligned}
@@
                 \nabla D = \pdv{q}(D(q)^T Q \phi) 
-                = -\frac{1}{q_s^2} \begin{bmatrix} -2 \frac{q_v}{q_s}^T Q \phi \;\;  & \phi^T Q \\
-                                   Q \phi \;\; & 0 \\
+                = -\frac{1}{q_s^2} \begin{bmatrix} 
+                    -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
+                                   Q \phi & 0 \\
                 \end{bmatrix}.
@@
 
-            % where $\delta q \in \R^4 = L(q_g)^T q$, $\psi \in \R^3 = \varphi(\delta q)$, and $\nabla \varphi^{-1}(q) \in \R^{3 \times 4} \equiv \pdv*{\varphi^{-1}}{q}$.
-
-            % \begin{equation}
-            %     \nabla_{x} = \psi^T Q \pdv{\varphi^{-1}}{q} L(\delta q) \pdv{\varphi}{\phi}
-            % \end{equation}
-
-            % \noindent We can obtain the Hessian through the chain rule:
-            % \begin{equation}
-            %     \begin{aligned}
-            %          \nabla^2_{xx} &= \pdv{\varphi}{\phi}^T L(\delta q)^T \left(\pdv{\varphi^{-1}}{q}\right)^T Q \left(\pdv{\varphi^{-1}}{q}\right) L(\delta q) \pdv{\varphi}{\phi} \\
-            %          &+ \pdv{\varphi}{\phi}^T L(\delta q)^T 
-            %          \nabla^2 \varphi^{-1} \left(\delta q, Q \psi \right) L(\delta q) \pdv{\varphi}{\phi} \\
-            %          &+ \nabla^2 \varphi \left(\delta q, L(\delta q)^T \pdv{\varphi^{-1}}{q} Q \psi \right)
-            %     \end{aligned}
-            % \end{equation}
-            % Evaluating at $\phi = 0$ we get the gradient
-
-        \subsubsection{Geodesic Distance}
-            Alternatively, we can use the geodesic distance between two quaternions \cite{Kuffner2004},
-            \begin{equation} \label{eq:quat_geodesic}
-                J_\text{geo} = (1-\abs{q_g^T q}) ,
-            \end{equation}
-            whose gradient and Hessian are,
-            \begin{align}
-                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
-                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
-            \end{align}
-            where the sign of the Hessian corresponds to the sign of $q_d^T q$.
-
-\section{Experiments} \label{sec:experiments}
-    Code code for all experiments is available on GitHub at \url{\texttt{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}.
-
-    \subsection{Trajectory Optimization}
-        In this section we present several trajectory optimization problems solved using our modified version of ALTRO. 
-        The two quadrotor examples use 101 knot points for 5 second trajectories, and the airplane example uses 51 knot points for a 1.25 second trajectory.
-All results were run on a desktop computer with an Intel i9-9900 3.10 GHz processor with 32GB of memory.
-Timing results for all three experiments are given in Fig.
-\ref{fig:timing_chart}.
-
-        In all examples, MLQR is implemented with a unit quaternion attitude state and Rodrigues parameters as the error state.
-The first example compares the two cost functions given in Sec.
-\ref{sec:cost_functions}, while the last two use the geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
-\ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention from \cite{michael2010grasp}, and quaternions (treating them na\"ively as vectors).
-All examples re-normalize the quaternion in the dynamics function and use a third-order explicit Runge-Kutta integrator.
-
-        \begin{figure}
-            \centering
-            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
-            \caption{Timing results for the airplane barrell roll, quadrotor flip, and quadrotor zig-zag examples using iLQR with roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR.
-The timing result for the Quadrotor flip with Euler angles is omitted since it failed to converge.}
-            \label{fig:timing_chart}
-        \end{figure}
-
-        \subsubsection{Quadrotor Zig-Zag}
-            \begin{figure}[t]
-                \centering
-                \begin{subfigure}[t]{0.55\textwidth}
-                    \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
-                    \caption{Zig-zag trajectory.
-The black quadrotors indicate the reference trajectory generated by iMLQR.
-The yellow quadrotors track the reference trajectory using time-varying MLQR, while the red uses the HFCA controller from \cite{watterson2020control}.}
-                    \label{fig:zig-zag}
-                \end{subfigure}
-                ~
-                \begin{subfigure}[t]{0.40\textwidth}
-                    \centering
-                    \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
-                    \caption{Snapshots of the quadrotor flip trajectory. The red-colored quadrotors represent the state near t=0 s and the green-colored quadrotors represent the state near t=5.0 s}
-                    \label{fig:quad_flip}
-                \end{subfigure}
-                \caption{Examples applying MLQR to trajectory optimization on quadrotors}
-            \end{figure}
-
-        % \begin{figure}[t]
-        %     \begin{minipage}{0.69\linewidth}
-        %         \centering
-        %         \includegraphics[height=7cm]{figures/zig_zag_tracking.png}
-        %         \caption{Zig-zag trajectory.
-The black quadrotors indicate the reference trajectory generated by iMLQR.
-The yellow quadrotors track the reference trajectory using time-varying MLQR, while the red uses the SE(3) controller from \cite{lee2010geometric}}
-        %         \label{fig:zig-zag}
-        %     \end{minipage}
-        %     \begin{minipage}{0.3\linewidth}
-        %         \centering
-        %         \begin{tabular}{cc}
-        %             \toprule
-        %             \textbf{Method} & \textbf{Success} \\
-        %             \midrule
-        %             Quat & 62\% \\
-        %             RPY  & 79\% \\
-        %             HFCA & 97\% \\
-        %             MLQR & 98\% \\
-        %             \bottomrule
-        %         \end{tabular}
-        %         \captionsetup{type=table}
-        %         \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization}
-        %         \label{tab:mlqr_success}
-        %     \end{minipage}
-        % \end{figure}
-
-        We set up a fairly simple trajectory optimization problem to achieve a ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of the form 
-        \begin{equation}
-                  \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
-                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
-                + \half \sum_{k=0} u_k^T R u_k
-        \end{equation}
-        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and $(0,10,0)$.
-We compared the two cost functions from Sec.
-\ref{sec:cost_functions}, using a standard quadratic penalty on the non-quaternion states.
-        \begin{align}
-            \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
-            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1 \pm \bar{q}^T q)
-        \end{align}
-        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular velocity.
-        We used the following weighting matrices: 
-        $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
-        $Q_w = \text{diag}(10^3 I_3,I_9)$, 
-        $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
-        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and terminal cost functions, respectively.
-
-        The geodesic cost function performed significantly better than the error quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29 iterations. 
-        In addition to better convergence behavior, the geodesic cost is about half as expensive to compute, which is unsurprising when comparing the gradients and Hessians in Sec.
-\ref{sec:cost_functions}.
-
-        \subsubsection{Quadrotor Flip}
-
-	    Similar to the previous example, we use a set of waypoints to encourage the quadrotor to do a flip. The cost weights for this example were 
-	    \begin{align*}
-	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \times 10^{-2}, 10^{-3} I_3, 10^{-3} I_3, 10^{-2} I_3) \\
-	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
-	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
-	    \end{align*}
-	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles of \ang{90},  \ang{180}, \ang{270}, and \ang{360}.
-The quadrotor was constrained to stay above the floor and move to a goal state 2 meters away in the $+y$ direction.
-The solver is initialized with a dynamically infeasible trajectory that linearly interpolates between the initial and final states, rotating the quad around the x-axis a full \ang{360}.
-
-	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using iMLQR within the ALTRO solver to handle constraints.
-The trajectory generated by iLQR with quaternions and quadratic costs was successful but had a slightly different trajectory.
-iLQR with Euler angles failed to solve the problem.
-        % \begin{figure}[t]
-        %     \centering
-        %     \includegraphics[height=6cm]{figures/quad_flip.png}
-        %     \caption{Snapshots of the quadrotor flip trajectory. The red-colored quadrotors represent the state near t=0 s and the green-colored quadrotors represent the state near t=5.0 s}
-        %     \label{fig:quad_flip}
-        % \end{figure}
-
-        \subsection{Airplane Barrel Roll}
-        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is tasked to do a barrel roll by setting a high terminal cost for being upside-down, see Fig.
-\ref{fig:barrellroll}.
-        The solver is initialized with level flight trim conditions. 
-        For both the quadrotor flip and the airplane barrel roll, iMLQR converged faster than the pure quaternion version.
-Despite the extra matrix multiplications and extra Hessian term, it also was faster per iteration than its iLQR counterpart.
-For these highly aerobatic maneuvers, we achieve, as expected, better performance by correctly leveraging the structure of the rotation group during the optimization routine.
-
-        \begin{figure}[h]
-            \centering
-            \includegraphics[height=3.5cm]{figures/barrellroll.png}
-            \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
-            \label{fig:barrellroll}
-        \end{figure}
-
-	    \begin{figure}[h]
-	        \centering
-	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
-	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller.
-Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics.
-}
-	        \label{fig:tracking_err}
-	    \end{figure}
-	    \vspace{-20pt}
-
-	\subsection{Tracking Control}
-	   % \begin{wrapfigure}{r}{0.5\textwidth}
-	   %     \centering
-	   %     \includegraphics{figures/mlqr_basin.tikz}
-	   %     \caption{Basin of attraction}
-	   %     \label{fig:mlqr_basin}
-	   % \end{wrapfigure}
-
-	    This section provides examples using MLQR for tracking and stabilizing control for a quadrotor.
-We compare our MLQR controller to HFCA, a state-of-the-art geometric tracking controller \cite{watterson2020control} based on the Hopf Fibration.
-
-	    \subsubsection{MLQR Stabilization}
-	    To test robustness and the domain of attraction of the MLQR controller, a quadrotor is linearized about hover at the origin and initialized from states uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular velocities, and uniformly over the entire space of rotations.
-The success rates for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is defined as the terminal error (calculated using \eqref{eq:state_error}) being within a unit ball of radius 0.1 after 10.0 seconds.
-	    To further analyze the basins of attraction, we uniformly sampled over only roll (which goes singular at \ang{90} for our Euler angle representation), and yaw.
-The results are shown in Fig.
-\ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the successfully stabilized initial conditions.
-
-	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also being more generally applicable, easier to tune, and easier to implement.
-It is also noteworthy that na\"ively using LQR with quaternions (incorrectly treating them as vectors) results in poorer performance than Euler angles, clearly demonstrating the value of the methodology presented in the current paper.
-
-	    \subsubsection{MLQR Tracking}
-	    Time-varying MLQR is used to track the zig-zag trajectory generated in the previous section.
-%The controller calculated the gains at $0.01 s$ intervals, and used zero-order-hold between intervals.
-	    To test tracking performance, zero-mean Gaussian noise was added as force and moments applied to the quadrotor.
-
-	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
-\ref{fig:zig-zag}, and \ref{fig:tracking_err}.
-While the MLQR controller dramatically out-performs HFCA, we found no significant difference in performance between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
-Since the tracking controller only deals with relatively small state and attitude errors, this is unsurprising.
-Additionally, we found the MLQR controller ran about 20x faster than the HFCA controller.
-
-	    \begin{figure}[h]
-	        \begin{minipage}{0.60\linewidth}
-    	        \centering
-    	        \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
-    	        \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black).
-All points within each contour were successfully stabilized by the controller.}
-    	        \label{fig:mlqr_basin}
-	        \end{minipage}
-	        \hspace{0.05\linewidth}
-            \begin{minipage}{0.34\linewidth}
-                \centering
-                \begin{tabular}{cc}
-                    \toprule
-                    \textbf{Method} & \textbf{Success} \\
-                    \midrule
-                    Quat & 61\% \\
-                    RPY  & 80\% \\
-                    HFCA & 97\% \\
-                    MLQR & 99\% \\
-                    \bottomrule
-                \end{tabular}
-                \captionsetup{type=table}
-                \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
-                \label{tab:mlqr_success}
-            \end{minipage}
-	    \end{figure}
-
-\section{Conclusions} \label{sec:conclusion}
-    We have presented a general, unified method for planning and control on rigid-body systems with arbitrary attitude using standard linear algebra and vector calculus.
-By applying these methods to LQR to correctly account for the group structure of rotations, we have matched or exceeded the performance of a state-of-the-art geometric controller designed specifically for quadrotors, while being more general, requiring less system-dependent tuning, having less computational overhead, and being easier to implement.
-
-    We have also demonstrated a straight-forward way to adapt nonlinear trajectory optimization techniques to work with quaternion-valued states.
-    For these problems, we found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be computationally efficient and provide excellent convergence in practice.
-We recommend the use of unit quaternions within planning and control algorithms for their simplicity, computational efficiency, and lack of singularities.
-We further recommend the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
-
-    Future areas of research include the application of these methodologies to more complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as more in-depth analysis of the convergence behavior of the algorithms proposed in the current work.
-
-\paragraph{Acknowledgements}
-This material is based upon work
-supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE-1656518.
-Any opinions, findings, and conclusions or recommendations
-expressed in this material are those of the author(s) and
-do not necessarily reflect the views of the National Science
-Foundation
-
-\printbibliography
 \end{document}

### 2020-09-09 paper/iros_abstract.tex
IEEE formatted paper

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 \maketitle
+
+\begin{abstract}
+Planning and controlling trajectories for floating-base robotic systems that experience
+large attitude changes is challenging due to the nontrivial group structure of 3D
+rotations. This paper introduces an accessible and unified approach for tracking control
+and optimization-based planning on the space of rotations based on vector calculus and
+linear algebra. The methodology is used to derive an extension of the Linear-Quadratic
+Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR
+(MLQR). We compare MLQR to a state-of-the-art geometric tracking controller designed
+specifically for quadrotors and demonstrate superior tracking and robustness. An
+iterative variant of MLQR is compared to existing methods on useveral constrained
+trajectory optimization problems, again demonstrating improved runtime performance. %to
+optimize trajectories for a variety of robotic systems, we provide benchmark comparisons
+between several of the most common attitude representations used for motion planning and
+control and find that the combination of unit quaternion state variables with Rodrigues
+parameters to represent attitude errors provides an excellent combination of performance
+and algorithmic simplicity.
+\keywords{
+    motion planning and control, quaternions, optimal control, linear quadratic regulator
+}
+\end{abstract}
 
@@
             \end{equation}
-            \begin{equation}
-                \begin{aligned}
-                     \nabla^2 J_\text{err} &= 
+            \begin{multline}
+                \nabla^2 J_\text{err} = 
                     G(\delta q)^T \! \left(
-                        D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q) 
+                    D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
                     + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
-                \end{aligned}
-            \end{equation}
+            \end{multline}
             where, for the Cayley map,
@@
 
+        \subsubsection{Geodesic Distance}
+            Alternatively, we can use the geodesic distance between two quaternions
+            \cite{Kuffner2004},
+            \begin{equation} \label{eq:quat_geodesic}
+                J_\text{geo} = (1-\abs{q_g^T q}) ,
+            \end{equation}
+            whose gradient and Hessian are,
+            \begin{align}
+                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
+                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
+            \end{align}
+            where the sign of the Hessian corresponds to the sign of $q_d^T q$.
+
+\section{Experiments} \label{sec:experiments}
+    Code for all experiments is available on 
+    \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
+    {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.
+
+    \subsection{Trajectory Optimization}
+        In this section we present several trajectory optimization problems solved using
+        our modified version of ALTRO. The two quadrotor examples use 101 knot points for
+        5 second trajectories, and the airplane example uses 51 knot points for a 1.25
+        second trajectory. All results were run on a desktop computer with an Intel
+        i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
+        experiments are given in Fig. \ref{fig:timing_chart}.
+
+        In all examples, MLQR is implemented with a unit quaternion attitude state and
+        Rodrigues parameters as the error state. The first example compares the two cost
+        functions given in Sec. \ref{sec:cost_functions}, while the last two use the
+        geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
+        \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
+        from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
+        vectors). All examples re-normalize the quaternion in the dynamics function and
+        use a third-order explicit Runge-Kutta integrator.
+
+        \begin{figure}
+            \centering
+            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
+            \caption{Timing results for the airplane barrell roll, quadrotor
+                flip, and quadrotor zig-zag examples using iLQR with
+                roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The
+                timing result for the Quadrotor flip with Euler angles is omitted
+                since it failed to converge.
+            }
+            \label{fig:timing_chart}
+        \end{figure}
+
+        \subsubsection{Quadrotor Zig-Zag}
+            \begin{figure}[t]
+                \centering
+                \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
+                \caption{Zig-zag trajectory. The black quadrotors indicate
+                    the reference trajectory generated by iMLQR. The yellow
+                    quadrotors track the reference trajectory using time-varying
+                    MLQR, while the red uses the HFCA controller from
+                    \cite{watterson2020control}.
+                }
+                \label{fig:zig-zag}
+            \end{figure}    
+            \begin{figure}[t]
+                \centering
+                \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
+                \caption{Snapshots of the quadrotor flip trajectory. The
+                    red-colored quadrotors represent the state near t=0 s and the
+                    green-colored quadrotors represent the state near t=5.0 s
+                }
+                \label{fig:quad_flip}
+            \end{figure}
+
+        We set up a fairly simple trajectory optimization problem to achieve a
+        ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of
+        the form
+        \begin{multline}
+                  \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
+                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) \\
+                + \half \sum_{k=1} u_k^T R u_k
+        \end{multline}
+        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus
+        \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and
+        $(0,10,0)$. We compared the two cost functions from Sec.
+        \ref{sec:cost_functions}, using a standard quadratic penalty on the
+        non-quaternion states.
+        \begin{align}
+            \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
+            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
+                \pm \bar{q}^T q)
+        \end{align}
+        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where
+        $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular
+        velocity.
+        We used the following weighting matrices: 
+        $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
+        $Q_w = \text{diag}(10^3 I_3,I_9)$, 
+        $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
+        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and
+        terminal cost functions, respectively.
+
+        The geodesic cost function performed significantly better than the error
+        quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29
+        iterations. In addition to better convergence behavior, the geodesic cost is
+        about half as expensive to compute, which is unsurprising when comparing the
+        gradients and Hessians in Sec. \ref{sec:cost_functions}.
+
+        \subsubsection{Quadrotor Flip}
+	    Similar to the previous example, we use a set of waypoints to encourage the
+	    quadrotor to do a flip. The cost weights for this example were
+	    \begin{align*}
+	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \!\times\! 10^{-2}, 10^{-3} I_3,
+    	        10^{-3} I_3, 10^{-2} I_3) \\
+	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
+	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
+	    \end{align*}
+	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
+	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
+	    to stay above the floor and move to a goal state 2 meters away in the $+y$
+	    direction. The solver is initialized with a dynamically infeasible trajectory
+	    that linearly interpolates between the initial and final states, rotating the
+	    quad around the x-axis a full \ang{360}.
+
+	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
+	    iMLQR within the ALTRO solver to handle constraints. The trajectory generated by
+	    iLQR with quaternions and quadratic costs was successful but had a slightly
+        different trajectory. iLQR with Euler angles failed to solve the problem.
+
+        \subsection{Airplane Barrel Roll}
+        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
+        tasked to do a barrel roll by setting a high terminal cost for being upside-down,
+        see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
+        conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
+        converged faster than the pure quaternion version. Despite the extra matrix
+        multiplications and extra Hessian term, it also was faster per iteration than its
+        iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
+        better performance by correctly leveraging the structure of the rotation group
+        during the optimization routine.
+
+        \begin{figure}[h]
+            \centering
+            \includegraphics[height=3.5cm]{figures/barrellroll.png}
+            \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
+            \label{fig:barrellroll}
+        \end{figure}
+
+	    \begin{figure}[h]
+	        \centering
+	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
+	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller.
+Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics.
+}
+	        \label{fig:tracking_err}
+	    \end{figure}
+        \vspace{-20pt}
+
+	\subsection{Tracking Control}
+	    This section provides examples using MLQR for tracking and stabilizing control
+	    for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art
+	    geometric tracking controller \cite{watterson2020control} based on the Hopf
+	    Fibration.
+
+	    \subsubsection{MLQR Stabilization}
+	    To test robustness and the domain of attraction of the MLQR controller, a
+	    quadrotor is linearized about hover at the origin and initialized from states
+	    uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular
+	    velocities, and uniformly over the entire space of rotations. The success rates
+	    for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is
+	    defined as the terminal error (calculated using \eqref{eq:state_error}) being
+	    within a unit ball of radius 0.1 after 10.0 seconds. To further analyze the
+	    basins of attraction, we uniformly sampled over only roll (which goes singular at
+	    \ang{90} for our Euler angle representation), and yaw. The results are shown in
+	    Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the
+	    successfully stabilized initial conditions.
+
+	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also
+	    being more generally applicable, easier to tune, and easier to implement. It is
+	    also noteworthy that na\"ively using LQR with quaternions (incorrectly treating
+	    them as vectors) results in poorer performance than Euler angles, clearly
+	    demonstrating the value of the methodology presented in the current paper.
+
+	    \subsubsection{MLQR Tracking}
+	    Time-varying MLQR is used to track the zig-zag trajectory generated in the
+	    previous section. %The controller calculated the gains at $0.01 s$ intervals, and
+	    used zero-order-hold between intervals. To test tracking performance, zero-mean
+	    Gaussian noise was added as force and moments applied to the quadrotor.
+
+	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
+	    \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller
+	    dramatically out-performs HFCA, we found no significant difference in performance
+	    between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
+	    Since the tracking controller only deals with relatively small state and attitude
+	    errors, this is unsurprising. Additionally, we found the MLQR controller ran
+        about 20x faster than the HFCA controller.
+
+	    \begin{figure*}[ht]
+	        \begin{minipage}{0.60\linewidth}
+                \centering
+                \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
+                \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black).
+All points within each contour were successfully stabilized by the controller.}
+                \label{fig:mlqr_basin}
+            \end{minipage}
+	        \hspace{0.05\linewidth}
+            \begin{minipage}{0.34\linewidth}
+                \centering
+                \begin{tabular}{cc}
+                    \toprule
+                    \textbf{Method} & \textbf{Success} \\
+                    \midrule
+                    Quat & 61\% \\
+                    RPY  & 80\% \\
+                    HFCA & 97\% \\
+                    MLQR & 99\% \\
+                    \bottomrule
+                \end{tabular}
+                \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
+                \label{tab:mlqr_success}
+            \end{minipage}
+        \end{figure*}
+
+\section{Conclusions} \label{sec:conclusion}
+    We have presented a general, unified method for planning and control on rigid-body
+    systems with arbitrary attitude using standard linear algebra and vector calculus. By
+    applying these methods to LQR to correctly account for the group structure of
+    rotations, we have matched or exceeded the performance of a state-of-the-art
+    geometric controller designed specifically for quadrotors, while being more general,
+    requiring less system-dependent tuning, having less computational overhead, and being
+    easier to implement.
+
+    We have also demonstrated a straight-forward way to adapt nonlinear trajectory
+    optimization techniques to work with quaternion-valued states. For these problems, we
+    found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
+    computationally efficient and provide excellent convergence in practice. We recommend
+    the use of unit quaternions within planning and control algorithms for their
+    simplicity, computational efficiency, and lack of singularities. We further recommend
+    the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
+
+    Future areas of research include the application of these methodologies to more
+    complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
+    more in-depth analysis of the convergence behavior of the algorithms proposed in the
+    current work.
+
+\paragraph{Acknowledgements}
+This material is based upon work supported by the National Science Foundation Graduate
+Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
+conclusions or recommendations expressed in this material are those of the author(s) and
+do not necessarily reflect the views of the National Science Foundation
+
+\printbibliography
+
 \end{document}

### 2020-09-09 paper/main.tex
Move IEEE version to main.tex

@@
 \usepackage{pifont}  % for x mark
-\usepackage{subcaption}
-\captionsetup{compatibility=false}
-\usepackage{wrapfig}
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 
-%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ BEGIN DOCUMENT ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~%
+\title{\LARGE \bf
+Planning with Attitude
+}
+
+\author{Brian Jackson$^1$ and Zachary Manchester$^1$%
+    \thanks{
+        $^1$Robotics Institute, 
+        Carnegie Mellon University, 
+        5000 Forbes Ave, Pittsburgh, PA, USA
+    }
+}
+
 \begin{document}
-\mainmatter              % start of a contribution
-
-\title{Planning and Control with Attitude}
-\author{Brian Jackson \and Zachary Manchester}
-
-\authorrunning{Jackson and Manchester}
-\institute{Stanford University, 496 Lomita Mall, Stanford, CA, USA,\\
-\email{\{bjack205,zacm\}@stanford.edu}}
-
 \maketitle
@@
 \begin{abstract}
-Planning and controlling trajectories for floating-base robotic systems that experience large attitude changes is challenging due to the nontrivial group structure of 3D rotations. 
-This paper introduces an accessible and unified approach for tracking control and optimization-based planning on the space of rotations based on vector calculus and linear algebra.
-The methodology is used to derive an extension of the Linear-Quadratic Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR (MLQR).
-We compare MLQR to a state-of-the-art geometric tracking controller designed specifically for quadrotors and demonstrate superior tracking and robustness. 
-An iterative variant of MLQR is compared to existing methods on useveral constrained trajectory optimization problems, again demonstrating improved runtime performance.
-%to optimize trajectories for a variety of robotic systems, we provide benchmark comparisons between several of the most common attitude representations used for motion planning and control and find that the combination of unit quaternion state variables with Rodrigues parameters to represent attitude errors provides an excellent combination of performance and algorithmic simplicity.
-\keywords{motion planning and control, quaternions, optimal control, linear quadratic regulator}
+Planning and controlling trajectories for floating-base robotic systems that experience
+large attitude changes is challenging due to the nontrivial group structure of 3D
+rotations. This paper introduces an accessible and unified approach for tracking control
+and optimization-based planning on the space of rotations based on vector calculus and
+linear algebra. The methodology is used to derive an extension of the Linear-Quadratic
+Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR
+(MLQR). We compare MLQR to a state-of-the-art geometric tracking controller designed
+specifically for quadrotors and demonstrate superior tracking and robustness. An
+iterative variant of MLQR is compared to existing methods on several constrained
+trajectory optimization problems, again demonstrating improved runtime performance. 
+\keywords{
+    motion planning and control, quaternions, optimal control, linear quadratic regulator
+}
 \end{abstract}
 
-\section{Introduction}
-    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional translations and rotations as part of their normal operation.
-    While representing translations is straightforward and intuitive, effectively representing the nontrivial group structure of 3D rotations has been a topic of study for many decades.
-Although we can intuitively deduce that rotations are three-dimensional, a globally non-singular three-parameter representation of the space of rotations does not exist \cite{stuelpnagel1964parametrization}.
-As a result, when parameterizing rotations, we must either a) pick a three-parameter representation and deal with discontinuities, or b) pick a higher-dimensional representation and deal with constraints between the parameters.
-While simply representing attitude is nontrivial, generating and tracking motion plans for floating-base systems is an even more challenging problem.
+\section{INTRODUCTION}
+    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous 
+    underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional 
+    translations and rotations as part of their normal operation. 
+    While representing translations is straightforward and intuitive, effectively 
+    representing the nontrivial group structure of 3D rotations has been a topic of study 
+    for many decades. Although we can intuitively deduce that rotations are three-dimensional, 
+    a globally non-singular three-parameter representation of the space of rotations does 
+    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing rotations, 
+    we must either a) pick a three-parameter representation and deal with discontinuities, 
+    or b) pick a higher-dimensional representation and deal with constraints between the 
+    parameters. While simply representing attitude is nontrivial, generating and tracking 
+    motion plans for floating-base systems is an even more challenging problem.
 
-    Early work on control problems involving the rotation group dates back to the 1970s, with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ \cite{Baillieul1978}.
-Effective attitude tracking controllers have been developed for satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude errors.
+    Early work on control problems involving the rotation group dates back to the 1970s, 
+    with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ 
+    \cite{Baillieul1978}. Effective attitude tracking controllers have been developed for 
+    satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,
+    Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum 
+    \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude 
+    errors.
 
-    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control.
-Approaches to optimal control on attitude problems include analytical methods applied to satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,Lee2008}, a combination of sampling-based planning and constrained trajectory optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators \cite{Saccon2013}, or more general theory for optimization on manifolds \cites{watterson2018trajectory}.
-Nearly all of these methods rely heavily on principles from differential geometry and Lie group theory; however, despite these works, many recent papers in the robotics community continue to apply traditional methods for motion planning and control with no regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,Williams2017,Geisert2016}.
+    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, 
+    sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control. 
+    Approaches to optimal control on attitude problems include analytical methods applied to
+    satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,
+    Lee2008}, a combination of sampling-based planning and constrained trajectory 
+    optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators 
+    \cite{Saccon2013}, or more general theory for optimization on manifolds 
+    \cites{watterson2018trajectory}. 
+    Nearly all of these methods rely heavily on principles from differential geometry and 
+    Lie group theory; however, despite these works, many recent papers in the robotics 
+    community continue to apply traditional methods for motion planning and control with no 
+    regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
+    Williams2017,Geisert2016}.
 
-    In this paper, we make a departure from previous approaches to geometric planning and control that rely heavily on ideas and notation from differential geometry, and instead use only basic mathematical tools from linear algebra and calculus that should be familiar to most roboticists.
-Similar to \cite{Mandic2011,Xu2016}, in Sec.
-\ref{sec:Quaternion_Calculus} we introduce a quaternion differential calculus, but take a significantly simpler and more general approach that unifies both planning and control, enabling straight-forward adaptation of existing algorithms to systems with quaternion states.
-To make this concrete, in Sec.
-\ref{sec:MLQR} we apply our approach to derive an extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), which is the control dual to the multiplicative extended Kalman filter (MEKF) from the state estimation literature \cite{markley2014fundamentals}.
-In Sec.
-\ref{sec:experiments} we provide several simulation results demonstrating the application of our method to tracking control and constrained trajectory optimization.
+    In this paper, we make a departure from previous approaches to geometric planning and 
+    control that rely heavily on ideas and notation from differential geometry, 
+    and instead use only basic mathematical tools from linear algebra and calculus that 
+    should be familiar to most roboticists. 
+    Similar to \cite{Mandic2011,Xu2016}, in Sec. \ref{sec:Quaternion_Calculus} we introduce 
+    a quaternion differential calculus, but take a significantly simpler and more general 
+    approach that unifies both planning and control, enabling straight-forward adaptation of 
+    existing algorithms to systems with quaternion states. 
+    To make this concrete, in Sec. \ref{sec:MLQR} we apply our approach to derive an 
+    extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), 
+    which is the control dual to the multiplicative extended Kalman filter (MEKF) from the 
+    state estimation literature \cite{markley2014fundamentals}. 
+    In Sec. \ref{sec:experiments} we provide several simulation results demonstrating the 
+    application of our method to tracking control and constrained trajectory optimization.
     In summary, our contributions include:
-    %\vspace{-10pt}
+
     \begin{itemize}
@@
 
-        Derivatives consider the effect an infinitesimal perturbation to the input has on an infinitesimal perturbation to the output.
-For vector spaces, the composition of the perturbation with the nominal value is simple addition and the infinitesimal perturbation lives in the same space as the original vector.
-For unit quaternions, however, neither of these are true; instead, they compose according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig.
+        Derivatives consider the effect an infinitesimal perturbation to the input has on
+        an infinitesimal perturbation to the output. For vector spaces, the composition
+        of the perturbation with the nominal value is simple addition and the
+        infinitesimal perturbation lives in the same space as the original vector. For
+        unit quaternions, however, neither of these are true; instead, they compose
+        according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to
+        first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig.
         \ref{fig:tangent_plane}).
-        %, and are related through a non-linear mapping $\varphi : \R^3 \to \Q$ )
-        The fact that differential unit quaternions are three-dimensional should make intuitive sense: Rotations are inherently three-dimensional and differential rotations should live in the same space as angular velocity, i.e.
-$\R^3$.
 
-        There are many possible three-parameter representations for small rotations in the literature.
-Many authors use the exponential map \cite{Baillieul1978,Zefran1998,Lee2008,Saccon2013,Sola2017,Fan2016,watterson2018trajectory}, while others have used the Cayley map (also known as Rodrigues parameters) \cite{Kobilarov2011,Kobilarov2014}, Modified Rodrigues Parameters (MRPs) \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}.
-        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are computationally efficient and do not inherit the sign ambiguity associated with unit quaternions.
-The mapping between a vector of Rodrigues parameters $\phi \in \R^3$ and a unit quaternion $q$ is known as the Cayley map:
-        \begin{equation} \label{eq:cayley}
+        The fact that differential unit quaternions are three-dimensional should make
+        intuitive sense: Rotations are inherently three-dimensional and differential
+        rotations should live in the same space as angular velocity, i.e. $\R^3$.
+
+        There are many possible three-parameter representations for small rotations in
+        the literature. Many authors use the exponential map \cite{Baillieul1978,
+        Zefran1998, Lee2008, Saccon2013, Sola2017, Fan2016, watterson2018trajectory},
+        while others have used the Cayley map (also known as Rodrigues parameters)
+        \cite{Kobilarov2011, Kobilarov2014}, Modified Rodrigues Parameters (MRPs)
+        \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}.
+        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are
+        computationally efficient and do not inherit the sign ambiguity associated with
+        unit quaternions. The mapping between a vector of Rodrigues parameters $\phi \in
+        \R^3$ and a unit quaternion $q$ is known as the Cayley map: \begin{equation}
+        \label{eq:cayley}
             q = \varphi(\phi) = \frac{1}{\sqrt{1 + \norm{\phi}^2}} \begin{bmatrix} 1 \\ \phi \end{bmatrix}.
@@
             \end{equation}
-            \begin{equation}
-                \begin{aligned}
-                     \nabla^2 J_\text{err} &= 
+            \begin{multline}
+                \nabla^2 J_\text{err} = 
                     G(\delta q)^T \! \left(
-                        D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q) 
+                    D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
                     + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
-                    %  &+ G(\delta q)^T \nabla D \left(\delta q, Q \phi \right) G(\delta q) \\
-                \end{aligned}
-            \end{equation}
+            \end{multline}
             where, for the Cayley map,
@@
 
-            % where $\delta q \in \R^4 = L(q_g)^T q$, $\psi \in \R^3 = \varphi(\delta q)$, and $\nabla \varphi^{-1}(q) \in \R^{3 \times 4} \equiv \pdv*{\varphi^{-1}}{q}$.
-
-            % \begin{equation}
-            %     \nabla_{x} = \psi^T Q \pdv{\varphi^{-1}}{q} L(\delta q) \pdv{\varphi}{\phi}
-            % \end{equation}
-
-            % \noindent We can obtain the Hessian through the chain rule:
-            % \begin{equation}
-            %     \begin{aligned}
-            %          \nabla^2_{xx} &= \pdv{\varphi}{\phi}^T L(\delta q)^T \left(\pdv{\varphi^{-1}}{q}\right)^T Q \left(\pdv{\varphi^{-1}}{q}\right) L(\delta q) \pdv{\varphi}{\phi} \\
-            %          &+ \pdv{\varphi}{\phi}^T L(\delta q)^T 
-            %          \nabla^2 \varphi^{-1} \left(\delta q, Q \psi \right) L(\delta q) \pdv{\varphi}{\phi} \\
-            %          &+ \nabla^2 \varphi \left(\delta q, L(\delta q)^T \pdv{\varphi^{-1}}{q} Q \psi \right)
-            %     \end{aligned}
-            % \end{equation}
-            % Evaluating at $\phi = 0$ we get the gradient
-
         \subsubsection{Geodesic Distance}
-            Alternatively, we can use the geodesic distance between two quaternions \cite{Kuffner2004},
+            Alternatively, we can use the geodesic distance between two quaternions
+            \cite{Kuffner2004},
             \begin{equation} \label{eq:quat_geodesic}
@@
 \section{Experiments} \label{sec:experiments}
-    Code code for all experiments is available on GitHub at \url{\texttt{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}.
+    Code for all experiments is available on 
+    \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
+    {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.
 
     \subsection{Trajectory Optimization}
-        In this section we present several trajectory optimization problems solved using our modified version of ALTRO. 
-        The two quadrotor examples use 101 knot points for 5 second trajectories, and the airplane example uses 51 knot points for a 1.25 second trajectory.
-All results were run on a desktop computer with an Intel i9-9900 3.10 GHz processor with 32GB of memory.
-Timing results for all three experiments are given in Fig.
-\ref{fig:timing_chart}.
+        In this section we present several trajectory optimization problems solved using
+        our modified version of ALTRO. The two quadrotor examples use 101 knot points for
+        5 second trajectories, and the airplane example uses 51 knot points for a 1.25
+        second trajectory. All results were run on a desktop computer with an Intel
+        i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
+        experiments are given in Fig. \ref{fig:timing_chart}.
 
-        In all examples, MLQR is implemented with a unit quaternion attitude state and Rodrigues parameters as the error state.
-The first example compares the two cost functions given in Sec.
-\ref{sec:cost_functions}, while the last two use the geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
-\ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention from \cite{michael2010grasp}, and quaternions (treating them na\"ively as vectors).
-All examples re-normalize the quaternion in the dynamics function and use a third-order explicit Runge-Kutta integrator.
+        In all examples, MLQR is implemented with a unit quaternion attitude state and
+        Rodrigues parameters as the error state. The first example compares the two cost
+        functions given in Sec. \ref{sec:cost_functions}, while the last two use the
+        geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
+        \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
+        from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
+        vectors). All examples re-normalize the quaternion in the dynamics function and
+        use a third-order explicit Runge-Kutta integrator.
 
@@
             \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
-            \caption{Timing results for the airplane barrell roll, quadrotor flip, and quadrotor zig-zag examples using iLQR with roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR.
-The timing result for the Quadrotor flip with Euler angles is omitted since it failed to converge.}
+            \caption{Timing results for the airplane barrell roll, quadrotor
+                flip, and quadrotor zig-zag examples using iLQR with
+                roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The
+                timing result for the Quadrotor flip with Euler angles is omitted
+                since it failed to converge.
+            }
             \label{fig:timing_chart}
@@
                 \centering
-                \begin{subfigure}[t]{0.55\textwidth}
                 \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
-                    \caption{Zig-zag trajectory.
-The black quadrotors indicate the reference trajectory generated by iMLQR.
-The yellow quadrotors track the reference trajectory using time-varying MLQR, while the red uses the HFCA controller from \cite{watterson2020control}.}
+                \caption{Zig-zag trajectory. The black quadrotors indicate
+                    the reference trajectory generated by iMLQR. The yellow
+                    quadrotors track the reference trajectory using time-varying
+                    MLQR, while the red uses the HFCA controller from
+                    \cite{watterson2020control}.
+                }
                 \label{fig:zig-zag}
-                \end{subfigure}
-                ~
-                \begin{subfigure}[t]{0.40\textwidth}
+            \end{figure}    
+            \begin{figure}[t]
                 \centering
                 \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
-                    \caption{Snapshots of the quadrotor flip trajectory. The red-colored quadrotors represent the state near t=0 s and the green-colored quadrotors represent the state near t=5.0 s}
+                \caption{Snapshots of the quadrotor flip trajectory. The
+                    red-colored quadrotors represent the state near t=0 s and the
+                    green-colored quadrotors represent the state near t=5.0 s
+                }
                 \label{fig:quad_flip}
-                \end{subfigure}
-                \caption{Examples applying MLQR to trajectory optimization on quadrotors}
             \end{figure}
 
-        % \begin{figure}[t]
-        %     \begin{minipage}{0.69\linewidth}
-        %         \centering
-        %         \includegraphics[height=7cm]{figures/zig_zag_tracking.png}
-        %         \caption{Zig-zag trajectory.
-The black quadrotors indicate the reference trajectory generated by iMLQR.
-The yellow quadrotors track the reference trajectory using time-varying MLQR, while the red uses the SE(3) controller from \cite{lee2010geometric}}
-        %         \label{fig:zig-zag}
-        %     \end{minipage}
-        %     \begin{minipage}{0.3\linewidth}
-        %         \centering
-        %         \begin{tabular}{cc}
-        %             \toprule
-        %             \textbf{Method} & \textbf{Success} \\
-        %             \midrule
-        %             Quat & 62\% \\
-        %             RPY  & 79\% \\
-        %             HFCA & 97\% \\
-        %             MLQR & 98\% \\
-        %             \bottomrule
-        %         \end{tabular}
-        %         \captionsetup{type=table}
-        %         \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization}
-        %         \label{tab:mlqr_success}
-        %     \end{minipage}
-        % \end{figure}
-
-        We set up a fairly simple trajectory optimization problem to achieve a ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of the form 
-        \begin{equation}
+        We set up a fairly simple trajectory optimization problem to achieve a
+        ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of
+        the form
+        \begin{multline}
                   \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
-                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
-                + \half \sum_{k=0} u_k^T R u_k
-        \end{equation}
-        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and $(0,10,0)$.
-We compared the two cost functions from Sec.
-\ref{sec:cost_functions}, using a standard quadratic penalty on the non-quaternion states.
+                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) \\
+                + \half \sum_{k=1} u_k^T R u_k
+        \end{multline}
+        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus
+        \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and
+        $(0,10,0)$. We compared the two cost functions from Sec.
+        \ref{sec:cost_functions}, using a standard quadratic penalty on the
+        non-quaternion states.
         \begin{align}
             \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
-            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1 \pm \bar{q}^T q)
+            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
+                \pm \bar{q}^T q)
         \end{align}
-        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular velocity.
+        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where
+        $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular
+        velocity.
         We used the following weighting matrices: 
@@
         $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
-        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and terminal cost functions, respectively.
+        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and
+        terminal cost functions, respectively.
 
-        The geodesic cost function performed significantly better than the error quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29 iterations. 
-        In addition to better convergence behavior, the geodesic cost is about half as expensive to compute, which is unsurprising when comparing the gradients and Hessians in Sec.
-\ref{sec:cost_functions}.
+        The geodesic cost function performed significantly better than the error
+        quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29
+        iterations. In addition to better convergence behavior, the geodesic cost is
+        about half as expensive to compute, which is unsurprising when comparing the
+        gradients and Hessians in Sec. \ref{sec:cost_functions}.
 
         \subsubsection{Quadrotor Flip}
-
-	    Similar to the previous example, we use a set of waypoints to encourage the quadrotor to do a flip. The cost weights for this example were 
+	    Similar to the previous example, we use a set of waypoints to encourage the
+	    quadrotor to do a flip. The cost weights for this example were
 	    \begin{align*}
-	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \times 10^{-2}, 10^{-3} I_3, 10^{-3} I_3, 10^{-2} I_3) \\
+	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \!\times\! 10^{-2}, 10^{-3} I_3,
+    	        10^{-3} I_3, 10^{-2} I_3) \\
 	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
@@
 	    \end{align*}
-	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles of \ang{90},  \ang{180}, \ang{270}, and \ang{360}.
-The quadrotor was constrained to stay above the floor and move to a goal state 2 meters away in the $+y$ direction.
-The solver is initialized with a dynamically infeasible trajectory that linearly interpolates between the initial and final states, rotating the quad around the x-axis a full \ang{360}.
+	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
+	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
+	    to stay above the floor and move to a goal state 2 meters away in the $+y$
+	    direction. The solver is initialized with a dynamically infeasible trajectory
+	    that linearly interpolates between the initial and final states, rotating the
+	    quad around the x-axis a full \ang{360}.
 
-	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using iMLQR within the ALTRO solver to handle constraints.
-The trajectory generated by iLQR with quaternions and quadratic costs was successful but had a slightly different trajectory.
-iLQR with Euler angles failed to solve the problem.
-        % \begin{figure}[t]
-        %     \centering
-        %     \includegraphics[height=6cm]{figures/quad_flip.png}
-        %     \caption{Snapshots of the quadrotor flip trajectory. The red-colored quadrotors represent the state near t=0 s and the green-colored quadrotors represent the state near t=5.0 s}
-        %     \label{fig:quad_flip}
-        % \end{figure}
+	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
+	    iMLQR within the ALTRO solver to handle constraints. The trajectory generated by
+	    iLQR with quaternions and quadratic costs was successful but had a slightly
+        different trajectory. iLQR with Euler angles failed to solve the problem.
 
         \subsection{Airplane Barrel Roll}
-        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is tasked to do a barrel roll by setting a high terminal cost for being upside-down, see Fig.
-\ref{fig:barrellroll}.
-        The solver is initialized with level flight trim conditions. 
-        For both the quadrotor flip and the airplane barrel roll, iMLQR converged faster than the pure quaternion version.
-Despite the extra matrix multiplications and extra Hessian term, it also was faster per iteration than its iLQR counterpart.
-For these highly aerobatic maneuvers, we achieve, as expected, better performance by correctly leveraging the structure of the rotation group during the optimization routine.
+        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
+        tasked to do a barrel roll by setting a high terminal cost for being upside-down,
+        see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
+        conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
+        converged faster than the pure quaternion version. Despite the extra matrix
+        multiplications and extra Hessian term, it also was faster per iteration than its
+        iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
+        better performance by correctly leveraging the structure of the rotation group
+        during the optimization routine.
 
@@
 	\subsection{Tracking Control}
-	   % \begin{wrapfigure}{r}{0.5\textwidth}
-	   %     \centering
-	   %     \includegraphics{figures/mlqr_basin.tikz}
-	   %     \caption{Basin of attraction}
-	   %     \label{fig:mlqr_basin}
-	   % \end{wrapfigure}
-
-	    This section provides examples using MLQR for tracking and stabilizing control for a quadrotor.
-We compare our MLQR controller to HFCA, a state-of-the-art geometric tracking controller \cite{watterson2020control} based on the Hopf Fibration.
+	    This section provides examples using MLQR for tracking and stabilizing control
+	    for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art
+	    geometric tracking controller \cite{watterson2020control} based on the Hopf
+	    Fibration.
 
 	    \subsubsection{MLQR Stabilization}
-	    To test robustness and the domain of attraction of the MLQR controller, a quadrotor is linearized about hover at the origin and initialized from states uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular velocities, and uniformly over the entire space of rotations.
-The success rates for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is defined as the terminal error (calculated using \eqref{eq:state_error}) being within a unit ball of radius 0.1 after 10.0 seconds.
-	    To further analyze the basins of attraction, we uniformly sampled over only roll (which goes singular at \ang{90} for our Euler angle representation), and yaw.
-The results are shown in Fig.
-\ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the successfully stabilized initial conditions.
+	    To test robustness and the domain of attraction of the MLQR controller, a
+	    quadrotor is linearized about hover at the origin and initialized from states
+	    uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular
+	    velocities, and uniformly over the entire space of rotations. The success rates
+	    for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is
+	    defined as the terminal error (calculated using \eqref{eq:state_error}) being
+	    within a unit ball of radius 0.1 after 10.0 seconds. To further analyze the
+	    basins of attraction, we uniformly sampled over only roll (which goes singular at
+	    \ang{90} for our Euler angle representation), and yaw. The results are shown in
+	    Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the
+	    successfully stabilized initial conditions.
 
-	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also being more generally applicable, easier to tune, and easier to implement.
-It is also noteworthy that na\"ively using LQR with quaternions (incorrectly treating them as vectors) results in poorer performance than Euler angles, clearly demonstrating the value of the methodology presented in the current paper.
+	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also
+	    being more generally applicable, easier to tune, and easier to implement. It is
+	    also noteworthy that na\"ively using LQR with quaternions (incorrectly treating
+	    them as vectors) results in poorer performance than Euler angles, clearly
+	    demonstrating the value of the methodology presented in the current paper.
 
 	    \subsubsection{MLQR Tracking}
-	    Time-varying MLQR is used to track the zig-zag trajectory generated in the previous section.
-%The controller calculated the gains at $0.01 s$ intervals, and used zero-order-hold between intervals.
-	    To test tracking performance, zero-mean Gaussian noise was added as force and moments applied to the quadrotor.
+	    Time-varying MLQR is used to track the zig-zag trajectory generated in the
+	    previous section. %The controller calculated the gains at $0.01 s$ intervals, and
+	    used zero-order-hold between intervals. To test tracking performance, zero-mean
+	    Gaussian noise was added as force and moments applied to the quadrotor.
 
 	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
-\ref{fig:zig-zag}, and \ref{fig:tracking_err}.
-While the MLQR controller dramatically out-performs HFCA, we found no significant difference in performance between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
-Since the tracking controller only deals with relatively small state and attitude errors, this is unsurprising.
-Additionally, we found the MLQR controller ran about 20x faster than the HFCA controller.
+	    \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller
+	    dramatically out-performs HFCA, we found no significant difference in performance
+	    between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
+	    Since the tracking controller only deals with relatively small state and attitude
+	    errors, this is unsurprising. Additionally, we found the MLQR controller ran
+        about 20x faster than the HFCA controller.
 
-	    \begin{figure}[h]
+	    \begin{figure*}[ht]
 	        \begin{minipage}{0.60\linewidth}
@@
                 \end{tabular}
-                \captionsetup{type=table}
                 \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
@@
             \end{minipage}
-	    \end{figure}
+        \end{figure*}
 
 \section{Conclusions} \label{sec:conclusion}
-    We have presented a general, unified method for planning and control on rigid-body systems with arbitrary attitude using standard linear algebra and vector calculus.
-By applying these methods to LQR to correctly account for the group structure of rotations, we have matched or exceeded the performance of a state-of-the-art geometric controller designed specifically for quadrotors, while being more general, requiring less system-dependent tuning, having less computational overhead, and being easier to implement.
+    We have presented a general, unified method for planning and control on rigid-body
+    systems with arbitrary attitude using standard linear algebra and vector calculus. By
+    applying these methods to LQR to correctly account for the group structure of
+    rotations, we have matched or exceeded the performance of a state-of-the-art
+    geometric controller designed specifically for quadrotors, while being more general,
+    requiring less system-dependent tuning, having less computational overhead, and being
+    easier to implement.
 
-    We have also demonstrated a straight-forward way to adapt nonlinear trajectory optimization techniques to work with quaternion-valued states.
-    For these problems, we found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be computationally efficient and provide excellent convergence in practice.
-We recommend the use of unit quaternions within planning and control algorithms for their simplicity, computational efficiency, and lack of singularities.
-We further recommend the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
+    We have also demonstrated a straight-forward way to adapt nonlinear trajectory
+    optimization techniques to work with quaternion-valued states. For these problems, we
+    found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
+    computationally efficient and provide excellent convergence in practice. We recommend
+    the use of unit quaternions within planning and control algorithms for their
+    simplicity, computational efficiency, and lack of singularities. We further recommend
+    the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
 
-    Future areas of research include the application of these methodologies to more complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as more in-depth analysis of the convergence behavior of the algorithms proposed in the current work.
+    Future areas of research include the application of these methodologies to more
+    complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
+    more in-depth analysis of the convergence behavior of the algorithms proposed in the
+    current work.
 
 \paragraph{Acknowledgements}
-This material is based upon work
-supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE-1656518.
-Any opinions, findings, and conclusions or recommendations
-expressed in this material are those of the author(s) and
-do not necessarily reflect the views of the National Science
-Foundation
+This material is based upon work supported by the National Science Foundation Graduate
+Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
+conclusions or recommendations expressed in this material are those of the author(s) and
+do not necessarily reflect the views of the National Science Foundation
 
 \printbibliography
+
 \end{document}

### 2020-09-10 paper/iros_abstract.tex
First draft of IROS abstract

--- 2020-09-09 paper/iros_abstract.tex
@@
 \section{INTRODUCTION}
+
     Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous
     underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
-    translations and rotations as part of their normal operation. 
-    While representing translations is straightforward and intuitive, effectively 
-    representing the nontrivial group structure of 3D rotations has been a topic of study 
-    for many decades. Although we can intuitively deduce that rotations are three-dimensional, 
-    a globally non-singular three-parameter representation of the space of rotations does 
-    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing rotations, 
-    we must either a) pick a three-parameter representation and deal with discontinuities, 
-    or b) pick a higher-dimensional representation and deal with constraints between the 
-    parameters. While simply representing attitude is nontrivial, generating and tracking 
+    translations and rotations as part of their normal operation. While simply
+    representing the attitude of these sytems is nontrivial, generating and tracking
     motion plans for floating-base systems is an even more challenging problem.
 
-    Early work on control problems involving the rotation group dates back to the 1970s, 
-    with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ 
-    \cite{Baillieul1978}. Effective attitude tracking controllers have been developed for 
-    satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,
-    Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum 
-    \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude 
-    errors.
+    Many approaches have been taken to address the problem of motion planning and control
+    on the $SE(3)$ manifold; most of these rely heavily on ideas and notation
+    from differential geometry. Despite some impressive work in this field, these ideas
+    have not been widely adopted and many continue to ignore the group structure of rotations,
+    especially in the field of optimal control and motion planning. One part of this problem
+    is that accounting for this structure requires low-level changes to the underlying
+    optimization algorithms, which is difficult or impossible to implement when relying on
+    commercial or even open-source optimization solvers.
 
-    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, 
-    sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control. 
-    Approaches to optimal control on attitude problems include analytical methods applied to
-    satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,
-    Lee2008}, a combination of sampling-based planning and constrained trajectory 
-    optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators 
-    \cite{Saccon2013}, or more general theory for optimization on manifolds 
-    \cites{watterson2018trajectory}. 
-    Nearly all of these methods rely heavily on principles from differential geometry and 
-    Lie group theory; however, despite these works, many recent papers in the robotics 
-    community continue to apply traditional methods for motion planning and control with no 
-    regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
-    Williams2017,Geisert2016}.
+    In the current abstract, we formulate a straightforward, generic method for adapting
+    existing gradient-based algorithms to correctly account for the group structure of
+    rotations. Based entirely on basic mathematical tools from linear algebra and
+    calculus, our method is computationally efficient and should be both accessible and
+    familiar to most roboticists. We then apply this method to our ALTRO solver
+    \cite{howell2019altro} and demonstrate its performance on several constrained trajectory
+    optimization problems.
 
-    In this paper, we make a departure from previous approaches to geometric planning and 
-    control that rely heavily on ideas and notation from differential geometry, 
-    and instead use only basic mathematical tools from linear algebra and calculus that 
-    should be familiar to most roboticists. 
-    Similar to \cite{Mandic2011,Xu2016}, in Sec. \ref{sec:Quaternion_Calculus} we introduce 
-    a quaternion differential calculus, but take a significantly simpler and more general 
-    approach that unifies both planning and control, enabling straight-forward adaptation of 
-    existing algorithms to systems with quaternion states. 
-    To make this concrete, in Sec. \ref{sec:MLQR} we apply our approach to derive an 
-    extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), 
-    which is the control dual to the multiplicative extended Kalman filter (MEKF) from the 
-    state estimation literature \cite{markley2014fundamentals}. 
-    In Sec. \ref{sec:experiments} we provide several simulation results demonstrating the 
-    application of our method to tracking control and constrained trajectory optimization.
-    In summary, our contributions include:
-
-    \begin{itemize}
-        \item A unified approach to quaternion differential calculus entirely based on standard vector calculus and linear algebraic operations
-        \item Derivation of multiplicative LQR, an adaptation of LQR to the control of systems with quaternion states
-        \item A set of benchmark motion-planning and trajectory-tracking problems in which we compare our approach to existing methods commonly used in the robotics community to account for 3D rotations
-    \end{itemize}
-
-\section{Background}
+\section{Notation and Convensions}
 
@@
 
-    \subsection{Unit Quaternions} \label{sec:quaternions}
-        We leverage the fact that quaternions are linear operators and that the space of 
-        quaternions $\mathbb{H}$ is isomorphic to $\R^4$ to explicitly 
-        represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as 
+    We represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as 
     a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and 
-        $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively.
+    $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively. We denote
+    the composition of two quaternions as $q_3 = q_2 \otimes q_1$.
 
-        Quaternion multiplication is defined as
-        \begin{equation} \label{eq:quat_mult}
-            \q_2 \otimes \q_1 = L(q_2) q_1 = R(q_1) q_2
-        \end{equation}
-        where $L(q)$ and $R(q)$ are orthonormal matrices defined as
-        \begin{align}
-            L(q) &:= \begin{bmatrix} 
-                q_s \;\; & -q_v^T \\ 
-                q_v \;\; & q_s I + \skewmat{q_v} 
-            \end{bmatrix} 
-            \label{eq:Lmult} \\
-            %= \begin{bmatrix} q & G(q) \end{bmatrix} \label{eq:Lmult} \\
-            R(q) &:=\begin{bmatrix} 
-                q_s \;\; & -q_v^T \\ 
-                q_v \;\; & q_s I - \skewmat{q_v} 
-            \end{bmatrix} \label{eq:Rmult},
-        \end{align}
-        and $\skewmat{x}$ is the skew-symmetric matrix operator
-        \begin{equation}
-            \skewmat{x} := \begin{bmatrix} 
-                0 & -x_3 & x_2 \\ 
-                x_3 & 0 & -x_1\\ 
-                -x_2 & x_1 & 0 
-            \end{bmatrix}.
-        \end{equation}
+        % Quaternion multiplication is defined as
+        % \begin{equation} \label{eq:quat_mult}
+        %     \q_2 \otimes \q_1 = L(q_2) q_1 = R(q_1) q_2
+        % \end{equation}
+        % where $L(q)$ and $R(q)$ are orthonormal matrices defined as
+        % \begin{align}
+        %     L(q) &:= \begin{bmatrix} 
+        %         q_s \;\; & -q_v^T \\ 
+        %         q_v \;\; & q_s I + \skewmat{q_v} 
+        %     \end{bmatrix} 
+        %     \label{eq:Lmult} \\
+        %     %= \begin{bmatrix} q & G(q) \end{bmatrix} \label{eq:Lmult} \\
+        %     R(q) &:=\begin{bmatrix} 
+        %         q_s \;\; & -q_v^T \\ 
+        %         q_v \;\; & q_s I - \skewmat{q_v} 
+        %     \end{bmatrix} \label{eq:Rmult},
+        % \end{align}
+        % and $\skewmat{x}$ is the skew-symmetric matrix operator
+        % \begin{equation}
+        %     \skewmat{x} := \begin{bmatrix} 
+        %         0 & -x_3 & x_2 \\ 
+        %         x_3 & 0 & -x_1\\ 
+        %         -x_2 & x_1 & 0 
+        %     \end{bmatrix}.
+        % \end{equation}
 
-        The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal 
-        to its conjugate $q^*$, which is simply the same quaternion with a negated vector 
-        part:
-        \begin{equation} \label{eq:T}
-            \q^* = T q := \begin{bmatrix} 
-                1 & \\ 
-                & -I_3 
-            \end{bmatrix} q
-        \end{equation}
-        The following identities, which are easily derived from 
-        \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
-        \begin{align}
-            &L(Tq) = L(q)^T = L(q)^{-1} \\
-            &R(Tq) = R(q)^T = R(q)^{-1} .
-        \end{align}
+        % The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal 
+        % to its conjugate $q^*$, which is simply the same quaternion with a negated vector 
+        % part:
+        % \begin{equation} \label{eq:T}
+        %     \q^* = T q := \begin{bmatrix} 
+        %         1 & \\ 
+        %         & -I_3 
+        %     \end{bmatrix} q
+        % \end{equation}
+        % The following identities, which are easily derived from 
+        % \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
+        % \begin{align}
+        %     &L(Tq) = L(q)^T = L(q)^{-1} \\
+        %     &R(Tq) = R(q)^T = R(q)^{-1} .
+        % \end{align}
 
-        We will sometimes find it helpful to create a quaternion with zero scalar part from 
-        a vector $r \in \R^3$. We denote this operation as,
-        \begin{equation}
-            \hat{r} = H r \equiv \begin{bmatrix} 0 \\ I_3 \end{bmatrix} r.
-        \end{equation}
-        Unit quaternions rotate a vector through the operation 
-        $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. 
-        This can be equivalently expressed using matrix multiplication as
-        \begin{align} 
-            r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation}
-        \end{align}
-        where $A(q)$ is the rotation matrix in terms of the elements of the quaternion 
-        \cite{markley2014fundamentals}.
-
-    \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
-        In the current work we will restrict our focus to rigid bodies moving freely in 3D 
-        space. That is, we consider systems with dynamics of the following form:
-        \begin{equation} \label{eq:rigid_body_dynamics}
-            x = \begin{bmatrix} r \\ R \\ v \\ \omega \end{bmatrix}, \quad 
-            \dot{x} = \begin{bmatrix} 
-                v \\ 
-                \texttt{kinematics}(R,\omega) \\ 
-                \frac{1}{m} F_G(x,u) \\ 
-                J^{-1}(\tau_L(x,u) - \omega \times J \omega) 
-            \end{bmatrix}
-        \end{equation}
-        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, 
-        $R \in SO(3)$ is the attitude, $v \in \R^3$ is the linear velocity, and 
-        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, 
-        $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the 
-        global frame, $\tau_L(x,u)$ are the moments in the local (body) frame, and 
-        $\texttt{kinematics}$ are the kinematics for the chosen attitude representation. The 
-        kinematics for unit quaternions are
-        \begin{equation}
-            \dot{q} = \half \q \otimes \hat{\omega} = \half L(q) H \omega.
-        \end{equation}
+        % We will sometimes find it helpful to create a quaternion with zero scalar part from 
+        % a vector $r \in \R^3$. We denote this operation as,
+        % \begin{equation}
+        %     \hat{r} = H r \equiv \begin{bmatrix} 0 \\ I_3 \end{bmatrix} r.
+        % \end{equation}
+        % Unit quaternions rotate a vector through the operation 
+        % $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. 
+        % This can be equivalently expressed using matrix multiplication as
+        % \begin{align} 
+        %     r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation}
+        % \end{align}
+        % where $A(q)$ is the rotation matrix in terms of the elements of the quaternion 
+        % \cite{markley2014fundamentals}.
 
 \section{Quaternion Differential Calculus} \label{sec:Quaternion_Calculus}
-    We now present a simple but powerful method for taking derivatives of functions 
-    involving quaternions based on the notation and linear algebraic operations outlined 
-    in Sec. \ref{sec:quaternions}.
+    Most modern methods for planning and control require derivatives with respect to the
+    state vector. We present a simple but powerful method for taking derivatives of 
+    quaternion-valued state vectors. The adaptation of this method to mixed state vectors,
+    such as $SE(3)$, is straightforward.
+
+    The key idea of this section is that differential quaternions live in three-dimensional
+    space instead of the four-dimensional space of quaternions. This can be seen graphically
+    (see Figure \ref{fig:tangent_plane}) or deduced intuitively, since rotations are
+    inherently three-dimensional and differential rotations should live in the same space as 
+    angular velocity, i.e. $\R^3$.
 
@@
     \end{figure}
+    In practice we have found Rodrigues Parameters to be a very effective representation
+    for three-dimensional differential rotations since they are computionally efficient
+    and do not inherit the sign ambiguity of quaternions.
 
-        Derivatives consider the effect an infinitesimal perturbation to the input has on
-        an infinitesimal perturbation to the output. For vector spaces, the composition
-        of the perturbation with the nominal value is simple addition and the
-        infinitesimal perturbation lives in the same space as the original vector. For
-        unit quaternions, however, neither of these are true; instead, they compose
-        according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to
-        first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig.
-        \ref{fig:tangent_plane}).
-
-        %, and are related through a non-linear mapping $\varphi : \R^3 \to \Q$ )
-        The fact that differential unit quaternions are three-dimensional should make
-        intuitive sense: Rotations are inherently three-dimensional and differential
-        rotations should live in the same space as angular velocity, i.e. $\R^3$.
-
-        skinny dog runs after skinny cat
-
-        There are many possible three-parameter representations for small rotations in
-        the literature. Many authors use the exponential map \cite{Baillieul1978,
-        Zefran1998, Lee2008, Saccon2013, Sola2017, Fan2016, watterson2018trajectory},
-        while others have used the Cayley map (also known as Rodrigues parameters)
-        \cite{Kobilarov2011, Kobilarov2014}, Modified Rodrigues Parameters (MRPs)
-        \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}.
-        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are
-        computationally efficient and do not inherit the sign ambiguity associated with
-        unit quaternions. The mapping between a vector of Rodrigues parameters $\phi \in
-        \R^3$ and a unit quaternion $q$ is known as the Cayley map: \begin{equation}
-        \label{eq:cayley}
+    The mapping between a vector of Rodrigues parameters $\phi \in \R^3$ and a unit
+    quaternion $q$ is known as the Cayley map:
+    \begin{equation} \label{eq:cayley}
         q = \varphi(\phi) = \frac{1}{\sqrt{1 + \norm{\phi}^2}} \begin{bmatrix} 1 \\ \phi \end{bmatrix}.
@@
     We will also make use of the inverse Cayley map:
-        \begin{equation}
+    \begin{equation} \label{eq:invcayley}
         \phi = \varphi^{-1}(q) = \frac{q_v}{q_s}.
@@
         Let $\phi \in \R^3$ be a differential rotation applied to a function with
-        quaternion inputs $y = h(q): \Q \to \R^p$, such that
+        unit quaternion inputs $y = h(q): \Q \to \R^p$, such that
         \begin{equation} \label{eq:vector_function}
-            y + \delta y = h(L(q) \varphi(\phi)) \approx h(q) +  \nabla h(q) \phi.
+            y + \delta y = h(q \otimes \varphi(\phi)) \approx h(q) +  \nabla h(q) \phi.
         \end{equation}
@@
         \begin{equation} \label{eq:quat_gradient}
-            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) 
-                        = \pdv{h}{q} \begin{bmatrix} 
+            \nabla h(q) = \pdv{h}{q} \begin{bmatrix} 
                             -q_v^T \\ 
@@
                         \end{bmatrix}
+                        := \pdv{h}{q} G(q) 
         \end{equation}
@@
 
-    \subsection{Hessian of Scalar-Valued Functions}
-	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by
-	    taking the Jacobian of \eqref{eq:quat_gradient} with respect to $\phi$ using the
-        product rule, again evaluated at $\phi = 0$:
-
+    \subsection{Other derivatives}
+        Using similar techniques, we find useful expressions for the Hessian of a
+        scalar-valued function (i.e. $p = 1$):
 	    \begin{equation} \label{eq:quat_hessian}
@@
         \end{equation}
-	    where the second term comes from the second derivative of $\varphi(\phi)$.
-	    Similar to $G(q)$, this ends up being the same (up to a scaling factor) for any
-        choice of three-parameter attitude representation.
-
-    \subsection{Jacobian of Quaternion-Valued Functions}
-        We now consider the case of a function that maps unit quaternions to unit
-        quaternions, $q' = f(q) : \Q \to \Q$. Here we must also consider the non-trivial
-        effect of a differential value applied to the output, i.e.:
-        \begin{equation} \label{eq:dqoutput}
-            L(q') \varphi(\phi') = f(L(q)\varphi(\phi)) .
-        \end{equation}
-        Solving \eqref{eq:dqoutput} for $\phi'$ we find,
-        \begin{equation} \label{eq:phiprime}
-            \phi' = \varphi^{-1} \left( L(q')^T f(L(q)\varphi(\phi)) \right) \approx \nabla f(q) \, \phi.
-        \end{equation}
-        Finally, the desired Jacobian is obtained by taking the derivative of
-        \eqref{eq:phiprime} with respect to $\phi$:
+        and the Jacobian of a quaternion-valued function $q' = f(q) : \Q \to \Q$:
         \begin{equation} \label{eq:quat_jacobian}
-            \nabla f(q) = H^T L(q')^T \pdv{f}{q} L(q) H = G(q')^T \pdv{f}{q} G(q).
-        \end{equation}
-        The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to
-        I_q$, where $I_q$ is the quaternion identity. Differentiating through the inverse
-        map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q}
-        \to H^T$ for any three-parameter attitude representation.
-
-\section{Multiplicative LQR} \label{sec:MLQR}
-    Leveraging the methods from the previous section, we derive multiplicative LQR
-    (MLQR), a variant of LQR that correctly accounts for the group structure of
-    rotations. For concreteness, we consider a system with rigid body dynamics, as
-    presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize
-    the system about a dynamically feasible discretized reference trajectory
-    $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
-    We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
-    \begin{equation} \label{eq:linearized_dynamics}
-        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k 
-    \end{equation}
-    where \begin{equation}
-        \begin{aligned}
-            A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
-            B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
-        \end{aligned}
-    \end{equation}
-    and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state error Jacobian:
-    \begin{equation} \label{eq:state_error}
-        \setlength\arraycolsep{1pt}
-        \delta x_k = \begin{bmatrix} 
-            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
-        \end{bmatrix}\!, \;
-        E(x) = \begin{bmatrix}
-            I_3 & & & \\
-            & G(q) & & \\
-            & & I_3 & \\
-            & & & I_3 \\
-        \end{bmatrix}\!.
+            \nabla f(q) = G(q')^T \pdv{f}{q} G(q).
         \end{equation}
 
-    With our linearized system, we can apply the traditional LQR Riccati recursion,
-    \begin{multline}
-         P_k = W_k +  A_k^T P_{k+1} A_k \\
-         - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
-    \end{multline}
-    which minimizes the quadratic cost function
+\section{Fast Constrained Trajectory Optimization on $SE(3)$}
+    \subsection{Algorithmic Modifications}
+    Here we briefly outline the changes to the ALTRO solver \cite{howell2019altro} to 
+    efficiently solve trajectory optimization problems for systems with 3D rotations in the
+    state vector. 
+%     The AL-iLQR algorithm at the heart of ALTRO requires linearization of the
+%     dynamics ($A \in \R^{n \times n}, B \in \R^{n \times m}$), the constraints ($\c_x \in 
+%     \R^{p \times n}, c_u \in \R^{p \times m}$), and a quadratic expansion of the cost
+%     function ($Q \in \R^{n \times n}, R \in \R^{m \times m}, q \in \R^n, r \in \R^m$). These
+%     can be computed as follows:
+    The linearization of the nonlinear discrete dynamics function $x_{k+1} = f(x_k,u_k)$ is 
+    ``corrected'' using \ref{eq:quat_jacobian}:
     \begin{equation}
-        \half \sum_{k=0}^N \delta x_k^T Q_k \delta x_k + \delta u_k^T R_k \delta u_k
+        A = G(f(x,u))^T \pdv{f}{x} G(x); \quad B = G(f(x,u))^T \pdv{f}{u}.
     \end{equation}
-    of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on
-    the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From
-    this, we get our non-linear feedback policy
-    \begin{equation} \label{eq:mlqr_control}
-        u = K_k \delta x_k + \bar{u}_k.
+    We can similarly compute the linearization of a nonlinear constraint function $c(x,u)$
+    using \ref{eq:quat_gradient}:
+    \begin{equation}
+        c_x = \pdv{c}{x} G(x); \quad c_u = \pdv{c}{x}.
     \end{equation}
-    where $\delta x_k$ is computed online using the non-linear error function
-    \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the
-    standard LQR formula:
-    \begin{equation} \label{eq:LQR_gain}
-         K_k := -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k.
-    \end{equation}
+    Lastly, we adjust the computation of the quadratic expansion of the augmented
+    Lagrangian cost function $J(x,u) + \lambda^T c(x,u) + \half c(x,u)^T I_\mu c(x,u)$
+    using \ref{eq:quat_gradient} and \ref{eq:quat_hessian}:
+    \begin{subequations}
+        \begin{align}
+            Q &= & G(x)^T & \pdv[2]{J}{x} G(x) + c_x^T I_\mu c_x + I_3 \pdv{J}{x} x \\
+            R &= &        & \pdv[2]{J}{u}      + c_u^T I_\mu c_u \\
+            H &= &        & \pdv{J}{u}{x} G(x) + c_u^T I_\mu c_x \\
+            q &= & G(x)^T & \pdv{J}{x} + c_x^T (I_\mu c(x,u) + \lambda) \\
+            r &= &        & \pdv{J}{u} + c_u^T (I_\mu c(x,u) + \lambda).
+        \end{align}
+    \end{subequations}
+    Here we define the \textit{state attitude Jacobian} $G(x)$ to be a block-diagonal
+    matrix where the block is an identity matrix for any vector-valued state and $G(q)$ for
+    any quaternion in the state vector. We refer the reader to the original ALTRO paper 
+    \cite{howell2019altro} for futher details on the algorithm.
 
-    The multiplicative Linear Quadratic Regulator is summarized as follows:
-    \begin{enumerate}
-        \item Pick $Q_k \in \R^{12 \times 12}$ to weight the state \textit{error}, and $R_k \in \R^{m \times m}$
-        \item Linearize the dynamics using the state error Jacobian \eqref{eq:state_error}
-        \item Compute $K_k$ using the standard LQR Riccati recursion
-        \item Online, compute the control using \eqref{eq:state_error} and \eqref{eq:mlqr_control}.
-    \end{enumerate}
+    These adjustments effectively modify the optimization so that it is performed on the
+    quaternion error state (or Rodrigues Parameters) instead of the space of unit
+    quaternions. In practice, this is analagous to the Multiplicative Extended Kalman
+    Filter \cite{markley2014fundamentals} in the state estimation community.
 
-    \subsection{Constrained Iterative MLQR}
-        As demonstrated above, it is very straight-forward to adapt LQR to use
-        quaternions. We similarly use this technique to adapt nonlinear trajectory
-        optimization algorithms. Here we present the modifications to the ALTRO solver
-        \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an
-        augmented Lagrangian approach to handle constraints.
+    The algorithm only needs one small modification to the ``forward pass,'' where the
+    trajectory is modified by simulating the system forward using the optimal feedback
+    policy, $\delta u = K \delta x + d$, computed during the ``backward'' pass. Instead of 
+    computing $\delta x$ using simple subtraction, we now compute the error state using the
+    inverse Cayley map \eqref{eq:invcayley} for the quaternion states. The rest of the
+    forward pass is unmodified.
 
-        For each iteration of iLQR within ALTRO, we can calculate a quadratic
-        approximation of the cost function using \eqref{eq:quat_gradient} and
-        \eqref{eq:quat_hessian}. Identical to \eqref{eq:linearized_dynamics}, we use
-        \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the
-        feedback and feedforward gains.
+    \subsection{Examples}
+        To demonstrate the effectiveness of the proposed approach, we ran the modified
+        version of ALTRO on several problems involving arbitrary 3D orientations. One 
+        required a quadrotor to do an aggresive ``zig-zag'' pattern, another to do a flip,
+        and the last tasked a model airplane model to do a barrell roll. The timing results
+        are shown in Figure \ref{fig:timing_chart}.
+        \begin{figure}
+            \centering
+            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
+            \caption{Timing results for the airplane barrell roll, quadrotor flip, and
+            quadrotor zig-zag examples using iLQR with roll-pitch-yaw Euler angles (RPY),
+            quaternions, and the current method. The timing result for the Quadrotor flip
+            with Euler angles is omitted since it failed to converge.
+            }
+            \label{fig:timing_chart}
+        \end{figure}
 
-        The only other modification to the algorithm is during the iLQR ``forward pass''
-        that simulates the system with the closed-loop MLQR controller, where we use
-        \eqref{eq:state_error} to calculate the nonlinear feedback policy during the
-        forward roll-outs.
-        For robustness, we may also add an additional criterion to the line search that
-        checks for singularities in the three-parameter error state which, for the Cayley
-        map, occur at \ang{180}, at which point the step size should be reduced.
+\section{Conclusion}
 
-    \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
-        In addition to the straight-forward modifications to the LQR algorithm itself, we
-        need to carefully consider the types of cost functions we minimize. We frequently
-        minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
-        (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We
-        propose two different cost functions that accomplish similar behavior.
-        For sake of clarity and space, we only consider the costs on the quaternion
-        variables: costs on the other states and the control variables remain unaffected.
+    We have proposed a general, accessible, and computationally efficient approach to
+    doing planning and control for systems with one or more unit quaternions in their
+    state vectors. The approach allows for straightforward adaptation of many
+    gradient-based methods for optimal control and motion planning, which we demonstrated
+    on the ALTRO solver. We have demonstrated that correctly leveraging the group
+    structure of rotations results in lower solve times and improved robustness of the
+    optimization algorithm.
 
-        \subsubsection{Error Quadratic}
-            Rather than simple subtraction, we can use a quadratic function on the
-            three-parameter error state \eqref{eq:state_error}:
-            \begin{equation} \label{eq:error_quadratic}
-                J_\text{err} = \half \phi^T Q \phi 
-                = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
-                        \left(\varphi^{-1}(\delta q)\right).
-            \end{equation}
-            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
-            and Hessian of \eqref{eq:error_quadratic} are
-            \begin{equation}
-                \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
-            \end{equation}
-            \begin{equation}
-                \begin{aligned}
-                     \nabla^2 J_\text{err} &= 
-                        G(\delta q)^T \! \left(
-                        D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q) 
-                      + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
-                \end{aligned}
-            \end{equation}
-            where, for the Cayley map,
-            \begin{equation}
-                D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
-                    q_v \;\; & -\frac{1}{q_s} I_3
-                \end{bmatrix}
-            \end{equation}
-            \begin{equation}
-                \nabla D = \pdv{q}(D(q)^T Q \phi) 
-                = -\frac{1}{q_s^2} \begin{bmatrix} 
-                    -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
-                                   Q \phi & 0 \\
-                \end{bmatrix}.
-            \end{equation}
+% \section{Multiplicative LQR} \label{sec:MLQR}
+%     Leveraging the methods from the previous section, we derive multiplicative LQR
+%     (MLQR), a variant of LQR that correctly accounts for the group structure of
+%     rotations. For concreteness, we consider a system with rigid body dynamics, as
+%     presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize
+%     the system about a dynamically feasible discretized reference trajectory
+%     $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
+%     We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
+%     \begin{equation} \label{eq:linearized_dynamics}
+%         \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k 
+%     \end{equation}
+%     where \begin{equation}
+%         \begin{aligned}
+%             A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
+%             B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
+%         \end{aligned}
+%     \end{equation}
+%     and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state error Jacobian:
+%     \begin{equation} \label{eq:state_error}
+%         \setlength\arraycolsep{1pt}
+%         \delta x_k = \begin{bmatrix} 
+%             r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
+%         \end{bmatrix}\!, \;
+%         E(x) = \begin{bmatrix}
+%             I_3 & & & \\
+%             & G(q) & & \\
+%             & & I_3 & \\
+%             & & & I_3 \\
+%         \end{bmatrix}\!.
+%     \end{equation}
+
+%     With our linearized system, we can apply the traditional LQR Riccati recursion,
+%     \begin{multline}
+%          P_k = W_k +  A_k^T P_{k+1} A_k \\
+%          - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
+%     \end{multline}
+%     which minimizes the quadratic cost function
+%     \begin{equation}
+%         \half \sum_{k=0}^N \delta x_k^T Q_k \delta x_k + \delta u_k^T R_k \delta u_k
+%     \end{equation}
+%     of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on
+%     the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From
+%     this, we get our non-linear feedback policy
+%     \begin{equation} \label{eq:mlqr_control}
+%         u = K_k \delta x_k + \bar{u}_k.
+%     \end{equation}
+%     where $\delta x_k$ is computed online using the non-linear error function
+%     \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the
+%     standard LQR formula:
+%     \begin{equation} \label{eq:LQR_gain}
+%          K_k := -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k.
+%     \end{equation}
+
+%     The multiplicative Linear Quadratic Regulator is summarized as follows:
+%     \begin{enumerate}
+%         \item Pick $Q_k \in \R^{12 \times 12}$ to weight the state \textit{error}, and $R_k \in \R^{m \times m}$
+%         \item Linearize the dynamics using the state error Jacobian \eqref{eq:state_error}
+%         \item Compute $K_k$ using the standard LQR Riccati recursion
+%         \item Online, compute the control using \eqref{eq:state_error} and \eqref{eq:mlqr_control}.
+%     \end{enumerate}
+
+%     \subsection{Constrained Iterative MLQR}
+%         As demonstrated above, it is very straight-forward to adapt LQR to use
+%         quaternions. We similarly use this technique to adapt nonlinear trajectory
+%         optimization algorithms. Here we present the modifications to the ALTRO solver
+%         \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an
+%         augmented Lagrangian approach to handle constraints.
+
+%         For each iteration of iLQR within ALTRO, we can calculate a quadratic
+%         approximation of the cost function using \eqref{eq:quat_gradient} and
+%         \eqref{eq:quat_hessian}. Identical to \eqref{eq:linearized_dynamics}, we use
+%         \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the
+%         feedback and feedforward gains.
+
+%         The only other modification to the algorithm is during the iLQR ``forward pass''
+%         that simulates the system with the closed-loop MLQR controller, where we use
+%         \eqref{eq:state_error} to calculate the nonlinear feedback policy during the
+%         forward roll-outs.
+%         For robustness, we may also add an additional criterion to the line search that
+%         checks for singularities in the three-parameter error state which, for the Cayley
+%         map, occur at \ang{180}, at which point the step size should be reduced.
+
+%     \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
+%         In addition to the straight-forward modifications to the LQR algorithm itself, we
+%         need to carefully consider the types of cost functions we minimize. We frequently
+%         minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
+%         (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We
+%         propose two different cost functions that accomplish similar behavior.
+%         For sake of clarity and space, we only consider the costs on the quaternion
+%         variables: costs on the other states and the control variables remain unaffected.
+
+%         \subsubsection{Error Quadratic}
+%             Rather than simple subtraction, we can use a quadratic function on the
+%             three-parameter error state \eqref{eq:state_error}:
+%             \begin{equation} \label{eq:error_quadratic}
+%                 J_\text{err} = \half \phi^T Q \phi 
+%                 = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
+%                         \left(\varphi^{-1}(\delta q)\right).
+%             \end{equation}
+%             where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
+%             and Hessian of \eqref{eq:error_quadratic} are
+%             \begin{equation}
+%                 \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
+%             \end{equation}
+%             \begin{multline}
+%                 \nabla^2 J_\text{err} = 
+%                     G(\delta q)^T \! \left(
+%                     D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
+%                     + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
+%             \end{multline}
+%             where, for the Cayley map,
+%             \begin{equation}
+%                 D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
+%                     q_v \;\; & -\frac{1}{q_s} I_3
+%                 \end{bmatrix}
+%             \end{equation}
+%             \begin{equation}
+%                 \nabla D = \pdv{q}(D(q)^T Q \phi) 
+%                 = -\frac{1}{q_s^2} \begin{bmatrix} 
+%                     -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
+%                                    Q \phi & 0 \\
+%                 \end{bmatrix}.
+%             \end{equation}
+
+%         \subsubsection{Geodesic Distance}
+%             Alternatively, we can use the geodesic distance between two quaternions
+%             \cite{Kuffner2004},
+%             \begin{equation} \label{eq:quat_geodesic}
+%                 J_\text{geo} = (1-\abs{q_g^T q}) ,
+%             \end{equation}
+%             whose gradient and Hessian are,
+%             \begin{align}
+%                 \nabla J_\text{geo} &= \pm q_g^T G(q) \\
+%                 \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
+%             \end{align}
+%             where the sign of the Hessian corresponds to the sign of $q_d^T q$.
+
+% \section{Experiments} \label{sec:experiments}
+%     Code for all experiments is available on 
+%     \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
+%     {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.
+
+%     \subsection{Trajectory Optimization}
+%         In this section we present several trajectory optimization problems solved using
+%         our modified version of ALTRO. The two quadrotor examples use 101 knot points for
+%         5 second trajectories, and the airplane example uses 51 knot points for a 1.25
+%         second trajectory. All results were run on a desktop computer with an Intel
+%         i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
+%         experiments are given in Fig. \ref{fig:timing_chart}.
+
+%         In all examples, MLQR is implemented with a unit quaternion attitude state and
+%         Rodrigues parameters as the error state. The first example compares the two cost
+%         functions given in Sec. \ref{sec:cost_functions}, while the last two use the
+%         geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
+%         \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
+%         from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
+%         vectors). All examples re-normalize the quaternion in the dynamics function and
+%         use a third-order explicit Runge-Kutta integrator.
+
+%         \begin{figure}
+%             \centering
+%             \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
+%             \caption{Timing results for the airplane barrell roll, quadrotor
+%                 flip, and quadrotor zig-zag examples using iLQR with
+%                 roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The
+%                 timing result for the Quadrotor flip with Euler angles is omitted
+%                 since it failed to converge.
+%             }
+%             \label{fig:timing_chart}
+%         \end{figure}
+
+%         \subsubsection{Quadrotor Zig-Zag}
+%             \begin{figure}[t]
+%                 \centering
+%                 \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
+%                 \caption{Zig-zag trajectory. The black quadrotors indicate
+%                     the reference trajectory generated by iMLQR. The yellow
+%                     quadrotors track the reference trajectory using time-varying
+%                     MLQR, while the red uses the HFCA controller from
+%                     \cite{watterson2020control}.
+%                 }
+%                 \label{fig:zig-zag}
+%             \end{figure}    
+%             \begin{figure}[t]
+%                 \centering
+%                 \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
+%                 \caption{Snapshots of the quadrotor flip trajectory. The
+%                     red-colored quadrotors represent the state near t=0 s and the
+%                     green-colored quadrotors represent the state near t=5.0 s
+%                 }
+%                 \label{fig:quad_flip}
+%             \end{figure}
+
+%         We set up a fairly simple trajectory optimization problem to achieve a
+%         ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of
+%         the form
+%         \begin{multline}
+%                   \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
+%                 + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) \\
+%                 + \half \sum_{k=1} u_k^T R u_k
+%         \end{multline}
+%         with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus
+%         \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and
+%         $(0,10,0)$. We compared the two cost functions from Sec.
+%         \ref{sec:cost_functions}, using a standard quadratic penalty on the
+%         non-quaternion states.
+%         \begin{align}
+%             \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
+%             \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
+%                 \pm \bar{q}^T q)
+%         \end{align}
+%         where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where
+%         $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular
+%         velocity.
+%         We used the following weighting matrices: 
+%         $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
+%         $Q_w = \text{diag}(10^3 I_3,I_9)$, 
+%         $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
+%         For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and
+%         terminal cost functions, respectively.
+
+%         The geodesic cost function performed significantly better than the error
+%         quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29
+%         iterations. In addition to better convergence behavior, the geodesic cost is
+%         about half as expensive to compute, which is unsurprising when comparing the
+%         gradients and Hessians in Sec. \ref{sec:cost_functions}.
+
+%         \subsubsection{Quadrotor Flip}
+% 	    Similar to the previous example, we use a set of waypoints to encourage the
+% 	    quadrotor to do a flip. The cost weights for this example were
+% 	    \begin{align*}
+% 	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \!\times\! 10^{-2}, 10^{-3} I_3,
+%     	        10^{-3} I_3, 10^{-2} I_3) \\
+% 	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
+% 	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
+% 	    \end{align*}
+% 	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
+% 	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
+% 	    to stay above the floor and move to a goal state 2 meters away in the $+y$
+% 	    direction. The solver is initialized with a dynamically infeasible trajectory
+% 	    that linearly interpolates between the initial and final states, rotating the
+% 	    quad around the x-axis a full \ang{360}.
+
+% 	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
+% 	    iMLQR within the ALTRO solver to handle constraints. The trajectory generated by
+% 	    iLQR with quaternions and quadratic costs was successful but had a slightly
+%         different trajectory. iLQR with Euler angles failed to solve the problem.
+
+%         \subsection{Airplane Barrel Roll}
+%         An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
+%         tasked to do a barrel roll by setting a high terminal cost for being upside-down,
+%         see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
+%         conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
+%         converged faster than the pure quaternion version. Despite the extra matrix
+%         multiplications and extra Hessian term, it also was faster per iteration than its
+%         iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
+%         better performance by correctly leveraging the structure of the rotation group
+%         during the optimization routine.
+
+%         \begin{figure}[h]
+%             \centering
+%             \includegraphics[height=3.5cm]{figures/barrellroll.png}
+%             \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
+%             \label{fig:barrellroll}
+%         \end{figure}
+
+% 	    \begin{figure}[h]
+% 	        \centering
+% 	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
+% 	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller.
+Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics.
+}
+% 	        \label{fig:tracking_err}
+% 	    \end{figure}
+%         \vspace{-20pt}
+
+% 	\subsection{Tracking Control}
+% 	    This section provides examples using MLQR for tracking and stabilizing control
+% 	    for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art
+% 	    geometric tracking controller \cite{watterson2020control} based on the Hopf
+% 	    Fibration.
+
+% 	    \subsubsection{MLQR Stabilization}
+% 	    To test robustness and the domain of attraction of the MLQR controller, a
+% 	    quadrotor is linearized about hover at the origin and initialized from states
+% 	    uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular
+% 	    velocities, and uniformly over the entire space of rotations. The success rates
+% 	    for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is
+% 	    defined as the terminal error (calculated using \eqref{eq:state_error}) being
+% 	    within a unit ball of radius 0.1 after 10.0 seconds. To further analyze the
+% 	    basins of attraction, we uniformly sampled over only roll (which goes singular at
+% 	    \ang{90} for our Euler angle representation), and yaw. The results are shown in
+% 	    Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the
+% 	    successfully stabilized initial conditions.
+
+% 	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also
+% 	    being more generally applicable, easier to tune, and easier to implement. It is
+% 	    also noteworthy that na\"ively using LQR with quaternions (incorrectly treating
+% 	    them as vectors) results in poorer performance than Euler angles, clearly
+% 	    demonstrating the value of the methodology presented in the current paper.
+
+% 	    \subsubsection{MLQR Tracking}
+% 	    Time-varying MLQR is used to track the zig-zag trajectory generated in the
+% 	    previous section. %The controller calculated the gains at $0.01 s$ intervals, and
+% 	    used zero-order-hold between intervals. To test tracking performance, zero-mean
+% 	    Gaussian noise was added as force and moments applied to the quadrotor.
+
+% 	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
+% 	    \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller
+% 	    dramatically out-performs HFCA, we found no significant difference in performance
+% 	    between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
+% 	    Since the tracking controller only deals with relatively small state and attitude
+% 	    errors, this is unsurprising. Additionally, we found the MLQR controller ran
+%         about 20x faster than the HFCA controller.
+
+% 	    \begin{figure*}[ht]
+% 	        \begin{minipage}{0.60\linewidth}
+%                 \centering
+%                 \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
+%                 \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black).
+All points within each contour were successfully stabilized by the controller.}
+%                 \label{fig:mlqr_basin}
+%             \end{minipage}
+% 	        \hspace{0.05\linewidth}
+%             \begin{minipage}{0.34\linewidth}
+%                 \centering
+%                 \begin{tabular}{cc}
+%                     \toprule
+%                     \textbf{Method} & \textbf{Success} \\
+%                     \midrule
+%                     Quat & 61\% \\
+%                     RPY  & 80\% \\
+%                     HFCA & 97\% \\
+%                     MLQR & 99\% \\
+%                     \bottomrule
+%                 \end{tabular}
+%                 \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
+%                 \label{tab:mlqr_success}
+%             \end{minipage}
+%         \end{figure*}
+
+% \section{Conclusions} \label{sec:conclusion}
+%     We have presented a general, unified method for planning and control on rigid-body
+%     systems with arbitrary attitude using standard linear algebra and vector calculus. By
+%     applying these methods to LQR to correctly account for the group structure of
+%     rotations, we have matched or exceeded the performance of a state-of-the-art
+%     geometric controller designed specifically for quadrotors, while being more general,
+%     requiring less system-dependent tuning, having less computational overhead, and being
+%     easier to implement.
+
+%     We have also demonstrated a straight-forward way to adapt nonlinear trajectory
+%     optimization techniques to work with quaternion-valued states. For these problems, we
+%     found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
+%     computationally efficient and provide excellent convergence in practice. We recommend
+%     the use of unit quaternions within planning and control algorithms for their
+%     simplicity, computational efficiency, and lack of singularities. We further recommend
+%     the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
+
+%     Future areas of research include the application of these methodologies to more
+%     complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
+%     more in-depth analysis of the convergence behavior of the algorithms proposed in the
+%     current work.
+
+% \paragraph{Acknowledgements}
+% This material is based upon work supported by the National Science Foundation Graduate
+% Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
+% conclusions or recommendations expressed in this material are those of the author(s) and
+% do not necessarily reflect the views of the National Science Foundation
+
+\printbibliography
 

### 2020-09-10 src/PlanningWithAttitude.jl
Start modifying examples

@@ -2,15 +2,10 @@ module PlanningWithAttitude
 
 using StaticArrays
 using LinearAlgebra
-using Parameters
-using ForwardDiff
-using TrajectoryOptimization
-const TO = TrajectoryOptimization
-using Statistics
-using ControlSystems
-using JLD2
-# using Logging
-# using Formatting
+using Rotations
+using RobotDynamics
+
+include("vecmodel.jl")
 
 # include("logger.jl")
 # include("knotpoint.jl")
@@ -21,8 +16,8 @@ using JLD2
 # include("ilqr.jl")
 # include("attitude_costs.jl")
 
-include("rbstate.jl")
-include("plotting.jl")
-include("tracking_control.jl")
-include("lqr_monte_carlo.jl")
+# include("rbstate.jl")
+# include("plotting.jl")
+# include("tracking_control.jl")
+# include("lqr_monte_carlo.jl")
 end

### 2020-09-10 src/vecmodel.jl

@@ -1,0 +1,8 @@
+struct VecModel{L} <: AbstractModel
+    model::L
+    VecModel(model::L) where L <: LieGroupModel = new{L}(model)
+end
+
+RobotDynamics.state_dim(model::VecModel) = state_dim(model.model)
+RobotDynamics.control_dim(model::VecModel) = control_dim(model.model)
+RobotDynamics.dynamics(model::VectorModel, args...) = dynamics(model.model, args...) 

### 2020-09-10 src/vecquat.jl

@@ -1,0 +1,335 @@
+struct VecQuat{T} <: Rotation{3,T}
+    w::T
+    x::T
+    y::T
+    z::T
+
+    @inline function VecQuat{T}(w, x, y, z, normalize::Bool = true) where T
+        if normalize
+            inorm = inv(sqrt(w*w + x*x + y*y + z*z))
+            new{T}(w*inorm, x*inorm, y*inorm, z*inorm)
+        else
+            new{T}(w, x, y, z)
+        end
+    end
+
+    VecQuat{T}(q::VecQuat) where T = new{T}(q.w, q.x, q.y, q.z)
+end
+
+# ~~~~~~~~~~~~~~~ Constructors ~~~~~~~~~~~~~~~ #
+# Use default map
+function VecQuat(w,x,y,z, normalize::Bool = true)
+    types = promote(w,x,y,z)
+    VecQuat{eltype(types)}(w,x,y,z, normalize)
+end
+
+# Pass in Vectors
+@inline function (::Type{Q})(q::AbstractVector, normalize::Bool = true) where Q <: VecQuat
+    check_length(q, 4)
+    Q(q[1], q[2], q[3], q[4], normalize)
+end
+@inline (::Type{Q})(q::StaticVector{4}, normalize::Bool = true) where Q <: VecQuat =
+    Q(q[1], q[2], q[3], q[4], normalize)
+
+# Copy constructors
+VecQuat(q::VecQuat) = q
+
+# Convert from UnitQuaternion
+(::Type{Q})(q::UnitQuaternion) where Q <: VecQuat = Q(q.w, q.x, q.y, q.z)
+(::Type{Q})(q::Vecquat) where Q <: UnitQuatenrion = Q(q.w, q.x, q.y, q.z)
+
+# ~~~~~~~~~~~~~~~ StaticArrays Interface ~~~~~~~~~~~~~~~ #
+function (::Type{Q})(t::NTuple{9}) where Q<:VecQuat
+    #=
+    This function solves the system of equations in Section 3.1
+    of https://arxiv.org/pdf/math/0701759.pdf. This cheap method
+    only works for matrices that are already orthonormal (orthogonal
+    and unit length columns). The nearest orthonormal matrix can
+    be found by solving Wahba's problem:
[…]

### 2020-09-11 src/PlanningWithAttitude.jl
Add quatslack types

@@ -3,9 +3,17 @@ module PlanningWithAttitude
 using StaticArrays
 using LinearAlgebra
 using Rotations
-using RobotDynamics
+import RobotDynamics
+import TrajectoryOptimization
+const RD = RobotDynamics
+const TO = TrajectoryOptimization
 
-include("vecmodel.jl")
+# include("vecmodel.jl")
+include("rotatedmodel.jl")
+
+export
+    QuatSlackModel,
+    UnitQuatConstraint
 
 # include("logger.jl")
 # include("knotpoint.jl")

### 2020-09-11 src/rotatedmodel.jl

@@ -1,0 +1,69 @@
+struct QuatSlackModel{R,M,L} <: RD.RigidBody{R} 
+    model::L
+    function QuatSlackModel(model::L) where L <: RD.RigidBody
+        M = RD.control_dim(model)
+        R = RD.rotation_type(model)
+        new{R,M,L}(model)
+    end
+end
+RobotDynamics.control_dim(::QuatSlackModel{<:Any,M}) where M = M+1 
+
+function RobotDynamics.dynamics(model::QuatSlackModel{<:Any,m}, x, u, t=0) where m
+    uinds = SVector{m}(1:m)
+    u0 = u[uinds]                    # original controls
+    s = u[m+1]                       # Quaternion slack
+    r,q,v,ω = RobotDynamics.parse_state(model, x)
+    q̄ = s*q                          # scale the quaternion by the extra control
+    x̄ = RobotDynamics.build_state(model, r, q̄, v, ω)
+    RD.dynamics(model.model, x̄, u, t)   # call the original dynamics with the scaled quaternion
+end
+
+"""
+    UnitQuatConstraint
+
+Enforce a that the unit quaternion in the state, at indices `qind`, has unit norm
+after being multiplied by a slack variable at index `sind` in the joint state-control vector
+`z = [x;u]`.
+"""
+struct UnitQuatConstraint <: TrajectoryOptimization.StageConstraint
+    n::Int
+    m::Int
+    qind::SVector{4,Int}
+    sind::Int
+end
+
+function UnitQuatConstraint(n::Int, m::Int, qind=4:7, sind=n+m)
+    UnitQuatConstraint(n, m, SVector{4}(qind), sind)
+end
+
+UnitQuatConstraint(model::QuatSlackModel) = UnitQuatConstraint(size(model)...)
+
+Base.copy(con::UnitQuatConstraint) = UnitQuatConstraint(con.n, con.m, con.qind, con.sind)
+
+@inline TO.sense(::UnitQuatConstraint) = Equality()
+@inline Base.length(::UnitQuatConstraint) = 1
+@inline RobotDynamics.state_dim(con::UnitQuatConstraint) = con.n
+@inline RobotDynamics.control_dim(con::UnitQuatConstraint) = con.m
+
+function TO.evaluate(con::UnitQuatConstraint, z::RD.AbstractKnotPoint) 
[…]

### 2020-09-11 src/vecmodel.jl

@@ -5,4 +5,4 @@ end
 
 RobotDynamics.state_dim(model::VecModel) = state_dim(model.model)
 RobotDynamics.control_dim(model::VecModel) = control_dim(model.model)
-RobotDynamics.dynamics(model::VectorModel, args...) = dynamics(model.model, args...) 
+RobotDynamics.dynamics(model::VecModel, args...) = dynamics(model.model, args...) 

### 2020-09-11 test/quatslack.jl

@@ -1,0 +1,41 @@
+using Test
+using PlanningWithAttitude
+using Rotations
+using RobotDynamics
+using TrajectoryOptimization
+using Altro
+using StaticArrays
+using ForwardDiff
+import RobotZoo.Quadrotor
+const TO = TrajectoryOptimization
+
+quad = Quadrotor()
+model = QuatSlackModel(quad)
+@test state_dim(quad) == 13 == state_dim(model)
+@test control_dim(model) == 5
+x,u = rand(quad)
+r,q,v,ω = RobotDynamics.parse_state(model, x)
+s = rand()
+x2 = RobotDynamics.build_state(model, r, q*(1/s), v, ω)
+x3 = RobotDynamics.build_state(model, r, q*s, v, ω)
+u2 = push(u, s) 
+@test dynamics(model, x2, u2) ≈ dynamics(quad, x, u)
+@test dynamics(model, x, u2) ≈ dynamics(quad, x3, u)
+
+con = UnitQuatConstraint(model)
+z = KnotPoint(x2,u2,0.1)
+z2 = KnotPoint(x,u2,0.1)
+@test TO.evaluate(con,z)[1] ≈ 0 atol=1e-12
+@test TO.evaluate(con,z2)[1] ≈ s^2 - 1 atol=1e-12
+
+c(x) = TO.evaluate(con, StaticKnotPoint(z, x))
+∇c = ForwardDiff.jacobian(c, z.z)
+∇c2 = zeros(1,18)
+TO.jacobian!(∇c2, con, z)
+@test ∇c[4:7] ≈ 2*Rotations.params(q)*s
+@test ∇c2 ≈ ∇c
+
+∇c = ForwardDiff.jacobian(c, z2.z)
+TO.jacobian!(∇c2, con, z2)
+@test ∇c[4:7] ≈ 2*Rotations.params(q)*s*s
+@test ∇c2 ≈ ∇c

### 2020-09-11 src/rotatedmodel.jl
Solve problem with quat slack

@@ -40,7 +40,7 @@ UnitQuatConstraint(model::QuatSlackModel) = UnitQuatConstraint(size(model)...)
 
 Base.copy(con::UnitQuatConstraint) = UnitQuatConstraint(con.n, con.m, con.qind, con.sind)
 
-@inline TO.sense(::UnitQuatConstraint) = Equality()
+@inline TO.sense(::UnitQuatConstraint) = TO.Equality()
 @inline Base.length(::UnitQuatConstraint) = 1
 @inline RobotDynamics.state_dim(con::UnitQuatConstraint) = con.n
 @inline RobotDynamics.control_dim(con::UnitQuatConstraint) = con.m

### 2020-09-11 test/quatslack.jl

@@ -4,7 +4,7 @@ using Rotations
 using RobotDynamics
 using TrajectoryOptimization
 using Altro
-using StaticArrays
+using StaticArrays, LinearAlgebra
 using ForwardDiff
 import RobotZoo.Quadrotor
 const TO = TrajectoryOptimization
@@ -39,3 +39,30 @@ TO.jacobian!(∇c2, con, z)
 TO.jacobian!(∇c2, con, z2)
 @test ∇c[4:7] ≈ 2*Rotations.params(q)*s*s
 @test ∇c2 ≈ ∇c
+
+
+n,m = size(model)
+N,tf = 101, 5.0
+
+Q = Diagonal(RobotDynamics.fill_state(model, 1,10,10,10.))
+R = Diagonal(SA[1,1,1,1,100.]*1e-2)
+x0 = zeros(model)[1] 
+xf = RobotDynamics.build_state(model, 
+    RBState([2,1,0.5], expm(SA[0,0,1]*pi/4), zeros(3), zeros(3))
+)
+utrim = push(zeros(quad)[2],1)
+obj = LQRObjective(Q, R, (N-1)*Q, xf, N, uf=utrim)
+
+cons = ConstraintList(n,m,N)
+slack = UnitQuatConstraint(model)
+add_constraint!(cons, slack, 1:N-1)
+
+prob = Problem(model, obj, xf, tf, x0=x0, constraints=cons)
+initial_controls!(prob, utrim) 
+controls(prob)
+rollout!(prob)
+solver = ALTROSolver(prob)
+set_options!(solver, verbose=2, show_summary=true)
+solve!(solver)
+RBState(model, states(solver)[end]) ⊖ RBState(model, xf)
+controls(solver)

### 2020-09-11 paper/iros_abstract.tex

@@
 
-    Many approaches have been taken to address the problem of motion planning and control
-    on the $SE(3)$ manifold; most of these rely heavily on ideas and notation
-    from differential geometry. Despite some impressive work in this field, these ideas
-    have not been widely adopted and many continue to ignore the group structure of rotations,
-    especially in the field of optimal control and motion planning. One part of this problem
-    is that accounting for this structure requires low-level changes to the underlying
-    optimization algorithms, which is difficult or impossible to implement when relying on
-    commercial or even open-source optimization solvers.
+    Many approaches \todo{add some citations here} have been taken to address the problem of motion planning and control
+    on the $SE(3)$ group; most of these rely heavily on ideas and notation
+    from differential geometry. Despite some impressive results, many of these ideas
+    have not been widely adopted by the robotics community, and many practitioners continue to ignore the group structure of rotations.
+One issue preventing wider adoption is that accounting for this structure requires low-level changes to the underlying
+    optimization algorithms, which is difficult or impossible when relying on existing
+    off-the-shelf solvers.
 
-    In the current abstract, we formulate a straightforward, generic method for adapting
-    existing gradient-based algorithms to correctly account for the group structure of
+    \todo{I'm worried that SE(3) and SO(3) results are getting confused/conflated here.
+I would explicitly talk about both, emphasize that we're handling SO(3) with the quaternion stuff, and that $SE(3) = SO(3) \times \mathbb{R}^3$ is a straightforward extension.}
+
+    To address this lack of solver support, we formulate a straightforward, generic method for adapting
+    existing Newton and gradient-based algorithms to correctly account for the group structure of
     rotations. Based entirely on basic mathematical tools from linear algebra and
     calculus, our method is computationally efficient and should be both accessible and
-    familiar to most roboticists. We then apply this method to our ALTRO solver
+    familiar to most roboticists. We apply this method to the ALTRO solver
     \cite{howell2019altro} and demonstrate its performance on several constrained trajectory
@@
     state vector. We present a simple but powerful method for taking derivatives of 
-    quaternion-valued state vectors. The adaptation of this method to mixed state vectors,
-    such as $SE(3)$, is straightforward.
+    quaternion-valued state vectors. The extension of this method to more general state vectors that contain Cartesian products of $SO(3)$, $SE(3)$, and $\mathbb{R}^n$ is straightforward.
 
@@
             When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in 
-            n-dimensional space, the tangent space $T$ is a plane living in $\R^{n-1}$, 
+            n-dimensional space, the tangent space $T$ is a hyper-plane $\R^{n-1}$, 
             illustrated here with $n=3$. Therefore, when linearizing about a unit 
@@
     \end{figure}
-    In practice we have found Rodrigues Parameters to be a very effective representation
+    In practice, we have found Rodrigues Parameters to be a very effective representation
     for three-dimensional differential rotations since they are computionally efficient
@@
         \end{equation}
-        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which
+        where $G(q) \in \R^{4 \times 3}$ is referred to as the \textit{attitude Jacobian}, which
         essentially becomes a ``conversion factor'' allowing us to apply results from
@@
     Lastly, we adjust the computation of the quadratic expansion of the augmented
-    Lagrangian cost function $J(x,u) + \lambda^T c(x,u) + \half c(x,u)^T I_\mu c(x,u)$
+    Lagrangian \todo{I would leave out the AL stuff and just talk about a generic cost function here to make it more general} cost function $J(x,u) + \lambda^T c(x,u) + \half c(x,u)^T I_\mu c(x,u)$
     using \ref{eq:quat_gradient} and \ref{eq:quat_hessian}:
@@
     inverse Cayley map \eqref{eq:invcayley} for the quaternion states. The rest of the
-    forward pass is unmodified.
+    forward pass is unmodified. \todo{I think you can safely omit some of these details and just refer to the ALTRO paper.}
 
@@
         \end{figure}
+
+        \todo{It would be great to compare to both the minimal iLQR stuff and SNOPT or IPOPT with quaternion norm constraints + slack variables (referred to as the ``embedding approach'').}
 

### 2020-09-11 paper/iros_abstract.tex
final draft of iros submission

@@
 
-    Many approaches \todo{add some citations here} have been taken to address the problem of motion planning and control
-    on the $SE(3)$ group; most of these rely heavily on ideas and notation
-    from differential geometry. Despite some impressive results, many of these ideas
-    have not been widely adopted by the robotics community, and many practitioners continue to ignore the group structure of rotations.
-One issue preventing wider adoption is that accounting for this structure requires low-level changes to the underlying
+    Many approaches have been taken to address the problem of motion planning and control
+    on the space of rigid body motions \cite{Kobilarov2011, Saccon2013,
+    watterson2020control}; most of these rely heavily on ideas and notation from
+    differential geometry. Despite some impressive results, many of these ideas have not
+    been widely adopted by the robotics community, and many practitioners continue to
+    ignore the group structure of rotations. One issue preventing wider adoption is that
+    accounting for this structure requires low-level changes to the underlying
     optimization algorithms, which is difficult or impossible when relying on existing
@@
 
-    \todo{I'm worried that SE(3) and SO(3) results are getting confused/conflated here.
-I would explicitly talk about both, emphasize that we're handling SO(3) with the quaternion stuff, and that $SE(3) = SO(3) \times \mathbb{R}^3$ is a straightforward extension.}
+    % \todo{I'm worried that SE(3) and SO(3) results are getting confused/conflated here. I
+    % would explicitly talk about both, emphasize that we're handling SO(3) with the
+    % quaternion stuff, and that $SE(3) = SO(3) \times \mathbb{R}^3$ is a straightforward
+    % extension.}
 
@@
 
-        % Quaternion multiplication is defined as
-        % \begin{equation} \label{eq:quat_mult}
-        %     \q_2 \otimes \q_1 = L(q_2) q_1 = R(q_1) q_2
-        % \end{equation}
-        % where $L(q)$ and $R(q)$ are orthonormal matrices defined as
-        % \begin{align}
-        %     L(q) &:= \begin{bmatrix} 
-        %         q_s \;\; & -q_v^T \\ 
-        %         q_v \;\; & q_s I + \skewmat{q_v} 
-        %     \end{bmatrix} 
-        %     \label{eq:Lmult} \\
-        %     %= \begin{bmatrix} q & G(q) \end{bmatrix} \label{eq:Lmult} \\
-        %     R(q) &:=\begin{bmatrix} 
-        %         q_s \;\; & -q_v^T \\ 
-        %         q_v \;\; & q_s I - \skewmat{q_v} 
-        %     \end{bmatrix} \label{eq:Rmult},
-        % \end{align}
-        % and $\skewmat{x}$ is the skew-symmetric matrix operator
-        % \begin{equation}
-        %     \skewmat{x} := \begin{bmatrix} 
-        %         0 & -x_3 & x_2 \\ 
-        %         x_3 & 0 & -x_1\\ 
-        %         -x_2 & x_1 & 0 
-        %     \end{bmatrix}.
-        % \end{equation}
-
-        % The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal 
-        % to its conjugate $q^*$, which is simply the same quaternion with a negated vector 
-        % part:
-        % \begin{equation} \label{eq:T}
-        %     \q^* = T q := \begin{bmatrix} 
-        %         1 & \\ 
-        %         & -I_3 
-        %     \end{bmatrix} q
-        % \end{equation}
-        % The following identities, which are easily derived from 
-        % \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
-        % \begin{align}
-        %     &L(Tq) = L(q)^T = L(q)^{-1} \\
-        %     &R(Tq) = R(q)^T = R(q)^{-1} .
-        % \end{align}
-
-        % We will sometimes find it helpful to create a quaternion with zero scalar part from 
-        % a vector $r \in \R^3$. We denote this operation as,
-        % \begin{equation}
-        %     \hat{r} = H r \equiv \begin{bmatrix} 0 \\ I_3 \end{bmatrix} r.
-        % \end{equation}
-        % Unit quaternions rotate a vector through the operation 
-        % $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. 
-        % This can be equivalently expressed using matrix multiplication as
-        % \begin{align} 
-        %     r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation}
-        % \end{align}
-        % where $A(q)$ is the rotation matrix in terms of the elements of the quaternion 
-        % \cite{markley2014fundamentals}.
+    \begin{figure}
+        \centering
+        \includegraphics[width=\columnwidth]{figures/quadflip_new.png}
+        \label{fig:quadflip}
+        \caption{Snapshots of quadrotor flip trajectory solved using a modified version of 
+        ALTRO that correctly accounts for the group structure of rotations. These modifications
+        make it easier to find trajectories that undergo large changes in attitude, such as
+        the loop-de-loop shown here.}
+    \end{figure}
 
@@
 
-    The key idea of this section is that differential quaternions live in three-dimensional
-    space instead of the four-dimensional space of quaternions. This can be seen graphically
-    (see Figure \ref{fig:tangent_plane}) or deduced intuitively, since rotations are
-    inherently three-dimensional and differential rotations should live in the same space as 
-    angular velocity, i.e. $\R^3$.
+    The key idea of this section is that differential quaternions live in
+    three-dimensional space instead of the four-dimensional space of quaternions; this
+    idea should match our intuition given rotations are inherently three-dimensional and
+    differential rotations should live in the same space as angular velocity, i.e.
+    $\R^3$.
 
-    \begin{figure}
-        \centering
-        \includegraphics[height=5cm]{figures/tangent_plane.tikz}
-        \caption{
-            When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in 
-            n-dimensional space, the tangent space $T$ is a hyper-plane $\R^{n-1}$, 
-            illustrated here with $n=3$. Therefore, when linearizing about a unit 
-            quaternion $q \in \Q$, the space of differential rotations lives in $\R^3$.
-        }
-        \label{fig:tangent_plane}
-    \end{figure}
+    % \begin{figure}
+    %     \centering
+    %     \includegraphics[height=5cm]{figures/tangent_plane.tikz}
+    %     \caption{
+    %         When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in 
+    %         n-dimensional space, the tangent space $T$ is a hyper-plane $\R^{n-1}$, 
+    %         illustrated here with $n=3$. Therefore, when linearizing about a unit 
+    %         quaternion $q \in \Q$, the space of differential rotations lives in $\R^3$.
+    %     }
+    %     \label{fig:tangent_plane}
+    % \end{figure}
     In practice, we have found Rodrigues Parameters to be a very effective representation
@@
 
-\section{Fast Constrained Trajectory Optimization on $SE(3)$}
+\section{Fast Constrained Trajectory Optimization with Attitude States}
     \subsection{Algorithmic Modifications}
@@
     state vector.
-%     The AL-iLQR algorithm at the heart of ALTRO requires linearization of the
-%     dynamics ($A \in \R^{n \times n}, B \in \R^{n \times m}$), the constraints ($\c_x \in 
-%     \R^{p \times n}, c_u \in \R^{p \times m}$), and a quadratic expansion of the cost
-%     function ($Q \in \R^{n \times n}, R \in \R^{m \times m}, q \in \R^n, r \in \R^m$). These
-%     can be computed as follows:
-    The linearization of the nonlinear discrete dynamics function $x_{k+1} = f(x_k,u_k)$ is 
-    ``corrected'' using \ref{eq:quat_jacobian}:
-    \begin{equation}
-        A = G(f(x,u))^T \pdv{f}{x} G(x); \quad B = G(f(x,u))^T \pdv{f}{u}.
-    \end{equation}
-    We can similarly compute the linearization of a nonlinear constraint function $c(x,u)$
-    using \ref{eq:quat_gradient}:
-    \begin{equation}
-        c_x = \pdv{c}{x} G(x); \quad c_u = \pdv{c}{x}.
-    \end{equation}
-    Lastly, we adjust the computation of the quadratic expansion of the augmented
-    Lagrangian \todo{I would leave out the AL stuff and just talk about a generic cost function here to make it more general} cost function $J(x,u) + \lambda^T c(x,u) + \half c(x,u)^T I_\mu c(x,u)$
-    using \ref{eq:quat_gradient} and \ref{eq:quat_hessian}:
-    \begin{subequations}
-        \begin{align}
-            Q &= & G(x)^T & \pdv[2]{J}{x} G(x) + c_x^T I_\mu c_x + I_3 \pdv{J}{x} x \\
-            R &= &        & \pdv[2]{J}{u}      + c_u^T I_\mu c_u \\
-            H &= &        & \pdv{J}{u}{x} G(x) + c_u^T I_\mu c_x \\
-            q &= & G(x)^T & \pdv{J}{x} + c_x^T (I_\mu c(x,u) + \lambda) \\
-            r &= &        & \pdv{J}{u} + c_u^T (I_\mu c(x,u) + \lambda).
-        \end{align}
-    \end{subequations}
-    Here we define the \textit{state attitude Jacobian} $G(x)$ to be a block-diagonal
-    matrix where the block is an identity matrix for any vector-valued state and $G(q)$ for
-    any quaternion in the state vector. We refer the reader to the original ALTRO paper 
+
+    The linearization of the nonlinear discrete dynamics function $x_{k+1} = f(x_k,u_k)$
+    is ``corrected'' using \ref{eq:quat_jacobian}: \begin{equation} A = E(f(x,u))^T
+    \pdv{f}{x} E(x); \quad B = E(f(x,u))^T \pdv{f}{u}. \end{equation} Here we define the
+    \textit{state attitude Jacobian} $E(x)$ to be a block-diagonal matrix where the block
+    is an identity matrix for any vector-valued state and $G(q)$ for any quaternion in
+    the state vector. Using \eqref{eq:quat_gradient}, \eqref{eq:quat_hessian},
+    \eqref{eq:quat_jacobian} we can derive similar modifications to the expansions of the
+    cost and constraint functions. We refer the reader to the original ALTRO paper
     \cite{howell2019altro} for futher details on the algorithm.
@@
     inverse Cayley map \eqref{eq:invcayley} for the quaternion states. The rest of the
-    forward pass is unmodified. \todo{I think you can safely omit some of these details and just refer to the ALTRO paper.}
+    forward pass is unmodified. 
+    %\todo{I think you can safely omit some of these details and just refer to the ALTRO paper.}
 
     \subsection{Examples}
-        To demonstrate the effectiveness of the proposed approach, we ran the modified
-        version of ALTRO on several problems involving arbitrary 3D orientations. One 
-        required a quadrotor to do an aggresive ``zig-zag'' pattern, another to do a flip,
-        and the last tasked a model airplane model to do a barrell roll. The timing results
-        are shown in Figure \ref{fig:timing_chart}.
-        \begin{figure}
-            \centering
-            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
-            \caption{Timing results for the airplane barrell roll, quadrotor flip, and
-            quadrotor zig-zag examples using iLQR with roll-pitch-yaw Euler angles (RPY),
-            quaternions, and the current method. The timing result for the Quadrotor flip
-            with Euler angles is omitted since it failed to converge.
-            }
-            \label{fig:timing_chart}
-        \end{figure}
+        To demonstrate the effectiveness of the proposed approach, we tested the modified
+        version of ALTRO by solving for a quadrotor flip trajectory (see Figure \ref{fig:quadflip}).
+        Four soft ``waypoint'' costs were placed along the trajectory, encouraging the
+        quadrotor to follow a loop. The trajectory was initialized with hover controls and a
+        state trajectory that smoothly rotated the quadrotor one full rotation about the 
+        x-axis while moving to the goal. 
+        The modified version of ALTRO solved the problem in 270 ms and 26 iterations, while
+        the original version that naively treated the quaternion as a vector in $\R^4$ 
+        solved in 870 ms and 91 iterations. The original version failed to provide a nice,
+        smooth loop and instead reversed direction halfway through the loop.
 
-        \todo{It would be great to compare to both the minimal iLQR stuff and SNOPT or IPOPT with quaternion norm constraints + slack variables (referred to as the ``embedding approach'').}
+        It's also worthwhile to note that this problem cannot be solved as-is with any
+        three-parameter representation because of the singularities that occur at 90, 180,
+        and 360 degrees for Euler angles, Rodrigues parameters, and MRP's, respectively.
+        A classic approach to overcome this issue is to do discrete switching of the 
+        reference orientation, which requires previous knowledge of the topology of the
+        optimal solution. 
+        % \begin{figure}
+        %     \centering
+        %     \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
+        %     \caption{Timing results for the airplane barrell roll, quadrotor flip, and
+        %     quadrotor zig-zag examples using iLQR with roll-pitch-yaw Euler angles (RPY),
+        %     quaternions, and the current method. The timing result for the Quadrotor flip
+        %     with Euler angles is omitted since it failed to converge.
+        %     }
+        %     \label{fig:timing_chart}
+        % \end{figure}
+
+        % \todo{It would be great to compare to both the minimal iLQR stuff and SNOPT or IPOPT with quaternion norm constraints + slack variables (referred to as the ``embedding approach'').}
 
@@
     gradient-based methods for optimal control and motion planning, which we demonstrated
-    on the ALTRO solver. We have demonstrated that correctly leveraging the group
-    structure of rotations results in lower solve times and improved robustness of the
-    optimization algorithm.
-
-% \section{Multiplicative LQR} \label{sec:MLQR}
-%     Leveraging the methods from the previous section, we derive multiplicative LQR
-%     (MLQR), a variant of LQR that correctly accounts for the group structure of
-%     rotations. For concreteness, we consider a system with rigid body dynamics, as
-%     presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize
-%     the system about a dynamically feasible discretized reference trajectory
-%     $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
-%     We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
-%     \begin{equation} \label{eq:linearized_dynamics}
-%         \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k 
-%     \end{equation}
-%     where \begin{equation}
-%         \begin{aligned}
-%             A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
-%             B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
-%         \end{aligned}
-%     \end{equation}
-%     and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state error Jacobian:
-%     \begin{equation} \label{eq:state_error}
-%         \setlength\arraycolsep{1pt}
-%         \delta x_k = \begin{bmatrix} 
-%             r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
-%         \end{bmatrix}\!, \;
-%         E(x) = \begin{bmatrix}
-%             I_3 & & & \\
-%             & G(q) & & \\
-%             & & I_3 & \\
-%             & & & I_3 \\
-%         \end{bmatrix}\!.
-%     \end{equation}
-
-%     With our linearized system, we can apply the traditional LQR Riccati recursion,
-%     \begin{multline}
-%          P_k = W_k +  A_k^T P_{k+1} A_k \\
-%          - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
-%     \end{multline}
-%     which minimizes the quadratic cost function
-%     \begin{equation}
-%         \half \sum_{k=0}^N \delta x_k^T Q_k \delta x_k + \delta u_k^T R_k \delta u_k
-%     \end{equation}
-%     of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on
-%     the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From
-%     this, we get our non-linear feedback policy
-%     \begin{equation} \label{eq:mlqr_control}
-%         u = K_k \delta x_k + \bar{u}_k.
-%     \end{equation}
-%     where $\delta x_k$ is computed online using the non-linear error function
-%     \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the
-%     standard LQR formula:
-%     \begin{equation} \label{eq:LQR_gain}
-%          K_k := -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k.
-%     \end{equation}
-
-%     The multiplicative Linear Quadratic Regulator is summarized as follows:
-%     \begin{enumerate}
-%         \item Pick $Q_k \in \R^{12 \times 12}$ to weight the state \textit{error}, and $R_k \in \R^{m \times m}$
-%         \item Linearize the dynamics using the state error Jacobian \eqref{eq:state_error}
-%         \item Compute $K_k$ using the standard LQR Riccati recursion
-%         \item Online, compute the control using \eqref{eq:state_error} and \eqref{eq:mlqr_control}.
-%     \end{enumerate}
-
-%     \subsection{Constrained Iterative MLQR}
-%         As demonstrated above, it is very straight-forward to adapt LQR to use
-%         quaternions. We similarly use this technique to adapt nonlinear trajectory
-%         optimization algorithms. Here we present the modifications to the ALTRO solver
-%         \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an
-%         augmented Lagrangian approach to handle constraints.
-
-%         For each iteration of iLQR within ALTRO, we can calculate a quadratic
-%         approximation of the cost function using \eqref{eq:quat_gradient} and
-%         \eqref{eq:quat_hessian}. Identical to \eqref{eq:linearized_dynamics}, we use
-%         \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the
-%         feedback and feedforward gains.
-
-%         The only other modification to the algorithm is during the iLQR ``forward pass''
-%         that simulates the system with the closed-loop MLQR controller, where we use
-%         \eqref{eq:state_error} to calculate the nonlinear feedback policy during the
-%         forward roll-outs.
-%         For robustness, we may also add an additional criterion to the line search that
-%         checks for singularities in the three-parameter error state which, for the Cayley
-%         map, occur at \ang{180}, at which point the step size should be reduced.
-
-%     \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
-%         In addition to the straight-forward modifications to the LQR algorithm itself, we
-%         need to carefully consider the types of cost functions we minimize. We frequently
-%         minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
-%         (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We
-%         propose two different cost functions that accomplish similar behavior.
-%         For sake of clarity and space, we only consider the costs on the quaternion
-%         variables: costs on the other states and the control variables remain unaffected.
-
-%         \subsubsection{Error Quadratic}
-%             Rather than simple subtraction, we can use a quadratic function on the
-%             three-parameter error state \eqref{eq:state_error}:
-%             \begin{equation} \label{eq:error_quadratic}
-%                 J_\text{err} = \half \phi^T Q \phi 
-%                 = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
-%                         \left(\varphi^{-1}(\delta q)\right).
-%             \end{equation}
-%             where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
-%             and Hessian of \eqref{eq:error_quadratic} are
-%             \begin{equation}
-%                 \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
-%             \end{equation}
-%             \begin{multline}
-%                 \nabla^2 J_\text{err} = 
-%                     G(\delta q)^T \! \left(
-%                     D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
-%                     + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
-%             \end{multline}
-%             where, for the Cayley map,
-%             \begin{equation}
-%                 D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
-%                     q_v \;\; & -\frac{1}{q_s} I_3
-%                 \end{bmatrix}
-%             \end{equation}
-%             \begin{equation}
-%                 \nabla D = \pdv{q}(D(q)^T Q \phi) 
-%                 = -\frac{1}{q_s^2} \begin{bmatrix} 
-%                     -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
-%                                    Q \phi & 0 \\
-%                 \end{bmatrix}.
-%             \end{equation}
-
-%         \subsubsection{Geodesic Distance}
-%             Alternatively, we can use the geodesic distance between two quaternions
-%             \cite{Kuffner2004},
-%             \begin{equation} \label{eq:quat_geodesic}
-%                 J_\text{geo} = (1-\abs{q_g^T q}) ,
-%             \end{equation}
-%             whose gradient and Hessian are,
-%             \begin{align}
-%                 \nabla J_\text{geo} &= \pm q_g^T G(q) \\
-%                 \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
-%             \end{align}
-%             where the sign of the Hessian corresponds to the sign of $q_d^T q$.
-
-% \section{Experiments} \label{sec:experiments}
-%     Code for all experiments is available on 
-%     \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
-%     {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.
-
-%     \subsection{Trajectory Optimization}
-%         In this section we present several trajectory optimization problems solved using
-%         our modified version of ALTRO. The two quadrotor examples use 101 knot points for
-%         5 second trajectories, and the airplane example uses 51 knot points for a 1.25
-%         second trajectory. All results were run on a desktop computer with an Intel
-%         i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
-%         experiments are given in Fig. \ref{fig:timing_chart}.
-
-%         In all examples, MLQR is implemented with a unit quaternion attitude state and
-%         Rodrigues parameters as the error state. The first example compares the two cost
-%         functions given in Sec. \ref{sec:cost_functions}, while the last two use the
-%         geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
-%         \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
-%         from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
-%         vectors). All examples re-normalize the quaternion in the dynamics function and
-%         use a third-order explicit Runge-Kutta integrator.
-
-%         \begin{figure}
-%             \centering
-%             \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
-%             \caption{Timing results for the airplane barrell roll, quadrotor
-%                 flip, and quadrotor zig-zag examples using iLQR with
-%                 roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The
-%                 timing result for the Quadrotor flip with Euler angles is omitted
-%                 since it failed to converge.
-%             }
-%             \label{fig:timing_chart}
-%         \end{figure}
-
-%         \subsubsection{Quadrotor Zig-Zag}
-%             \begin{figure}[t]
-%                 \centering
-%                 \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
-%                 \caption{Zig-zag trajectory. The black quadrotors indicate
-%                     the reference trajectory generated by iMLQR. The yellow
-%                     quadrotors track the reference trajectory using time-varying
-%                     MLQR, while the red uses the HFCA controller from
-%                     \cite{watterson2020control}.
-%                 }
-%                 \label{fig:zig-zag}
-%             \end{figure}    
-%             \begin{figure}[t]
-%                 \centering
-%                 \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
-%                 \caption{Snapshots of the quadrotor flip trajectory. The
-%                     red-colored quadrotors represent the state near t=0 s and the
-%                     green-colored quadrotors represent the state near t=5.0 s
-%                 }
-%                 \label{fig:quad_flip}
-%             \end{figure}
-
-%         We set up a fairly simple trajectory optimization problem to achieve a
-%         ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of
-%         the form
-%         \begin{multline}
-%                   \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
-%                 + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) \\
-%                 + \half \sum_{k=1} u_k^T R u_k
-%         \end{multline}
-%         with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus
-%         \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and
-%         $(0,10,0)$. We compared the two cost functions from Sec.
-%         \ref{sec:cost_functions}, using a standard quadratic penalty on the
-%         non-quaternion states.
-%         \begin{align}
-%             \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
-%             \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
-%                 \pm \bar{q}^T q)
-%         \end{align}
-%         where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where
-%         $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular
-%         velocity.
-%         We used the following weighting matrices: 
-%         $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
-%         $Q_w = \text{diag}(10^3 I_3,I_9)$, 
-%         $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
-%         For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and
-%         terminal cost functions, respectively.
-
-%         The geodesic cost function performed significantly better than the error
-%         quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29
-%         iterations. In addition to better convergence behavior, the geodesic cost is
-%         about half as expensive to compute, which is unsurprising when comparing the
-%         gradients and Hessians in Sec. \ref{sec:cost_functions}.
-
-%         \subsubsection{Quadrotor Flip}
-% 	    Similar to the previous example, we use a set of waypoints to encourage the
-% 	    quadrotor to do a flip. The cost weights for this example were
-% 	    \begin{align*}
-% 	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \!\times\! 10^{-2}, 10^{-3} I_3,
-%     	        10^{-3} I_3, 10^{-2} I_3) \\
-% 	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
-% 	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
-% 	    \end{align*}
-% 	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
-% 	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
-% 	    to stay above the floor and move to a goal state 2 meters away in the $+y$
-% 	    direction. The solver is initialized with a dynamically infeasible trajectory
-% 	    that linearly interpolates between the initial and final states, rotating the
-% 	    quad around the x-axis a full \ang{360}.
-
-% 	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
-% 	    iMLQR within the ALTRO solver to handle constraints. The trajectory generated by
-% 	    iLQR with quaternions and quadratic costs was successful but had a slightly
-%         different trajectory. iLQR with Euler angles failed to solve the problem.
-
-%         \subsection{Airplane Barrel Roll}
-%         An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
-%         tasked to do a barrel roll by setting a high terminal cost for being upside-down,
-%         see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
-%         conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
-%         converged faster than the pure quaternion version. Despite the extra matrix
-%         multiplications and extra Hessian term, it also was faster per iteration than its
-%         iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
-%         better performance by correctly leveraging the structure of the rotation group
-%         during the optimization routine.
-
-%         \begin{figure}[h]
-%             \centering
-%             \includegraphics[height=3.5cm]{figures/barrellroll.png}
-%             \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
-%             \label{fig:barrellroll}
-%         \end{figure}
-
-% 	    \begin{figure}[h]
-% 	        \centering
-% 	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
-% 	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller.
-Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics.
-}
-% 	        \label{fig:tracking_err}
-% 	    \end{figure}
-%         \vspace{-20pt}
-
-% 	\subsection{Tracking Control}
-% 	    This section provides examples using MLQR for tracking and stabilizing control
-% 	    for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art
-% 	    geometric tracking controller \cite{watterson2020control} based on the Hopf
-% 	    Fibration.
-
-% 	    \subsubsection{MLQR Stabilization}
-% 	    To test robustness and the domain of attraction of the MLQR controller, a
-% 	    quadrotor is linearized about hover at the origin and initialized from states
-% 	    uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular
-% 	    velocities, and uniformly over the entire space of rotations. The success rates
-% 	    for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is
-% 	    defined as the terminal error (calculated using \eqref{eq:state_error}) being
-% 	    within a unit ball of radius 0.1 after 10.0 seconds. To further analyze the
-% 	    basins of attraction, we uniformly sampled over only roll (which goes singular at
-% 	    \ang{90} for our Euler angle representation), and yaw. The results are shown in
-% 	    Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the
-% 	    successfully stabilized initial conditions.
-
-% 	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also
-% 	    being more generally applicable, easier to tune, and easier to implement. It is
-% 	    also noteworthy that na\"ively using LQR with quaternions (incorrectly treating
-% 	    them as vectors) results in poorer performance than Euler angles, clearly
-% 	    demonstrating the value of the methodology presented in the current paper.
-
-% 	    \subsubsection{MLQR Tracking}
-% 	    Time-varying MLQR is used to track the zig-zag trajectory generated in the
-% 	    previous section. %The controller calculated the gains at $0.01 s$ intervals, and
-% 	    used zero-order-hold between intervals. To test tracking performance, zero-mean
-% 	    Gaussian noise was added as force and moments applied to the quadrotor.
-
-% 	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
-% 	    \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller
-% 	    dramatically out-performs HFCA, we found no significant difference in performance
-% 	    between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
-% 	    Since the tracking controller only deals with relatively small state and attitude
-% 	    errors, this is unsurprising. Additionally, we found the MLQR controller ran
-%         about 20x faster than the HFCA controller.
-
-% 	    \begin{figure*}[ht]
-% 	        \begin{minipage}{0.60\linewidth}
-%                 \centering
-%                 \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
-%                 \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black).
-All points within each contour were successfully stabilized by the controller.}
-%                 \label{fig:mlqr_basin}
-%             \end{minipage}
-% 	        \hspace{0.05\linewidth}
-%             \begin{minipage}{0.34\linewidth}
-%                 \centering
-%                 \begin{tabular}{cc}
-%                     \toprule
-%                     \textbf{Method} & \textbf{Success} \\
-%                     \midrule
-%                     Quat & 61\% \\
-%                     RPY  & 80\% \\
-%                     HFCA & 97\% \\
-%                     MLQR & 99\% \\
-%                     \bottomrule
-%                 \end{tabular}
-%                 \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
-%                 \label{tab:mlqr_success}
-%             \end{minipage}
-%         \end{figure*}
-
-% \section{Conclusions} \label{sec:conclusion}
-%     We have presented a general, unified method for planning and control on rigid-body
-%     systems with arbitrary attitude using standard linear algebra and vector calculus. By
-%     applying these methods to LQR to correctly account for the group structure of
-%     rotations, we have matched or exceeded the performance of a state-of-the-art
-%     geometric controller designed specifically for quadrotors, while being more general,
-%     requiring less system-dependent tuning, having less computational overhead, and being
-%     easier to implement.
-
-%     We have also demonstrated a straight-forward way to adapt nonlinear trajectory
-%     optimization techniques to work with quaternion-valued states. For these problems, we
-%     found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
-%     computationally efficient and provide excellent convergence in practice. We recommend
-%     the use of unit quaternions within planning and control algorithms for their
-%     simplicity, computational efficiency, and lack of singularities. We further recommend
-%     the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
-
-%     Future areas of research include the application of these methodologies to more
-%     complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
-%     more in-depth analysis of the convergence behavior of the algorithms proposed in the
-%     current work.
-
-% \paragraph{Acknowledgements}
-% This material is based upon work supported by the National Science Foundation Graduate
-% Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
-% conclusions or recommendations expressed in this material are those of the author(s) and
-% do not necessarily reflect the views of the National Science Foundation
+    on the ALTRO solver. 
+    Initial results suggest that our approach can significantly reduce solve times by 
+    improving convergence rates for problems that require large changes in attitude over
+    the course of the trajectory. Future work will extend these ideas to Newton-based methods
+    such as direct collocation and other domains outside of trajectory optimization.
 

### 2020-09-11 examples/ipopt_test.jl

@@ -1,0 +1,203 @@
+using TrajectoryOptimization
+using RobotDynamics, Rotations
+using Ipopt
+using LinearAlgebra, StaticArrays
+using ForwardDiff
+using SparseArrays
+using BenchmarkTools
+import RobotZoo.Quadrotor
+const TO = TrajectoryOptimization
+
+## Define the problem
+model = Quadrotor{UnitQuaternion}()
+n,m = size(model)
+N,tf = 51,5.0
+dt = tf/(N-1)
+
+# Objective
+Q = Diagonal(RobotDynamics.fill_state(model, 1,1,10,10.)*1e-3)
+R = Diagonal(SA[1,1,1,1.])
+Qf = (N-1)*Q*1e3
+x0 = zeros(model)[1] 
+xf = RobotDynamics.build_state(model, 
+    RBState([2,1,0.5], expm(SA[0,0,1]*pi/4), zeros(3), zeros(3))
+)
+utrim = zeros(model)[2]
+obj = LQRObjective(Q, R, Qf, xf, N, uf=utrim)
+prob = Problem(model, obj, xf, tf, x0=x0)
+initial_controls!(prob, utrim)
+rollout!(prob)
+initial_controls!(prob, utrim .+ 1)
+J0 = cost(prob)
+
+# Trajectory
+X = states(prob)
+U = controls(prob)
+Z = Float64[]
+for k = 1:N-1
+    append!(Z, X[k])
+    append!(Z, U[k])
+end
+append!(Z, X[N])
+
+## Solve with Altro
+using Altro
+solver = ALTROSolver(prob, show_summary=true)
+solve!(solver)
+
+## Solve with Ipopt
+function dynamics_constraint_structure(n,m,N)
+    P = (N-1)*n
+    NN = N*n + (N-1)*m
+    Gstruct = spzeros(Int,P,NN) 
+    ind1 = 1:n
+    ind2 = 1:n+m
+    blk = collect(reshape(1:(n+m)*n, n, n+m))
+    for k = 1:N-1
+        Gstruct[ind1,ind2] .= blk
+        blk .+= n*(n+m)
+        ind1 = ind1 .+ n
+        ind2 = ind2 .+ (n+m)
+    end
+    ind1 = 1:n
+    ind2 = ind1 .+ (n+m) 
+    idx = (N-1)*n*(n+m)
+    for k = 1:N-1
+        for (i,j) in zip(ind1,ind2)
+            idx += 1
[…]

### 2020-09-11 examples/quadrotor_flip.jl

@@ -1,52 +1,28 @@
-
+import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+using RobotDynamics
+import RobotZoo.Quadrotor
 using TrajectoryOptimization
+using PlanningWithAttitude
+using Altro
+using TrajOptPlots
+using BenchmarkTools
+const TO = TrajectoryOptimization
+
 using Random
 using StaticArrays
 using LinearAlgebra
 using MeshCat
-using TrajOptPlots
 using Rotations
-import TrajectoryOptimization: Rotation
-const TO = TrajectoryOptimization
-
-max_con_viol = 1.0e-8
-T = Float64
-verbose = true
-
-opts_ilqr = iLQRSolverOptions{T}(verbose=verbose,
-    cost_tolerance=1e-4,
-    iterations=300)
-
-opts_al = AugmentedLagrangianSolverOptions{T}(verbose=verbose,
-    opts_uncon=opts_ilqr,
-    iterations=40,
-    cost_tolerance=1.0e-5,
-    cost_tolerance_intermediate=1.0e-4,
-    constraint_tolerance=max_con_viol,
-    penalty_scaling=10.,
-    penalty_initial=1.)
-
-opts_altro = ALTROSolverOptions{T}(verbose=verbose,
-    opts_al=opts_al,
-    R_inf=1.0e-8,
-    resolve_feasible_problem=false,
-    projected_newton=false,
-    projected_newton_tolerance=1.0e-3)
 
 
-model = Dynamics.Quadrotor2{UnitQuaternion{Float64,VectorPart}}()
-n,m = 13,4
-
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
-end
-delete!(vis)
-set_mesh!(vis, model)
-
-function gen_quad_flip(Rot=UnitQuaternion{Float64,CayleyMap}; use_rot=Rot<:UnitQuaternion,
-        costfun=:auto)
-    model = Dynamics.Quadrotor2{Rot}(use_rot=use_rot)
+function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=false, 
+        renorm::Bool=false)
+    model = Quadrotor{Rot}()
+    if renorm
+        model = QuatRenorm(model)
+    end
     n,m = size(model)
+    m += slack
     rsize = n-9
 
     # discretization
[…]

### 2020-09-11 src/PlanningWithAttitude.jl

@@ -8,12 +8,14 @@ import TrajectoryOptimization
 const RD = RobotDynamics
 const TO = TrajectoryOptimization
 
-# include("vecmodel.jl")
 include("rotatedmodel.jl")
+include("vecmodel.jl")
 
 export
     QuatSlackModel,
-    UnitQuatConstraint
+    UnitQuatConstraint,
+    VecModel,
+    QuatRenorm
 
 # include("logger.jl")
 # include("knotpoint.jl")

### 2020-09-11 src/rotatedmodel.jl

@@ -1,3 +1,19 @@
+struct QuatRenorm{R,RK,L} <: RD.RigidBody{R}
+    model::L
+    function QuatRenorm(model::RD.RigidBody{R}, rk::Type{<:RD.QuadratureRule}=RD.RK4) where R
+        new{R,rk,typeof(model)}(model)
+    end
+end
+RD.control_dim(model::QuatRenorm) = RD.control_dim(model.model)
+
+function RD.discrete_dynamics(::Type{RD.RK4}, 
+        model::QuatRenorm{<:Any,RD.RK4}, x::StaticVector, u::StaticVector, t, dt)
+    x2 = RD.discrete_dynamics(RD.RK4, model.model, x, u, t, dt)
+    # renormalize the quaternion
+    r,q,v,ω = RobotDynamics.parse_state(model.model, x2, true)
+    RobotDynamics.build_state(model.model, r,q,v,ω)
+end
+
 struct QuatSlackModel{R,M,L} <: RD.RigidBody{R} 
     model::L
     function QuatSlackModel(model::L) where L <: RD.RigidBody
@@ -48,18 +64,18 @@ Base.copy(con::UnitQuatConstraint) = UnitQuatConstraint(con.n, con.m, con.qind, 
 function TO.evaluate(con::UnitQuatConstraint, z::RD.AbstractKnotPoint) 
     q = z.z[con.qind]
     s = z.z[con.sind]
-    q̄ = q*s
+    q̄ = q
     return SA[dot(q̄,q̄) - 1]
 end
 
 function TO.jacobian!(∇c, con::UnitQuatConstraint, z::RD.AbstractKnotPoint)
     q = z.z[con.qind]
     s = z.z[con.sind]
-    q̄ = q*s
+    q̄ = q
     for (i,j) in enumerate(con.qind)
-        ∇c[1,j] = 2*q̄[i]*s
+        ∇c[1,j] = 2*q̄[i]
     end
-    ∇c[con.sind] = 2s*dot(q,q)
+    # ∇c[con.sind] = 2s*dot(q,q)
     return false
 end
 

### 2020-09-11 src/vecmodel.jl

@@ -1,8 +1,8 @@
-struct VecModel{L} <: AbstractModel
+struct VecModel{L} <: RD.AbstractModel
     model::L
-    VecModel(model::L) where L <: LieGroupModel = new{L}(model)
+    VecModel(model::L) where L <: RD.LieGroupModel = new{L}(model)
 end
 
-RobotDynamics.state_dim(model::VecModel) = state_dim(model.model)
-RobotDynamics.control_dim(model::VecModel) = control_dim(model.model)
-RobotDynamics.dynamics(model::VecModel, args...) = dynamics(model.model, args...) 
+RobotDynamics.state_dim(model::VecModel) = RD.state_dim(model.model)
+RobotDynamics.control_dim(model::VecModel) = RD.control_dim(model.model)
+RobotDynamics.dynamics(model::VecModel, args...) = RD.dynamics(model.model, args...) 

### 2020-09-14 paper/main.tex
IROS submission

@@
 \section{INTRODUCTION}
+
     Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous
     underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
-    translations and rotations as part of their normal operation. 
-    While representing translations is straightforward and intuitive, effectively 
-    representing the nontrivial group structure of 3D rotations has been a topic of study 
-    for many decades. Although we can intuitively deduce that rotations are three-dimensional, 
-    a globally non-singular three-parameter representation of the space of rotations does 
-    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing rotations, 
-    we must either a) pick a three-parameter representation and deal with discontinuities, 
-    or b) pick a higher-dimensional representation and deal with constraints between the 
-    parameters. While simply representing attitude is nontrivial, generating and tracking 
-    motion plans for floating-base systems is an even more challenging problem.
+    translations and rotations as part of their normal operation. While representing
+    translations is straightforward and intuitive, effectively representing the
+    nontrivial group structure of 3D rotations has been a topic of study for many
+    decades. Although we can intuitively deduce that rotations are three-dimensional, a
+    globally non-singular three-parameter representation of the space of rotations does
+    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing
+    rotations, we must either a) pick a three-parameter representation and deal with
+    discontinuities, or b) pick a higher-dimensional representation and deal with
+    constraints between the parameters. While simply representing attitude is nontrivial,
+    generating and tracking motion plans for floating-base systems is an even more
+    challenging problem.
 
@@
     \cite{Baillieul1978}. Effective attitude tracking controllers have been developed for
-    satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,
+    satellites \cite{wie1985quaternion}, quadrotors
+    \cite{Fresk2013,Liu2015,lee2010geometric,
     Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum
@@
 
-    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, 
-    sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control. 
-    Approaches to optimal control on attitude problems include analytical methods applied to
-    satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,
-    Lee2008}, a combination of sampling-based planning and constrained trajectory 
-    optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators 
-    \cite{Saccon2013}, or more general theory for optimization on manifolds 
-    \cites{watterson2018trajectory}. 
-    Nearly all of these methods rely heavily on principles from differential geometry and 
-    Lie group theory; however, despite these works, many recent papers in the robotics 
-    community continue to apply traditional methods for motion planning and control with no 
-    regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
+    More recently, these ideas have been extended to trajectory generation
+    \cite{Zefran1998}, sample-based motion planning \cite{Zefran1999,Kuffner2004}, and
+    optimal control. Approaches to optimal control on attitude problems include
+    analytical methods applied to satellites \cite{Spindler1998}, discrete mechanics
+    \cite{Kobilarov2011,Kobilarov2014, Lee2008}, a combination of sampling-based planning
+    and constrained trajectory optimization for satellite formations \cite{Garcia2005,
+    Aoude2008}, projection operators \cite{Saccon2013}, or more general theory for
+    optimization on manifolds \cites{watterson2018trajectory}. Nearly all of these
+    methods rely heavily on principles from differential geometry and Lie group theory;
+    however, despite these works, many recent papers in the robotics community continue
+    to apply traditional methods for motion planning and control with no regard for the
+    group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
     Williams2017,Geisert2016}.

### 2020-09-14 test/quatslack.jl

@@ -9,6 +9,20 @@ using ForwardDiff
 import RobotZoo.Quadrotor
 const TO = TrajectoryOptimization
 
+## Test renorm
+quad = Quadrotor()
+x,u = rand(quad)
+x2 = discrete_dynamics(RK4,quad,x,u,0.0,0.1)
+q  = orientation(quad, x)
+q2 = orientation(quad, x2)
+norm(q)
+norm(q2)
+
+model = QuatRenorm(quad,RK4)
+x3 = discrete_dynamics(RK4, model, x, u, 0.0, 0.1)
+q3 = orientation(quad, x3)
+
+## Test model
 quad = Quadrotor()
 model = QuatSlackModel(quad)
 @test state_dim(quad) == 13 == state_dim(model)
@@ -22,25 +36,34 @@ u2 = push(u, s)
 @test dynamics(model, x2, u2) ≈ dynamics(quad, x, u)
 @test dynamics(model, x, u2) ≈ dynamics(quad, x3, u)
 
+## Test constraint
 con = UnitQuatConstraint(model)
 z = KnotPoint(x2,u2,0.1)
 z2 = KnotPoint(x,u2,0.1)
-@test TO.evaluate(con,z)[1] ≈ 0 atol=1e-12
-@test TO.evaluate(con,z2)[1] ≈ s^2 - 1 atol=1e-12
+@test TO.evaluate(con,z)[1] ≈ 1/s^2-1 atol=1e-9
+@test TO.evaluate(con,z2)[1] ≈ 0 atol=1e-12
 
 c(x) = TO.evaluate(con, StaticKnotPoint(z, x))
 ∇c = ForwardDiff.jacobian(c, z.z)
 ∇c2 = zeros(1,18)
 TO.jacobian!(∇c2, con, z)
-@test ∇c[4:7] ≈ 2*Rotations.params(q)*s
+q1 = z.z[4:7]
+@test ∇c2[4:7] ≈ 2*q1
 @test ∇c2 ≈ ∇c
 
 ∇c = ForwardDiff.jacobian(c, z2.z)
 TO.jacobian!(∇c2, con, z2)
-@test ∇c[4:7] ≈ 2*Rotations.params(q)*s*s
+@test ∇c2[4:7] ≈ 2*Rotations.params(q)
 @test ∇c2 ≈ ∇c
 
+## Test VecModel
+model2 = VecModel(model)
+@test dynamics(model2, x2, u2) ≈ dynamics(model, x2, u2)
[…]

### 2020-10-08 examples/Flexible_Spacecraft/flexible_spacecraft_dynamics.jl
added flexible spacecraft example

@@ -1,0 +1,120 @@
+# post processing
+function post_process(solver)
+    X = states(solver)
+    U = controls(solver)
+    X = Vector.(X)
+    U = Vector.(U)
+
+    pts = length(X)-1
+    q_hist = zeros(4,pts)
+    ω_hist = zeros(3,pts)
+    u_hist = zeros(4,pts)
+    η_hist = zeros(pts)
+    ηd_hist = zeros(pts)
+    θ_hist = zeros(pts)
+    r_hist = zeros(4,pts)
+    for i = 1:pts
+        ω_hist[:,i] = X[i][1:3]
+        q_hist[:,i] = X[i][4:7]
+        η_hist[i] = X[i][8]
+        ηd_hist[i] = X[i][9]
+        r_hist[:,i] = X[i][10:13]
+        u_hist[:,i] = U[i][1:4]
+
+        θ_hist[i] = 2*atan(norm(q_hist[2:4,i]),q_hist[1,i])
+    end
+
+    return ω_hist, q_hist, u_hist,η_hist, ηd_hist, r_hist, θ_hist
+end
+
+import RobotDynamics: dynamics, forces, moments, wrenches, mass_matrix, inertia, inertia_inv, orientation
+import RobotDynamics: state_dim, control_dim
+
+
+
+
+import Rotations: lmult, vmat, hmat
+
+struct FlexSatellite <: LieGroupModel
+    J::SArray{Tuple{3,3},Float64,2,9}
+    B::SArray{Tuple{3,4},Float64,2,12}
+    C::SArray{Tuple{3,3},Float64,2,9}
+    K::SArray{Tuple{3,3},Float64,2,9}
+    δ::SArray{Tuple{3,3},Float64,2,9}
+    first_inv::SArray{Tuple{3,3},Float64,2,9}
+    ϕMΣ::SArray{Tuple{3,3},Float64,2,9}
+end
+
+
+# units of kg-m
+ϕMΣ  = [0 1 0;
+       1 0 0;
+       0 .2 -.8];
+
+ϕMΣ = SMatrix{3,3}(ϕMΣ)
+
+
+# units of kg-m^2
+δ =  [0 0 1;
+         0 1 0;
+        -.7 .1 .1]
+δ = copy(transpose(δ))
+δ = SMatrix{3,3}(δ)
+
+J = diagm([1;2;3])
+
+J = SMatrix{3,3}(J)
+
+mass = 28.54*14.5939
+
[…]

### 2020-10-08 examples/Flexible_Spacecraft/solve_con.jl

@@ -1,0 +1,161 @@
+cd(dirname(@__FILE__))
+Pkg.activate(".")
+using MATLAB
+using StaticArrays
+using Parameters
+using RobotDynamics
+using Rotations
+using LinearAlgebra
+# using Plots
+using TrajectoryOptimization
+using Altro
+
+
+include(joinpath(dirname(@__FILE__),"flexible_spacecraft_dynamics.jl"))
+
+const TO = TrajectoryOptimization
+
+
+# Discretization
+tf = 150.0
+N = 501
+dt = tf/(N-1)
+
+# Model
+model = FlexSatellite()
+n,m = size(model)
+
+# quaternion to dcm functions
+function skew(v)
+	"""Skew-symmetric matrix from 3 element array"""
+    return @SMatrix [0 -v[3] v[2]; v[3] 0 -v[1]; -v[2] v[1] 0]
+end
+function dcm_from_q(q)
+	"""DCM from quaternion, scalar first"""
+	s = @views q[1]
+	v = @views q[2:4]
+    return I + 2*skew(v)*(s*I + skew(v))
+end
+
+
+struct AttitudeKeepOut{T} <: TO.StateConstraint
+	n::Int
+	keepoutdir::SVector{3,T}
+	bodyvec::SVector{3,T}
+	keepoutangle::Float64
+	function AttitudeKeepOut(n::Int, keepoutdir::SVector, bodyvec::SVector, keepoutangle::T) where T
+		new{T}(n,keepoutdir,bodyvec,keepoutangle)
+	end
+end
+
+TO.state_dim(con::AttitudeKeepOut) = con.n
+TO.sense(::AttitudeKeepOut) = Inequality()
+Base.length(::AttitudeKeepOut) = 1
+TO.evaluate(con::AttitudeKeepOut, x::SVector) = (
+SA[ dot(dcm_from_q(x[4:7])*con.bodyvec,con.keepoutdir)-cosd(con.keepoutangle)])
+
+# Initial and final states
+ω = @SVector zeros(3)
+# q0 = Rotations.params(UnitQuaternion(I))
+q0 = Rotations.params(expm(deg2rad(150) * normalize(@SVector [1,2,3])))
+# qf = Rotations.params(expm(deg2rad(90) * normalize(@SVector rand(3))))
+qf = Rotations.params(UnitQuaternion(I))
[…]

### 2020-10-08 examples/Flexible_Spacecraft/solve_uncon.jl

@@ -1,0 +1,129 @@
+cd(dirname(@__FILE__))
+Pkg.activate(".")
+
+using StaticArrays
+using Parameters
+using RobotDynamics
+using Rotations
+using LinearAlgebra
+# using Plots
+using TrajectoryOptimization
+using Altro
+using MATLAB
+
+include(joinpath(dirname(@__FILE__),"flexible_spacecraft_dynamics.jl"))
+
+const TO = TrajectoryOptimization
+
+
+# Discretization
+tf = 150.0
+N = 501
+dt = tf/(N-1)
+
+# Model
+model = FlexSatellite()
+n,m = size(model)
+
+# quaternion to dcm functions
+function skew(v)
+	"""Skew-symmetric matrix from 3 element array"""
+    return @SMatrix [0 -v[3] v[2]; v[3] 0 -v[1]; -v[2] v[1] 0]
+end
+function dcm_from_q(q)
+	"""DCM from quaternion, scalar first"""
+	s = @views q[1]
+	v = @views q[2:4]
+    return I + 2*skew(v)*(s*I + skew(v))
+end
+
+
+struct AttitudeKeepOut{T} <: TO.StateConstraint
+	n::Int
+	keepoutdir::SVector{3,T}
+	bodyvec::SVector{3,T}
+	keepoutangle::Float64
+	function AttitudeKeepOut(n::Int, keepoutdir::SVector, bodyvec::SVector, keepoutangle::T) where T
+		new{T}(n,keepoutdir,bodyvec,keepoutangle)
+	end
+end
+
+TO.state_dim(con::AttitudeKeepOut) = con.n
+TO.sense(::AttitudeKeepOut) = Inequality()
+Base.length(::AttitudeKeepOut) = 1
+TO.evaluate(con::AttitudeKeepOut, x::SVector) = (
+SA[ dot(dcm_from_q(x[4:7])*con.bodyvec,con.keepoutdir)-cosd(con.keepoutangle)])
+
+# Initial and final states
+ω = @SVector zeros(3)
+q0 = Rotations.params(expm(deg2rad(150) * normalize(@SVector [1,2,3])))
+qf = Rotations.params(UnitQuaternion(I))
+
+r0 = zeros(4)
+x0 = [ω; q0;zeros(3);zeros(3);r0]
+xf = [ω; qf;zeros(3);zeros(3);r0]
+
+# Objective
[…]

### 2020-10-09 examples/yak_barrellroll.jl
Barrell Roll examples

@@ -1,40 +1,147 @@
-
+import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+using RobotDynamics
+import RobotZoo.Quadrotor
 using TrajectoryOptimization
-const TO = TrajectoryOptimization
+using Altro
 using BenchmarkTools
-using TrajOptPlots
-using LinearAlgebra
+const TO = TrajectoryOptimization
+const RD = RobotDynamics
+
+using Random
 using StaticArrays
-using FileIO
-using DataFrames
-using MeshCat
-using GeometryTypes
-using CoordinateTransformations
+using LinearAlgebra
+using Rotations
+
 using PlanningWithAttitude
+using JLD2
+# using TrajOptPlots
+# using MeshCat
 
-# Start visualizer
-Rot = MRP{Float64}
-model = Dynamics.YakPlane(Rot)
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
+args = (
+    integration = RK3,
+    termcon = :quatvec,
+    projected_newton = false,
+    show_summary=true
+)
+
+##
+"""
+Compare Cost Functions
+"""
+
+function run_example(;vecstate=false, costfun=:Quadratic, args...)
+    prob,opts = YakProblems(vecstate=vecstate, costfun=costfun; args...) 
+    solver = ALTROSolver(prob, opts)
+    b = benchmark_solve!(solver)
+    push!(cost_comparison[:costfun], costfun)
+    push!(cost_comparison[:errstate], !vecstate)
+    push!(cost_comparison[:time], median(b).time)
+    push!(cost_comparison[:iters], iterations(solver))
+    push!(cost_comparison[:cost], solver.stats.cost)
+    push!(cost_comparison[:c_max], solver.stats.c_max)
+    push!(cost_comparison[:is_outer], solver.stats.iteration_outer)
 end
-delete!(vis)
-set_mesh!(vis, model)
 
-function gen_barrellroll(Rot; kwargs...)
-    prob = Problems.YakProblems(Rot, scenario=:barrellroll; kwargs...)
-    solver = iLQRSolver(prob)
-    solver.opts.verbose = false
-    solver
+cost_comparison = Dict(
+    :costfun => Symbol[], 
+    :errstate => Bool[], 
+    :time => Float64[], 
+    :iters => Int[],
+    :cost => Vector{Float64}[],
[…]

### 2020-10-09 src/PlanningWithAttitude.jl

@@ -3,19 +3,30 @@ module PlanningWithAttitude
 using StaticArrays
 using LinearAlgebra
 using Rotations
+using RobotZoo
+using Altro
+using TrajectoryOptimization
 import RobotDynamics
-import TrajectoryOptimization
 const RD = RobotDynamics
 const TO = TrajectoryOptimization
 
-include("rotatedmodel.jl")
+# include("rotatedmodel.jl")
 include("vecmodel.jl")
+include("airplane_problem.jl")
+include("quat_cons.jl")
+include("quat_costs.jl")
+include("quat_norm.jl")
 
 export
-    QuatSlackModel,
-    UnitQuatConstraint,
     VecModel,
-    QuatRenorm
+    YakProblems,
+    QuatGeoCon,
+    QuatErr,
+    QuatVecEq,
+    ErrorQuadratic,
+    LieLQR
+
+
 
 # include("logger.jl")
 # include("knotpoint.jl")

### 2020-10-09 src/airplane_problem.jl

@@ -1,0 +1,109 @@
+using TrajectoryOptimization: LieLQRCost
+using ForwardDiff
+
+function YakProblems(;
+        integration=RD.RK4,
+        N = 101,
+        vecstate=false,
+        scenario=:barrellroll, 
+        costfun=:Quadratic, 
+        termcon=:goal,
+        quatnorm=:none,
+        kwargs...
+    )
+    model = RobotZoo.YakPlane(UnitQuaternion)
+
+    opts = SolverOptions(
+        cost_tolerance_intermediate = 1e-1,
+        penalty_scaling = 100.,
+        penalty_initial = 0.01;
+        kwargs...
+    )
+
+    s = RD.LieState(model)
+    n,m = size(model)
+    rsize = size(model)[1] - 9
+    vinds = SA[1,2,3,8,9,10,11,12,13]
+
+    # Discretization
+    tf = 1.25
+    dt = tf/(N-1)
+
+    if scenario == :barrellroll
+        ey = @SVector [0,1,0.]
+
+        # Initial and final condition
+        p0 = MRP(0.997156, 0., 0.075366) # initial orientation
+        pf = MRP(0., -0.0366076, 0.) # final orientation (upside down)
+
+        x0 = RD.build_state(model, [-3,0,1.5], p0, [5,0,0], [0,0,0])
+        utrim  = @SVector  [41.6666, 106, 74.6519, 106]
+        xf = RD.build_state(model, [3,0,1.5], pf, [5,0,0], [0,0,0])
+
+        # Objective
+        Qf_diag = RD.fill_state(model, 100, 500, 100, 100.)
+        Q_diag = RD.fill_state(model, 0.1, 0.1, 0.1, 0.1)
+        Qf = Diagonal(Qf_diag)
+        Q = Diagonal(Q_diag)
+        R = Diagonal(@SVector fill(1e-3,4))
+        if quatnorm == :slack
+            m += 1
+            R = Diagonal(push(R.diag, 1e-6))
+            utrim = push(utrim, 0)
+        end
+        if costfun == :Quadratic
+            costfun = LQRCost(Q, R, xf, utrim)
+            costterm = LQRCost(Qf, R, xf, utrim)
+        elseif costfun == :QuatLQR
+            # costfun = LieLQRCost(s, Q, R, xf, utrim; w=0.1)
+            # costterm = LieLQRCost(s, Qf, R, xf, utrim; w=200.0)
+            costfun = QuatLQRCost(Q, R, xf, utrim; w=0.1)
[…]

### 2020-10-09 src/quat_costs.jl

@@ -1,0 +1,394 @@
+using TrajectoryOptimization: QuadraticCostFunction, CostFunction
+import RobotDynamics: state_dim, control_dim
+import TrajectoryOptimization: is_blockdiag, is_diag, 
+    stage_cost, gradient!, hessian!, change_dimension
+import Base: +
+
+function LieLQR(s::RD.LieState{Rot,P}, Q::Diagonal{<:Any,<:StaticVector}, R::Diagonal, 
+        xf, uf=zeros(size(R,1)); kwargs...) where {Rot,P}
+    if Rot <: UnitQuaternion && length(Q.diag) == 13
+        Qd = deleteat(Q.diag, 4)
+        Q = Diagonal(Qd)
+    end
+    n = length(s)
+    n̄ = RD.state_diff_size(s) 
+    G = zeros(n,n̄)
+    RD.state_diff_jacobian!(G, s, xf)
+    Q_ = G*Q*G' 
+    # q = -Q_*xf
+    # r = -R*uf
+    # c = 0.5*xf'Q_*xf + 0.5*ur'R*uf
+    LQRCost(Q_, R, xf, uf; kwargs...)
+end
+
+############################################################################################
+#                        QUADRATIC QUATERNION COST FUNCTION
+############################################################################################
+
+struct DiagonalQuatCost{N,M,T,N4} <: QuadraticCostFunction{N,M,T}
+    Q::Diagonal{T,SVector{N,T}}
+    R::Diagonal{T,SVector{M,T}}
+    q::SVector{N,T}
+    r::SVector{M,T}
+    c::T
+    w::T
+    q_ref::SVector{4,T}
+    q_ind::SVector{4,Int}
+    Iq::SMatrix{N,4,T,N4}
+    function DiagonalQuatCost(Q::Diagonal{T,SVector{N,T}}, R::Diagonal{T,SVector{M,T}},
+            q::SVector{N,T}, r::SVector{M,T}, c::T, w::T,
+            q_ref::SVector{4,T}, q_ind::SVector{4,Int}) where {T,N,M}
+        Iq = @MMatrix zeros(N,4)
+        for i = 1:4
+            Iq[q_ind[i],i] = 1
+        end
+        Iq = SMatrix{N,4}(Iq)
[…]

### 2020-10-09 src/rotatedmodel.jl

@@ -1,85 +1,0 @@
-struct QuatRenorm{R,RK,L} <: RD.RigidBody{R}
-    model::L
-    function QuatRenorm(model::RD.RigidBody{R}, rk::Type{<:RD.QuadratureRule}=RD.RK4) where R
-        new{R,rk,typeof(model)}(model)
-    end
-end
-RD.control_dim(model::QuatRenorm) = RD.control_dim(model.model)
-
-function RD.discrete_dynamics(::Type{RD.RK4}, 
-        model::QuatRenorm{<:Any,RD.RK4}, x::StaticVector, u::StaticVector, t, dt)
-    x2 = RD.discrete_dynamics(RD.RK4, model.model, x, u, t, dt)
-    # renormalize the quaternion
-    r,q,v,ω = RobotDynamics.parse_state(model.model, x2, true)
-    RobotDynamics.build_state(model.model, r,q,v,ω)
-end
-
-struct QuatSlackModel{R,M,L} <: RD.RigidBody{R} 
-    model::L
-    function QuatSlackModel(model::L) where L <: RD.RigidBody
-        M = RD.control_dim(model)
-        R = RD.rotation_type(model)
-        new{R,M,L}(model)
-    end
-end
-RobotDynamics.control_dim(::QuatSlackModel{<:Any,M}) where M = M+1 
-
-function RobotDynamics.dynamics(model::QuatSlackModel{<:Any,m}, x, u, t=0) where m
-    uinds = SVector{m}(1:m)
-    u0 = u[uinds]                    # original controls
-    s = u[m+1]                       # Quaternion slack
-    r,q,v,ω = RobotDynamics.parse_state(model, x)
-    q̄ = s*q                          # scale the quaternion by the extra control
-    x̄ = RobotDynamics.build_state(model, r, q̄, v, ω)
-    RD.dynamics(model.model, x̄, u, t)   # call the original dynamics with the scaled quaternion
-end
-
-"""
-    UnitQuatConstraint
-
-Enforce a that the unit quaternion in the state, at indices `qind`, has unit norm
-after being multiplied by a slack variable at index `sind` in the joint state-control vector
-`z = [x;u]`.
-"""
-struct UnitQuatConstraint <: TrajectoryOptimization.StageConstraint
-    n::Int
-    m::Int
[…]

### 2020-10-09 examples/quadrotor_flip.jl
Quadrotor flip working well

@@ -14,9 +14,9 @@ using LinearAlgebra
 using MeshCat
 using Rotations
 
-
+##
 function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=false, 
-        renorm::Bool=false)
+        renorm::Bool=false, costfun=LQRCost)
     model = Quadrotor{Rot}()
     if renorm
         model = QuatRenorm(model)
@@ -60,6 +60,11 @@ function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=fal
     ]
     # times = [35, 41, 47, 51, 55, 61, 70, 101]
     times = [45, 51, 55, 75, 101]
+
+    """
+    Costs
+    """
+    # intermediate costs
     Qw_diag = RobotDynamics.build_state(model, 
         [1e3,1e1,1e3], 
         (@SVector fill(5e4,rsize)), 
@@ -67,8 +72,9 @@ function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=fal
     )
     Qf_diag = RobotDynamics.fill_state(model, 10., 100, 10, 10)
     xf = RobotDynamics.build_state(model, wpts[end][1], wpts[end][2], zeros(3), zeros(3))
-    cost_nom = LQRCost(Diagonal(Q_diag), R, xf)
+    cost_nom = costfun(Diagonal(Q_diag), R, xf)
 
+    # waypoint costs
     costs = map(1:length(wpts)) do i
         r,q = wpts[i]
         xg = RobotDynamics.build_state(model, r, q, v_nom, [2pi/3, 0, 0])
@@ -79,7 +85,7 @@ function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=fal
             Q = Diagonal(1e-3*Qw_diag)
             w = 100.
         end
-        LQRCost(Q, R, xg)
+        costfun(Q, R, xg)
     end
 
     costs_all = map(1:N) do k
@@ -90,13 +96,11 @@ function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=fal
             cost_nom
         end
     end
-
     obj = Objective(costs_all)
 
     # Constraints
     conSet = ConstraintList(n,m,N)
-    noquat = deleteat(SVector{13}(1:13),4) 
[…]

### 2020-10-09 src/PlanningWithAttitude.jl

@@ -24,7 +24,9 @@ export
     QuatErr,
     QuatVecEq,
     ErrorQuadratic,
-    LieLQR
+    QuatLQRCost,
+    LieLQR,
+    QuatRenorm
 
 
 

### 2020-10-09 src/quat_costs.jl

@@ -35,15 +35,16 @@ struct DiagonalQuatCost{N,M,T,N4} <: QuadraticCostFunction{N,M,T}
     q_ref::SVector{4,T}
     q_ind::SVector{4,Int}
     Iq::SMatrix{N,4,T,N4}
+    terminal::Bool
     function DiagonalQuatCost(Q::Diagonal{T,SVector{N,T}}, R::Diagonal{T,SVector{M,T}},
             q::SVector{N,T}, r::SVector{M,T}, c::T, w::T,
-            q_ref::SVector{4,T}, q_ind::SVector{4,Int}) where {T,N,M}
+            q_ref::SVector{4,T}, q_ind::SVector{4,Int}; terminal::Bool=false) where {T,N,M}
         Iq = @MMatrix zeros(N,4)
         for i = 1:4
             Iq[q_ind[i],i] = 1
         end
         Iq = SMatrix{N,4}(Iq)
-        return new{N,M,T,N*4}(Q, R, q, r, c, w, q_ref, q_ind, Iq)
+        return new{N,M,T,N*4}(Q, R, q, r, c, w, q_ref, q_ind, Iq, terminal)
     end
 end
 
@@ -102,20 +103,20 @@ function change_dimension(cost::DiagonalQuatCost, n, m, ix, iu)
     q[ix] = cost.q
     r[iu] = cost.r
     qind = (1:n)[ix[cost.q_ind]]
-    DiagonalQuatCost(Diagonal(Q_diag), Diagonal(R_diag), q, r, cost.c, cost.w,
-        cost.q_ref, q_ind)
+    DiagonalQuatCost(Diagonal(SVector{n}(Qd)), Diagonal(SVector{m}(Rd)), 
+        SVector{n}(q), SVector{m}(r), cost.c, cost.w, cost.q_ref, qind)
 end
 
-function (+)(cost1::DiagonalQuatCost, cost2::QuadraticCost)
+function (+)(cost1::DiagonalQuatCost, cost2::TO.QuadraticCostFunction)
     @assert state_dim(cost1) == state_dim(cost2)
     @assert control_dim(cost1) == control_dim(cost2)
-    @assert norm(cost2.H) ≈ 0
+    is_diag(cost2) || @assert norm(cost2.H) ≈ 0
     DiagonalQuatCost(cost1.Q + cost2.Q, cost1.R + cost2.R,
[…]

### 2020-10-09 examples/quadrotor_flip.jl
Quadrotor flip example timing results

@@ -14,164 +14,21 @@ using LinearAlgebra
 using MeshCat
 using Rotations
 
-##
-function gen_quad_flip(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=false, 
-        renorm::Bool=false, costfun=LQRCost)
-    model = Quadrotor{Rot}()
-    if renorm
-        model = QuatRenorm(model)
-    end
-    n,m = size(model)
-    m += slack
-    rsize = n-9
-
-    # discretization
-    N = 101 # number of knot points
-    tf = 5.0
-    dt = tf/(N-1) # total time
-
-    # Initial condition
-    x0_pos = @SVector [0., -1., 1.]
-    x0 = RobotDynamics.build_state(model, x0_pos, UnitQuaternion(I), zeros(3), zeros(3))
-
-    # cost
-    Q_diag = RobotDynamics.build_state(model, 
-        [1e-2,1e-2,5e-2], 
-        fill(1e-5,rsize), 
-        fill(1e-3,3), 
-        fill(1e-2,3)
-    )
-    R = Diagonal(@SVector fill(1e-3,m))
-    q_nom = UnitQuaternion(I)
-    v_nom, ω_nom = zeros(3), zeros(3)
-    x_nom = RobotDynamics.build_state(model, zeros(3), q_nom, v_nom, ω_nom)
-
-    # waypoints
-    ex = @SVector [1.0, 0, 0]
-    wpts = [
-        ((@SVector [0, 0.5, 1.5,]),   expm(ex*deg2rad(90))),
-        # ((@SVector [0, 0.2, 1.5]),    expm(ex*deg2rad(90))),
-        # ((@SVector [0, 0.0, 2.0]),    expm(ex*deg2rad(135))),
-        ((@SVector [0, 0.0, 2.5]),    expm(ex*deg2rad(180))),
-        # ((@SVector [0, 0.0, 2.0]),    expm(ex*deg2rad(225))),
-        ((@SVector [0,-0.5, 1.5]),    expm(ex*deg2rad(-90))),
-        ((@SVector [0, 0.65, 1.0]),    expm(ex*deg2rad(360))),
-        ((@SVector [0, 1.0, 1.0]),    expm(ex*deg2rad(360))),
-    ]
-    # times = [35, 41, 47, 51, 55, 61, 70, 101]
-    times = [45, 51, 55, 75, 101]
-
-    """
-    Costs
-    """
-    # intermediate costs
[…]

### 2020-10-09 examples/yak_barrellroll.jl

@@ -4,6 +4,7 @@ import RobotZoo.Quadrotor
 using TrajectoryOptimization
 using Altro
 using BenchmarkTools
+using LaTeXTabulars
 const TO = TrajectoryOptimization
 const RD = RobotDynamics
 
@@ -64,6 +65,7 @@ run_example(costfun=:ErrorQuadratic; args...)
 ## Plots
 @load "cost_comparison.jld2" cost_comparison
 using Plots
+using PGFPlotsX 
 pgfplotsx()
 
 # Append initial constraint violation for plot
@@ -77,12 +79,55 @@ end
 is_outer = map(cost_comparison[:is_outer]) do iters
     Bool.(insert!(diff(iters),1,0))
 end
+y = cost_comparison[:c_max][1]
+y2 = cost_comparison[:c_max][3]
+x = 1:length(y)
 
-p = plot(cost_comparison[:c_max][1], label="original")
-plot!(cost_comparison[:c_max][3], label="modified",
-    xlabel="iteration", ylabel="constraint violation", yscale=:log10
+p = @pgf Axis(
+    {
+        xlabel="iterations",
+        ylabel="contraint satisfaction",
+        "ymode=log",
+        xmajorgrids,
+        ymajorgrids,
+    },
+    Plot(
+        {
+            color="cyan",
+            no_marks,
+            "very thick"
+        },
+        Coordinates(x,y)
+    ),
+    PlotInc(
+        {
+            color="orange",
+            no_marks,
+            "very thick"
+        },
+        Coordinates(1:length(y2), y2)
+    ),
+    Legend("original","modified")
 )
-savefig(p, "figs/c_max_convergence.tikz")
+pgfsave("paper/figures/c_max_convergence.tikz", p, include_preamble=false)
+
+# Export the table
+using DataFrames
+df = DataFrame(cost_comparison)
+df.time .= round.(df.time/1e6, digits=2)
+tab = df[:,[:costfun,:iters,:time]]
+tab.costfun = String.(tab.costfun)
+latex_tabular("paper/figures/cost_comparison.tex",
+    Tabular("lll"),
+    [
+        Rule(:top),
+        ["Cost Function", "Iterations", "time"],
+        Rule(:mid),
+        Matrix(tab),
+        Rule(:bottom)
+    ]
+)
+# savefig(p, "figs/c_max_convergence.tikz")
 
 ## 
 """

### 2020-10-09 src/PlanningWithAttitude.jl

@@ -12,14 +12,16 @@ const TO = TrajectoryOptimization
 
 # include("rotatedmodel.jl")
 include("vecmodel.jl")
-include("airplane_problem.jl")
 include("quat_cons.jl")
 include("quat_costs.jl")
 include("quat_norm.jl")
+include("airplane_problem.jl")
+include("quadflip_problem.jl")
 
 export
     VecModel,
     YakProblems,
+    QuadFlipProblem,
     QuatGeoCon,
     QuatErr,
     QuatVecEq,

### 2020-10-09 src/quadflip_problem.jl

@@ -1,0 +1,138 @@
+using RobotZoo: Quadrotor
+
+function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=false, 
+        renorm::Bool=false, costfun=LQRCost, integration=RD.RK4)
+    model = Quadrotor{Rot}()
+    if renorm
+        model = QuatRenorm(model)
+    end
+    n,m = size(model)
+    m += slack
+    rsize = n-9
+
+    # discretization
+    N = 101 # number of knot points
+    tf = 5.0
+    dt = tf/(N-1) # total time
+
+    # Initial condition
+    x0_pos = @SVector [0., -1., 1.]
+    x0 = RobotDynamics.build_state(model, x0_pos, UnitQuaternion(I), zeros(3), zeros(3))
+
+    # cost
+    Q_diag = RobotDynamics.build_state(model, 
+        [1e-2,1e-2,5e-2], 
+        fill(1e-5,rsize), 
+        fill(1e-3,3), 
+        fill(1e-2,3)
+    )
+    R = Diagonal(@SVector fill(1e-3,m))
+    q_nom = UnitQuaternion(I)
+    v_nom, ω_nom = zeros(3), zeros(3)
+    x_nom = RobotDynamics.build_state(model, zeros(3), q_nom, v_nom, ω_nom)
+
+    # waypoints
+    ex = @SVector [1.0, 0, 0]
+    wpts = [
+        ((@SVector [0, 0.5, 1.5,]),   expm(ex*deg2rad(90))),
+        # ((@SVector [0, 0.2, 1.5]),    expm(ex*deg2rad(90))),
+        # ((@SVector [0, 0.0, 2.0]),    expm(ex*deg2rad(135))),
+        ((@SVector [0, 0.0, 2.5]),    expm(ex*deg2rad(180))),
+        # ((@SVector [0, 0.0, 2.0]),    expm(ex*deg2rad(225))),
+        ((@SVector [0,-0.5, 1.5]),    expm(ex*deg2rad(-90))),
+        ((@SVector [0, 0.65, 1.0]),    expm(ex*deg2rad(360))),
+        ((@SVector [0, 1.0, 1.0]),    expm(ex*deg2rad(360))),
+    ]
+    # times = [35, 41, 47, 51, 55, 61, 70, 101]
+    times = [45, 51, 55, 75, 101]
+
+    """
+    Costs
+    """
+    # intermediate costs
[…]

### 2020-10-09 examples/Flexible_Spacecraft/solve_con.jl
updated flexible spacecraft example
Applied Brian's recommend change (lowering quaternion cost)

@@ -17,8 +17,8 @@ const TO = TrajectoryOptimization
 
 
 # Discretization
-tf = 150.0
-N = 501
+tf = 100.0
+N = 401
 dt = tf/(N-1)
 
 # Model
@@ -56,43 +56,36 @@ SA[ dot(dcm_from_q(x[4:7])*con.bodyvec,con.keepoutdir)-cosd(con.keepoutangle)])
 
 # Initial and final states
 ω = @SVector zeros(3)
-# q0 = Rotations.params(UnitQuaternion(I))
 q0 = Rotations.params(expm(deg2rad(150) * normalize(@SVector [1,2,3])))
-# qf = Rotations.params(expm(deg2rad(90) * normalize(@SVector rand(3))))
 qf = Rotations.params(UnitQuaternion(I))
-# q0 = @SVector [.5,.5,.5,.5]
-# qf = normalize(@SVector [0.0258,0.0258,0.9990,0.0258])
 r0 = zeros(4)
 x0 = [ω; q0;zeros(3);zeros(3);r0]
 xf = [ω; qf;zeros(3);zeros(3);r0]
 
 # Objective
-Q = Diagonal(@SVector [10,10,10,100000,100000,100000,100000,10,10.0,10,10,10,10,0,0,0,0])
-R = Diagonal(@SVector fill(0.00001, m))
+Q = Diagonal(@SVector [10,10,10,100,100,100,100,10,10.0,10,10,10,10,0,0,0,0])
+R = Diagonal(@SVector fill(1.0, m))
 Qf = Q
 obj = LQRObjective(Q,R,Qf,xf,N)
 
 # constaint
 cons =  ConstraintList(n,m,N)
-bnd = BoundConstraint(n,m, u_min=-0.2, u_max=.2)
-add_constraint!(cons,bnd,1:N)
+bnd = BoundConstraint(n,m, u_min=-.5, u_max=.5)
+add_constraint!(cons,bnd,1:N-1)
 
 ## keep out
 keepoutdir = @SVector [1.0,0,0]
 bodyvec = @SVector [0.360019,-0.92957400,0.07920895]
-# bodyvec = @SVector [0.661490,-0.67104348,0.334858278]
-# keepoutdir = @SVector [0,0,1.0]
-# bodyvec = @SVector [0.75,0.433,0.5]
 keepout = AttitudeKeepOut( n, keepoutdir, bodyvec, 40.0 )
 
 add_constraint!(cons,keepout,1:N)
 
 
 
[…]

### 2020-10-09 examples/Flexible_Spacecraft/solve_uncon.jl

@@ -1,6 +1,6 @@
 cd(dirname(@__FILE__))
 Pkg.activate(".")
-
+using MATLAB
 using StaticArrays
 using Parameters
 using RobotDynamics
@@ -9,7 +9,7 @@ using LinearAlgebra
 # using Plots
 using TrajectoryOptimization
 using Altro
-using MATLAB
+
 
 include(joinpath(dirname(@__FILE__),"flexible_spacecraft_dynamics.jl"))
 
@@ -17,8 +17,8 @@ const TO = TrajectoryOptimization
 
 
 # Discretization
-tf = 150.0
-N = 501
+tf = 100.0
+N = 401
 dt = tf/(N-1)
 
 # Model
@@ -58,35 +58,32 @@ SA[ dot(dcm_from_q(x[4:7])*con.bodyvec,con.keepoutdir)-cosd(con.keepoutangle)])
 ω = @SVector zeros(3)
 q0 = Rotations.params(expm(deg2rad(150) * normalize(@SVector [1,2,3])))
 qf = Rotations.params(UnitQuaternion(I))
-
 r0 = zeros(4)
 x0 = [ω; q0;zeros(3);zeros(3);r0]
 xf = [ω; qf;zeros(3);zeros(3);r0]
 
 # Objective
-Q = Diagonal(@SVector [10,10,10,100000,100000,100000,100000,10,10.0,10,10,10,10,0,0,0,0])
-R = Diagonal(@SVector fill(0.00001, m))
+Q = Diagonal(@SVector [10,10,10,100,100,100,100,10,10.0,10,10,10,10,0,0,0,0])
+R = Diagonal(@SVector fill(1.0, m))
 Qf = Q
 obj = LQRObjective(Q,R,Qf,xf,N)
 
 # constaint
 cons =  ConstraintList(n,m,N)
-bnd = BoundConstraint(n,m, u_min=-0.2, u_max=.2)
-add_constraint!(cons,bnd,1:N)
+bnd = BoundConstraint(n,m, u_min=-.5, u_max=.5)
+add_constraint!(cons,bnd,1:N-1)
 
 ## keep out
 keepoutdir = @SVector [1.0,0,0]
 bodyvec = @SVector [0.360019,-0.92957400,0.07920895]
 
 
-
-
-prob = Problem(model, obj, xf, tf, x0=x0,constraints = cons,N=N)
+prob = Problem(model, obj, xf, tf, x0=x0,constraints = cons,N=N;integration=RK4)
 
[…]

### 2020-10-09 paper/main.tex

@@
 
-\author{Brian Jackson$^1$ and Zachary Manchester$^1$%
+\author{Brian Jackson$^1$, Kevin Tracy$^1$, and Zachary Manchester$^1$%
     \thanks{
@@
 large attitude changes is challenging due to the nontrivial group structure of 3D
-rotations. This paper introduces an accessible and unified approach for tracking control
-and optimization-based planning on the space of rotations based on vector calculus and
-linear algebra. The methodology is used to derive an extension of the Linear-Quadratic
-Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR
-(MLQR). We compare MLQR to a state-of-the-art geometric tracking controller designed
-specifically for quadrotors and demonstrate superior tracking and robustness. An
-iterative variant of MLQR is compared to existing methods on several constrained
-trajectory optimization problems, again demonstrating improved runtime performance. 
+rotations. This paper introduces an accessible approach for 
+optimization-based planning on the space of rotations based on vector calculus and
+linear algebra. 
 \keywords{
@@
     a quaternion differential calculus, but take a significantly simpler and more general 
-    approach that unifies both planning and control, enabling straight-forward adaptation of 
+    approach, enabling straight-forward adaptation of 
     existing algorithms to systems with quaternion states. 
-    To make this concrete, in Sec. \ref{sec:MLQR} we apply our approach to derive an 
-    extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), 
-    which is the control dual to the multiplicative extended Kalman filter (MEKF) from the 
-    state estimation literature \cite{markley2014fundamentals}. 
-    In Sec. \ref{sec:experiments} we provide several simulation results demonstrating the 
-    application of our method to tracking control and constrained trajectory optimization.
+    To make this concrete, in Sec. \ref{sec:Wahbas} we apply this method to the canonical
+    Wahbas problem and demonstrate dramatically superior convergence to approaches that
+    fail to properly account for the group structure. 
+    In Sec. \ref{sec:trajopt} we extend these ideas to the problem of trajectory optimization,
+    and detail modifications to ALTRO, a state-of-the-art constrained trajectory optimization
+    solver, and demonstrate the performance gains on several constrained benchmark problems.
+    To our knowledge, there does not currently exist a solver that is capable of leveraging
+    the unique Markovian structure of the fixed-horizon trajectory optimization problem while
+    correctly accounting for the group structure of 3D rotations.
     In summary, our contributions include:
@@
     \begin{itemize}
-        \item A unified approach to quaternion differential calculus entirely based on standard vector calculus and linear algebraic operations
-        \item Derivation of multiplicative LQR, an adaptation of LQR to the control of systems with quaternion states
-        \item A set of benchmark motion-planning and trajectory-tracking problems in which we compare our approach to existing methods commonly used in the robotics community to account for 3D rotations
+        \item A unified approach to quaternion differential calculus entirely based on
+        standard vector calculus and linear algebraic operations control of systems with
+        quaternion states
+        \item A fast and efficient solver for trajectory optimization problems with 
+        attitude dynamics and nonlinear constraints that correctly accounts for the group
+        structure of 3D rotations during the solve
     \end{itemize}
@@
         where $A(q)$ is the rotation matrix in terms of the elements of the quaternion 
-        \cite{markley2014fundamentals}.
+        \cite{kane1983spacecraftdynamics}.
 
@@
                 v \\ 
-                \texttt{kinematics}(R,\omega) \\ 
+                \half \q \otimes \hat{\omega} = \half L(q) H \omega \\ 
                 \frac{1}{m} F_G(x,u) \\ 
@@
         $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the 
-        global frame, $\tau_L(x,u)$ are the moments in the local (body) frame, and 
-        $\texttt{kinematics}$ are the kinematics for the chosen attitude representation. The 
-        kinematics for unit quaternions are
-        \begin{equation}
-            \dot{q} = \half \q \otimes \hat{\omega} = \half L(q) H \omega.
-        \end{equation}
+        global frame, and $\tau_L(x,u)$ are the moments in the local (body) frame.
 
@@
 
-\section{Multiplicative LQR} \label{sec:MLQR}
-    Leveraging the methods from the previous section, we derive multiplicative LQR
-    (MLQR), a variant of LQR that correctly accounts for the group structure of
-    rotations. For concreteness, we consider a system with rigid body dynamics, as
-    presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize
-    the system about a dynamically feasible discretized reference trajectory
-    $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
-    We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
+\section{Modifying Newton's Method} \label{sec:Wahbas}
+    Brief background and description of the problem.
+
+    \subsection{Methodology}
+    Brief description of using Newton's method to solve the batch version of Wahbas problem.
+
+    \subsection{Results}
+
+\section{Trajectory Optimization on $\R^n \times SO(3)$}
+    Here we outline the modifications to the ALTRO solver \cite{howell2019altro}, to
+    solve trajectory optimization problems for rigid bodies, which extends easily to
+    arbitrary systems in $R^n \times SO(3)$. ALTRO is an efficient solver for constrained
+    nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented
+    Lagrangian framework.
+
+    We consider a trajectory optimization problem of the form:
+    \begin{mini}[2]
+        {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
+        \addConstraint{x_{k+1} = f(x_k,u_k)}
+        \addConstraint{g_k(x_k,u_k)}{\leq 0}
+        \addConstraint{h_k(x_k,u_k)}{=0}
+        \label{discrete_trajopt},
+    \end{mini}
+    where $x$ and $u$ are the state and control vectors as described in Sec. \ref{sec:rigidbody_dynamics},
+    $f$ are the dynamics as defined in \eqref{eq:rigid_body_dynamics}, $\ell$ is a general
+    nonlinear cost function,
+    $N$ is the number of time steps, and $g_k$, $h_k$ are general nonlinear inequality and 
+    equality constraints.
+
+    Like most gradient or Newton-based methods for optimization, ALTRO approximates the 
+    nonlinear functions $f, \ell, g,$ and $h$ with their first or second-order Taylor series
+    expansions. Leveraging the methods from Sec: \ref{sec:Quaternion_Calculus}, we adapt the
+    algorithm to optimize directly on the error state $\delta x \in \R^{12}$.
+
+    We begin by linearizing the dynamics about the reference trajectory using
+    \eqref{eq:quat_jacobian}. Our linearized error dynamics become
     \begin{equation} \label{eq:linearized_dynamics}
@@
 
-    With our linearized system, we can apply the traditional LQR Riccati recursion,
+    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to our nonlinear cost
+    functions $\ell$ and 
+    \eqref{eq:quat_jacobian} to the nonlinear constraint functions $g_k$ and $h_k$, we can calculate
+    the second-order expansion 
     \begin{multline}
-         P_k = W_k +  A_k^T P_{k+1} A_k \\
-         - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
+        \delta \ell(x,u) \approx  \half \delta x^T \ell_xx \delta x
+            + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta u \\
+            + \ell_x^T \delta x^T + \ell_u^T \delta u 
     \end{multline}
-    which minimizes the quadratic cost function
+    of the augmented Lagrangian cost function:
     \begin{equation}
-        \half \sum_{k=0}^N \delta x_k^T Q_k \delta x_k + \delta u_k^T R_k \delta u_k
+        \mathcal{L}_A = \mathcal{L}_N(x_N,\lambda_N,\mu_N) + 
+            \sum_{k=0}^{N-1} \mathcal{L}_k(x_k,u_k,\lambda_k,\mu_k)
     \end{equation}
-    of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on
-    the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From
-    this, we get our non-linear feedback policy
+    where
+    \begin{equation}
+        \mathcal{L}_k(x,u,\lambda,\mu) = \ell(x,u) + 
+            (\lambda + \half I_\mu c(x,u))^T c(x,u),
+    \end{equation}
+    with $c(x,u)$ being the concatenation of the constraints $f,g$, and $h$ at a given time step,
+    and $I_\mu$ the penalty matrix.
+
+    With this expansion, we calculate the expansion of the ``action-value function'' $Q(x,u)$
+    as normal:
+    \begin{align}
+        Q_{xx} &= \ell_{xx} + A_{k}^T P_{k+1} A_{k} \label{Qxx_exp}\\
+        Q_{uu} &= \ell_{uu} + B_{k}^T P_{k+1} B_{k} \label{Quu_exp}\\
+        Q_{ux} &= \ell_{ux} + B_{k}^T P_{k+1} A_{k} \label{Qux_exp}\\
+        Q_x &= \ell_x  + A_{k}^T p_{k+1} \label{Qx_exp}\\
+        Q_u &= \ell_u + B_{k}^T p_{k+1} \label{Qu_exp},
+    \end{align}
+    from which we can calculate the quadratic expansion of the cost-to-go 
+    $P_k \in \R^{12 \times 12}$, $p_k \in \R^12$, and optimal linearized feedback gains 
+    $K_k \in \R^{m \times 12}$ and $d_k \in \R^m$ by starting at the terminal state and
+    resursing backward in time along the trajectory during the ``backward pass'' of the iLQR
+    algorithm. During the ``forward pass'', the dynamics are simulated forward in time using
+    the feedback gains computed during the backward pass. At each time step, the control is 
+    calculated using the linear feedback controller: 
     \begin{equation} \label{eq:mlqr_control}
-        u = K_k \delta x_k + \bar{u}_k.
+        u_k = K_k \delta x_k + \bar{u}_k.
     \end{equation}
-    where $\delta x_k$ is computed online using the non-linear error function
-    \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the
-    standard LQR formula:
-    \begin{equation} \label{eq:LQR_gain}
-         K_k := -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k.
-    \end{equation}
-
-    The multiplicative Linear Quadratic Regulator is summarized as follows:
-    \begin{enumerate}
-        \item Pick $Q_k \in \R^{12 \times 12}$ to weight the state \textit{error}, and $R_k \in \R^{m \times m}$
-        \item Linearize the dynamics using the state error Jacobian \eqref{eq:state_error}
-        \item Compute $K_k$ using the standard LQR Riccati recursion
-        \item Online, compute the control using \eqref{eq:state_error} and \eqref{eq:mlqr_control}.
-    \end{enumerate}
-
-    \subsection{Constrained Iterative MLQR}
-        As demonstrated above, it is very straight-forward to adapt LQR to use
-        quaternions. We similarly use this technique to adapt nonlinear trajectory
-        optimization algorithms. Here we present the modifications to the ALTRO solver
-        \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an
-        augmented Lagrangian approach to handle constraints.
-
-        For each iteration of iLQR within ALTRO, we can calculate a quadratic
-        approximation of the cost function using \eqref{eq:quat_gradient} and
-        \eqref{eq:quat_hessian}. Identical to \eqref{eq:linearized_dynamics}, we use
-        \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the
-        feedback and feedforward gains.
-
-        The only other modification to the algorithm is during the iLQR ``forward pass''
-        that simulates the system with the closed-loop MLQR controller, where we use
-        \eqref{eq:state_error} to calculate the nonlinear feedback policy during the
-        forward roll-outs.
-        For robustness, we may also add an additional criterion to the line search that
-        checks for singularities in the three-parameter error state which, for the Cayley
-        map, occur at \ang{180}, at which point the step size should be reduced.
+    where $\bar{u}_k$ is the control value from the previous iteration, and $\delta x$ is 
+    computed using \eqref{eq:state_error}, with $x_k$ the current state estimate and $\bar{x}_k$
+    the state from the previous iteration. The rest of the algorithm is left unchanged. 
+    For more details on the ALTRO algorithm, the reader is encouraged to refer to the original
+    paper \cite{howell2019altro}.
 
     \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
-        In addition to the straight-forward modifications to the LQR algorithm itself, we
+        In addition to the straight-forward modifications to the ALTRO algorithm itself, we
         need to carefully consider the types of cost functions we minimize. We frequently
@@
 
-        \begin{figure}
-            \centering
-            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
-            \caption{Timing results for the airplane barrell roll, quadrotor
-                flip, and quadrotor zig-zag examples using iLQR with
-                roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The
-                timing result for the Quadrotor flip with Euler angles is omitted
-                since it failed to converge.
-            }
-            \label{fig:timing_chart}
-        \end{figure}
-
-        \subsubsection{Quadrotor Zig-Zag}
-            \begin{figure}[t]
-                \centering
-                \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
-                \caption{Zig-zag trajectory. The black quadrotors indicate
-                    the reference trajectory generated by iMLQR. The yellow
-                    quadrotors track the reference trajectory using time-varying
-                    MLQR, while the red uses the HFCA controller from
-                    \cite{watterson2020control}.
-                }
-                \label{fig:zig-zag}
-            \end{figure}    
-            \begin{figure}[t]
-                \centering
-                \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
-                \caption{Snapshots of the quadrotor flip trajectory. The
-                    red-colored quadrotors represent the state near t=0 s and the
-                    green-colored quadrotors represent the state near t=5.0 s
-                }
-                \label{fig:quad_flip}
-            \end{figure}
-
-        We set up a fairly simple trajectory optimization problem to achieve a
-        ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of
-        the form
-        \begin{multline}
-                  \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
-                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) \\
-                + \half \sum_{k=1} u_k^T R u_k
-        \end{multline}
-        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus
-        \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and
-        $(0,10,0)$. We compared the two cost functions from Sec.
-        \ref{sec:cost_functions}, using a standard quadratic penalty on the
-        non-quaternion states.
-        \begin{align}
-            \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
-            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
-                \pm \bar{q}^T q)
-        \end{align}
-        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where
-        $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular
-        velocity.
-        We used the following weighting matrices: 
-        $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
-        $Q_w = \text{diag}(10^3 I_3,I_9)$, 
-        $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
-        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and
-        terminal cost functions, respectively.
-
-        The geodesic cost function performed significantly better than the error
-        quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29
-        iterations. In addition to better convergence behavior, the geodesic cost is
-        about half as expensive to compute, which is unsurprising when comparing the
-        gradients and Hessians in Sec. \ref{sec:cost_functions}.
-
         \subsubsection{Quadrotor Flip}
@@
 
+            \begin{figure}[t]
+                \centering
+                \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
+                \caption{Snapshots of the quadrotor flip trajectory. The
+                    red-colored quadrotors represent the state near t=0 s and the
+                    green-colored quadrotors represent the state near t=5.0 s
+                }
+                \label{fig:quad_flip}
+            \end{figure}
+
         \subsection{Airplane Barrel Roll}
@@
         \end{figure}
-
-	    \begin{figure}[h]
-	        \centering
-	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
-	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller.
-Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics.
-}
-	        \label{fig:tracking_err}
-	    \end{figure}
-        \vspace{-20pt}
-
-	\subsection{Tracking Control}
-	    This section provides examples using MLQR for tracking and stabilizing control
-	    for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art
-	    geometric tracking controller \cite{watterson2020control} based on the Hopf
-	    Fibration.
-
-	    \subsubsection{MLQR Stabilization}
-	    To test robustness and the domain of attraction of the MLQR controller, a
-	    quadrotor is linearized about hover at the origin and initialized from states
-	    uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular
-	    velocities, and uniformly over the entire space of rotations. The success rates
-	    for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is
-	    defined as the terminal error (calculated using \eqref{eq:state_error}) being
-	    within a unit ball of radius 0.1 after 10.0 seconds. To further analyze the
-	    basins of attraction, we uniformly sampled over only roll (which goes singular at
-	    \ang{90} for our Euler angle representation), and yaw. The results are shown in
-	    Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the
-	    successfully stabilized initial conditions.
-
-	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also
-	    being more generally applicable, easier to tune, and easier to implement. It is
-	    also noteworthy that na\"ively using LQR with quaternions (incorrectly treating
-	    them as vectors) results in poorer performance than Euler angles, clearly
-	    demonstrating the value of the methodology presented in the current paper.
-
-	    \subsubsection{MLQR Tracking}
-	    Time-varying MLQR is used to track the zig-zag trajectory generated in the
-	    previous section. %The controller calculated the gains at $0.01 s$ intervals, and
-	    used zero-order-hold between intervals. To test tracking performance, zero-mean
-	    Gaussian noise was added as force and moments applied to the quadrotor.
-
-	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
-	    \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller
-	    dramatically out-performs HFCA, we found no significant difference in performance
-	    between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
-	    Since the tracking controller only deals with relatively small state and attitude
-	    errors, this is unsurprising. Additionally, we found the MLQR controller ran
-        about 20x faster than the HFCA controller.
-
-	    \begin{figure*}[ht]
-	        \begin{minipage}{0.60\linewidth}
-                \centering
-                \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
-                \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black).
-All points within each contour were successfully stabilized by the controller.}
-                \label{fig:mlqr_basin}
-            \end{minipage}
-	        \hspace{0.05\linewidth}
-            \begin{minipage}{0.34\linewidth}
-                \centering
-                \begin{tabular}{cc}
-                    \toprule
-                    \textbf{Method} & \textbf{Success} \\
-                    \midrule
-                    Quat & 61\% \\
-                    RPY  & 80\% \\
-                    HFCA & 97\% \\
-                    MLQR & 99\% \\
-                    \bottomrule
-                \end{tabular}
-                \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
-                \label{tab:mlqr_success}
-            \end{minipage}
-        \end{figure*}
 

### 2020-10-09 paper/main.tex
Update timing results and plots

@@
 \usepackage{tikzscale}
+% \pgfplotsset{every axis plot post/.style={
+%     very thick,
+% }
+% }
 
@@
     \subsection{Results}
+    \begin{figure}
+        \centering
+        \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
+        \caption{Convergence comparison for Wahbas problem. By performing Newton's method
+        on the error quaternion and applying the result to the full quaternion we achieve
+        quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
+        error.}
+    \end{figure}
 
@@
 
+    \begin{table}
+        \centering
+        \input{figures/timing_results.tex}
+        \label{tab:timing_results}
+        \caption{Trajectory Optimization Timing Results}
+    \end{table}
+
     \subsection{Trajectory Optimization}
@@
 
-        \subsubsection{Quadrotor Flip}
-	    Similar to the previous example, we use a set of waypoints to encourage the
-	    quadrotor to do a flip. The cost weights for this example were
-	    \begin{align*}
-	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \!\times\! 10^{-2}, 10^{-3} I_3,
-    	        10^{-3} I_3, 10^{-2} I_3) \\
-	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
-	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
-	    \end{align*}
-	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
-	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
-	    to stay above the floor and move to a goal state 2 meters away in the $+y$
-	    direction. The solver is initialized with a dynamically infeasible trajectory
-	    that linearly interpolates between the initial and final states, rotating the
-	    quad around the x-axis a full \ang{360}.
-
-	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
-	    iMLQR within the ALTRO solver to handle constraints. The trajectory generated by
-	    iLQR with quaternions and quadratic costs was successful but had a slightly
-        different trajectory. iLQR with Euler angles failed to solve the problem.
-
-            \begin{figure}[t]
+        \subsection{Satellite Attitude Keep-Out}
+        \begin{figure}
             \centering
-                \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
-                \caption{Snapshots of the quadrotor flip trajectory. The
-                    red-colored quadrotors represent the state near t=0 s and the
-                    green-colored quadrotors represent the state near t=5.0 s
-                }
-                \label{fig:quad_flip}
+            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
+            \caption{Attitude keep out}
         \end{figure}
@@
             \centering
-            \includegraphics[height=3.5cm]{figures/barrellroll.png}
+            \includegraphics[width=\columnwidth]{figures/barrellroll.png}
             \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
             \label{fig:barrellroll}
+        \end{figure}
+
+        \begin{figure}[h]
+            \centering
+            \includegraphics[height=4cm,width=\columnwidth]{figures/c_max_convergence.tikz}
+            \caption{Constraint convergence when solving the barrell roll problem. Compares 
+            the convergence of the original version of ALTRO versus the new, modified version
+            that optimizing on the error state.}
+            \label{fig:c_max_convergence}
+        \end{figure}
+
+        \subsubsection{Quadrotor Flip}
+        We sucessfully optimized a 360 degree flip trajectory for a quadrotor using the 
+        modified version of ALTRO.
+	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
+	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
+	    to stay above the floor and move to a goal state 2 meters away in the $+y$
+	    direction. The solver is initialized with a dynamically infeasible trajectory
+	    that linearly interpolates between the initial and final states, rotating the
+	    quad around the x-axis a full \ang{360}.
+
+	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
+	    ALTRO.  t
+
+            \begin{figure}[t]
+                \centering
+                \includegraphics[height=5.2cm,trim={0 0cm 0 0},clip]{figures/quadflip.png}
+                \caption{Snapshots of the quadrotor flip trajectory. The
+                    red-colored quadrotors represent the state near t=0 s and the
+                    green-colored quadrotors represent the state near t=5.0 s
+                }
+                \label{fig:quad_flip}
             \end{figure}

### 2020-10-09 examples/Flexible_Spacecraft/solve_con.jl

@@ -1,6 +1,5 @@
-cd(dirname(@__FILE__))
-Pkg.activate(".")
-using MATLAB
+import Pkg; Pkg.activate(@__DIR__)
+# using MATLAB
 using StaticArrays
 using Parameters
 using RobotDynamics
@@ -8,92 +7,20 @@ using Rotations
 using LinearAlgebra
 # using Plots
 using TrajectoryOptimization
+using PlanningWithAttitude
 using Altro
 
 
-include(joinpath(dirname(@__FILE__),"flexible_spacecraft_dynamics.jl"))
-
 const TO = TrajectoryOptimization
 
-
-# Discretization
-tf = 100.0
-N = 401
-dt = tf/(N-1)
-
-# Model
-model = FlexSatellite()
-n,m = size(model)
-
-# quaternion to dcm functions
-function skew(v)
-	"""Skew-symmetric matrix from 3 element array"""
-    return @SMatrix [0 -v[3] v[2]; v[3] 0 -v[1]; -v[2] v[1] 0]
-end
-function dcm_from_q(q)
-	"""DCM from quaternion, scalar first"""
-	s = @views q[1]
-	v = @views q[2:4]
-    return I + 2*skew(v)*(s*I + skew(v))
-end
-
-
-struct AttitudeKeepOut{T} <: TO.StateConstraint
-	n::Int
-	keepoutdir::SVector{3,T}
-	bodyvec::SVector{3,T}
-	keepoutangle::Float64
-	function AttitudeKeepOut(n::Int, keepoutdir::SVector, bodyvec::SVector, keepoutangle::T) where T
-		new{T}(n,keepoutdir,bodyvec,keepoutangle)
-	end
-end
-
-TO.state_dim(con::AttitudeKeepOut) = con.n
-TO.sense(::AttitudeKeepOut) = Inequality()
-Base.length(::AttitudeKeepOut) = 1
-TO.evaluate(con::AttitudeKeepOut, x::SVector) = (
-SA[ dot(dcm_from_q(x[4:7])*con.bodyvec,con.keepoutdir)-cosd(con.keepoutangle)])
-
-# Initial and final states
-ω = @SVector zeros(3)
-q0 = Rotations.params(expm(deg2rad(150) * normalize(@SVector [1,2,3])))
-qf = Rotations.params(UnitQuaternion(I))
-r0 = zeros(4)
-x0 = [ω; q0;zeros(3);zeros(3);r0]
[…]

### 2020-10-09 examples/trajopt_examples.jl

@@ -36,6 +36,7 @@ function add_result!(name, errstate::Bool, solver)
     push!(results[:c_max], [c0; solver.stats.c_max])
     return b
 end
+results_path = joinpath(@__DIR__, "timing_results.jld2")
 
 """
 Barrell Roll
@@ -57,23 +58,96 @@ prob,opts = YakProblems(vecstate=false, costfun=:QuatLQR; args...)
 solver = ALTROSolver(prob, opts)
 add_result!(:barrellroll, true, solver)
 
-@save "timing_results.jld2" results 
+@save results_path results 
 
 """
 Quadrotor Flip
 """
 ## Original Method
 prob,opts = QuadFlipProblem(vecmodel=true)
-solver = ALTROSolver(prob2, opts, R_inf=1e-4, infeasible=true, show_summary=true)
-solve!(solver)
+solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true)
+add_result!(:quadflip, false, solver)
 
 ## Modified Method
 prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost)
 solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true)
-solve!(solver)
+add_result!(:quadflip, true, solver)
 
-@save "timing_results.jld2" results
+@save results_path results
 
 """
 Flexible Satellite
 """
+## Original Method
+prob,opts = SatelliteKeepOutProblem(vecstate=true, costfun=LQRCost)
+solver = ALTROSolver(prob, opts)
+add_result!(:satellite, false, solver)
+
+## Modified Method
+prob,opts = SatelliteKeepOutProblem(vecstate=false, costfun=QuatLQRCost)
+solver = ALTROSolver(prob, opts)
+add_result!(:satellite, true, solver)
+
+@save results_path results
+
+
+## Save the table
+import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+using LaTeXTabulars, JLD2, DataFrames, PGFPlotsX
+results_path = joinpath(@__DIR__, "timing_results.jld2")
+@load results_path results
+names = Dict()
+
+df = DataFrame(results)
+df.time .= round.(df.time, digits=2)
+df.prob = String.(df.problem)
+
+
+mod = df[df.errstate,:]
+org = df[.!df.errstate,:]
+tab = vcat(map(1:3) do i
[…]

### 2020-10-09 src/PlanningWithAttitude.jl

@@ -17,11 +17,13 @@ include("quat_costs.jl")
 include("quat_norm.jl")
 include("airplane_problem.jl")
 include("quadflip_problem.jl")
+include("flexible_spacecraft_dynamics.jl")
 
 export
     VecModel,
     YakProblems,
     QuadFlipProblem,
+    SatelliteKeepOutProblem,
     QuatGeoCon,
     QuatErr,
     QuatVecEq,

### 2020-10-09 src/quat_norm.jl

@@ -13,13 +13,13 @@ function RD.discrete_dynamics(::Type{Q},
     r,q,v,ω = RobotDynamics.parse_state(model.model, x2, true)
     RobotDynamics.build_state(model.model, r,q,v,ω)
 end
-RD.wrenches(model::QuatRenorm, z::RD.AbstractKnotPoint) = RD.wrenches(model.model, z)
-RD.forces(model::QuatRenorm, x, u) = RD.forces(model.model, x, u)
-RD.moments(model::QuatRenorm, x::StaticVector, u::StaticVector) = RD.moments(model.model, x, u)
-RD.mass(model::QuatRenorm) = RD.mass(model.model)
-RD.inertia(model::QuatRenorm) = RD.inertia(model.model)
-RD.inertia_inv(model::QuatRenorm) = RD.inertia_inv(model.model)
-RD.velocity_frame(model::QuatRenorm) = RD.velocity_frame(model.model)
+# RD.wrenches(model::QuatRenorm, z::RD.AbstractKnotPoint) = RD.wrenches(model.model, z)
+# RD.forces(model::QuatRenorm, x, u) = RD.forces(model.model, x, u)
+# RD.moments(model::QuatRenorm, x::StaticVector, u::StaticVector) = RD.moments(model.model, x, u)
+# RD.mass(model::QuatRenorm) = RD.mass(model.model)
+# RD.inertia(model::QuatRenorm) = RD.inertia(model.model)
+# RD.inertia_inv(model::QuatRenorm) = RD.inertia_inv(model.model)
+# RD.velocity_frame(model::QuatRenorm) = RD.velocity_frame(model.model)
 
 
 

### 2020-10-09 paper/main.tex
Updates from Overleaf

--- 2020-10-09 paper/main.tex
@@
     To make this concrete, in Sec. \ref{sec:Wahbas} we apply this method to the canonical
-    Wahbas problem and demonstrate dramatically superior convergence to approaches that
+    Wahba's problem and demonstrate dramatically superior convergence to approaches that
     fail to properly account for the group structure. 
@@
 \section{Modifying Newton's Method} \label{sec:Wahbas}
-    Brief background and description of the problem.
+    Newton's method uses derivative information about a function to iteratively approximate it's roots.
+For unconstrained systems, this method is highly effective, and can exhibit quadratic convergence.
+For constrained systems, the updated parameter can be projected back onto the feasible set at each iteration, but without the same convergence guarantees.
+For the constraints on $SO(3)$, Newton's method struggles to converge past a certain threshold due to this projection.
+By leveraging the quaternion calculus introduced, Newton's method can be modified to implicitly account for these constraints.
+To demonstrate this, we will examine Wahba's Problem.
+In 1965, Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's attitude from vector measurements \cite{markley2014fundamentals}.
+We will solve this problem using a standard nonlinear least squares method, as well as a method that exploits the true group structure of $SO(3)$ using the quaternion calculus presented here.
 
     \subsection{Methodology}
-    Brief description of using Newton's method to solve the batch version of Wahbas problem.
+
+    Given known vectors in some inertial frame, $N$, and measurements of these vectors in some body fixed frame, $B$, we can define Wahba's loss function as the following:
+    \begin{equation}
+        L = \sum_i w_i \| ^N v_i -  {}^N A(q){}^B\,\,  ^Bv_i \|_2^2.
+    \end{equation}
+    This is a nonlinear function of the attitude, which will be stored as a quaternion. For convenience, the following residual vector function will be defined:
+    \begin{equation}
+        r_i = \sqrt{w_i} (^N v_i -  {}^N A(q){}^B\,\,  ^Bv_i ),
+    \end{equation}
+    such that when these vectors are stacked in $r$, the loss function is a sum of squares. This problem now looks like a standard nonlinear least squares problem:
+    \begin{mini*}
+    {q}{ \|r(q)\|_2^2 }{}{}
+    \addConstraint { q\in SO(3).}
+    \end{mini*}
+     There are two methods for keeping the solution in the $SO(3)$, group, a projected and a multiplicative method.
+In a projected Gauss-Newton method, the Newton step is applied additively, and the resulting estimate is projected back onto $SO(3)$ with a normalization.
+This algorithm can be seen in algorithm \ref{alg:pgn}.
+    \begin{algorithm} 
+    	\begin{algorithmic}[1]
+    		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
+    		\State $k = 0$
+    		\While{significant progress}
+    		    \State $J = \frac{\partial r(q_k)}{ \partial q}$ 
+    		    \State $ q_{step} = -(J^TJ)^{-1}J^T r(q_k)$ 
+    		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
+    		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + q_{step})$ 
+    		    \State $k = k + 1$
+    		\EndWhile
+    	\end{algorithmic}
+    \end{algorithm}
+    Where the $SO(3)$ projection operator is just a normalization.
+This method can successfully lower the loss function initially, but the normalization errors will prevent convergence.
+To get around this, the quaternion calculus presented in this paper can be used to generate a Newton step that is a three parameter rotation, one that can be applied multiplicatively.
+This method of applying the Newton step preserves the constraint satisfaction, and doesn't introduce any normalization errors.
+    \begin{algorithm} 
+    	\begin{algorithmic}[1]
+    		\caption{Multiplicative Gauss-Newton Method}\label{alg:mgn}
+    		\State $k = 0$
+    		\While{significant progress}
+    		    \State $J = [\frac{\partial r(q_k)}{ \partial q}]G(q_k)$ \Comment{quaternion adjusted Jacobian}
+    		    \State $ \phi_{step} = -(J^TJ)^{-1}J^T r(q_k)$ 
+    		    \State $q_{step} = \varphi(\phi_{step})$ 
+    		    \State $q_{k+1} = q_n \otimes q_{step}$ \Comment{apply step multiplicatively}
+    		    \State $k = k + 1$
+    		\EndWhile
+    	\end{algorithmic}
+    \end{algorithm}
 
     \subsection{Results}
-
+    % \begin{figure}
+    %     \centering
+    %     \input{paper/figures/kevins_plots/convergence_plot}
+    %     \caption{Caption}
+    %     \label{fig:my_label}
+    % \end{figure}
 \section{Trajectory Optimization on $\R^n \times SO(3)$}

### 2020-10-09 paper/main.tex

@@
     \subsection{Results}
-    % \begin{figure}
-    %     \centering
-    %     \input{paper/figures/kevins_plots/convergence_plot}
-    %     \caption{Caption}
-    %     \label{fig:my_label}
-    % \end{figure}
+    \begin{figure}
+        \centering
+        \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
+        \caption{Convergence comparison for Wahbas problem. By performing Newton's method
+        on the error quaternion and applying the result to the full quaternion we achieve
+        quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
+        error.}
+    \end{figure}
+
 \section{Trajectory Optimization on $\R^n \times SO(3)$}

### 2020-10-09 paper/main.tex
Clean up results section

--- 2020-10-09 paper/main.tex
@@
         error.}
+        \label{fig:wahba_convergence}
     \end{figure}
 
-\section{Trajectory Optimization on $\R^n \times SO(3)$}
+\section{Trajectory Optimization on $\R^n \times SO(3)$} \label{sec:trajopt}
     Here we outline the modifications to the ALTRO solver \cite{howell2019altro}, to
@@
 
-        \subsubsection{Geodesic Distance}
+        \subsubsection{Geodesic Distance} \label{sec:geodesic}
             Alternatively, we can use the geodesic distance between two quaternions
@@
 \section{Experiments} \label{sec:experiments}
+    In this section we present several trajectory optimization problems for systems that
+    undergo large changes in attitude: an airplane barrell roll, a quadrotor flip, and a
+    satellite with flexible modes that must slew to a new orientation while avoiding a 
+    ``keep out'' orientation zone. All results were run on a desktop computer with an 
+    AMD Ryzen 2950x processor with 40 GB of RAM. 
+    All problems are run using ALTRO, first without any of the modification presented 
+    in the current paper, labeled ``original'', and then using the modifications listed
+    in Sec. \ref{sec:trajopt} and the geodesic cost funtion described in Sec. \ref{sec:geodesic}.
+    Timing results are summarized in Table \ref{tab:timing_results}.
+    All experiments were solved to a constraint satisfaction tolerance of $10^{-5}$.
     Code for all experiments is available on 
@@
 
-    \subsection{Trajectory Optimization}
-        In this section we present several trajectory optimization problems solved using
-        our modified version of ALTRO. The two quadrotor examples use 101 knot points for
-        5 second trajectories, and the airplane example uses 51 knot points for a 1.25
-        second trajectory. All results were run on a desktop computer with an Intel
-        i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
-        experiments are given in Fig. \ref{fig:timing_chart}.
-
-        In all examples, MLQR is implemented with a unit quaternion attitude state and
-        Rodrigues parameters as the error state. The first example compares the two cost
-        functions given in Sec. \ref{sec:cost_functions}, while the last two use the
-        geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
-        \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
-        from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
-        vectors). All examples re-normalize the quaternion in the dynamics function and
-        use a third-order explicit Runge-Kutta integrator.
-
     \subsection{Satellite Attitude Keep-Out}
@@
     \subsection{Airplane Barrel Roll}
+
         An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
@@
         see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
-        conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
-        converged faster than the pure quaternion version. Despite the extra matrix
-        multiplications and extra Hessian term, it also was faster per iteration than its
-        iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
-        better performance by correctly leveraging the structure of the rotation group
-        during the optimization routine.
+        conditions. The convergence of the different versions is compared in Fig.
+        \ref{fig:c_max_convergence}. For both the quadrotor flip and the airplane barrel
+        roll, the modified version of ALTRO converged faster than original version.
+        For these highly aerobatic maneuvers, we achieve, as expected, better
+        performance by correctly leveraging the structure of the rotation group during
+        the optimization routine.
 
@@
 
-        \subsubsection{Quadrotor Flip}
+    \subsection{Quadrotor Flip}
         We sucessfully optimized a 360 degree flip trajectory for a quadrotor using the 
@@
 	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
-	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
+        of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular arc.
+        The quadrotor was constrained
 	    to stay above the floor and move to a goal state 2 meters away in the $+y$
-	    direction. The solver is initialized with a dynamically infeasible trajectory
+	    direction. The solver was initialized with a dynamically infeasible trajectory
 	    that linearly interpolates between the initial and final states, rotating the
@@
 	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
-	    ALTRO.  t
+        ALTRO. The original version of ALTRO, even after significant tuning efforts,
+        would not converge to the desired solution. It instead got halfway and then ``unwound''
+        360 degrees and then continued rotating to the final orientation. This behavior
+        is common and expected when attempting optimization that does not properly account 
+        for the grou structure of rotations.
 
@@
     We have presented a general, unified method for planning and control on rigid-body
-    systems with arbitrary attitude using standard linear algebra and vector calculus. By
-    applying these methods to LQR to correctly account for the group structure of
-    rotations, we have matched or exceeded the performance of a state-of-the-art
-    geometric controller designed specifically for quadrotors, while being more general,
-    requiring less system-dependent tuning, having less computational overhead, and being
-    easier to implement.
+    systems with arbitrary attitude using standard linear algebra and vector calculus.
+    We have demonstrated that the application of this methodology is straightforward and
+    yields substantion improvements in the convergence of Newton-based methods (see Fig. 
+    \ref{fig:wahba_convergence}) while also offering improvements for nonlinear constrained
+    trajectory optimization for floating-base systems (see Table \ref{tab:timing_results}).
 
-    We have also demonstrated a straight-forward way to adapt nonlinear trajectory
-    optimization techniques to work with quaternion-valued states. For these problems, we
-    found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
-    computationally efficient and provide excellent convergence in practice. We recommend
-    the use of unit quaternions within planning and control algorithms for their
-    simplicity, computational efficiency, and lack of singularities. We further recommend
-    the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.
+    With the modifications presented, ALTRO can solve problems few other methods for 
+    trajectory optimization can. Many state-of-the-art methods such as direct 
+    collocation or sequential convex programming rely on commercial, proprietary, or 
+    general-purpose solvers whose internal numerics often abstracted away from the user. 
+    By exploiting both the unique Markovian structure of the trajectory optimization problem
+    and the group structure of rotations, ALTRO will likely be able to solve new classes of 
+    problems with at or near real-time performance. Future work will focus on continued 
+    refinement of the implementation and benchmarking against current state-of-the-art 
+    methods for trajectory optimization.
 
-    Future areas of research include the application of these methodologies to more
-    complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
-    more in-depth analysis of the convergence behavior of the algorithms proposed in the
-    current work.
+    Additionally, the methods presented here can easily be leveraged to adapt other classes
+    of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future
+    work may include adaptation of methods for state estimation, localization, design, or 
+    other methods for motion planning such as direct collocation.
+
+    Future work will also include the application of these methods to multi-body robotic
+    systems, such as humanoids or quadrupeds, especially in methods that exploit 
+    ``maximal'' coordinates that include the 3D orientation of each body \cite{brudigam2020linear}.
 

### 2020-10-09 paper/main.tex

@@
 \section{Modifying Newton's Method} \label{sec:Wahbas}
-    Brief background and description of the problem.
+    Newton's method uses derivative information about a function to iteratively approximate it's roots.
+For unconstrained systems, this method is highly effective, and can exhibit quadratic convergence.
+For constrained systems, the updated parameter can be projected back onto the feasible set at each iteration, but without the same convergence guarantees.
+For the constraints on $SO(3)$, Newton's method struggles to converge past a certain threshold due to this projection.
+By leveraging the quaternion calculus introduced, Newton's method can be modified to implicitly account for these constraints.
+To demonstrate this, we will examine Wahba's Problem.
+In 1965, Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's attitude from vector measurements \cite{markley2014fundamentals}.
+We will solve this problem using a standard nonlinear least squares method, as well as a method that exploits the true group structure of $SO(3)$ using the quaternion calculus presented here.
 
     \subsection{Methodology}
-    Brief description of using Newton's method to solve the batch version of Wahbas problem.
+
+    Given known vectors in some inertial frame, $N$, and measurements of these vectors in some body fixed frame, $B$, we can define Wahba's loss function as the following:
+    \begin{equation}
+        L = \sum_i w_i \| ^N v_i -  {}^N A(q){}^B\,\,  ^Bv_i \|_2^2.
+    \end{equation}
+    This is a nonlinear function of the attitude, which will be stored as a quaternion. For convenience, the following residual vector function will be defined:
+    \begin{equation}
+        r_i = \sqrt{w_i} (^N v_i -  {}^N A(q){}^B\,\,  ^Bv_i ),
+    \end{equation}
+    such that when these vectors are stacked in $r$, the loss function is a sum of squares. This problem now looks like a standard nonlinear least squares problem:
+    \begin{mini*}
+    {q}{ \|r(q)\|_2^2 }{}{}
+    \addConstraint { q\in SO(3).}
+    \end{mini*}
+     There are two methods for keeping the solution in the $SO(3)$, group, a projected and a multiplicative method.
+In a projected Gauss-Newton method, the Newton step is applied additively, and the resulting estimate is projected back onto $SO(3)$ with a normalization.
+This algorithm can be seen in algorithm \ref{alg:pgn}.
+    \begin{algorithm} 
+    	\begin{algorithmic}[1]
+    		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
+    		\State $k = 0$
+    		\While{significant progress}
+    		    \State $J = \frac{\partial r(q_k)}{ \partial q}$ 
+    		    \State $ q_{step} = -(J^TJ)^{-1}J^T r(q_k)$ 
+    		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
+    		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + q_{step})$ 
+    		    \State $k = k + 1$
+    		\EndWhile
+    	\end{algorithmic}
+    \end{algorithm}
+    Where the $SO(3)$ projection operator is just a normalization.
+This method can successfully lower the loss function initially, but the normalization errors will prevent convergence.
+To get around this, the quaternion calculus presented in this paper can be used to generate a Newton step that is a three parameter rotation, one that can be applied multiplicatively.
+This method of applying the Newton step preserves the constraint satisfaction, and doesn't introduce any normalization errors.
+    \begin{algorithm} 
+    	\begin{algorithmic}[1]
+    		\caption{Multiplicative Gauss-Newton Method}\label{alg:mgn}
+    		\State $k = 0$
+    		\While{significant progress}
+    		    \State $J = [\frac{\partial r(q_k)}{ \partial q}]G(q_k)$ \Comment{quaternion adjusted Jacobian}
+    		    \State $ \phi_{step} = -(J^TJ)^{-1}J^T r(q_k)$ 
+    		    \State $q_{step} = \varphi(\phi_{step})$ 
+    		    \State $q_{k+1} = q_n \otimes q_{step}$ \Comment{apply step multiplicatively}
+    		    \State $k = k + 1$
+    		\EndWhile
+    	\end{algorithmic}
+    \end{algorithm}
 

### 2020-10-09 paper/main.tex
Updates from Overleaf

--- 2020-10-09 paper/main.tex
@@
     \begin{equation}
-        L = \sum_i w_i \| ^N v_i -  {}^N A(q){}^B\,\,  ^Bv_i \|_2^2.
+        L = \sum_i w_i \| ^N v_i -  {}^N A^B(q)\,\,  ^Bv_i \|_2^2.
     \end{equation}
@@
     \begin{equation}
-        r_i = \sqrt{w_i} (^N v_i -  {}^N A(q){}^B\,\,  ^Bv_i ),
+        r_i = \sqrt{w_i} (^N v_i -  {}^N A{}^B(q)\,\,  ^Bv_i ),
     \end{equation}
@@
     \end{mini*}
-     There are two methods for keeping the solution in the $SO(3)$, group, a projected and a multiplicative method.
+     There are two ways for keeping the solution in the $SO(3)$ group, a projected method and a multiplicative method.
 In a projected Gauss-Newton method, the Newton step is applied additively, and the resulting estimate is projected back onto $SO(3)$ with a normalization.
@@
     \end{algorithm}
-    Where the $SO(3)$ projection operator is just a normalization.
    This method can successfully lower the loss function initially, but the normalization errors will prevent convergence.
-To get around this, the quaternion calculus presented in this paper can be used to generate a Newton step that is a three parameter rotation, one that can be applied multiplicatively.
+To get around this, the quaternion calculus presented in this paper can be used to generate a Newton step that is a rotation instead of an addition.
 This method of applying the Newton step preserves the constraint satisfaction, and doesn't introduce any normalization errors.
@@
         \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
-        \caption{Convergence comparison for Wahbas problem. By performing Newton's method
+        \caption{Convergence comparison for Wahba's problem. By performing Newton's method
         on the error quaternion and applying the result to the full quaternion we achieve

### 2020-10-10 paper/main.tex
Modify the abstract

@@
 \begin{abstract}
-Planning and controlling trajectories for floating-base robotic systems that experience
+Planning trajectories for floating-base robotic systems that experience
 large attitude changes is challenging due to the nontrivial group structure of 3D
-rotations. This paper introduces an accessible approach for 
+rotations. This paper introduces a powerful and accessible approach for 
 optimization-based planning on the space of rotations based on vector calculus and
-linear algebra. 
+linear algebra. We demonstrate the effectiveness of the approach by adapting Newton's
+method to solve the canonical Wahba's problem, and modifying the ALTRO solver for 
+constrained nonlinear trajectory optimization to plan directly on the space of unit 
+quaternions, achieving superior convergence on problems involving significant changes 
+in attitude.
 \keywords{
@@
 \section{Modifying Newton's Method} \label{sec:Wahbas}
-    Newton's method uses derivative information about a function to iteratively approximate it's roots.
-For unconstrained systems, this method is highly effective, and can exhibit quadratic convergence.
-For constrained systems, the updated parameter can be projected back onto the feasible set at each iteration, but without the same convergence guarantees.
-For the constraints on $SO(3)$, Newton's method struggles to converge past a certain threshold due to this projection.
-By leveraging the quaternion calculus introduced, Newton's method can be modified to implicitly account for these constraints.
-To demonstrate this, we will examine Wahba's Problem.
-In 1965, Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's attitude from vector measurements \cite{markley2014fundamentals}.
-We will solve this problem using a standard nonlinear least squares method, as well as a method that exploits the true group structure of $SO(3)$ using the quaternion calculus presented here.
+
+    Newton's method uses derivative information about a function to iteratively
+    approximate its roots. For unconstrained systems, this method is highly effective,
+    and can exhibit quadratic convergence. For constrained systems, the updated parameter
+    can be projected back onto the feasible set at each iteration, but without the same
+    convergence guarantees. For the constraints on $SO(3)$, Newton's method struggles to
+    converge past a certain threshold due to this projection. By leveraging the
+    quaternion calculus introduced, Newton's method can be modified to implicitly account
+    for these constraints. To demonstrate this, we will examine Wahba's Problem. In 1965,
+    Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's
+    attitude from vector measurements \cite{markley2014fundamentals}. We will solve this
+    problem using a standard nonlinear least squares method, as well as a method that
+    exploits the true group structure of $SO(3)$ using the quaternion calculus presented
+    here.
 
@@
 
-    Given known vectors in some inertial frame, $N$, and measurements of these vectors in some body fixed frame, $B$, we can define Wahba's loss function as the following:
-    \begin{equation}
-        L = \sum_i w_i \| ^N v_i -  {}^N A^B(q)\,\,  ^Bv_i \|_2^2.
-    \end{equation}
-    This is a nonlinear function of the attitude, which will be stored as a quaternion. For convenience, the following residual vector function will be defined:
-    \begin{equation}
-        r_i = \sqrt{w_i} (^N v_i -  {}^N A{}^B(q)\,\,  ^Bv_i ),
-    \end{equation}
-    such that when these vectors are stacked in $r$, the loss function is a sum of squares. This problem now looks like a standard nonlinear least squares problem:
+    Given known vectors in some inertial frame, ${}^N v_i$, and measurements of these
+    vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
+    rotation from the body to inertial frames ${}^N q^N$, expressed as a quaternion. We
+    can define Wahba's loss function as the following: 
+    \begin{align} \label{eq:wahba_loss}
+        L &= \sum_i w_i \norm{{}^N v_i - {}^N q^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
+    \end{align} 
+    where $r_i(q)$ is the residual vector.
+
+    We can solve for ${}^N q^N$ using a nonlinear least squares method minimizing
+    Wahba's loss function:
     \begin{mini*}
-    {q}{ \|r(q)\|_2^2 }{}{}
+        {q}{ \norm{r(q)}_2^2 }{}{}
         \addConstraint { q\in SO(3).}
     \end{mini*}
-     There are two ways for keeping the solution in the $SO(3)$ group, a projected method and a multiplicative method.
-In a projected Gauss-Newton method, the Newton step is applied additively, and the resulting estimate is projected back onto $SO(3)$ with a normalization.
-This algorithm can be seen in algorithm \ref{alg:pgn}.
+
+    Following the typical approach for Newton's method, we minimize \eqref{eq:wahba_loss}
+    by setting the gradient to zero:
+    \begin{equation} \label{eq:newton_foc}
+        \begin{aligned}
+            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i{q} := J^T r(q) = 0 \\
+              &= \sum_i (-2 H^T R(q)^T R({}^B \hat{v}_i)^T r(q).
+        \end{aligned}
+    \end{equation}
+    which can be obtained from the chain rule and \eqref{eq:quaternion_rotation}.
+
+    Treating $q$ as a vector in $\R^4$, we obtain a solution to \eqref{eq:newton_foc}
+    using the Moore-Penrose pseudoinverse, $\delta q = (J^T J)^{-1} J^T$, and our
+    next candidate quaternion via simple addition, $q_{k+1} = q_k + \delta q$. Since
+    $q_{k+1}$ will no longer be unit norm, we project it back on the unit sphere via
+    the projection operator $\Pi(q) = q / \norm{q}$. This ``projected'' Newton approach
+    is summarized in Algorithm \ref{alg:pgn}.
+
     \begin{algorithm} 
@@
     		\While{significant progress}
-    		    \State $J = \frac{\partial r(q_k)}{ \partial q}$ 
-    		    \State $ q_{step} = -(J^TJ)^{-1}J^T r(q_k)$ 
+    		    \State $J = \pdv{r(q_k)}{ \partial q}$ 
+    		    \State $ \delta q = -(J^TJ)^{-1}J^T r(q_k)$ 
     		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
-    		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + q_{step})$ 
+    		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + \delta q)$ 
     		    \State $k = k + 1$
@@
     \end{algorithm}
-   This method can successfully lower the loss function initially, but the normalization errors will prevent convergence.
-To get around this, the quaternion calculus presented in this paper can be used to generate a Newton step that is a rotation instead of an addition.
-This method of applying the Newton step preserves the constraint satisfaction, and doesn't introduce any normalization errors.
+
+    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$,
+    we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
+    \eqref{eq:quat_jacobian}:
+    \begin{equation} 
+        \bar{J} = \pdv{r(q \otimes \varphi{\phi})}{\phi} = \pdv{r(q)}{q} G(q).
+    \end{equation}
+    We obtain our step---this time in the actual tangent space---as before:
+    $\phi = (\bar{J}^T \bar{J})^{-1} \bar{J}^T)$. To obtain our next iterate, we ``add'' the
+    step using the correct notion of composition for the group: 
+    $q_{k+1} = q_k \otimes \varphi{\phi}$. This ``multiplicative'' Newton algorithm is 
+    summarized in Algorithm \ref{alg:mgn}.
+
     \begin{algorithm} 
@@
     		\While{significant progress}
-    		    \State $J = [\frac{\partial r(q_k)}{ \partial q}]G(q_k)$ \Comment{quaternion adjusted Jacobian}
-    		    \State $ \phi_{step} = -(J^TJ)^{-1}J^T r(q_k)$ 
-    		    \State $q_{step} = \varphi(\phi_{step})$ 
-    		    \State $q_{k+1} = q_n \otimes q_{step}$ \Comment{apply step multiplicatively}
+    		    \State $\bar{J} = \pdv{r(q_k)}{q} G(q_k)$ \Comment{quaternion adjusted Jacobian}
+    		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
+    		    \State $q_{k+1} = q_n \otimes \varphi{\phi}$ \Comment{apply step multiplicatively}
     		    \State $k = k + 1$
@@
     \subsection{Results}
+    As illustrated in Figure \ref{fig:wahba_convergence}, it is clear that by projecting 
+    back onto the unit quaternion at each iteration make initial progress, but fails to 
+    exhibit the quadratic convergence typical of a Newton method. By optimizing directly 
+    in the space of differential quaternions, we achieve the expected quadratic convergence.
+    It should also be clear from the previous section that the adaptations to the original
+    Newton method are simple and straightforward, highlighting both the effectiveness and
+    value of the proposed approach.
     \begin{figure}
@@
         quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
-        error.}
+        error. The angle error is calculated relative to the true analytical solution obtained
+        via an SVD decomposition.}
         \label{fig:wahba_convergence}
@@
 
-\paragraph{Acknowledgements}
+\paragraph*{Acknowledgements}
 This material is based upon work supported by the National Science Foundation Graduate

### 2020-10-10 paper/main.tex
Updates from Overleaf

--- 2020-10-09 paper/main.tex
@@
     \subsection{Results}
-    \begin{figure}
+
+    As visible in figure \ref{fig:wahba_convergence}, the original Newton's method makes progress early on in the iterations, but stops making progress as the normalization errors grow.
+The modified Newton's method, that applies the newton step as a rotation instead of an addition, is able to converge to floating point accuracy.
+
+    \begin{figure}[H]
         \centering
@@
     \end{figure}
+    % \begin{figure}[H]
+    %     \centering
+    %     \includegraphics[width=\columnwidth, height=5cm]{figures/kevins_plots/convergence_plot.tikz}
+    %     \caption{Convergence comparison for Wahba's problem. By performing Newton's method
+    %     on the error quaternion and applying the result to the full quaternion we achieve
+    %     quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
+    %     error.}
+    %     \label{fig:wahba_convergence}
+    % \end{figure}
 

### 2020-10-10 paper/main.tex
Added flexible sat stuff.

--- 2020-10-10 paper/main.tex
@@
     		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
-    		    \State $q_{k+1} = q_n \otimes \varphi{\phi}$ \Comment{apply step multiplicatively}
+    		    \State $q_{k+1} = q_k \otimes \varphi(\phi)$ \Comment{apply step multiplicatively}
     		    \State $k = k + 1$
@@
     \begin{figure}
+
         \centering
@@
     \subsection{Satellite Attitude Keep-Out}
+    A spacecraft with flexible appendages modes must perform a 150 degree slew, while making sure that a startracker does not look within 40 degrees of the sun.
+The flexible body spacecraft dynamics are based on the truncated hybrid coordinate model, with the three largest modes captured \cite{Tracy2020}.
+We can guarantee the startracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
+    \begin{equation}
+        \langle ^N A ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
+    \end{equation}
+    From this constraint, an eigen-axis slew trajectory planner may violate this constraint \cite{markley2014fundamentals}.
+ALTRO is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the startracker line of sight constraint.
         \begin{figure}
@@
             \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
-            \caption{Attitude keep out}
+            \caption{Visualization of the flexible spacecraft slew with a dedicated keep out zone.
+Attitude is parameterized with a Rodrigues parameter to visualize in 3D.
+The constraint surface represents attitudes where the startracker is within 40$^\circ$ of the sun.
+The unconstrained solution violates this constraint, while the constrained solution is able reason about the keep out zone.}
         \end{figure}
@@
             \includegraphics[height=4cm,width=\columnwidth]{figures/c_max_convergence.tikz}
-            \caption{Constraint convergence when solving the barrell roll problem. Compares 
+            \caption{Constraint convergence when solving the barrel roll problem. Compares 
             the convergence of the original version of ALTRO versus the new, modified version

### 2020-10-10 paper/main.tex
More updates to flexible sat.

@@
     		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
-    		    \State $q_{k+1} = q_k \otimes \varphi(\phi)$ \Comment{apply step multiplicatively}
+    		    \State $q_{k+1} = L(q_k) \varphi(\phi)$ \Comment{apply step multiplicatively}
     		    \State $k = k + 1$
@@
     \subsection{Satellite Attitude Keep-Out}
-    A spacecraft with flexible appendages modes must perform a 150 degree slew, while making sure that a startracker does not look within 40 degrees of the sun.
+    A spacecraft with flexible appendages modes must perform a 150 degree slew, while ensuring that a star tracker does not look within 40 degrees of the sun.
+The attitudes that result in the star tracker looking too close to the sun will be referred to as a keep-out zone.
 The flexible body spacecraft dynamics are based on the truncated hybrid coordinate model, with the three largest modes captured \cite{Tracy2020}.
-We can guarantee the startracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
+We can guarantee the star tracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
     \begin{equation}
@@
     \end{equation}
-    From this constraint, an eigen-axis slew trajectory planner may violate this constraint \cite{markley2014fundamentals}.
-ALTRO is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the startracker line of sight constraint.
+    Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep out zone \cite{markley2014fundamentals}.
+ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
         \begin{figure}
@@
             \caption{Visualization of the flexible spacecraft slew with a dedicated keep out zone.
-Attitude is parameterized with a Rodrigues parameter to visualize in 3D.
-The constraint surface represents attitudes where the startracker is within 40$^\circ$ of the sun.
+Attitude is parameterized with a Rodrigues parameter to visualize in three dimensions.
+The constraint surface represents attitudes where the star tracker is within 40$^\circ$ of the sun.
 The unconstrained solution violates this constraint, while the constrained solution is able reason about the keep out zone.}

### 2020-10-10 paper/main.tex
Updated flexible sat section.

@@
     \begin{align} \label{eq:wahba_loss}
-        L &= \sum_i w_i \norm{{}^N v_i - {}^N q^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
+        L &= \sum_i w_i \norm{{}^N v_i - {}^N q^B \,\, ^Bv_i }_2^2 = \norm{r_i(q)}_2^2, 
     \end{align} 
@@
     \subsection{Satellite Attitude Keep-Out}
-    A spacecraft with flexible appendages modes must perform a 150 degree slew, while ensuring that a star tracker does not look within 40 degrees of the sun.
-The attitudes that result in the star tracker looking too close to the sun will be referred to as a keep-out zone.
+    A spacecraft with flexible appendages must perform a 150 degree slew, while ensuring that a star tracker line-of-sight does not cross within 40 degrees of the sun.
+The attitudes that result in the star tracker pointing too close to the sun will be referred to as a keep-out zone.
 The flexible body spacecraft dynamics are based on the truncated hybrid coordinate model, with the three largest modes captured \cite{Tracy2020}.
@@
     \begin{equation}
-        \langle ^N A ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
+        \langle ^N q ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
     \end{equation}
-    Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep out zone \cite{markley2014fundamentals}.
+    Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep-out zone \cite{markley2014fundamentals}.
 ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
@@
             \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
-            \caption{Visualization of the flexible spacecraft slew with a dedicated keep out zone.
-Attitude is parameterized with a Rodrigues parameter to visualize in three dimensions.
-The constraint surface represents attitudes where the star tracker is within 40$^\circ$ of the sun.
-The unconstrained solution violates this constraint, while the constrained solution is able reason about the keep out zone.}
+            \caption{Visualization of the flexible spacecraft slew with a dedicated keep-out zone.
+Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions.
+The constraint surface represents attitudes where the star tracker line-of-sight is within 40$^\circ$ of the sun.
+The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
         \end{figure}

### 2020-10-10 paper/main.tex
Minor quaternion notation problems

--- 2020-10-10 paper/main.tex
@@
     \begin{align} \label{eq:wahba_loss}
-        L &= \sum_i w_i \norm{{}^N v_i - {}^N q^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
+        L &= \sum_i w_i \norm{{}^N v_i - {}^N A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
     \end{align} 
@@
     step using the correct notion of composition for the group: 
-    $q_{k+1} = q_k \otimes \varphi{\phi}$. This ``multiplicative'' Newton algorithm is 
+    $q_{k+1} = L(q_k) \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
     summarized in Algorithm \ref{alg:mgn}.
@@
 
-        \subsubsection{Error Quadratic}
+        \subsubsection{Error Quadratic} \label{sec:error_quadratic}
             Rather than simple subtraction, we can use a quadratic function on the
@@
     \subsection{Satellite Attitude Keep-Out}
-    A spacecraft with flexible appendages modes must perform a 150 degree slew, while making sure that a startracker does not look within 40 degrees of the sun.
-The flexible body spacecraft dynamics are based on the truncated hybrid coordinate model, with the three largest modes captured \cite{Tracy2020}.
-We can guarantee the startracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
+    A spacecraft with flexible appendages must perform a 150 degree slew, while
+    making sure that a startracker does not look within 40 degrees of the sun. The
+    flexible body spacecraft dynamics are based on the truncated hybrid coordinate model,
+    with the three largest modes captured \cite{Tracy2020}. We can guarantee the
+    startracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the
+    inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
     \begin{equation}
@@
     \end{equation}
-    From this constraint, an eigen-axis slew trajectory planner may violate this constraint \cite{markley2014fundamentals}.
-ALTRO is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the startracker line of sight constraint.
+    From this constraint, an eigen-axis slew trajectory planner may violate this
+    constraint \cite{markley2014fundamentals}. ALTRO is able to reason about the
+    nonlinear flexible spacecraft dynamics, actuator constraints, and the startracker
+    line of sight constraint. The resulting attitude trajectory is shown Figure
+    \ref{{fig:satellite}}
         \begin{figure}
@@
 The unconstrained solution violates this constraint, while the constrained solution is able reason about the keep out zone.}
+            \label{fig:satellite}
         \end{figure}
@@
         the optimization routine.
+
+        We also compared using a quadratic cost on the 3-dimensional quaternion error
+        described in Section \ref{sec:error_quadratic} to the geodesic cost function in
+        Section \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of 
+        iterations and computation time, taking only 41 iterations and
 

### 2020-10-10 paper/main.tex

--- 2020-10-10 paper/main.tex
@@
     \begin{align} \label{eq:wahba_loss}
-        L &= \sum_i w_i \norm{{}^N v_i - {}^N q^B \,\, ^Bv_i }_2^2 = \norm{r_i(q)}_2^2, 
+        L &= \sum_i w_i \norm{{}^N v_i - {}^N A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
     \end{align} 
@@
     \subsection{Satellite Attitude Keep-Out}
-    A spacecraft with flexible appendages must perform a 150 degree slew, while ensuring that a star tracker line-of-sight does not cross within 40 degrees of the sun.
-The attitudes that result in the star tracker pointing too close to the sun will be referred to as a keep-out zone.
-The flexible body spacecraft dynamics are based on the truncated hybrid coordinate model, with the three largest modes captured \cite{Tracy2020}.
-We can guarantee the star tracker body fixed vector ($^Br_{st}$) does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by formulating the following constraint:
+    A spacecraft with flexible appendages must perform a 150 degree slew, while ensuring
+    that a star tracker line-of-sight does not cross within 40 degrees of the sun. The
+    attitudes that result in the star tracker pointing too close to the sun will be
+    referred to as a keep-out zone. The flexible body spacecraft dynamics are based on
+    the truncated hybrid coordinate model, with the three largest modes captured
+    \cite{Tracy2020}. We can guarantee the star tracker body fixed vector ($^Br_{st}$)
+    does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by
+    formulating the following constraint:
     \begin{equation}
@@
     \end{equation}
-    Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep-out zone \cite{markley2014fundamentals}.
-ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
+    Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory
+    planner may end up violating this keep-out zone \cite{markley2014fundamentals}. ALTRO
+    can generate a feasible trajectory for this problem since it is able to reason about
+    the nonlinear flexible spacecraft dynamics, actuator constraints, and the star
+    tracker line of sight constraint.
         \begin{figure}

### 2020-10-10 paper/main.tex
Update barrell roll example and refine paper

@@
 \usepackage{amssymb}
+\usepackage{gensymb}
 
@@
 in attitude.
-\keywords{
-    motion planning and control, quaternions, optimal control, linear quadratic regulator
+\keywords{motion planning and control, quaternions, optimal control, 
+linear quadratic regulator
 }
@@
         \item A unified approach to quaternion differential calculus entirely based on
-        standard vector calculus and linear algebraic operations control of systems with
-        quaternion states
-        \item A fast and efficient solver for trajectory optimization problems with 
+        standard vector calculus and linear algebraic operations, and
+        \item a fast and efficient solver for trajectory optimization problems with 
         attitude dynamics and nonlinear constraints that correctly accounts for the group
-        structure of 3D rotations during the solve
+        structure of 3D rotations during the solve.
     \end{itemize}
@@
     vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
-    rotation from the body to inertial frames ${}^N q^N$, expressed as a quaternion. We
+    rotation from the body to inertial frame ${}^N q^B$, expressed as a quaternion. We
     can define Wahba's loss function as the following: 
@@
 
-    We can solve for ${}^N q^N$ using a nonlinear least squares method minimizing
+    We can solve for ${}^N q^B$ using a nonlinear least squares method minimizing
     Wahba's loss function:
@@
     \begin{equation} 
-        \bar{J} = \pdv{r(q \otimes \varphi{\phi})}{\phi} = \pdv{r(q)}{q} G(q).
+        \bar{J} = \pdv{r(\mathbf{q} \otimes \varphi(\phi))}{\phi} = \pdv{r(q)}{q} G(q).
     \end{equation}
@@
     step using the correct notion of composition for the group: 
-    $q_{k+1} = L(q_k) \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
+    $\q_{k+1} = \q_k \otimes \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
     summarized in Algorithm \ref{alg:mgn}.
@@
     		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
-    		    \State $q_{k+1} = L(q_k) \varphi(\phi)$ \Comment{apply step multiplicatively}
+    		    \State $\q_{k+1} = \q_k \otimes \varphi(\phi)$ \Comment{apply step multiplicatively}
     		    \State $k = k + 1$
@@
     solve trajectory optimization problems for rigid bodies, which extends easily to
-    arbitrary systems in $R^n \times SO(3)$. ALTRO is an efficient solver for constrained
+    arbitrary systems in $\R^n \times SO(3)$. ALTRO is an efficient solver for constrained
     nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented
@@
     from which we can calculate the quadratic expansion of the cost-to-go 
-    $P_k \in \R^{12 \times 12}$, $p_k \in \R^12$, and optimal linearized feedback gains 
+    $P_k \in \R^{12 \times 12}$, $p_k \in \R^{12}$, and optimal linearized feedback gains 
     $K_k \in \R^{m \times 12}$ and $d_k \in \R^m$ by starting at the terminal state and
@@
     where $\bar{u}_k$ is the control value from the previous iteration, and $\delta x$ is
-    computed using \eqref{eq:state_error}, with $x_k$ the current state estimate and $\bar{x}_k$
-    the state from the previous iteration. The rest of the algorithm is left unchanged. 
-    For more details on the ALTRO algorithm, the reader is encouraged to refer to the original
-    paper \cite{howell2019altro}.
+    computed using \eqref{eq:state_error}, with $x_k$ being the current state estimate
+    and $\bar{x}_k$ the state from the previous iteration. The rest of the algorithm is
+    left unchanged. For more details on the ALTRO algorithm, the reader is encouraged to
+    refer to the original paper \cite{howell2019altro}.
 
@@
     AMD Ryzen 2950x processor with 40 GB of RAM. 
-    All problems are run using ALTRO, first without any of the modification presented 
+    All problems are run using ALTRO, first without any of the modifications presented 
     in the current paper, labeled ``original'', and then using the modifications listed
-    in Sec. \ref{sec:trajopt} and the geodesic cost funtion described in Sec. \ref{sec:geodesic}.
+    in Sec. \ref{sec:trajopt} and the geodesic cost funtion described in Sec. \ref{sec:geodesic},
+    labeled ``modified''.
     Timing results are summarized in Table \ref{tab:timing_results}.
-    All experiments were solved to a constraint satisfaction tolerance of $10^{-5}$.
+    All experiments were solved to a constraint satisfaction tolerance of $10^{-5}$ and 
+    discretized with a 4th order Runge-Kutta integrator.
     Code for all experiments is available on 
@@
 
-        We also compared using a quadratic cost on the 3-dimensional quaternion error
-        described in Section \ref{sec:error_quadratic} to the geodesic cost function in
-        Section \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of 
-        iterations and computation time, taking only 41 iterations and
+        We also compared the ``error quadratic'' cost function described in Section
+        \ref{sec:error_quadratic} to the geodesic cost function in Section
+        \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of
+        iterations and computation time, taking only 36 iterations and 1.8 ms/iteration
+        vs 76 iterations and 7.6 ms/iteration for the error quadratic. Not only is the
+        geodesic cost function much cheaper to compute, it also converges much faster
+        than the geodesic cost function. This could be due to the strong nonlinearities
+        introduced when computing the quaternion error, making it more difficult to
+        optimize.
 
@@
 	    ALTRO. The original version of ALTRO, even after significant tuning efforts,
-        would not converge to the desired solution. It instead got halfway and then ``unwound''
-        360 degrees and then continued rotating to the final orientation. This behavior
-        is common and expected when attempting optimization that does not properly account 
-        for the grou structure of rotations.
+	    would not converge to the desired solution. It instead got halfway and then
+	    ``unwound'' 360 degrees and then continued rotating to the final orientation.
+	    This behavior is common and expected when attempting optimization that does not
+	    properly account for the group structure of rotations. It is also worth noting
+	    that this problem could not be solved using any three-parameter attitude
+	    representation, since it passes through the singularities at $90\degree,
+	    180\degree$, and $360\degree$ for Euler angles, Rodrigues parameters, and
+	    Modified Rodrigues Parameters, respectively.
 
@@
             \centering
-                \includegraphics[height=5.2cm,trim={0 0cm 0 0},clip]{figures/quadflip.png}
+            \includegraphics[width=\columnwidth]{figures/quadflip.png}
             \caption{Snapshots of the quadrotor flip trajectory. The
-                    red-colored quadrotors represent the state near t=0 s and the
-                    green-colored quadrotors represent the state near t=5.0 s
+                gree-colored quadrotors represent the state near t=0 s and the
+                red-colored quadrotors represent the state near t=5.0 s
             }
@@
     collocation or sequential convex programming rely on commercial, proprietary, or 
-    general-purpose solvers whose internal numerics often abstracted away from the user. 
+    general-purpose solvers whose internal numerics often are abstracted away from the user. 
     By exploiting both the unique Markovian structure of the trajectory optimization problem
@@
     refinement of the implementation and benchmarking against current state-of-the-art 
-    methods for trajectory optimization.
+    methods for trajectory optimization and other specialized solvers for optimization on
+    manifolds.
 

### 2020-10-10 examples/trajopt_examples.jl

@@ -93,10 +93,9 @@ add_result!(:satellite, true, solver)
 
 ## Save the table
 import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
-using LaTeXTabulars, JLD2, DataFrames, PGFPlotsX
+using LaTeXTabulars, JLD2, DataFrames, PGFPlotsX, Printf
 results_path = joinpath(@__DIR__, "timing_results.jld2")
 @load results_path results
-names = Dict()
 
 df = DataFrame(results)
 df.time .= round.(df.time, digits=2)

### 2020-10-10 examples/yak_barrellroll.jl

@@ -19,7 +19,8 @@ using JLD2
 # using MeshCat
 
 args = (
-    integration = RK3,
+    constraint_tolerance=1e-5,
+    integration = RK4,
     termcon = :quatvec,
     projected_newton = false,
     show_summary=true
@@ -29,6 +30,11 @@ args = (
 """
 Compare Cost Functions
 """
+
+prob,opts = YakProblems(vecstate=false, costfun=:ErrorQuadratic; args...) 
+solver = ALTROSolver(prob, opts)
+solve!(solver)
+ilqr = Altro.get_ilqr(solver)
 
 function run_example(;vecstate=false, costfun=:Quadratic, args...)
     prob,opts = YakProblems(vecstate=vecstate, costfun=costfun; args...) 
@@ -57,7 +63,7 @@ cost_comparison = Dict(
 run_example(vecstate=true; args...)
 run_example(; args...)
 run_example(costfun=:QuatLQR; args...)
-run_example(costfun=:LieLQR; args...)
+# run_example(costfun=:LieLQR; args...)
 run_example(costfun=:ErrorQuadratic; args...)
 
 @save "cost_comparison.jld2" cost_comparison

### 2020-10-10 src/airplane_problem.jl

@@ -1,5 +1,6 @@
 using TrajectoryOptimization: LieLQRCost
 using ForwardDiff
+using RobotDynamics
 
 function YakProblems(;
         integration=RD.RK4,
@@ -15,8 +16,8 @@ function YakProblems(;
 
     opts = SolverOptions(
         cost_tolerance_intermediate = 1e-1,
-        penalty_scaling = 100.,
-        penalty_initial = 0.01;
+        penalty_scaling = 10.,
+        penalty_initial = 10.;
         kwargs...
     )
 
@@ -40,6 +41,13 @@ function YakProblems(;
         utrim  = @SVector  [41.6666, 106, 74.6519, 106]
         xf = RD.build_state(model, [3,0,1.5], pf, [5,0,0], [0,0,0])
 
+        # Xref trajectory
+        x̄0 = RBState(model, x0)
+        x̄f = RBState(model, xf)
+        Xref = map(1:N) do k
+            x̄0 + (x̄f - x̄0)*((k-1)/(N-1))
+        end
+
         # Objective
         Qf_diag = RD.fill_state(model, 100, 500, 100, 100.)
         Q_diag = RD.fill_state(model, 0.1, 0.1, 0.1, 0.1)
@@ -52,21 +60,31 @@ function YakProblems(;
             utrim = push(utrim, 0)
         end
         if costfun == :Quadratic
+            costfuns = map(Xref) do xref
+                LQRCost(Q, R, xf, utrim)
+            end
             costfun = LQRCost(Q, R, xf, utrim)
             costterm = LQRCost(Qf, R, xf, utrim)
+            costfuns[end] = costterm
         elseif costfun == :QuatLQR
-            # costfun = LieLQRCost(s, Q, R, xf, utrim; w=0.1)
-            # costterm = LieLQRCost(s, Qf, R, xf, utrim; w=200.0)
+            costfuns = map(Xref) do xref
+                QuatLQRCost(Q, R, xf, utrim, w=0.1)
+            end
             costfun = QuatLQRCost(Q, R, xf, utrim; w=0.1)
             costterm = QuatLQRCost(Qf, R, xf, utrim; w=200.0)
+            costfuns[end] = costterm
         elseif costfun == :LieLQR
             costfun = LieLQR(s, Q, R, xf, utrim)
             costterm = LieLQR(s, Qf, R, xf, utrim)
         elseif costfun == :ErrorQuadratic
+            costfuns = map(Xref) do xref
[…]

### 2020-10-11 paper/main.tex
edits + comments

@@
 
+\DeclareMathOperator{\sign}{sign}
+
 \newcommand{\dR}{\delta \mathcal{R}}
@@
 rotations. This paper introduces a powerful and accessible approach for 
-optimization-based planning on the space of rotations based on vector calculus and
-linear algebra. We demonstrate the effectiveness of the approach by adapting Newton's
-method to solve the canonical Wahba's problem, and modifying the ALTRO solver for 
-constrained nonlinear trajectory optimization to plan directly on the space of unit 
+optimization-based planning on the space of rotations using only standard linear algebra and vector calculus. We demonstrate the effectiveness of the approach by adapting Newton's
+method to solve the canonical Wahba's problem, and modifying the trajectory optimization solver ALTRO to plan directly on the space of unit 
 quaternions, achieving superior convergence on problems involving significant changes 
 in attitude.
-\keywords{motion planning and control, quaternions, optimal control, 
-linear quadratic regulator
-}
+%\keywords{trajectory optimization, motion planning, control, quaternions}
+
 \end{abstract}
@@
 
-    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous
+    Many robotic systems---including quadrotors, airplanes, satellites, autonomous
     underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
@@
     not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing
-    rotations, we must either a) pick a three-parameter representation and deal with
-    discontinuities, or b) pick a higher-dimensional representation and deal with
+    rotations, we must either a) choose a three-parameter representation and deal with
+    singularities and discontinuities, or b) choose a higher-dimensional representation and deal with
     constraints between the parameters. While simply representing attitude is nontrivial,
@@
     \cite{Zefran1998}, sample-based motion planning \cite{Zefran1999,Kuffner2004}, and
-    optimal control. Approaches to optimal control on attitude problems include
+    optimal control. Approaches to optimal control with attitude states include
     analytical methods applied to satellites \cite{Spindler1998}, discrete mechanics
@@
     however, despite these works, many recent papers in the robotics community continue
-    to apply traditional methods for motion planning and control with no regard for the
+    to naively apply standard methods for motion planning and control with no regard for the
     group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
@@
     control that rely heavily on ideas and notation from differential geometry, 
-    and instead use only basic mathematical tools from linear algebra and calculus that 
-    should be familiar to most roboticists. 
-    Similar to \cite{Mandic2011,Xu2016}, in Sec. \ref{sec:Quaternion_Calculus} we introduce 
-    a quaternion differential calculus, but take a significantly simpler and more general 
-    approach, enabling straight-forward adaptation of 
+    and instead use only basic mathematical tools from linear algebra and vector calculus that 
+    should be familiar to most roboticists. In Sec. \ref{sec:Quaternion_Calculus} we introduce 
+    an approach to quaternion differential calculus similar to \cite{Mandic2011,Xu2016}, but significantly simpler and more general, enabling straight-forward adaptation of 
     existing algorithms to systems with quaternion states. 
-    To make this concrete, in Sec. \ref{sec:Wahbas} we apply this method to the canonical
-    Wahba's problem and demonstrate dramatically superior convergence to approaches that
+    For concreteness, we then apply our method to the canonical
+    Wahba's problem~\cite{Wahba1965} in Sec. \ref{sec:Wahbas}, and demonstrate superior convergence to approaches that
     fail to properly account for the group structure. 
@@
     and detail modifications to ALTRO, a state-of-the-art constrained trajectory optimization
-    solver, and demonstrate the performance gains on several constrained benchmark problems.
-    To our knowledge, there does not currently exist a solver that is capable of leveraging
-    the unique Markovian structure of the fixed-horizon trajectory optimization problem while
-    correctly accounting for the group structure of 3D rotations.
+    solver, and demonstrate performance gains on several benchmark problems.
+    %To our knowledge, there does not currently exist a solver that is capable of leveraging the unique Markovian structure of the fixed-horizon trajectory optimization problem while correctly accounting for the group structure of 3D rotations.
     In summary, our contributions include:
@@
     \begin{itemize}
-        \item A unified approach to quaternion differential calculus entirely based on
-        standard vector calculus and linear algebraic operations, and
+        \item A unified approach to quaternion differential calculus entirely based on standard linear algebra and vector calculus.
+        \item Derivation of a Newton-based algorithm for nonlinear optimization directly on the space of unit quaternions using our notation.
         \item a fast and efficient solver for trajectory optimization problems with 
         attitude dynamics and nonlinear constraints that correctly accounts for the group
-        structure of 3D rotations during the solve.
+        structure of 3D rotations.
     \end{itemize}
@@
         a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and 
-        $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively.
+        $q_v \in \R^3$ are referred to as the scalar and vector parts of the quaternion, respectively.
 
@@
     \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
-        In the current work we will restrict our focus to rigid bodies moving freely in 3D 
+        For clarity, we will restrict our attention to rigid bodies moving freely in 3D 
         space. That is, we consider systems with dynamics of the following form:
@@
         where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, 
-        $R \in SO(3)$ is the attitude, $v \in \R^3$ is the linear velocity, and 
+        $R \in SO(3)$ is the attitude \todo{why didn't we just make this $q$?}, $v \in \R^3$ is the linear velocity, and 
         $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, 
         $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the 
-        global frame, and $\tau_L(x,u)$ are the moments in the local (body) frame.
+        global frame, and $\tau_L(x,u)$ are the moments in the local (body) frame \todo{Can we stick with just the term ``body'' throughout the paper (and maybe change $L$ to $B$ here)?}.
 
@@
         intuitive sense: Rotations are inherently three-dimensional and differential
-        rotations should live in the same space as angular velocity, i.e. $\R^3$.
+        rotations should live in the same space as angular velocities, i.e. $\R^3$.
 
@@
         particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be
-        obtained using finite difference or automatic differentiation.
+        obtained using finite differences or automatic differentiation.
         As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the
-        same (up to a constant scaling factor) for any choice of three-parameter attitude
+        same (up to a constant scalar factor) for any choice of three-parameter attitude
         representation.
@@
 	    where the second term comes from the second derivative of $\varphi(\phi)$.
-	    Similar to $G(q)$, this ends up being the same (up to a scaling factor) for any
+	    Similar to $G(q)$, this expression is the same (up to a constant scalar factor) for any
         choice of three-parameter attitude representation.
@@
         We now consider the case of a function that maps unit quaternions to unit
-        quaternions, $q' = f(q) : \Q \to \Q$. Here we must also consider the non-trivial
-        effect of a differential value applied to the output, i.e.:
+        quaternions, $q' = f(q) : \Q \to \Q$. \todo{Let's make sure we're bing consistent with our usage of $\Q$ vs. $\mathbb{H}$.} Here we must also consider the non-trivial
+        effect of a differential rotation applied to the output, i.e.:
         \begin{equation} \label{eq:dqoutput}
@@
         The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to
-        I_q$, where $I_q$ is the quaternion identity. Differentiating through the inverse
-        map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q}
-        \to H^T$ for any three-parameter attitude representation.
+        I_q$, where $I_q$ is the quaternion identity. Once again, \eqref{eq:quat_jacobian} holds (up to a constant) for any three-parameter attitude representation.
+
+        %Differentiating through the inverse map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q}\to H^T$ for any three-parameter attitude representation.
 
 \section{Modifying Newton's Method} \label{sec:Wahbas}
+
+\todo{I would re-work this section to only talk about the details of doing it the ``correct'' way (i.e.
+consistent with the previous section).
+Then just call the standard version a ``naive Newton method in which the quaternion is re-normalized at each step'' and don't bother writing any of the mathematical details for that version.
+I think it will get confusing if you mix derivative types/notation.}
 
@@
     Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's
-    attitude from vector measurements \cite{markley2014fundamentals}. We will solve this
+    attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will solve this
     problem using a standard nonlinear least squares method, as well as a method that
@@
     \subsection{Methodology}
-
+\todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
     Given known vectors in some inertial frame, ${}^N v_i$, and measurements of these
@@
     \end{align} 
+    \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
     where $r_i(q)$ is the residual vector.
@@
     \end{mini*}
-
+    \todo{This should be $q \in \Q$ here, since unit quaternions are not isomorphic to $SO(3)$ (rotation matrices).}
     Following the typical approach for Newton's method, we minimize \eqref{eq:wahba_loss}
@@
     exhibit the quadratic convergence typical of a Newton method. By optimizing directly 
-    in the space of differential quaternions, we achieve the expected quadratic convergence.
+    in the space of unit quaternions, we achieve the expected quadratic convergence.
     It should also be clear from the previous section that the adaptations to the original
@@
     solve trajectory optimization problems for rigid bodies, which extends easily to
-    arbitrary systems in $\R^n \times SO(3)$. ALTRO is an efficient solver for constrained
-    nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented
-    Lagrangian framework.
+    arbitrary systems whose state is in $\R^n \times SO(3)$.
+%ALTRO is an efficient solver for constrained nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented Lagrangian framework.
 
-    We consider a trajectory optimization problem of the form:
+    We consider trajectory optimization problems of the form,
     \begin{mini}[2]
@@
     \end{equation}
-    and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state error Jacobian:
+    and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state-error Jacobian, respectively:
     \begin{equation} \label{eq:state_error}
@@
     \begin{multline}
-        \delta \ell(x,u) \approx  \half \delta x^T \ell_xx \delta x
+        \delta \ell(x,u) \approx  \half \delta x^T \ell_{xx} \delta x
             + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta u \\
@@
     \end{multline}
-    of the augmented Lagrangian cost function:
+    of the augmented Lagrangian cost function: \todo{I think you can safely leave out any references to the augmented Lagrangian, etc. and just talk about the cost function.}
     \begin{equation}
@@
         minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
-        (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We
+        (x-x_g)$; However, na\"ive substraction of unit quaternions is ill-defined. We
         propose two different cost functions that accomplish similar behavior.
@@
         \subsubsection{Geodesic Distance} \label{sec:geodesic}
-            Alternatively, we can use the geodesic distance between two quaternions
+            Alternatively, we can use the geodesic distance between two quaternions:
             \cite{Kuffner2004},
@@
             \begin{align}
-                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
-                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
+                \nabla J_\text{geo} &= \sign(q_d^T q) q_g^T G(q) \\
+                \nabla^2 J_\text{geo} &= \sign(q_d^T q) I_3 q_g^T q ,
             \end{align}
-            where the sign of the Hessian corresponds to the sign of $q_d^T q$.
+where $\sign$ denotes the signum function.
+
+\todo{If the geodesic distance works better on all the examples, I vote for leaving out the error quadratic stuff.}
 
@@
     In this section we present several trajectory optimization problems for systems that
-    undergo large changes in attitude: an airplane barrell roll, a quadrotor flip, and a
-    satellite with flexible modes that must slew to a new orientation while avoiding a 
-    ``keep out'' orientation zone. All results were run on a desktop computer with an 
+    undergo large changes in attitude: an airplane barrel roll, a quadrotor flip, and a
+    satellite with flexible solar panels that must slew to a new orientation while avoiding a 
+    keep-out zone. All results were run on a desktop computer with an 
     AMD Ryzen 2950x processor with 40 GB of RAM. 
     All problems are run using ALTRO, first without any of the modifications presented 
-    in the current paper, labeled ``original'', and then using the modifications listed
-    in Sec. \ref{sec:trajopt} and the geodesic cost funtion described in Sec. \ref{sec:geodesic},
+    in the current paper, analagous to the naive Newton's method in section \ref{sec:Wahbas} and labeled ``naive'', and then using the modifications listed
+    in Sec. \ref{sec:trajopt} and the geodesic cost function described in Sec. \ref{sec:geodesic},
     labeled ``modified''.
@@
         \label{tab:timing_results}
-        \caption{Trajectory Optimization Timing Results}
+        \caption{Trajectory Optimization Timing Results (naive/modified)}
     \end{table}
@@
     \subsection{Satellite Attitude Keep-Out}
-    A spacecraft with flexible appendages must perform a 150 degree slew, while ensuring
+    \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes.
+Also, let's just call the start tracker a camera or sensor.}
+    A spacecraft with flexible appendages must perform a 150 degree slew while ensuring
     that a star tracker line-of-sight does not cross within 40 degrees of the sun. The
@@
     formulating the following constraint:
+    \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
     \begin{equation}
@@
     \end{equation}
-    Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory
-    planner may end up violating this keep-out zone \cite{markley2014fundamentals}. ALTRO
-    can generate a feasible trajectory for this problem since it is able to reason about
-    the nonlinear flexible spacecraft dynamics, actuator constraints, and the star
-    tracker line of sight constraint.
+    %Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep-out zone \cite{markley2014fundamentals}.
+ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
         \begin{figure}
@@
             \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
-            \caption{Visualization of the flexible spacecraft slew with a dedicated keep-out zone.
+            \caption{Visualization of the flexible spacecraft slew with a keep-out zone.
 Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions.
@@
         \end{figure}
+\todo{Talk a little about the results in the plot here.}
 
     \subsection{Airplane Barrel Roll}
+\todo{Reference \eqref{eq:rigid_body_dynamics} and then say that the forces and torques due to lift and drag are fit from wind tunnel data. Also, cite that paper.}
 
@@
         see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
-        conditions. The convergence of the different versions is compared in Fig.
-        \ref{fig:c_max_convergence}. For both the quadrotor flip and the airplane barrel
+        conditions. The convergence of the different versions of ALTRO is compared in Fig.
+        \ref{fig:c_max_convergence}. For both the quadrotor flip \todo{put the quadrotor comments in the quadrotor section.} and the airplane barrel 
         roll, the modified version of ALTRO converged faster than original version.
@@
 
+		\todo{If we're not going to plot it and actually do an experimental comparison, I vote for leaving this stuff out.}
         We also compared the ``error quadratic'' cost function described in Section
@@
         geodesic cost function much cheaper to compute, it also converges much faster
-        than the geodesic cost function. This could be due to the strong nonlinearities
+        than the error quadratic cost function. This could be due to the strong nonlinearities
         introduced when computing the quaternion error, making it more difficult to
@@
     \subsection{Quadrotor Flip}
-        We sucessfully optimized a 360 degree flip trajectory for a quadrotor using the 
+    \todo{Cite the Penn quadrotor paper for the dynamics model.}
+        We optimized a 360 degree flip trajectory for a quadrotor using the 
         modified version of ALTRO.
-	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
+	    Four intermediary ``keyframes'' were used to encourage the quadrotor to be at angles
         of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular arc.
@@
 	    ALTRO. The original version of ALTRO, even after significant tuning efforts,
-	    would not converge to the desired solution. It instead got halfway and then
-	    ``unwound'' 360 degrees and then continued rotating to the final orientation.
+	    was not able to converge to the desired solution.
+	    %It instead got halfway and then ``unwound'' 360 degrees and then continued rotating to the final orientation.
 	    This behavior is common and expected when attempting optimization that does not
@@
 	    representation, since it passes through the singularities at $90\degree,
-	    180\degree$, and $360\degree$ for Euler angles, Rodrigues parameters, and
+	    180\degree$, and $360\degree$ associated with Euler angles, Rodrigues parameters, and
 	    Modified Rodrigues Parameters, respectively.
@@
 \section{Conclusions} \label{sec:conclusion}
-    We have presented a general, unified method for planning and control on rigid-body
+    We have presented a general, unified method for optimization-based planning and control for rigid-body
     systems with arbitrary attitude using standard linear algebra and vector calculus.
     We have demonstrated that the application of this methodology is straightforward and
-    yields substantion improvements in the convergence of Newton-based methods (see Fig. 
-    \ref{fig:wahba_convergence}) while also offering improvements for nonlinear constrained
+    yields substantial improvements in the convergence of Newton-based methods (see Fig. 
+    \ref{fig:wahba_convergence}), while also offering improvements for nonlinear constrained
     trajectory optimization for floating-base systems (see Table \ref{tab:timing_results}).
@@
     With the modifications presented, ALTRO can solve problems few other methods for 
-    trajectory optimization can. Many state-of-the-art methods such as direct 
-    collocation or sequential convex programming rely on commercial, proprietary, or 
+    trajectory optimization can. Many state-of-the-art methods, such as direct 
+    collocation or sequential convex programming, rely on commercial, proprietary, or 
     general-purpose solvers whose internal numerics often are abstracted away from the user. 
-    By exploiting both the unique Markovian structure of the trajectory optimization problem
-    and the group structure of rotations, ALTRO will likely be able to solve new classes of 
-    problems with at or near real-time performance. Future work will focus on continued 
-    refinement of the implementation and benchmarking against current state-of-the-art 
-    methods for trajectory optimization and other specialized solvers for optimization on
-    manifolds.
+    By exploiting the unique structure of both the trajectory optimization problem
+    and the rotation group, ALTRO will likely be able to solve more challenging problems with performance.
 
-    Additionally, the methods presented here can easily be leveraged to adapt other classes
+    The methods presented here can easily be leveraged to adapt other classes
     of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future
@@
 
-    Future work will also include the application of these methods to multi-body robotic
-    systems, such as humanoids or quadrupeds, especially in methods that exploit 
+    Future work will focus on continued 
+    refinement of ALTRO and performance improvements for use in real-time model-predictive control applications, as well as extensions to multi-body robotic
+    systems, such as humanoids or quadrupeds, represented in
     ``maximal'' coordinates that include the 3D orientation of each body \cite{brudigam2020linear}.
@@
 
+\todo{Add NASA acknowledgements (JPL + ECF)}
+
 \printbibliography

### 2020-10-12 paper/main.tex
Addressed Zac's comments.

@@
     attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will solve this
-    problem using a standard nonlinear least squares method, as well as a method that
+    problem using a modified version of Newton's method that
     exploits the true group structure of $SO(3)$ using the quaternion calculus presented
@@
     \subsection{Methodology}
-\todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
-    Given known vectors in some inertial frame, ${}^N v_i$, and measurements of these
+% \todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
+    Given known vectors in some world frame, ${}^W v_i$, and measurements of these
     vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
-    rotation from the body to inertial frame ${}^N q^B$, expressed as a quaternion. We
+    rotation from the body to world frame ${}^W q^B$, expressed as a quaternion. We
     can define Wahba's loss function as the following: 
     \begin{align} \label{eq:wahba_loss}
-        L &= \sum_i w_i \norm{{}^N v_i - {}^N A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
+        L &= \sum_i \norm{{}^W v_i - {}^W A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
     \end{align} 
-    \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
+    % \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
     where $r_i(q)$ is the residual vector.
 
-    We can solve for ${}^N q^B$ using a nonlinear least squares method minimizing
+    We can solve for ${}^W q^B$ using a nonlinear least squares method minimizing
     Wahba's loss function:
@@
         {q}{ \norm{r(q)}_2^2 }{}{}
-        \addConstraint { q\in SO(3).}
+        \addConstraint { q\in \Q.}
     \end{mini*}
-    \todo{This should be $q \in \Q$ here, since unit quaternions are not isomorphic to $SO(3)$ (rotation matrices).}
+    % \todo{This should be $q \in \Q$ here, since unit quaternions are not isomorphic to $SO(3)$ (rotation matrices).}
     Following the typical approach for Newton's method, we minimize \eqref{eq:wahba_loss}
@@
 
-    \begin{algorithm} 
-    	\begin{algorithmic}[1]
-    		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
-    		\State $k = 0$
-    		\While{significant progress}
-    		    \State $J = \pdv{r(q_k)}{ \partial q}$ 
-    		    \State $ \delta q = -(J^TJ)^{-1}J^T r(q_k)$ 
-    		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
-    		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + \delta q)$ 
-    		    \State $k = k + 1$
-    		\EndWhile
-    	\end{algorithmic}
-    \end{algorithm}
+    % \begin{algorithm} 
+    % 	\begin{algorithmic}[1]
+    % 		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
+    % 		\State $k = 0$
+    % 		\While{significant progress}
+    % 		    \State $J = \pdv{r(q_k)}{ \partial q}$ 
+    % 		    \State $ \delta q = -(J^TJ)^{-1}J^T r(q_k)$ 
+    % 		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
+    % 		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + \delta q)$ 
+    % 		    \State $k = k + 1$
+    % 		\EndWhile
+    % 	\end{algorithmic}
+    % \end{algorithm}
 
-    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$,
+    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$ \todo{Brian: is this supposed to be "differential rotation"?},
     we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
@@
     formulating the following constraint:
-    \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
+    % \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
     \begin{equation}
-        \langle ^N q ^B {}^Br_{st} , ^Nr_{sun}\rangle \leq \cos(40 ^\circ).
+         [{}^W r_{sun}]^T[{}^W A(q) {}^B\, {}^Br_{st} ] \leq \cos(40 ^\circ).
     \end{equation}
@@
 The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
+            \label{fig:keepout}
         \end{figure}
-\todo{Talk a little about the results in the plot here.}
+% \todo{Talk a little about the results in the plot here.}
+From Figure \ref{fig:keepout} we can see that without the pointing constraint present, the trajectory will pass through the keep-out zone, making the trajectory infeasible.
+With the keep-out zone constraint present, ALTRO is able to converge on a trajectory that .
 

### 2020-10-12 paper/main.tex
Adressed Zac's comments on flexible sat and Wahba sections.

@@
         \begin{aligned}
-            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i{q} := J^T r(q) = 0 \\
+            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i{q} := \bar{J}^T r(q) = 0 \\
               &= \sum_i (-2 H^T R(q)^T R({}^B \hat{v}_i)^T r(q).
@@
 
-    Treating $q$ as a vector in $\R^4$, we obtain a solution to \eqref{eq:newton_foc}
-    using the Moore-Penrose pseudoinverse, $\delta q = (J^T J)^{-1} J^T$, and our
-    next candidate quaternion via simple addition, $q_{k+1} = q_k + \delta q$. Since
-    $q_{k+1}$ will no longer be unit norm, we project it back on the unit sphere via
-    the projection operator $\Pi(q) = q / \norm{q}$. This ``projected'' Newton approach
-    is summarized in Algorithm \ref{alg:pgn}.
+    % Treating $q$ as a vector in $\R^4$, we obtain a solution to \eqref{eq:newton_foc}
+    % using the Moore-Penrose pseudoinverse, $\delta q = (J^T J)^{-1} J^T$, and our
+    % next candidate quaternion via simple addition, $q_{k+1} = q_k + \delta q$. Since
+    % $q_{k+1}$ will no longer be unit norm, we project it back on the unit sphere via
+    % the projection operator $\Pi(q) = q / \norm{q}$. This ``projected'' Newton approach
+    % is summarized in Algorithm \ref{alg:pgn}.
 
@@
 
-    Alternatively, if we instead minimize with respect to a differential quaternion $\phi$ \todo{Brian: is this supposed to be "differential rotation"?},
-    we adapt the algorithm by simply ``correcting'' our Jacobian $J$ using 
+    We can minimize this loss function with respect to a differential rotation $\phi$,
+    we do this by simply ``correcting'' our Jacobian $\bar{J}$ using 
     \eqref{eq:quat_jacobian}:
     \begin{equation} 
-        \bar{J} = \pdv{r(\mathbf{q} \otimes \varphi(\phi))}{\phi} = \pdv{r(q)}{q} G(q).
+        {J} = \pdv{r(\mathbf{q} \otimes \varphi(\phi))}{\phi} = \pdv{r(q)}{q} G(q).
     \end{equation}
-    We obtain our step---this time in the actual tangent space---as before:
-    $\phi = (\bar{J}^T \bar{J})^{-1} \bar{J}^T)$. To obtain our next iterate, we ``add'' the
+    We then use this to obtain our Newton step by using the Moore-Penrose psuedoinverse:
+    $\phi = ({J}^T {J})^{-1} {J}^T$. To obtain our next iterate, we ``add'' the
     step using the correct notion of composition for the group: 
@@
     		\While{significant progress}
-    		    \State $\bar{J} = \pdv{r(q_k)}{q} G(q_k)$ \Comment{quaternion adjusted Jacobian}
-    		    \State $ \phi = -(\bar{J}^T \bar{J})^{-1} \bar{J}^T r(q_k)$ 
+    		    \State ${J} = \pdv{r(q_k)}{q} G(q_k)$ \Comment{quaternion adjusted Jacobian}
+    		    \State $ \phi = -({J}^T {J})^{-1} {J}^T r(q_k)$ 
     		    \State $\q_{k+1} = \q_k \otimes \varphi(\phi)$ \Comment{apply step multiplicatively}
@@
     \subsection{Results}
-    As illustrated in Figure \ref{fig:wahba_convergence}, it is clear that by projecting 
-    back onto the unit quaternion at each iteration make initial progress, but fails to 
+    We compared this modified Newton's method with a na\"ive Newton's method where the quaternion is simply re-normalized at every iteration. 
+    As illustrated in Figure \ref{fig:wahba_convergence}, it is clear that the naive original method make progress initially, but fails to
     exhibit the quadratic convergence typical of a Newton method. By optimizing directly 
-    in the space of unit quaternions, we achieve the expected quadratic convergence.
+    in the space of unit quaternions, the modified method is able to achieve the expected quadratic convergence.
     It should also be clear from the previous section that the adaptations to the original
@@
     \subsection{Satellite Attitude Keep-Out}
-    \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes.
+    % \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes.
 Also, let's just call the start tracker a camera or sensor.}
     A spacecraft with flexible appendages must perform a 150 degree slew while ensuring
-    that a star tracker line-of-sight does not cross within 40 degrees of the sun. The
-    attitudes that result in the star tracker pointing too close to the sun will be
-    referred to as a keep-out zone. The flexible body spacecraft dynamics are based on
-    the truncated hybrid coordinate model, with the three largest modes captured
-    \cite{Tracy2020}. We can guarantee the star tracker body fixed vector ($^Br_{st}$)
-    does not look within 40 degrees of the inertial sun vector ($^Nr_{sun}$) by
-    formulating the following constraint:
+    that a body-mounted camera does not cross within 40 degrees of the sun. The
+    attitudes that result in the camera pointing too close to the sun will be
+    referred to as a keep-out zone.
+The flexible body spacecraft dynamics are based on equation \eqref{eq:rigid_body_dynamics}, with the addition of four reaction wheels, and the three largest flexible modes \cite{Tracy2020}.
+In order to ensure no damage to the camera, we must formulate a constraint that guarantees that the camera line-of-sight body-fixed vector ($^Br_{cam}$)
+    does not look within 40 degrees of the world-fixed sun vector ($^Wr_{sun}$). This angle constraint can be represented as a dot product between the two vectors expressed in the world frame:
     % \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
     \begin{equation}
-         [{}^W r_{sun}]^T[{}^W A(q) {}^B\, {}^Br_{st} ] \leq \cos(40 ^\circ).
+         [{}^W r_{sun}]^T[{}^W A(q) {}^B\, {}^Br_{cam} ] \leq \cos(40 ^\circ).
     \end{equation}
-    %Due to this constraint on the feasible attitudes, an eigen-axis slew trajectory planner may end up violating this keep-out zone \cite{markley2014fundamentals}.
-ALTRO can generate a feasible trajectory for this problem since it is able to reason about the nonlinear flexible spacecraft dynamics, actuator constraints, and the star tracker line of sight constraint.
-        \begin{figure}
+    The attitudes that satisfy this constraint comprise a non-convex set, with the constraint itself being nonlinear in the attitude quaternion. 
+        \begin{figure}[H]
             \centering
@@
 Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions.
-The constraint surface represents attitudes where the star tracker line-of-sight is within 40$^\circ$ of the sun.
+The constraint surface represents attitudes where the camera line-of-sight is within 40$^\circ$ of the sun.
 The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
@@
 % \todo{Talk a little about the results in the plot here.}
-From Figure \ref{fig:keepout} we can see that without the pointing constraint present, the trajectory will pass through the keep-out zone, making the trajectory infeasible.
-With the keep-out zone constraint present, ALTRO is able to converge on a trajectory that .
+ALTRO is able to reason about this nonlinear constraint, and account for the group structure present in the quaternion.
+From Figure \ref{fig:keepout} we can see that without the camera constraint present, the trajectory will pass through the keep-out zone, making the trajectory infeasible.
+With the keep-out zone constraint present, ALTRO is able to converge on a trajectory that performs the slew in an efficient manner, while keeping the camera sufficiently far from the sun.
 

### 2020-10-12 Wahbas_problem/NLLS_Wahba.jl
PGF plots for Wahba's problem

@@ -1,8 +1,9 @@
 # activate the virtual environment
-cd(dirname(@__FILE__))
-Pkg.activate(".")
+# cd(dirname(@__FILE__))
+import Pkg; Pkg.activate(@__DIR__)
 
-using ForwardDiff, LinearAlgebra, MATLAB, Statistics
+using ForwardDiff, LinearAlgebra, Statistics, JLD2
+# using MATLAB
 
 function randq()
     """Random quaternion"""
@@ -348,15 +349,23 @@ function projected_newton_wahba()
 
 end
 
-# run trials
-trials = 250
-mn = zeros(trials,21)
-pn = zeros(trials,21)
+## run trials
+trials = 25
+max_iters = 21
+mn = zeros(trials,max_iters)
+pn = zeros(trials,max_iters)
 
 for i = 1:trials
-q, qsvd, mn[i,:] = newton_wahba()
-q, qsvd, pn[i,:] = projected_newton_wahba()
+    q, qsvd, mn[i,:] = newton_wahba()
+    q, qsvd, pn[i,:] = projected_newton_wahba()
 end
+@save joinpath(@__DIR__, "wahba_results.jld2") mn pn
+
+
+
+## Generate plot
+using PGFPlotsX, Statistics
+@load joinpath(@__DIR__, "wahba_results.jld2") mn pn
 
 # average the results
 p_average = zeros(5)
@@ -366,6 +375,63 @@ for i = 1:5
     p_average[i] = (mean(pn[:,i]))
     m_average[i] = (mean(mn[:,i]))
 end
+
+p_plots = map(eachrow(pn)) do p
+    @pgf PlotInc(
+        {
+            color="cyan",
+            no_marks,
+            "line width=0.1pt",
+            "forget plot",
+            opacity=0.3
+            # "dashed"
+        },
+        Coordinates(1:5, p[1:5])
+    )
+end
+m_plots = map(eachrow(mn)) do p
+    @pgf PlotInc(
+        {
+            color="orange",
+            no_marks,
+            "line width=0.1pt",
+            # "dashed",
+            "forget plot",
+            opacity=0.3
+        },
+        Coordinates(1:5, p[1:5])
+    )
+end
+p = @pgf Axis(
+    {
+        xlabel="iterations",
+        ylabel="angle error (degrees)",
+        "ymode=log",
+        xmajorgrids,
+        ymajorgrids,
+        "legend style={at={(0.1,0.1)},anchor=south west}"
+    },
+    m_plots...,
+    p_plots...,
[…]

### 2020-10-12 paper/main.tex
Address Zac's comments

@@
 \usepackage{gensymb}
+\usepackage{tensor}
 
@@
 \newcommand{\fail}{{\color{red} \ding{55}}}
+\newcommand{\inframe}[2]{{}^{#1}\!#2}
+\newcommand{\toframe}[3]{\inframe{#1}{#3}^{#2}}
 
@@
     We begin by defining some useful conventions and notation. 
-    Attitude is defined as the rotation from the robot's body frame to a global inertial 
-        frame. 
+    Attitude is defined as the rotation from the robot's body frame to the world frame.
     We also define gradients to be row vectors, that is, for 
@@
         quaternions $\mathbb{H}$ is isomorphic to $\R^4$ to explicitly
-        represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as 
-        a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and 
-        $q_v \in \R^3$ are referred to as the scalar and vector parts of the quaternion, respectively.
+        represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$
+        as a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and
+        $q_v \in \R^3$ are referred to as the scalar and vector parts of the quaternion,
+        respectively. The space of unit quaternions, $\Q = \{q : \norm{q}_2 = 1\}$, is a 
+        a double-cover of the rotation group $SO(3)$, since $q$ and $-q$ represent the same 
+        rotation \cite{markley2014fundamentals}.
 
@@
 
-        The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal 
-        to its conjugate $q^*$, which is simply the same quaternion with a negated vector 
+        The inverse of a unit quaternion $\q^{-1}$, giving the opposite rotation, is equal 
+        to its conjugate $\q^*$, which is simply the same quaternion with a negated vector 
         part:
@@
                 \half \q \otimes \hat{\omega} = \half L(q) H \omega \\ 
-                \frac{1}{m} F_G(x,u) \\ 
-                J^{-1}(\tau_L(x,u) - \omega \times J \omega) 
+                \frac{1}{m} \inframe{W}{F(x,u)} \\ 
+                J^{-1}(\inframe{B}{\tau(x,u)} - \omega \times J \omega) 
             \end{bmatrix}
         \end{equation}
-        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, 
-        $R \in SO(3)$ is the attitude \todo{why didn't we just make this $q$?}, $v \in \R^3$ is the linear velocity, and 
-        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, 
-        $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the 
-        global frame, and $\tau_L(x,u)$ are the moments in the local (body) frame \todo{Can we stick with just the term ``body'' throughout the paper (and maybe change $L$ to $B$ here)?}.
+        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the
+        position, $\q \in \Q$ is the attitude, $v \in \R^3$ is the linear velocity, and
+        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, $J \in
+        \R^{3\times3}$ is the inertia matrix, $\inframe{W}{F(x,u)} \in \R^3$ are the
+        forces in the world frame, and $\inframe{B}{\tau(x,u)}$ are the moments in the
+        body frame.
 
@@
         We now consider the case of a function that maps unit quaternions to unit
-        quaternions, $q' = f(q) : \Q \to \Q$. \todo{Let's make sure we're bing consistent with our usage of $\Q$ vs. $\mathbb{H}$.} Here we must also consider the non-trivial
+        quaternions, $q' = f(q) : \Q \to \Q$. 
+        % \todo{Let's make sure we're bing consistent with our usage of $\Q$ vs. $\mathbb{H}$.} 
+        Here we must also consider the non-trivial
         effect of a differential rotation applied to the output, i.e.:
@@
 
-\todo{I would re-work this section to only talk about the details of doing it the ``correct'' way (i.e.
+% \todo{I would re-work this section to only talk about the details of doing it the ``correct'' way (i.e.
 consistent with the previous section).
@@
     Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's
-    attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will solve this
-    problem using a modified version of Newton's method that
-    exploits the true group structure of $SO(3)$ using the quaternion calculus presented
-    here.
+    attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will
+    solve this problem using a modified version of Newton's method that exploits the true
+    group structure of $SO(3)$ using the quaternion calculus presented here.
 
@@
 % \todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
-    Given known vectors in some world frame, ${}^W v_i$, and measurements of these
-    vectors in some body fixed frame, ${}^B v_i$, our goal is to determine the relative
-    rotation from the body to world frame ${}^W q^B$, expressed as a quaternion. We
+    Given known vectors in some world frame, $\inframe{W}{v_i}$, and measurements of these
+    vectors in some body fixed frame, $\inframe{B}{v_i}$, our goal is to determine the relative
+    rotation from the body to world frame $\toframe{W}{B}{A(q)}$, expressed as a quaternion. We
     can define Wahba's loss function as the following: 
-    \begin{align} \label{eq:wahba_loss}
-        L &= \sum_i \norm{{}^W v_i - {}^W A(q)^B \,\, ^Bv_i }_2^2. = \norm{r_i(q)}_2^2 
-    \end{align} 
+    \begin{equation} \label{eq:wahba_loss}
+        L = \sum_i \norm{\inframe{W}{v_i} - \toframe{W}{B}{A(q)} \,\, \inframe{B}{v_i} }_2^2 
+           = \norm{r_i(q)}_2^2 
+    \end{equation} 
     % \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
@@
 
-    We can solve for ${}^W q^B$ using a nonlinear least squares method minimizing
+    We can solve for $\toframe{W}{B}{q}$ using a nonlinear least squares method minimizing
     Wahba's loss function:
@@
         \begin{aligned}
-            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i{q} := \bar{J}^T r(q) = 0 \\
+            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i(q) := \bar{J}^T r(q) = 0 \\
               &= \sum_i (-2 H^T R(q)^T R({}^B \hat{v}_i)^T r(q).
@@
 
-    We can minimize this loss function with respect to a differential rotation $\phi$,
-    we do this by simply ``correcting'' our Jacobian $\bar{J}$ using 
+    Applying the methodology from the previous section, we actually want to minimize 
+    \eqref{eq:wahba_loss} with respect to a differential rotation $\phi$, which 
+    we do by simply ``correcting'' our Jacobian $\bar{J}$ using 
     \eqref{eq:quat_jacobian}:
@@
     We then use this to obtain our Newton step by using the Moore-Penrose psuedoinverse:
-    $\phi = ({J}^T {J})^{-1} {J}^T$. To obtain our next iterate, we ``add'' the
-    step using the correct notion of composition for the group: 
-    $\q_{k+1} = \q_k \otimes \varphi(\phi)$. This ``multiplicative'' Newton algorithm is 
-    summarized in Algorithm \ref{alg:mgn}.
+    $\phi = ({J}^T {J})^{-1} {J}^T$. To obtain our next iterate, we ``add'' the step
+    using the correct notion of composition for the group: $\q_{k+1} = \q_k \otimes
+    \varphi(\phi)$. This ``multiplicative'' Newton algorithm is summarized in Algorithm
+    \ref{alg:mgn}.
 
@@
     \subsection{Results}
-    We compared this modified Newton's method with a na\"ive Newton's method where the quaternion is simply re-normalized at every iteration. 
-    As illustrated in Figure \ref{fig:wahba_convergence}, it is clear that the naive original method make progress initially, but fails to
-    exhibit the quadratic convergence typical of a Newton method. By optimizing directly 
-    in the space of unit quaternions, the modified method is able to achieve the expected quadratic convergence.
-    It should also be clear from the previous section that the adaptations to the original
-    Newton method are simple and straightforward, highlighting both the effectiveness and
-    value of the proposed approach.
+    We compared this modified Newton's method with a na\"ive Newton's method where the
+    quaternion is simply re-normalized at every iteration. As illustrated in Figure
+    \ref{fig:wahba_convergence}, it is clear that the naive original method makes progress
+    initially, but fails to exhibit the quadratic convergence typical of a Newton method.
+    By optimizing directly in the space of unit quaternions, the modified method is able
+    to achieve the expected quadratic convergence. It should also be clear from the 
+    previous section that the adaptations to the original Newton method are simple and
+    straightforward, highlighting both the effectiveness and value of the proposed
+    approach.
     \begin{figure}
@@
     \begin{mini}[2]
-        {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
+        {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell_k(x_k,u_k) }{}{}
         \addConstraint{x_{k+1} = f(x_k,u_k)}
@@
     where $x$ and $u$ are the state and control vectors as described in Sec. \ref{sec:rigidbody_dynamics},
-    $f$ are the dynamics as defined in \eqref{eq:rigid_body_dynamics}, $\ell$ is a general
-    nonlinear cost function,
+    $f$ are the dynamics as defined in \eqref{eq:rigid_body_dynamics}, $\ell_k$ is a general
+    nonlinear cost function at a single time step,
     $N$ is the number of time steps, and $g_k$, $h_k$ are general nonlinear inequality and 
@@
     Like most gradient or Newton-based methods for optimization, ALTRO approximates the
-    nonlinear functions $f, \ell, g,$ and $h$ with their first or second-order Taylor series
-    expansions. Leveraging the methods from Sec: \ref{sec:Quaternion_Calculus}, we adapt the
-    algorithm to optimize directly on the error state $\delta x \in \R^{12}$.
+    nonlinear functions $f, \ell, g,$ and $h$ with their first or second-order Taylor
+    series expansions. Leveraging the methods from Sec: \ref{sec:Quaternion_Calculus}, we
+    adapt the algorithm to optimize directly on the error state $\delta x \in \R^{12}$.
 
@@
 
-    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to our nonlinear cost
-    functions $\ell$ and 
-    \eqref{eq:quat_jacobian} to the nonlinear constraint functions $g_k$ and $h_k$, we can calculate
-    the second-order expansion 
+    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to our nonlinear
+    cost functions $\ell$ and \eqref{eq:quat_jacobian} to the nonlinear constraint
+    functions $g_k$ and $h_k$, we can calculate the second-order expansion of the cost 
+    function:
     \begin{multline}
@@
             + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta u \\
-            + \ell_x^T \delta x^T + \ell_u^T \delta u 
+            + \ell_x^T \delta x^T + \ell_u^T \delta u.
     \end{multline}
-    of the augmented Lagrangian cost function: \todo{I think you can safely leave out any references to the augmented Lagrangian, etc. and just talk about the cost function.}
-    \begin{equation}
-        \mathcal{L}_A = \mathcal{L}_N(x_N,\lambda_N,\mu_N) + 
-            \sum_{k=0}^{N-1} \mathcal{L}_k(x_k,u_k,\lambda_k,\mu_k)
-    \end{equation}
-    where
-    \begin{equation}
-        \mathcal{L}_k(x,u,\lambda,\mu) = \ell(x,u) + 
-            (\lambda + \half I_\mu c(x,u))^T c(x,u),
-    \end{equation}
-    with $c(x,u)$ being the concatenation of the constraints $f,g$, and $h$ at a given time step,
-    and $I_\mu$ the penalty matrix.
+    % \todo{I think you can safely leave out any references to the augmented Lagrangian,
+    % etc. and just talk about the cost function.}
+    % \begin{equation}
+    %     \mathcal{L}_A = \mathcal{L}_N(x_N,\lambda_N,\mu_N) + 
+    %         \sum_{k=0}^{N-1} \mathcal{L}_k(x_k,u_k,\lambda_k,\mu_k)
+    % \end{equation}
+    % where
+    % \begin{equation}
+    %     \mathcal{L}_k(x,u,\lambda,\mu) = \ell(x,u) + 
+    %         (\lambda + \half I_\mu c(x,u))^T c(x,u),
+    % \end{equation}
+    % with $c(x,u)$ being the concatenation of the constraints $f,g$, and $h$ at a given time step,
+    % and $I_\mu$ the penalty matrix.
 
     With this expansion, we calculate the expansion of the ``action-value function'' $Q(x,u)$
-    as normal:
+    as normal, including the extra terms from the augmented Lagrangian cost:
     \begin{align}
@@
     \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
-        In addition to the straight-forward modifications to the ALTRO algorithm itself, we
-        need to carefully consider the types of cost functions we minimize. We frequently
-        minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
-        (x-x_g)$; However, na\"ive substraction of unit quaternions is ill-defined. We
-        propose two different cost functions that accomplish similar behavior.
-        For sake of clarity and space, we only consider the costs on the quaternion
-        variables: costs on the other states and the control variables remain unaffected.
+        In addition to the straight-forward modifications to the ALTRO algorithm itself,
+        we need to carefully consider the types of cost functions we minimize. We
+        frequently minimize costs that penalize distance from a goal state, e.g. $\half
+        (x-x_g)^T Q (x-x_g)$; However, na\"ive substraction of unit quaternions is
+        ill-defined. We propose the following cost function which penalizes the geodesic
+        distance between two unit quaternions \cite{Kuffner2004}, which we have found to
+        work well in practice. For sake of clarity and space, we only consider the costs
+        on the quaternion variables: costs on the other states and the control variables
+        remain unaffected.
 
-        \subsubsection{Error Quadratic} \label{sec:error_quadratic}
-            Rather than simple subtraction, we can use a quadratic function on the
-            three-parameter error state \eqref{eq:state_error}:
-            \begin{equation} \label{eq:error_quadratic}
-                J_\text{err} = \half \phi^T Q \phi 
-                = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
-                        \left(\varphi^{-1}(\delta q)\right).
-            \end{equation}
-            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
-            and Hessian of \eqref{eq:error_quadratic} are
-            \begin{equation}
-                \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
-            \end{equation}
-            \begin{multline}
-                \nabla^2 J_\text{err} = 
-                    G(\delta q)^T \! \left(
-                    D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
-                    + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
-            \end{multline}
-            where, for the Cayley map,
-            \begin{equation}
-                D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
-                    q_v \;\; & -\frac{1}{q_s} I_3
-                \end{bmatrix}
-            \end{equation}
-            \begin{equation}
-                \nabla D = \pdv{q}(D(q)^T Q \phi) 
-                = -\frac{1}{q_s^2} \begin{bmatrix} 
-                    -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
-                                   Q \phi & 0 \\
-                \end{bmatrix}.
-            \end{equation}
-
-        \subsubsection{Geodesic Distance} \label{sec:geodesic}
-            Alternatively, we can use the geodesic distance between two quaternions:
-            \cite{Kuffner2004},
         \begin{equation} \label{eq:quat_geodesic}
@@
 
-\todo{If the geodesic distance works better on all the examples, I vote for leaving out the error quadratic stuff.}
+        % \subsubsection{Error Quadratic} \label{sec:error_quadratic}
+        %     Rather than simple subtraction, we can use a quadratic function on the
+        %     three-parameter error state \eqref{eq:state_error}:
+        %     \begin{equation} \label{eq:error_quadratic}
+        %         J_\text{err} = \half \phi^T Q \phi 
+        %         = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
+        %                 \left(\varphi^{-1}(\delta q)\right).
+        %     \end{equation}
+        %     where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
+        %     and Hessian of \eqref{eq:error_quadratic} are
+        %     \begin{equation}
+        %         \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
+        %     \end{equation}
+        %     \begin{multline}
+        %         \nabla^2 J_\text{err} = 
+        %             G(\delta q)^T \! \left(
+        %             D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
+        %             + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
+        %     \end{multline}
+        %     where, for the Cayley map,
+        %     \begin{equation}
+        %         D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
+        %             q_v \;\; & -\frac{1}{q_s} I_3
+        %         \end{bmatrix}
+        %     \end{equation}
+        %     \begin{equation}
+        %         \nabla D = \pdv{q}(D(q)^T Q \phi) 
+        %         = -\frac{1}{q_s^2} \begin{bmatrix} 
+        %             -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
+        %                            Q \phi & 0 \\
+        %         \end{bmatrix}.
+        %     \end{equation}
+
+        % \subsubsection{Geodesic Distance} \label{sec:geodesic}
+
+% \todo{If the geodesic distance works better on all the examples, I vote for leaving out the error quadratic stuff.}
 
@@
     \begin{equation}
-         [{}^W r_{sun}]^T[{}^W A(q) {}^B\, {}^Br_{cam} ] \leq \cos(40 ^\circ).
+         \left(\inframe{W}{r_{sun}}\right)^T 
+         \left(\toframe{W}{B}{A(q)} \, \inframe{B}{r_{cam}} \right) \leq \cos(40 ^\circ).
     \end{equation}
@@
     \subsection{Airplane Barrel Roll}
-\todo{Reference \eqref{eq:rigid_body_dynamics} and then say that the forces and torques due to lift and drag are fit from wind tunnel data. Also, cite that paper.}
+% \todo{Reference \eqref{eq:rigid_body_dynamics} and then say that the forces and torques due to lift and drag are fit from wind tunnel data. Also, cite that paper.}
 
-        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
-        tasked to do a barrel roll by setting a high terminal cost for being upside-down,
-        see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
-        conditions. The convergence of the different versions of ALTRO is compared in Fig.
-        \ref{fig:c_max_convergence}. For both the quadrotor flip \todo{put the quadrotor comments in the quadrotor section.} and the airplane barrel 
-        roll, the modified version of ALTRO converged faster than original version.
-        For these highly aerobatic maneuvers, we achieve, as expected, better
-        performance by correctly leveraging the structure of the rotation group during
-        the optimization routine.
+        Our airplane model is a simple rigid body as defined in Section
+        \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag fit
+        from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do a
+        barrell roll by setting a high terminal cost for being upside-down, see Fig.
+        \ref{fig:barrellroll}. The solver is initialized with level flight trim
+        conditions. The convergence of the different versions of ALTRO is compared in
+        Fig. \ref{fig:c_max_convergence}. As expected, the modified version achieves
+        better convergence and faster solve times compared to the na\"ive version since
+        the expansions being provided to the algorithm more accurately capture the
+        relationship between the attitude state and the goal and constraints.
 
-		\todo{If we're not going to plot it and actually do an experimental comparison, I vote for leaving this stuff out.}
-        We also compared the ``error quadratic'' cost function described in Section
-        \ref{sec:error_quadratic} to the geodesic cost function in Section
-        \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of
-        iterations and computation time, taking only 36 iterations and 1.8 ms/iteration
-        vs 76 iterations and 7.6 ms/iteration for the error quadratic. Not only is the
-        geodesic cost function much cheaper to compute, it also converges much faster
-        than the error quadratic cost function. This could be due to the strong nonlinearities
-        introduced when computing the quaternion error, making it more difficult to
-        optimize.
+		% \todo{If we're not going to plot it and actually do an experimental comparison, I vote for leaving this stuff out.}
+        % We also compared the ``error quadratic'' cost function described in Section
+        % \ref{sec:error_quadratic} to the geodesic cost function in Section
+        % \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of
+        % iterations and computation time, taking only 36 iterations and 1.8 ms/iteration
+        % vs 76 iterations and 7.6 ms/iteration for the error quadratic. Not only is the
+        % geodesic cost function much cheaper to compute, it also converges much faster
+        % than the error quadratic cost function. This could be due to the strong nonlinearities
+        % introduced when computing the quaternion error, making it more difficult to
+        % optimize.
 
@@
     \subsection{Quadrotor Flip}
-    \todo{Cite the Penn quadrotor paper for the dynamics model.}
-        We optimized a 360 degree flip trajectory for a quadrotor using the 
-        modified version of ALTRO.
-	    Four intermediary ``keyframes'' were used to encourage the quadrotor to be at angles
-        of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular arc.
-        The quadrotor was constrained
-	    to stay above the floor and move to a goal state 2 meters away in the $+y$
-	    direction. The solver was initialized with a dynamically infeasible trajectory
-	    that linearly interpolates between the initial and final states, rotating the
-        quad around the x-axis a full \ang{360}.
+        We optimized a 360 degree flip trajectory for a quadrotor with dynamics adapated
+        from \cite{mellinger2012trajectory} using the modified version of ALTRO. Four
+        intermediary ``keyframes'' were used to encourage the quadrotor to be at angles
+        of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular
+        arc. The quadrotor was constrained to stay above the floor and move to a goal
+        state 2 meters away in the $+y$ direction. The solver was initialized with a
+        dynamically infeasible trajectory that linearly interpolates between the initial
+        and final states, rotating the quad around the x-axis a full \ang{360}.
 

### 2020-10-12 examples/trajopt_examples.jl

@@ -147,6 +147,6 @@ p = @pgf Axis(
         },
         Coordinates(1:length(c_maxes[2]),c_maxes[2])
     ),
-    Legend("original","modified")
+    Legend("naive","modified")
 )
 pgfsave("paper/figures/c_max_convergence.tikz", p, include_preamble=false)

### 2020-10-12 paper/main.tex
Add some more discussion on flexsat

@@
     undergo large changes in attitude: an airplane barrel roll, a quadrotor flip, and a
-    satellite with flexible solar panels that must slew to a new orientation while avoiding a 
-    keep-out zone. All results were run on a desktop computer with an 
-    AMD Ryzen 2950x processor with 40 GB of RAM. 
-    All problems are run using ALTRO, first without any of the modifications presented 
-    in the current paper, analagous to the naive Newton's method in section \ref{sec:Wahbas} and labeled ``naive'', and then using the modifications listed
-    in Sec. \ref{sec:trajopt} and the geodesic cost function described in Sec. \ref{sec:geodesic},
-    labeled ``modified''.
-    Timing results are summarized in Table \ref{tab:timing_results}.
-    All experiments were solved to a constraint satisfaction tolerance of $10^{-5}$ and 
-    discretized with a 4th order Runge-Kutta integrator.
-    Code for all experiments is available on 
+    satellite with flexible solar panels that must slew to a new orientation while
+    avoiding a keep-out zone. All results were run on a desktop computer with an AMD
+    Ryzen 2950x processor with 40 GB of RAM. All problems are run using ALTRO, first
+    without any of the modifications presented in the current paper, analagous to the
+    naive Newton's method in section \ref{sec:Wahbas} and labeled ``naive'', and then
+    using the modifications listed in Sec. \ref{sec:trajopt} and the geodesic cost
+    function described in Sec. \ref{sec:cost_functions}, labeled ``modified''. Timing
+    results are summarized in Table \ref{tab:timing_results}. All experiments were solved
+    to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 4th order
+    Runge-Kutta integrator. Code for all experiments is available on
     \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
@@
 Also, let's just call the start tracker a camera or sensor.}
-    A spacecraft with flexible appendages must perform a 150 degree slew while ensuring
-    that a body-mounted camera does not cross within 40 degrees of the sun. The
-    attitudes that result in the camera pointing too close to the sun will be
-    referred to as a keep-out zone.
-The flexible body spacecraft dynamics are based on equation \eqref{eq:rigid_body_dynamics}, with the addition of four reaction wheels, and the three largest flexible modes \cite{Tracy2020}.
-In order to ensure no damage to the camera, we must formulate a constraint that guarantees that the camera line-of-sight body-fixed vector ($^Br_{cam}$)
-    does not look within 40 degrees of the world-fixed sun vector ($^Wr_{sun}$). This angle constraint can be represented as a dot product between the two vectors expressed in the world frame:
+        A spacecraft with flexible appendages must perform a 150 degree slew while
+        ensuring that a body-mounted camera does not cross within 40 degrees of the sun.
+        The attitudes that result in the camera pointing too close to the sun will be
+        referred to as a keep-out zone. The flexible body spacecraft dynamics are based
+        on equation \eqref{eq:rigid_body_dynamics}, with the addition of four reaction
+        wheels, and the three largest flexible modes \cite{Tracy2020}. In order to ensure
+        no damage to the camera, we must formulate a constraint that guarantees that the
+        camera line-of-sight body-fixed vector ($^Br_{cam}$) does not look within 40
+        degrees of the world-fixed sun vector ($^Wr_{sun}$). This angle constraint can be
+        represented as a dot product between the two vectors expressed in the world
+        frame:
         % \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
@@
         \end{equation}
-    The attitudes that satisfy this constraint comprise a non-convex set, with the constraint itself being nonlinear in the attitude quaternion. 
+        The attitudes that satisfy this constraint comprise a non-convex set, with the
+        constraint itself being nonlinear in the attitude quaternion.
         \begin{figure}[H]
@@
         % \todo{Talk a little about the results in the plot here.}
-ALTRO is able to reason about this nonlinear constraint, and account for the group structure present in the quaternion.
-From Figure \ref{fig:keepout} we can see that without the camera constraint present, the trajectory will pass through the keep-out zone, making the trajectory infeasible.
-With the keep-out zone constraint present, ALTRO is able to converge on a trajectory that performs the slew in an efficient manner, while keeping the camera sufficiently far from the sun.
+        ALTRO is able to reason about this nonlinear constraint, and account for the group
+        structure present in the quaternion. From Figure \ref{fig:keepout} we can see that
+        without the camera constraint present, the trajectory will pass through the keep-out
+        zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
+        is able to converge on a trajectory that performs the slew in an efficient manner, while
+        keeping the camera sufficiently far from the sun.
+
+        The modifications didn't have a significant impact on runtime performance for this
+        example. In our experience, the computational benefits of the proposed methodology are
+        problem-dependent, especially for highly nonlinear systems. The following examples result
+        in more dynamic and aerobatic behaviors, where the computational benefit of the proposed
+        methodology is significant, including an example where the naive method failed to find a
+        good solution.
 
@@
 
-        \begin{figure}[h]
+        \begin{figure}[ht]
             \centering

### 2020-10-12 paper/main.tex
Update IROS abstract

@@
                             -q_v^T \\ 
-                            sI_3 + \skewmat{q_v}
+                            q_s I_3 + \skewmat{q_v}
                         \end{bmatrix}
@@
 % \todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
-    Given known vectors in some world frame, $\inframe{W}{v_i}$, and measurements of these
+    Given known vectors in some world frame, $\inframe{W}{w_i}$, and measurements of these
     vectors in some body fixed frame, $\inframe{B}{v_i}$, our goal is to determine the relative
@@
     \begin{equation} \label{eq:wahba_loss}
-        L = \sum_i \norm{\inframe{W}{v_i} - \toframe{W}{B}{A(q)} \,\, \inframe{B}{v_i} }_2^2 
+        L = \sum_i \norm{\inframe{W}{w_i} - \toframe{W}{B}{A(q)} \,\, \inframe{B}{v_i} }_2^2 
            = \norm{r_i(q)}_2^2 
@@
     collocation or sequential convex programming, rely on commercial, proprietary, or
-    general-purpose solvers whose internal numerics often are abstracted away from the user. 
-    By exploiting the unique structure of both the trajectory optimization problem
-    and the rotation group, ALTRO will likely be able to solve more challenging problems with performance.
+    general-purpose solvers whose internal numerics often are abstracted away from the
+    user. By exploiting the unique structure of both the trajectory optimization problem
+    and the rotation group, ALTRO will likely be able to solve more challenging problems
+    with performance.
 

### 2020-10-12 examples/terminal_quats.jl

@@ -15,54 +15,6 @@ using LinearAlgebra
 # using MeshCat
 using Rotations
 
-## Quaternion Goal Constraints
-struct QuatGeoCon{T} <: TO.StateConstraint
-    n::Int
-    qf::UnitQuaternion{T}
-    qind::SVector{4,Int}
-end
-function TO.evaluate(con::QuatGeoCon, x::SVector)
-    q = x[con.qind]
-    qf = Rotations.params(con.qf)
-    dq = qf'q
-    return SA[min(1-dq, 1+dq)]
-end
-TO.sense(::QuatGeoCon) = TO.Equality()
-TO.state_dim(con::QuatGeoCon) = con.n
-Base.length(::QuatGeoCon) = 1
-
-
-struct QuatErr{T} <: TO.StateConstraint
-    n::Int
-    qf::UnitQuaternion{T}
-    qind::SVector{4,Int}
-end
-function TO.evaluate(con::QuatErr, x::StaticVector)
-    qf = con.qf
-    q = UnitQuaternion(x[con.qind])
-    return qf ⊖ q
-end
-TO.sense(::QuatErr) = TO.Equality()
-TO.state_dim(con::QuatErr) = con.n
-Base.length(con::QuatErr) = 3
-
-struct QuatVecEq{T} <: TO.StateConstraint
-    n::Int
-    qf::UnitQuaternion{T}
-    qind::SVector{4,Int}
-end
-function TO.evaluate(con::QuatVecEq, x::StaticVector)
-    qf = Rotations.params(con.qf)
-    q = normalize(x[con.qind])
-    dq = qf'q
-    if dq < 0
-        qf *= -1
-    end
-    return SA[qf[1] - q[1], qf[2] - q[2], qf[3] - q[3], qf[4] - q[4]] 
-end
-TO.sense(::QuatVecEq) = TO.Equality()
-TO.state_dim(con::QuatVecEq) = con.n
-Base.length(con::QuatVecEq) = 4
 
 
 ## Test constraints

### 2020-10-13 examples/trajopt_examples.jl
Get quad videos

@@ -107,7 +107,7 @@ org = df[.!df.errstate,:]
 tab = vcat(map(1:3) do i
     [mod.prob[i] @sprintf("%i / %i", org.iters[i], mod.iters[i]) @sprintf("%.2f / %.2f", org.time[i], mod.time[i])]
 end...)
-latex_tabular("paper/figures/timing_results.tex",
+latex_tabular(joinpath(@__DIR__,"..","paper/figures/timing_results.tex"),
     Tabular("lll"),
     [
         Rule(:top),
@@ -149,4 +149,4 @@ p = @pgf Axis(
     ),
     Legend("naive","modified")
 )
-pgfsave("paper/figures/c_max_convergence.tikz", p, include_preamble=false)
+pgfsave(joinpath(@__DIR__, "..", "paper/figures/c_max_convergence.tikz"), p, include_preamble=false)

### 2020-10-13 src/quadflip_problem.jl

@@ -125,14 +125,15 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
     opts = SolverOptions(
         cost_tolerance=1e-5,
         cost_tolerance_intermediate=1e-5,
-        constraint_tolerance=1e-4,
+        constraint_tolerance=1e-5,
         projected_newton_tolerance=1e-2,
         iterations_outer=40,
         penalty_scaling=10.,
         penalty_initial=0.1,
         show_summary=false,
         verbose_pn=false,
-        verbose=0
+        verbose=0,
+        projected_newton=false
     )
     return prob, opts
 end

### 2020-10-13 paper/iros_abstract.tex
IROS submission
Update IROS abstract

@@
 
+% Planning and controlling trajectories for floating-base robotic systems that
+% experience large attitude changes is challenging due to the nontrivial group
+% structure of 3D rotations. This paper introduces a powerful, computationally
+% efficient, and accessible approach for adapting existing Newton and
+% gradient-based methods to correctly account for the group structure of
+% rotations. We demonstrate the effectiveness of the approach by modifying the
+% ALTRO solver to optimize a loop-de-loop trajectory for a quadrotor, enabling
+% the solver to find a dramatically better solution in less than a third the
+% time.
+
 \section{INTRODUCTION}
 
-    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous
+    Many robotic systems---including quadrotors, airplanes, satellites, 
     underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
@@
 
-    To address this lack of solver support, we formulate a straightforward, generic method for adapting
-    existing Newton and gradient-based algorithms to correctly account for the group structure of
-    rotations. Based entirely on basic mathematical tools from linear algebra and
-    calculus, our method is computationally efficient and should be both accessible and
-    familiar to most roboticists. We apply this method to the ALTRO solver
-    \cite{howell2019altro} and demonstrate its performance on several constrained trajectory
-    optimization problems.
+    To address this lack of solver support, we formulate a straightforward, generic
+    method for adapting existing Newton and gradient-based algorithms to correctly
+    account for the group structure of rotations. Based entirely on basic mathematical
+    tools from linear algebra and calculus, our method is computationally efficient and
+    should be both accessible and familiar to most roboticists. We apply this method to
+    the ALTRO solver \cite{howell2019altro} and demonstrate its performance on a 
+    constrained trajectory optimization problem.
 
@@
 
-    We represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as 
-    a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and 
-    $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively. We denote
-    the composition of two quaternions as $q_3 = q_2 \otimes q_1$.
+    We represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$
+    as a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and $q_v
+    \in \R^3$ are the scalar and vector part of the quaternion, respectively. We denote
+    the composition of two quaternions as $\q_3 = \q_2 \otimes \q_1$. We use
+    $\skewmat{x}$ to denote the skew-symmetric matrix such that $\skewmat{x} y = x \times
+    y$.
 
@@
         \centering
-        \includegraphics[width=\columnwidth]{figures/quadflip_new.png}
+        \includegraphics[width=\columnwidth]{figures/quadflip.png}
         \label{fig:quadflip}
-        \caption{Snapshots of quadrotor flip trajectory solved using a modified version of 
-        ALTRO that correctly accounts for the group structure of rotations. These modifications
-        make it easier to find trajectories that undergo large changes in attitude, such as
-        the loop-de-loop shown here.}
+        \caption{Snapshots of quadrotor flip trajectory solved using a modified version
+        of ALTRO that correctly accounts for the group structure of rotations. These
+        modifications make it easier to find trajectories that undergo large changes in
+        attitude, such as the loop-de-loop shown here.}
     \end{figure}
@@
     state vector. We present a simple but powerful method for taking derivatives of
-    quaternion-valued state vectors. The extension of this method to more general state vectors that contain Cartesian products of $SO(3)$, $SE(3)$, and $\mathbb{R}^n$ is straightforward.
+    quaternion-valued state vectors. The extension of this method to more general state
+    vectors that contain Cartesian products of $SO(3)$, $SE(3)$, and $\mathbb{R}^n$ is
+    straightforward.
 
@@
     idea should match our intuition given rotations are inherently three-dimensional and
-    differential rotations should live in the same space as angular velocity, i.e.
+    differential rotations should live in the same space as angular velocities, i.e.
     $\R^3$.
@@
 
-    These adjustments effectively modify the optimization so that it is performed on the
-    quaternion error state (or Rodrigues Parameters) instead of the space of unit
+    These adjustments effectively modify the optimization so that it is performed on 
+    differential rotations (or Rodrigues Parameters) instead of the space of unit
     quaternions. In practice, this is analagous to the Multiplicative Extended Kalman
@@
 
-    The algorithm only needs one small modification to the ``forward pass,'' where the
-    trajectory is modified by simulating the system forward using the optimal feedback
-    policy, $\delta u = K \delta x + d$, computed during the ``backward'' pass. Instead of 
-    computing $\delta x$ using simple subtraction, we now compute the error state using the
-    inverse Cayley map \eqref{eq:invcayley} for the quaternion states. The rest of the
-    forward pass is unmodified. 
-    %\todo{I think you can safely omit some of these details and just refer to the ALTRO paper.}
+    The iLQR algorithm derives a locally-optimal feedback policy of the form $\delta u =
+    K \delta x + d$ during the ``backward pass'' of each iteration, which is then used to
+    simulate the system forward during the ``forward pass''. Instead of computing $\delta
+    x$ using simple substraction as we would in the original algorithm, we now compute the
+    error using the inverse Cayley map \eqref{eq:invcayley} for the quaternion states. The
+    rest of the forward pass is unmodified.
 
@@
         x-axis while moving to the goal. 
-        The modified version of ALTRO solved the problem in 270 ms and 26 iterations, while
-        the original version that naively treated the quaternion as a vector in $\R^4$ 
-        solved in 870 ms and 91 iterations. The original version failed to provide a nice,
+        The modified version of ALTRO solved the problem in 140 ms and 22 iterations, while
+        the original version that na\"ively treated the quaternion as a vector in $\R^4$ 
+        solved in 420 ms and 58 iterations. The original version failed to provide a nice,
         smooth loop and instead reversed direction halfway through the loop.
@@
         and 360 degrees for Euler angles, Rodrigues parameters, and MRP's, respectively.
-        A classic approach to overcome this issue is to do discrete switching of the 
-        reference orientation, which requires previous knowledge of the topology of the
-        optimal solution. 
+        % A classic approach to overcome this issue is to do discrete switching of the 
+        % reference orientation, which requires previous knowledge of the topology of the
+        % optimal solution. 
         % \begin{figure}
@@
     state vectors. The approach allows for straightforward adaptation of many
-    gradient-based methods for optimal control and motion planning, which we demonstrated
-    on the ALTRO solver. 
-    Initial results suggest that our approach can significantly reduce solve times by 
-    improving convergence rates for problems that require large changes in attitude over
-    the course of the trajectory. Future work will extend these ideas to Newton-based methods
-    such as direct collocation and other domains outside of trajectory optimization.
+    gradient and Newton-based methods for optimal control and motion planning, which we demonstrated
+    on the ALTRO solver. By exploiting the unique structure of both the trajectory
+    optimization problem and the rotation group, the newly modified version of ALTRO will
+    likely be able to solve more challenging problems with performance.
+    Future work will extend these ideas to Newton-based methods such as direct
+    collocation and other domains outside of trajectory optimization.
 

### 2020-10-14 paper/main.tex
Adjust margins
more comments

@@
 \overrideIEEEmargins                                      % Needed to meet printer requirements.
+\usepackage[top=60pt, left=48pt, right=48pt, bottom=43pt]{geometry}
 
@@
 
-\author{Brian Jackson$^1$, Kevin Tracy$^1$, and Zachary Manchester$^1$%
+\author{Brian E. Jackson$^1$, Kevin Tracy$^1$, and Zachary Manchester$^1$%
     \thanks{
@@
 optimization-based planning on the space of rotations using only standard linear algebra and vector calculus. We demonstrate the effectiveness of the approach by adapting Newton's
-method to solve the canonical Wahba's problem, and modifying the trajectory optimization solver ALTRO to plan directly on the space of unit 
+method to solve the canonical Wahba's problem, and modify the trajectory optimization solver ALTRO to plan directly on the space of unit 
 quaternions, achieving superior convergence on problems involving significant changes 
@@
     methods rely heavily on principles from differential geometry and Lie group theory;
-    however, despite these works, many recent papers in the robotics community continue
+    However, despite these works, many recent papers in the robotics community continue
     to naively apply standard methods for motion planning and control with no regard for the
-    group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
-    Williams2017,Geisert2016}.
+    group structure of rigid body motion
+    %\cite{Alothman2016,deCrousaz2015, Williams2017,Geisert2016}
+    .
 
@@
         \item A unified approach to quaternion differential calculus entirely based on standard linear algebra and vector calculus.
-        \item Derivation of a Newton-based algorithm for nonlinear optimization directly on the space of unit quaternions using our notation.
-        \item a fast and efficient solver for trajectory optimization problems with 
+        \item Derivation of a Newton-based algorithm for nonlinear optimization directly on the space of unit quaternions.
+        \item Implementation of a fast and efficient solver for trajectory optimization problems with 
         attitude dynamics and nonlinear constraints that correctly accounts for the group
@@
         \end{equation}
-        The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to
-        I_q$, where $I_q$ is the quaternion identity. Once again, \eqref{eq:quat_jacobian} holds (up to a constant) for any three-parameter attitude representation.
+        %The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to
+        %I_q$, where $I_q$ is the quaternion identity.
+        Once again, \eqref{eq:quat_jacobian} holds (up to a constant) for any three-parameter attitude representation.
 
@@
     approximate its roots. For unconstrained systems, this method is highly effective,
-    and can exhibit quadratic convergence. For constrained systems, the updated parameter
+    and can exhibit quadratic convergence rates. For constrained systems, the updates 
     can be projected back onto the feasible set at each iteration, but without the same
-    convergence guarantees. For the constraints on $SO(3)$, Newton's method struggles to
-    converge past a certain threshold due to this projection. By leveraging the
-    quaternion calculus introduced, Newton's method can be modified to implicitly account
-    for these constraints. To demonstrate this, we will examine Wahba's Problem. In 1965,
-    Grace Wahba proposed the criterion for a least squares estimate of a spacecraft's
-    attitude from vector measurements \cite{Wahba1965, markley2014fundamentals}. We will
-    solve this problem using a modified version of Newton's method that exploits the true
-    group structure of $SO(3)$ using the quaternion calculus presented here.
+    convergence guarantees.
+    %For the constraints on $SO(3)$, Newton's method struggles to
+    %converge past a certain threshold due to this projection.
+
+    In this section, we will leverage the quaternion calculus results introduced in the previous section to modify Newton's method so that it implicitly accounts for the quaternion unit-norm constraint.
+Unlike the projection approach, this modified form of Newton's method retains the fast convergence rates associated with the unconstrained method.
+We will demonstrate this behavior on Wahba's Problem, a least-squares attitude estimation problem~\cite{Wahba1965, markley2014fundamentals}.
 
@@
 % \todo{Let's keep our notation and language consistent for body vs. inertial frames. I propose ``body'' (B) and ``inertial'' or ``world'' (W), which is the standard in robotics.}
-    Given known vectors in some world frame, $\inframe{W}{w_i}$, and measurements of these
-    vectors in some body fixed frame, $\inframe{B}{v_i}$, our goal is to determine the relative
-    rotation from the body to world frame $\toframe{W}{B}{A(q)}$, expressed as a quaternion. We
-    can define Wahba's loss function as the following: 
+    Given a set of known vectors in the world frame, $\inframe{W}{w_i}$, and measurements of these vectors in the body frame, $\inframe{B}{v_i}$, we seek the
+    rotation from the body to the world frame $\toframe{W}{B}{A(q)}$ that solves the following optimization problem, \todo{Kevin: Don't use $L$ here because it conflicts with \eqref{eq:Lmult}:}
+    \begin{mini*}
+        {q}{ L(q) }{}{}
+        \addConstraint { q\in \Q,}
+    \end{mini*}
+    where Wahba's loss function $L(q)$ is,
     \begin{equation} \label{eq:wahba_loss}
-        L = \sum_i \norm{\inframe{W}{w_i} - \toframe{W}{B}{A(q)} \,\, \inframe{B}{v_i} }_2^2 
-           = \norm{r_i(q)}_2^2 
+        L(q) = \sum_i \norm{\inframe{W}{w_i} - A(q) \,\, \inframe{B}{v_i} }_2^2 
+           = \norm{r(q)}_2^2 ,
     \end{equation} 
-    % \todo{Let's leave the $w_i$ off for clarity since you're not using it later.}
-    where $r_i(q)$ is the residual vector.
+    and $r(q)$ is the residual vector.
 
-    We can solve for $\toframe{W}{B}{q}$ using a nonlinear least squares method minimizing
-    Wahba's loss function:
-    \begin{mini*}
-        {q}{ \norm{r(q)}_2^2 }{}{}
-        \addConstraint { q\in \Q.}
-    \end{mini*}
-    % \todo{This should be $q \in \Q$ here, since unit quaternions are not isomorphic to $SO(3)$ (rotation matrices).}
-    Following the typical approach for Newton's method, we minimize \eqref{eq:wahba_loss}
-    by setting the gradient to zero:
+    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_jacobian} and \eqref{eq:quaternion_rotation}: \todo{Kevin: fix the second line of this equation:}
     \begin{equation} \label{eq:newton_foc}
         \begin{aligned}
-            \pdv{L}{q}^T &= \sum_i \pdv{r(q)}{q}^T r_i(q) := \bar{J}^T r(q) = 0 \\
-              &= \sum_i (-2 H^T R(q)^T R({}^B \hat{v}_i)^T r(q).
+            \nabla r(q) &= \pdv{r}{q} G(q) \\
+              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) \right).
         \end{aligned}
     \end{equation}
-    which can be obtained from the chain rule and \eqref{eq:quaternion_rotation}.
-
-    % Treating $q$ as a vector in $\R^4$, we obtain a solution to \eqref{eq:newton_foc}
-    % using the Moore-Penrose pseudoinverse, $\delta q = (J^T J)^{-1} J^T$, and our
-    % next candidate quaternion via simple addition, $q_{k+1} = q_k + \delta q$. Since
-    % $q_{k+1}$ will no longer be unit norm, we project it back on the unit sphere via
-    % the projection operator $\Pi(q) = q / \norm{q}$. This ``projected'' Newton approach
-    % is summarized in Algorithm \ref{alg:pgn}.
-
-    % \begin{algorithm} 
-    % 	\begin{algorithmic}[1]
-    % 		\caption{Projected Gauss-Newton Method}\label{alg:pgn}
-    % 		\State $k = 0$
-    % 		\While{significant progress}
-    % 		    \State $J = \pdv{r(q_k)}{ \partial q}$ 
-    % 		    \State $ \delta q = -(J^TJ)^{-1}J^T r(q_k)$ 
-    % 		  %  \State $q_{n+1} = \text{normalize}(q_n + q_{step})$ 
-    % 		  \State $q_{k+1} = \Pi_{SO(3)}(q_k + \delta q)$ 
-    % 		    \State $k = k + 1$
-    % 		\EndWhile
-    % 	\end{algorithmic}
-    % \end{algorithm}
-
-    Applying the methodology from the previous section, we actually want to minimize 
-    \eqref{eq:wahba_loss} with respect to a differential rotation $\phi$, which 
-    we do by simply ``correcting'' our Jacobian $\bar{J}$ using 
-    \eqref{eq:quat_jacobian}:
+    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
     \begin{equation}
-        {J} = \pdv{r(\mathbf{q} \otimes \varphi(\phi))}{\phi} = \pdv{r(q)}{q} G(q).
+    	\phi_k = ({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
     \end{equation}
-    We then use this to obtain our Newton step by using the Moore-Penrose psuedoinverse:
-    $\phi = ({J}^T {J})^{-1} {J}^T$. To obtain our next iterate, we ``add'' the step
-    using the correct notion of composition for the group: $\q_{k+1} = \q_k \otimes
-    \varphi(\phi)$. This ``multiplicative'' Newton algorithm is summarized in Algorithm
-    \ref{alg:mgn}.
+	The update is then applied using the composition for the group:
+	\begin{equation}
+		\q_{k+1} = \q_k \otimes \varphi(\phi_k)
+	\end{equation}.
+	This ``multiplicative'' Gauss-Newton method is summarized in Algorithm \ref{alg:mgn}.
 
@@
     		\State $k = 0$
-    		\While{significant progress}
-    		    \State ${J} = \pdv{r(q_k)}{q} G(q_k)$ \Comment{quaternion adjusted Jacobian}
-    		    \State $ \phi = -({J}^T {J})^{-1} {J}^T r(q_k)$ 
-    		    \State $\q_{k+1} = \q_k \otimes \varphi(\phi)$ \Comment{apply step multiplicatively}
+    		\While{$\norm{\phi} > $ tolerance}
+    		    \State ${\nabla r} = \pdv{r(q_k)}{q} G(q_k)$ 
+    		    	\Comment{Compute Jacobian}
+    		    \State $ \phi = -({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k)$ \Comment{Compute update step}
+    		    \State $q_{k+1} = L(q_k) \varphi(\phi)$ \Comment{Apply update step}
     		    \State $k = k + 1$
@@
     \subsection{Results}
-    We compared this modified Newton's method with a na\"ive Newton's method where the
-    quaternion is simply re-normalized at every iteration. As illustrated in Figure
-    \ref{fig:wahba_convergence}, it is clear that the naive original method makes progress
-    initially, but fails to exhibit the quadratic convergence typical of a Newton method.
-    By optimizing directly in the space of unit quaternions, the modified method is able
-    to achieve the expected quadratic convergence. It should also be clear from the 
-    previous section that the adaptations to the original Newton method are simple and
-    straightforward, highlighting both the effectiveness and value of the proposed
-    approach.
+    Figure \ref{fig:wahba_convergence} compares the multiplicative Gauss-Newton method with a na\"ive Newton's method in which the quaternion is simply projected back onto the unit sphere at every iteration.
+That the na\"ive method makes progress
+    initially, but quickly stalls.
+    By correctly handling the group structure of unit quaternions, the multiplicative method is able
+    to maintain the fast convergence rates typical of Newton's method.
     \begin{figure}
@@
         \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
-        \caption{Convergence comparison for Wahba's problem. By performing Newton's method
-        on the error quaternion and applying the result to the full quaternion we achieve
-        quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
-        error. The angle error is calculated relative to the true analytical solution obtained
-        via an SVD decomposition.}
+        \caption{Convergence comparison for Wahba's problem.
+By modfying Newton's method with the methods of section \ref{sec:Quaternion_Calculus},        quadratic convergence rates are achieved, while a na\"ive approach stalls after only a few itterations.}
         \label{fig:wahba_convergence}
@@
 \section{Trajectory Optimization on $\R^n \times SO(3)$} \label{sec:trajopt}
-    Here we outline the modifications to the ALTRO solver \cite{howell2019altro}, to
+    Here we outline the modifications to the ALTRO solver \cite{howell2019altro} to
     solve trajectory optimization problems for rigid bodies, which extends easily to
@@
     series expansions. Leveraging the methods from Sec: \ref{sec:Quaternion_Calculus}, we
-    adapt the algorithm to optimize directly on the error state $\delta x \in \R^{12}$.
+    adapt the algorithm to optimize directly on the error state $\delta x \in \R^{12}$:
+    \begin{equation} \label{eq:state_error}
+    	\delta x_k = \begin{bmatrix} 
+            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
+        \end{bmatrix} .
+    \end{equation}
 
-    We begin by linearizing the dynamics about the reference trajectory using
-    \eqref{eq:quat_jacobian}. Our linearized error dynamics become
+    We begin by linearizing the dynamics about the reference state and input trajectories, $\bar{x}$ and $\bar{u}$, using
+    \eqref{eq:quat_jacobian}. The linearized error dynamics become,
     \begin{equation} \label{eq:linearized_dynamics}
-        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k 
+        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k ,
     \end{equation}
@@
     \end{equation}
-    and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state-error Jacobian, respectively:
-    \begin{equation} \label{eq:state_error}
-        \setlength\arraycolsep{1pt}
-        \delta x_k = \begin{bmatrix} 
-            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
-        \end{bmatrix}\!, \;
+    and $E(x_k) \in \R^{12 \times 13}$ is the error-state Jacobian:
+    \begin{equation} \label{eq:state_error_jacobian}
         E(x) = \begin{bmatrix}
@@
     \end{equation}
-
-    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to our nonlinear
+    By applying \eqref{eq:quat_gradient} and \eqref{eq:quat_hessian} to the nonlinear
     cost functions $\ell$ and \eqref{eq:quat_jacobian} to the nonlinear constraint
@@
 
-    With this expansion, we calculate the expansion of the ``action-value function'' $Q(x,u)$
-    as normal, including the extra terms from the augmented Lagrangian cost:
+	With these results, we can apply standard Newton and quasi-Newton techniques along the lines of Section \ref{sec:Wahbas}.
+We can also calculate a second-order expansion of the ``action-value function'' $Q(x,u)$ needed in differential dynamic programming (DDP) and LQR-based methods,
     \begin{align}
@@
     from which we can calculate the quadratic expansion of the cost-to-go 
-    $P_k \in \R^{12 \times 12}$, $p_k \in \R^{12}$, and optimal linearized feedback gains 
-    $K_k \in \R^{m \times 12}$ and $d_k \in \R^m$ by starting at the terminal state and
-    resursing backward in time along the trajectory during the ``backward pass'' of the iLQR
-    algorithm. During the ``forward pass'', the dynamics are simulated forward in time using
-    the feedback gains computed during the backward pass. At each time step, the control is 
-    calculated using the linear feedback controller: 
+    $P_k \in \R^{12 \times 12}$, $p_k \in \R^{12}$, and optimal linear feedback gains 
+    $K_k \in \R^{m \times 12}$ and feed-forward corrections $d_k \in \R^m$ by starting at the terminal state and
+    performing a backward Riccati recursion as usual~\cite{li2004iterative,howell2019altro}.
+
+    During the ``forward rollout'' of these methods, the dynamics are simulated forward in time using updated control inputs: 
     \begin{equation} \label{eq:mlqr_control}
-        u_k = K_k \delta x_k + \bar{u}_k.
+        u_k = \bar{u}_k - d_k - K_k \delta x_k.
     \end{equation}
@@
     computed using \eqref{eq:state_error}, with $x_k$ being the current state estimate
-    and $\bar{x}_k$ the state from the previous iteration. The rest of the algorithm is
-    left unchanged. For more details on the ALTRO algorithm, the reader is encouraged to
-    refer to the original paper \cite{howell2019altro}.
+    and $\bar{x}_k$ the state from the previous iteration.
+For more details on the ALTRO algorithm, which leverages ideas from both DDP and direct Newton methods, we refer the reader to~\cite{howell2019altro}.
 
     \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
-        In addition to the straight-forward modifications to the ALTRO algorithm itself,
-        we need to carefully consider the types of cost functions we minimize. We
+        In addition to the straight-forward modifications to the ALTRO algorithm itself, some care must be taken in designing cost functions that are well-suited to unit quaternions. We
         frequently minimize costs that penalize distance from a goal state, e.g. $\half
-        (x-x_g)^T Q (x-x_g)$; However, na\"ive substraction of unit quaternions is
-        ill-defined. We propose the following cost function which penalizes the geodesic
-        distance between two unit quaternions \cite{Kuffner2004}, which we have found to
-        work well in practice. For sake of clarity and space, we only consider the costs
-        on the quaternion variables: costs on the other states and the control variables
-        remain unaffected.
-
+        (x-x_g)^T Q (x-x_g)$; However, na\"ive substraction of unit quaternions does not respect their group structure, and often results in undesired behavior.
+Instead, we have found the following cost function, which penalizes the geodesic
+        distance between two unit quaternions \cite{Kuffner2004}, to work well in practice:
         \begin{equation} \label{eq:quat_geodesic}
-            J_\text{geo} = (1-\abs{q_g^T q}) ,
+            J_\text{geo} = (1-\abs{q_g^T q}) .
         \end{equation}
-        whose gradient and Hessian are,
+        Its gradient and Hessian are,
         \begin{align}
@@
 
-    \subsection{Satellite Attitude Keep-Out}
-        % \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes.
-Also, let's just call the start tracker a camera or sensor.}
-        A spacecraft with flexible appendages must perform a 150 degree slew while
-        ensuring that a body-mounted camera does not cross within 40 degrees of the sun.
-        The attitudes that result in the camera pointing too close to the sun will be
-        referred to as a keep-out zone. The flexible body spacecraft dynamics are based
-        on equation \eqref{eq:rigid_body_dynamics}, with the addition of four reaction
-        wheels, and the three largest flexible modes \cite{Tracy2020}. In order to ensure
-        no damage to the camera, we must formulate a constraint that guarantees that the
-        camera line-of-sight body-fixed vector ($^Br_{cam}$) does not look within 40
-        degrees of the world-fixed sun vector ($^Wr_{sun}$). This angle constraint can be
-        represented as a dot product between the two vectors expressed in the world
-        frame:
-        % \todo{Let's use standard matrix notation here rather than the angle brackets, and let's also write the rotation correctly using $A(q)$ like in the other sections.}
-        \begin{equation}
-            \left(\inframe{W}{r_{sun}}\right)^T 
-            \left(\toframe{W}{B}{A(q)} \, \inframe{B}{r_{cam}} \right) \leq \cos(40 ^\circ).
-        \end{equation}
-        The attitudes that satisfy this constraint comprise a non-convex set, with the
-        constraint itself being nonlinear in the attitude quaternion.
-        \begin{figure}[H]
-            \centering
-            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
-            \caption{Visualization of the flexible spacecraft slew with a keep-out zone.
-Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions.
-The constraint surface represents attitudes where the camera line-of-sight is within 40$^\circ$ of the sun.
-The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
-            \label{fig:keepout}
-        \end{figure}
-        % \todo{Talk a little about the results in the plot here.}
-        ALTRO is able to reason about this nonlinear constraint, and account for the group
-        structure present in the quaternion. From Figure \ref{fig:keepout} we can see that
-        without the camera constraint present, the trajectory will pass through the keep-out
-        zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
-        is able to converge on a trajectory that performs the slew in an efficient manner, while
-        keeping the camera sufficiently far from the sun.
-
-        The modifications didn't have a significant impact on runtime performance for this
-        example. In our experience, the computational benefits of the proposed methodology are
-        problem-dependent, especially for highly nonlinear systems. The following examples result
-        in more dynamic and aerobatic behaviors, where the computational benefit of the proposed
-        methodology is significant, including an example where the naive method failed to find a
-        good solution.
-
     \subsection{Airplane Barrel Roll}
@@
 
-        Our airplane model is a simple rigid body as defined in Section
+        A 180 degree barrel roll trajectory for a fixed-wing airplane was optimized. The airplane's dynamics model consists of the a simple rigid body as defined in Section
         \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag fit
         from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do a
-        barrell roll by setting a high terminal cost for being upside-down, see Fig.
-        \ref{fig:barrellroll}. The solver is initialized with level flight trim
+        barrel roll by setting a high terminal cost for deviations from an upside-down attitude (see Fig.
+        \ref{fig:barrellroll}). The solver was initialized with level flight trim
         conditions. The convergence of the different versions of ALTRO is compared in
@@
         relationship between the attitude state and the goal and constraints.
+\todo{Brian: you mention constraints in the text and have it on the y-axis in Fig. 4, but you don't state what these constraints are in the text.}
+
+\todo{Brian: comment on the performance numbers in the text.}
 
@@
             \includegraphics[width=\columnwidth]{figures/barrellroll.png}
-            \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
+            \caption{Barrel roll trajectory computed by ALTRO using a terminal cost to encourage an upside-down attitude.}
             \label{fig:barrellroll}
@@
             \includegraphics[height=4cm,width=\columnwidth]{figures/c_max_convergence.tikz}
-            \caption{Constraint convergence when solving the barrel roll problem. Compares 
-            the convergence of the original version of ALTRO versus the new, modified version
-            that optimizing on the error state.}
+            \caption{Constraint satisfaction as a function of iteration when solving the barrel roll problem using ALTRO both with and without the modifications for optimizing unit quaternions.}
             \label{fig:c_max_convergence}
@@
     \subsection{Quadrotor Flip}
-        We optimized a 360 degree flip trajectory for a quadrotor with dynamics adapated
-        from \cite{mellinger2012trajectory} using the modified version of ALTRO. Four
-        intermediary ``keyframes'' were used to encourage the quadrotor to be at angles
+        A 360 degree flip trajectory for a quadrotor was optimized with dynamics adapted
+        from \cite{mellinger2012trajectory}. Four
+        intermediary ``keyframes'' were used to encourage the quadrotor to reacj angles
         of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular
@@
         dynamically infeasible trajectory that linearly interpolates between the initial
-        and final states, rotating the quad around the x-axis a full \ang{360}.
+        and final states, rotating the quadrotor around the x-axis a full \ang{360}.
+
+\todo{Brian: little more detail on cost weights.
+Any control cost?
+If we used the same cost function form for all of these (i.e.
+geodesic on q + quadratic on everything else), let's write it at the beginning of the experiment section.}
 
@@
 
+	    \todo{Brian: you say the original version didn't work at all, but there are timing results for it in the table???}
+
         \begin{figure}[t]
@@
 
+    \subsection{Satellite Attitude Keep-Out}
+        % \todo{Let's add a little more detail here: How about saying the dynamics are \eqref{eq:rigid_body_dynamics} plus a few extra states to capture the bending modes.
+Also, let's just call the start tracker a camera or sensor.}
+        A spacecraft with flexible appendages was tasked to perform a 150 degree slew maneuver while
+        ensuring that a body-mounted camera did not point within 40 degrees ``keep-out zone'' around the sun vector. The spacecraft dynamics are presented in detail in~\cite{Tracy2020}, and are based
+        on equation \eqref{eq:rigid_body_dynamics} with the addition of six states to account for three flexible modes. Control torques are generated by four reaction wheels.
+
+        \todo{Kevin: add some info on cost function, initial guess (if any), etc.}
+
+        We enforce the camera keep-out zone can with the following constraint,
+        \begin{equation}
+            \left(\inframe{W}{r_{sun}}\right)^T 
+            \left(\toframe{W}{B}{A(q)} \, \inframe{B}{r_{cam}} \right) \leq \cos(40 ^\circ) ,
+        \end{equation}
+        where $^Br_{cam}$ is the camera line-of-sight unit vector in the body frame and $^Wr_{sun}$ is the unit vector pointing to the sun in the world frame.
+The attitudes that satisfy this constraint comprise a non-convex set, with the
+        constraint itself being nonlinear in $q$.
+        \begin{figure}[H]
+            \centering
+            \includegraphics{figures/kevins_plots/attitude_slew_plot.tikz}
+            \caption{Visualization of the flexible spacecraft slew with a keep-out zone.
+Attitude is parameterized with a Rodrigues parameter to visualize the trajectory in three dimensions.
+The constraint surface represents attitudes where the camera line-of-sight is within 40$^\circ$ of the sun.
+The unconstrained solution violates this constraint, while the constrained solution is able to avoid the keep out zone.}
+            \label{fig:keepout}
+        \end{figure}
+        % \todo{Talk a little about the results in the plot here.}
+
+        \todo{Kevin: looks like both versions of ALTRO worked fine on this problem, so we should probably not say all this group stuff.
+Maybe just comment on the timing results and try to offer an explanation.}
+
+        ALTRO is able to reason about this nonlinear constraint, and account for the group
+        structure present in the quaternion. From Figure \ref{fig:keepout} we can see that
+        without the camera constraint present, the trajectory will pass through the keep-out
+        zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
+        is able to converge on a trajectory that performs the slew in an efficient manner, while
+        keeping the camera sufficiently far from the sun.
+
+        The modifications didn't have a significant impact on runtime performance for this
+        example. In our experience, the computational benefits of the proposed methodology are
+        problem-dependent, especially for highly nonlinear systems. The following examples result
+        in more dynamic and aerobatic behaviors, where the computational benefit of the proposed
+        methodology is significant, including an example where the naive method failed to find a
+        good solution.
+
 \section{Conclusions} \label{sec:conclusion}
@@
 
+    \todo{This is a really nice pitch. Can we say something like this in the intro?}
     With the modifications presented, ALTRO can solve problems few other methods for
@@
 
-\paragraph*{Acknowledgements}
+\section*{Acknowledgements}
 This material is based upon work supported by the National Science Foundation Graduate

### 2020-10-14 examples/quadrotor_flip.jl
Tune examples and use projected newton

@@ -13,6 +13,7 @@ using StaticArrays
 using LinearAlgebra
 using MeshCat
 using Rotations
+using Colors
 
 ## Try ALTRO
 # prob = gen_quad_flip(UnitQuaternion, slack=false, vecmodel=false, renorm=true,
@@ -21,13 +22,14 @@ using Rotations
 prob,opts = QuadFlipProblem(UnitQuaternion, slack=false, vecmodel=false, renorm=true,
     costfun=QuatLQRCost
 )
-solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true)
+solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true, 
+    projected_newton=true)
 solve!(solver)
 
 ## Original Method
-prob2 = gen_quad_flip(UnitQuaternion, slack=false, vecmodel=true)
 prob2,opts = QuadFlipProblem(UnitQuaternion, slack=false, vecmodel=true)
-solver2 = ALTROSolver(prob2, opts, R_inf=1e-4, infeasible=true, show_summary=true)
+solver2 = ALTROSolver(prob2, opts, R_inf=1e-4, infeasible=true, show_summary=true,
+    projected_newton=true)
 solve!(solver2)
 
 ## Visualize
@@ -41,10 +43,6 @@ quad = Quadrotor()
 TrajOptPlots.set_mesh!(vis, quad)
 visualize!(vis, quad, get_trajectory(solver))
 visualize!(vis, quad, get_trajectory(solver2))
-
-## Plot y-z trajectory
-using Plots
-RobotDynamics.traj2(states(solver), xind=2, yind=3)
 
 ## Waypoints
 delete!(vis)

### 2020-10-14 examples/trajopt_examples.jl

@@ -44,13 +44,14 @@ Barrell Roll
 args = (
     integration = RK3,
     termcon = :quatvec,
-    projected_newton = false,
+    # projected_newton = false,
     show_summary=true
 )
 
 ## Original Method
 prob,opts = YakProblems(vecstate=true, costfun=:Quadratic; args...) 
 solver = ALTROSolver(prob, opts)
+solve!(solver)
 add_result!(:barrellroll, false, solver)
 
 ## Modified Method
@@ -65,12 +66,12 @@ Quadrotor Flip
 """
 ## Original Method
 prob,opts = QuadFlipProblem(vecmodel=true)
-solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true)
+solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true)
 add_result!(:quadflip, false, solver)
 
 ## Modified Method
 prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost)
-solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true)
+solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true)
 add_result!(:quadflip, true, solver)
 
 @save results_path results
@@ -107,6 +108,9 @@ org = df[.!df.errstate,:]
 tab = vcat(map(1:3) do i
     [mod.prob[i] @sprintf("%i / %i", org.iters[i], mod.iters[i]) @sprintf("%.2f / %.2f", org.time[i], mod.time[i])]
 end...)
+
+tab[2,2] = "— / $(mod.iters[2])"  # remove unsucessful quadrotor results
+tab[2,3] = "— / $(mod.time[2])"
 latex_tabular(joinpath(@__DIR__,"..","paper/figures/timing_results.tex"),
     Tabular("lll"),
     [
@@ -130,6 +134,7 @@ p = @pgf Axis(
         "ymode=log",
         xmajorgrids,
         ymajorgrids,
+        "legend style={at={(0.1,0.1)},anchor=south west}"
     },
     Plot(
         {

### 2020-10-14 src/flexible_spacecraft_dynamics.jl

@@ -180,11 +180,11 @@ function SatelliteKeepOutProblem(constrained::Bool=true; vecstate::Bool=false, c
 	xf = SVector{17}([ω; qf;zeros(3);zeros(3);r0])
 
 	# Objective
-	Q = Diagonal(@SVector [10,10,10,100,100,100,100,10,10.0,10,10,10,10,0,0,0,0])
+	Q = Diagonal(@SVector [10,10,10, 1,1,1,1, 10,10.0,10, 10,10,10,1,1,1,1.])
 	R = Diagonal(@SVector fill(10.0, m))
 	Qf = Q * N
-	cost0 = costfun(Q, R, xf, w=10.)
-	cost_term = costfun(Qf, R, xf, w=100.)
+	cost0 = costfun(Q, R, xf, w=0.1)
+	cost_term = costfun(Qf, R, xf, w=1.0)
 	obj = Objective(cost0, cost_term, N)
 	# obj = LQRObjective(Q,R,Qf,xf,N)
 
@@ -216,7 +216,7 @@ function SatelliteKeepOutProblem(constrained::Bool=true; vecstate::Bool=false, c
 	opts = SolverOptions(
 		verbose=1,
 		constraint_tolerance = 1e-5,
-		penalty_scaling = 10.0,
+		penalty_scaling = 200.0,
 		penalty_initial = 1e3,
 		cost_tolerance = 1e-4,
 		cost_tolerance_intermediate = 1e-4,

### 2020-10-14 src/quadflip_problem.jl

@@ -133,7 +133,7 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
         show_summary=false,
         verbose_pn=false,
         verbose=0,
-        projected_newton=false
+        projected_newton=true
     )
     return prob, opts
 end

### 2020-10-14 src/quat_cons.jl

@@ -41,10 +41,10 @@ function TO.evaluate(con::QuatVecEq, x::StaticVector)
     if dq < 0
         qf *= -1
     end
-    return SA[qf[1] - q[1], qf[2] - q[2], qf[3] - q[3], qf[4] - q[4]] 
+    return -SA[qf[2] - q[2], qf[3] - q[3], qf[4] - q[4]] 
 end
 TO.sense(::QuatVecEq) = TO.Equality()
 TO.state_dim(con::QuatVecEq) = con.n
-Base.length(con::QuatVecEq) = 4
+Base.length(con::QuatVecEq) = 3
 
 

### 2020-10-14 paper/main.tex
Address Zac's comments

@@
     methods rely heavily on principles from differential geometry and Lie group theory;
-    However, despite these works, many recent papers in the robotics community continue
+    however, despite these works, many recent papers in the robotics community continue
     to naively apply standard methods for motion planning and control with no regard for the
-    group structure of rigid body motion
+    group structure of rigid body motion.
     %\cite{Alothman2016,deCrousaz2015, Williams2017,Geisert2016}
-    .
 
     In this paper, we make a departure from previous approaches to geometric planning and
-    control that rely heavily on ideas and notation from differential geometry, 
-    and instead use only basic mathematical tools from linear algebra and vector calculus that 
-    should be familiar to most roboticists. In Sec. \ref{sec:Quaternion_Calculus} we introduce 
-    an approach to quaternion differential calculus similar to \cite{Mandic2011,Xu2016}, but significantly simpler and more general, enabling straight-forward adaptation of 
-    existing algorithms to systems with quaternion states. 
-    For concreteness, we then apply our method to the canonical
-    Wahba's problem~\cite{Wahba1965} in Sec. \ref{sec:Wahbas}, and demonstrate superior convergence to approaches that
-    fail to properly account for the group structure. 
-    In Sec. \ref{sec:trajopt} we extend these ideas to the problem of trajectory optimization,
-    and detail modifications to ALTRO, a state-of-the-art constrained trajectory optimization
-    solver, and demonstrate performance gains on several benchmark problems.
+    control that rely heavily on ideas and notation from differential geometry, and
+    instead use only basic mathematical tools from linear algebra and vector calculus
+    that should be familiar to most roboticists. In Sec. \ref{sec:Quaternion_Calculus} we
+    introduce an approach to quaternion differential calculus similar to
+    \cite{Mandic2011,Xu2016}, but significantly simpler and more general, enabling
+    straight-forward adaptation of existing algorithms to systems with quaternion states.
+    For concreteness, we then apply our method to the canonical Wahba's
+    problem~\cite{Wahba1965} in Sec. \ref{sec:Wahbas}, and demonstrate superior
+    convergence to approaches that fail to properly account for the group structure. In
+    Sec. \ref{sec:trajopt} we extend these ideas to the problem of trajectory
+    optimization, and detail modifications to ALTRO, a state-of-the-art constrained
+    trajectory optimization solver, and demonstrate performance gains on several
+    benchmark problems. With the modifications presented in this paper, ALTRO explicitly 
+    leverages both the structure of the trajectory optimization problem as well as the 
+    group structure of 3D rotations, making it uniquely well-suited to solving challenging
+    problems with near real-time performance.
+
     %To our knowledge, there does not currently exist a solver that is capable of leveraging the unique Markovian structure of the fixed-horizon trajectory optimization problem while correctly accounting for the group structure of 3D rotations.
@@
     \begin{itemize}
-        \item A unified approach to quaternion differential calculus entirely based on standard linear algebra and vector calculus.
-        \item Derivation of a Newton-based algorithm for nonlinear optimization directly on the space of unit quaternions.
-        \item Implementation of a fast and efficient solver for trajectory optimization problems with 
-        attitude dynamics and nonlinear constraints that correctly accounts for the group
-        structure of 3D rotations.
+        \item A unified approach to quaternion differential calculus entirely based on
+        standard linear algebra and vector calculus.
+        \item Derivation of a Newton-based algorithm for nonlinear optimization directly
+        on the space of unit quaternions.
+        \item Implementation of a fast and efficient solver for trajectory optimization
+        problems with attitude dynamics and nonlinear constraints that correctly accounts
+        for the group structure of 3D rotations.
     \end{itemize}
@@
         %I_q$, where $I_q$ is the quaternion identity.
-        Once again, \eqref{eq:quat_jacobian} holds (up to a constant) for any three-parameter attitude representation.
+        Once again, \eqref{eq:quat_jacobian} holds (up to a constant) for any
+        three-parameter attitude representation.
 
@@
         \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
-        \caption{Convergence comparison for Wahba's problem.
-By modfying Newton's method with the methods of section \ref{sec:Quaternion_Calculus},        quadratic convergence rates are achieved, while a na\"ive approach stalls after only a few itterations.}
+        \caption{Convergence comparison for Wahba's problem. By modfying Newton's method
+        with the methods of section \ref{sec:Quaternion_Calculus}, quadratic convergence
+        rates are achieved, while a na\"ive approach stalls after only a few
+        iterations.}
         \label{fig:wahba_convergence}
     \end{figure}
-    % \begin{figure}[H]
-    %     \centering
-    %     \includegraphics[width=\columnwidth, height=5cm]{figures/kevins_plots/convergence_plot.tikz}
-    %     \caption{Convergence comparison for Wahba's problem. By performing Newton's method
-    %     on the error quaternion and applying the result to the full quaternion we achieve
-    %     quadratic convergence, whereas the more na\"ive approach doesn't converge to zero
-    %     error.}
-    %     \label{fig:wahba_convergence}
-    % \end{figure}
 
@@
     and $\bar{x}_k$ the state from the previous iteration. 
-For more details on the ALTRO algorithm, which leverages ideas from both DDP and direct Newton methods, we refer the reader to~\cite{howell2019altro}.
+    For more details on the ALTRO algorithm, which leverages ideas from both DDP and
+    direct Newton methods, we refer the reader to~\cite{howell2019altro}.
 
@@
     satellite with flexible solar panels that must slew to a new orientation while
-    avoiding a keep-out zone. All results were run on a desktop computer with an AMD
-    Ryzen 2950x processor with 40 GB of RAM. All problems are run using ALTRO, first
-    without any of the modifications presented in the current paper, analagous to the
-    naive Newton's method in section \ref{sec:Wahbas} and labeled ``naive'', and then
-    using the modifications listed in Sec. \ref{sec:trajopt} and the geodesic cost
-    function described in Sec. \ref{sec:cost_functions}, labeled ``modified''. Timing
-    results are summarized in Table \ref{tab:timing_results}. All experiments were solved
-    to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 4th order
-    Runge-Kutta integrator. Code for all experiments is available on
+    avoiding a keep-out zone.    
+    All problems are run using ALTRO, first without any of the modifications presented in
+    the current paper, analagous to the naive Newton's method in section \ref{sec:Wahbas}
+    and labeled ``naive'', and then using the modifications listed in Sec.
+    \ref{sec:trajopt} and the geodesic cost function described in Sec.
+    \ref{sec:cost_functions}, labeled ``modified''. All cost functions were of the
+    following form:
+    \begin{align}
+        \ell_\text{naive}(x, \bar{x}, Q) &= \half (x - \bar{x})^T Q (x - \bar{x}) \\
+        \ell_\text{modified}(x, \bar{x}, Q) &= \half (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
+            \pm \bar{q}^T q)
+    \end{align}
+    where $\bar{x}$ is the reference state and $\bar{Q} = \text{diag}(Q_r, \vec{0}_4,
+    Q_v, Q_\omega)$, with $Q_r,Q_v,Q_\omega$ being the weights of $Q$ for position, and
+    linear and angular velocity, respectively.
+
+    Timing results are summarized in Table \ref{tab:timing_results}. All experiments were
+    solved to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 3rd
+    order Runge-Kutta integrator. All results were run on a desktop computer with an AMD
+    Ryzen 2950x processor with 40 GB of RAM. Code for all experiments is available on
     \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
@@
 
-        A 180 degree barrel roll trajectory for a fixed-wing airplane was optimized. The airplane's dynamics model consists of the a simple rigid body as defined in Section
-        \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag fit
-        from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do a
-        barrel roll by setting a high terminal cost for deviations from an upside-down attitude (see Fig.
-        \ref{fig:barrellroll}). The solver was initialized with level flight trim
+        A 180 degree barrel roll trajectory for a fixed-wing airplane was optimized. The
+        airplane's dynamics model consists of the a simple rigid body as defined in
+        Section \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag
+        fit from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do
+        a barrel roll by constraining the terminal state to upside-down (see Fig.
+        \ref{fig:barrellroll}). To avoid issues with the double-cover and integration error,
+        the following was used to enforce a terminal orientation to $\bar{q}$:
+        \begin{equation}
+            \frac{q_v}{\norm{q}} - \bar{q}_v
+                \sign\left(\bar{q}^T \frac{q }{ \norm{q}}\right) = 0
+        \end{equation}
+
+        The solver was initialized with level flight trim
         conditions. The convergence of the different versions of ALTRO is compared in
@@
         the expansions being provided to the algorithm more accurately capture the
-        relationship between the attitude state and the goal and constraints.
-\todo{Brian: you mention constraints in the text and have it on the y-axis in Fig. 4, but you don't state what these constraints are in the text.}
-
-\todo{Brian: comment on the performance numbers in the text.}
-
-		% \todo{If we're not going to plot it and actually do an experimental comparison, I vote for leaving this stuff out.}
-        % We also compared the ``error quadratic'' cost function described in Section
-        % \ref{sec:error_quadratic} to the geodesic cost function in Section
-        % \ref{sec:geodesic}. The geodesic cost was more efficient, both in term of
-        % iterations and computation time, taking only 36 iterations and 1.8 ms/iteration
-        % vs 76 iterations and 7.6 ms/iteration for the error quadratic. Not only is the
-        % geodesic cost function much cheaper to compute, it also converges much faster
-        % than the error quadratic cost function. This could be due to the strong nonlinearities
-        % introduced when computing the quaternion error, making it more difficult to
-        % optimize.
+        relationship between the attitude state and the goal and constraints. For this 
+        relatively simple problem, we gained a modest 7\% improvement in runtime, despite
+        the additional matrix multiplications when calculating the cost and constraint
+        expansions.
 
@@
         A 360 degree flip trajectory for a quadrotor was optimized with dynamics adapted
-        from \cite{mellinger2012trajectory}. Four
-        intermediary ``keyframes'' were used to encourage the quadrotor to reacj angles
-        of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around an approximately circular
-        arc. The quadrotor was constrained to stay above the floor and move to a goal
-        state 2 meters away in the $+y$ direction. The solver was initialized with a
-        dynamically infeasible trajectory that linearly interpolates between the initial
-        and final states, rotating the quadrotor around the x-axis a full \ang{360}.
-
-\todo{Brian: little more detail on cost weights.
-Any control cost?
-If we used the same cost function form for all of these (i.e.
-geodesic on q + quadratic on everything else), let's write it at the beginning of the experiment section.}
+        from \cite{mellinger2012trajectory}. To encourage the flip, we specified a 
+        ``waypoint'' cost function of the following form:
+        \begin{multline}
+                  \sum_{k \in \mathcal{N}} \ell(x_k, \hat{x}, \hat{Q}) 
+                + \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
+                + \half \sum_{k=1} u_k^T R u_k
+        \end{multline}
+        with $\mathcal{W} = \{45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
+        \mathcal{W}$. The four intermediary ``waypoints'' were used to encourage the
+        quadrotor to reach angles of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around
+        an approximately circular arc. The last waypoint was used to encourage the
+        quadrotor to move towards the final goal. The quadrotor was constrained to stay
+        above the floor and move to a goal state 2 meters away in the $+y$ direction. The
+        solver was provided a dynamically infeasible initial trajectory that linearly
+        interpolates between the initial and final states, rotating the quadrotor around
+        the x-axis a full \ang{360}.
 
 	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
-	    ALTRO. The original version of ALTRO, even after significant tuning efforts,
-	    was not able to converge to the desired solution.
-	    %It instead got halfway and then ``unwound'' 360 degrees and then continued rotating to the final orientation.
+	    ALTRO. The original version of ALTRO, even after significant tuning efforts, was
+        not able to converge to the desired solution. 
 	    This behavior is common and expected when attempting optimization that does not
@@
 	    representation, since it passes through the singularities at $90\degree,
-	    180\degree$, and $360\degree$ associated with Euler angles, Rodrigues parameters, and
-	    Modified Rodrigues Parameters, respectively.
+	    180\degree$, and $360\degree$ associated with Euler angles, Rodrigues parameters,
+	    and Modified Rodrigues Parameters, respectively.
 
-	    \todo{Brian: you say the original version didn't work at all, but there are timing results for it in the table???}
-
-        \begin{figure}[t]
+        \begin{figure}[ht]
             \centering
@@
         constraint itself being nonlinear in $q$.
-        \begin{figure}[H]
+        \begin{figure}[ht]
             \centering

### 2020-10-14 paper/main.tex
tiny edits

@@
     \end{equation}
-    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
+    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update $\phi_k$ via the Moore-Penrose psuedoinverse:
     \begin{equation}
@@
 %ALTRO is an efficient solver for constrained nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented Lagrangian framework.
-
     We consider trajectory optimization problems of the form,
@@
 
-    Like most gradient or Newton-based methods for optimization, ALTRO approximates the
+    ALTRO combines techniques from both differential dynamic programming (DDP) and direct transcription methods to achieve high performance on challenging constrained nonlinear trajectory optimization problems.
+Like most methods for nonlinear optimization, ALTRO iteratively approximates the
     nonlinear functions $f, \ell, g,$ and $h$ with their first or second-order Taylor
@@
 	With these results, we can apply standard Newton and quasi-Newton techniques along the lines of Section \ref{sec:Wahbas}.
-We can also calculate a second-order expansion of the ``action-value function'' $Q(x,u)$ needed in differential dynamic programming (DDP) and LQR-based methods,
+We can also calculate a second-order expansion of the ``action-value function'' $Q(x,u)$ needed in DDP and LQR-based methods,
     \begin{align}
@@
     where $\bar{u}_k$ is the control value from the previous iteration, and $\delta x$ is
-    computed using \eqref{eq:state_error}, with $x_k$ being the current state estimate
-    and $\bar{x}_k$ the state from the previous iteration. 
-    For more details on the ALTRO algorithm, which leverages ideas from both DDP and
-    direct Newton methods, we refer the reader to~\cite{howell2019altro}.
+    computed using \eqref{eq:state_error}.%, with $x_k$ being the current state estimate and $\bar{x}_k$ the state from the previous iteration. 
+    For more details on the ALTRO algorithm, we refer the reader to~\cite{howell2019altro}.
 
@@
         quadrotor to move towards the final goal. The quadrotor was constrained to stay
-        above the floor and move to a goal state 2 meters away in the $+y$ direction. The
+        above the floor and move to a goal state two meters away in the $+y$ direction. The
         solver was provided a dynamically infeasible initial trajectory that linearly
@@
 \section*{Acknowledgements}
-This material is based upon work supported by the National Science Foundation Graduate
+This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503). This material is based upon work supported by the National Science Foundation Graduate
 Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
 conclusions or recommendations expressed in this material are those of the author(s) and
-do not necessarily reflect the views of the National Science Foundation
-
-\todo{Add NASA acknowledgements (JPL + ECF)}
+do not necessarily reflect the views of the National Science Foundation.
 

### 2020-10-14 paper/main.tex
Updates from Overleaf

--- 2020-10-14 paper/main.tex
@@
     Given a set of known vectors in the world frame, $\inframe{W}{w_i}$, and measurements of these vectors in the body frame, $\inframe{B}{v_i}$, we seek the
-    rotation from the body to the world frame $\toframe{W}{B}{A(q)}$ that solves the following optimization problem, \todo{Kevin: Don't use $L$ here because it conflicts with \eqref{eq:Lmult}:}
+    rotation from the body to the world frame $\toframe{W}{B}{A(q)}$ that solves the following optimization problem,
     \begin{mini*}
-        {q}{ L(q) }{}{}
+        {q}{ W(q) }{}{}
         \addConstraint { q\in \Q,}
     \end{mini*}
-    where Wahba's loss function $L(q)$ is,
+    where Wahba's loss function $W(q)$ is,
     \begin{equation} \label{eq:wahba_loss}
-        L(q) = \sum_i \norm{\inframe{W}{w_i} - A(q) \,\, \inframe{B}{v_i} }_2^2 
+        W(q) = \sum_i \norm{\inframe{W}{w_i} - A(q) \,\, \inframe{B}{v_i} }_2^2 
            = \norm{r(q)}_2^2 ,
@@
 
-    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_jacobian} and \eqref{eq:quaternion_rotation}: \todo{Kevin: fix the second line of this equation:}
-    \begin{equation} \label{eq:newton_foc}
-        \begin{aligned}
+    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
+    \begin{align} \label{eq:newton_foc}
             \nabla r(q) &= \pdv{r}{q} G(q) \\
-              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) \right).
-        \end{aligned}
-    \end{equation}
+              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) G(q)\right).
+    \end{align}
     Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:

### 2020-10-14 paper/main.tex

--- 2020-10-14 paper/main.tex
@@
 
-    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_jacobian} and \eqref{eq:quaternion_rotation}: \todo{Kevin: fix the second line of this equation:}
-    \begin{equation} \label{eq:newton_foc}
-        \begin{aligned}
+    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
+    \begin{align} \label{eq:newton_foc}
             \nabla r(q) &= \pdv{r}{q} G(q) \\
-              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) \right).
-        \end{aligned}
-    \end{equation}
-    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update $\phi_k$ via the Moore-Penrose psuedoinverse:
+              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) G(q)\right).
+    \end{align}
+    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
+
     \begin{equation}

### 2020-10-14 paper/main.tex
added JPL acknowledgement
Tweak the costs and some other minor edits

--- 2020-10-14 paper/main.tex
@@
     Given a set of known vectors in the world frame, $\inframe{W}{w_i}$, and measurements of these vectors in the body frame, $\inframe{B}{v_i}$, we seek the
-    rotation from the body to the world frame $\toframe{W}{B}{A(q)}$ that solves the following optimization problem,
+    rotation from the body to the world frame $\toframe{W}{B}{A(q)}$ that solves the following optimization problem, \todo{Kevin: Don't use $L$ here because it conflicts with \eqref{eq:Lmult}:}
     \begin{mini*}
-        {q}{ W(q) }{}{}
+        {q}{ L(q) }{}{}
         \addConstraint { q\in \Q,}
     \end{mini*}
-    where Wahba's loss function $W(q)$ is,
+    where Wahba's loss function $L(q)$ is,
     \begin{equation} \label{eq:wahba_loss}
-        W(q) = \sum_i \norm{\inframe{W}{w_i} - A(q) \,\, \inframe{B}{v_i} }_2^2 
+        L(q) = \sum_i \norm{\inframe{W}{w_i} - A(q) \,\, \inframe{B}{v_i} }_2^2 
            = \norm{r(q)}_2^2 ,
@@
 
-    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
-    \begin{align} \label{eq:newton_foc}
+    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_jacobian} and \eqref{eq:quaternion_rotation}: \todo{Kevin: fix the second line of this equation:}
+    \begin{equation} \label{eq:newton_foc}
+        \begin{aligned}
             \nabla r(q) &= \pdv{r}{q} G(q) \\
-              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) G(q)\right).
-    \end{align}
-    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
+              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) \right).
+        \end{aligned}
+    \end{equation}
+    Given a guess solution, $q_k$, the standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
     \begin{equation}
@@
 	\begin{equation}
-		\q_{k+1} = \q_k \otimes \varphi(\phi_k)
-	\end{equation}.
+		\q_{k+1} = \q_k \otimes \varphi(\phi_k).
+	\end{equation}
 	This ``multiplicative'' Gauss-Newton method is summarized in Algorithm \ref{alg:mgn}.
@@
     \subsection{Results}
-    Figure \ref{fig:wahba_convergence} compares the multiplicative Gauss-Newton method with a na\"ive Newton's method in which the quaternion is simply projected back onto the unit sphere at every iteration.
-That the na\"ive method makes progress
-    initially, but quickly stalls.
-    By correctly handling the group structure of unit quaternions, the multiplicative method is able
-    to maintain the fast convergence rates typical of Newton's method.
+    Figure \ref{fig:wahba_convergence} compares the multiplicative Gauss-Newton method
+    with a na\"ive Newton's method in which the quaternion is simply projected back onto
+    the unit sphere at every iteration. The na\"ive method makes progress initially, but
+    quickly stalls. By correctly handling the group structure of unit quaternions, the
+    multiplicative method is able to maintain the fast convergence rates typical of
+    Newton's method.
     \begin{figure}
@@
     and $\bar{x}_k$ the state from the previous iteration. 
-    For more details on the ALTRO algorithm, which leverages ideas from both DDP and
-    direct Newton methods, we refer the reader to~\cite{howell2019altro}.
+    For more details on the ALTRO algorithm, which leverages ideas from both DDP
+    and direct Newton methods, we refer the reader to~\cite{howell2019altro}.
 
@@
         frequently minimize costs that penalize distance from a goal state, e.g. $\half
-        (x-x_g)^T Q (x-x_g)$; However, na\"ive substraction of unit quaternions does not respect their group structure, and often results in undesired behavior.
+        (x-x_g)^T Q (x-x_g)$; however, na\"ive substraction of unit quaternions does not respect their group structure, and often results in undesired behavior.
 Instead, we have found the following cost function, which penalizes the geodesic
@@
     All problems are run using ALTRO, first without any of the modifications presented in
-    the current paper, analagous to the naive Newton's method in section \ref{sec:Wahbas}
+    the current paper, analagous to the na\"ive Newton's method in section \ref{sec:Wahbas}
     and labeled ``naive'', and then using the modifications listed in Sec.
     \ref{sec:trajopt} and the geodesic cost function described in Sec.
-    \ref{sec:cost_functions}, labeled ``modified''. All cost functions were of the
+    \ref{sec:cost_functions}, labeled ``modified''. All cost functions are of the
     following form:
     \begin{align}
-        \ell_\text{naive}(x, \bar{x}, Q) &= \half (x - \bar{x})^T Q (x - \bar{x}) \\
-        \ell_\text{modified}(x, \bar{x}, Q) &= \half (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
-            \pm \bar{q}^T q)
+        \ell_\text{naive}(x, u, \bar{x}, Q, R) &= \half (x - \bar{x})^T Q (x - \bar{x}) + \half u^T R u \\
+        \ell_\text{modified}(x, u, \bar{x}, Q, R) 
+            &= \ell_\text{naive}(x, u, \bar{x}, \bar{Q}, R) + w (1 \pm \bar{q}^T q)
     \end{align}
@@
 
-    Timing results are summarized in Table \ref{tab:timing_results}. All experiments were
+    Timing results are summarized in Table \ref{tab:timing_results}. All experiments are 
     solved to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 3rd
-    order Runge-Kutta integrator. All results were run on a desktop computer with an AMD
+    order Runge-Kutta integrator. The results were run on a desktop computer with an AMD
     Ryzen 2950x processor with 40 GB of RAM. Code for all experiments is available on
@@
         fit from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do
-        a barrel roll by constraining the terminal state to upside-down (see Fig.
+        a barrel roll by constraining the terminal state to be upside-down (see Fig.
         \ref{fig:barrellroll}). To avoid issues with the double-cover and integration error,
-        the following was used to enforce a terminal orientation to $\bar{q}$:
+        the following was used to constrain a terminal orientation to $\bar{q}$:
         \begin{equation}
@@
 
-        \begin{figure}[ht]
+        \begin{figure}[t]
             \centering
@@
         \begin{multline}
-                  \sum_{k \in \mathcal{N}} \ell(x_k, \hat{x}, \hat{Q}) 
-                + \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) 
-                + \half \sum_{k=1} u_k^T R u_k
+                  \sum_{k \in \mathcal{N}} \ell(x_k, u_k, \hat{x}, \hat{Q}, R) 
+                + \sum_{k \in \mathcal{W}} \ell(x_k, u_k, \bar{x}_k, Q_w, R) 
         \end{multline}
-        with $\mathcal{W} = \{45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
-        \mathcal{W}$. The four intermediary ``waypoints'' were used to encourage the
+        where $\hat{x}$, $\hat{Q}$ are the nominal state and state weight matrix, $Q_w$ is 
+        the weight matrix for the waypoints, and
+        $\mathcal{W} = \{45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
+        \mathcal{W}$.
+        Four intermediary ``waypoints'' were used to encourage the
         quadrotor to reach angles of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around

### 2020-10-14 paper/main.tex

--- 2020-10-14 paper/main.tex
@@
 \section*{Acknowledgements}
-This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503). This material is based upon work supported by the National Science Foundation Graduate
+This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503).
+This research was carried out in part at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration and funded through JPL’s Strategic University Research Partnerships (SURP) program.
+This material is based upon work supported by the National Science Foundation Graduate
 Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and

### 2020-10-14 paper/main.tex
more small tweaks

--- 2020-10-14 paper/main.tex
@@
             \nabla r(q) &= \pdv{r}{q} G(q) \\
-              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) G(q)\right).
+              &= -2 H^T R(q)^T \left( \sum_i R({}^B \hat{v}_i) \right) G(q) .
     \end{align}
@@
         a barrel roll by constraining the terminal state to upside-down (see Fig.
-        \ref{fig:barrellroll}). To avoid issues with the double-cover and integration error,
-        the following was used to enforce a terminal orientation to $\bar{q}$:
+        \ref{fig:barrellroll}). To mitigate issues with integration error and drift in the magnitude of the quaternion,
+        the following constraint function was used to enforce a terminal orientation of $\bar{q}$:
         \begin{equation}
@@
     systems with arbitrary attitude using standard linear algebra and vector calculus.
-    We have demonstrated that the application of this methodology is straightforward and
-    yields substantial improvements in the convergence of Newton-based methods (see Fig. 
-    \ref{fig:wahba_convergence}), while also offering improvements for nonlinear constrained
-    trajectory optimization for floating-base systems (see Table \ref{tab:timing_results}).
+    The application of this methodology is straightforward and
+    yields substantial improvements in the convergence of Newton and DDP-based methods, while also offering improvements for nonlinear constrained
+    trajectory optimization for floating-base systems.
 
-    \todo{This is a really nice pitch. Can we say something like this in the intro?}
-    With the modifications presented, ALTRO can solve problems few other methods for
-    trajectory optimization can. Many state-of-the-art methods, such as direct
-    collocation or sequential convex programming, rely on commercial, proprietary, or
-    general-purpose solvers whose internal numerics often are abstracted away from the
-    user. By exploiting the unique structure of both the trajectory optimization problem
-    and the rotation group, ALTRO will likely be able to solve more challenging problems
-    with performance.
+    Many state-of-the-art trajectory optimization methods, including direct collocation and sequential convex programming, rely on general-purpose optimization solvers whose internal numerical methods are not exposed to the user.
+Therefore, these methods are unable to exploit the full structure of both the trajectory optimization problem and the rotation group at a low level.
+In contrast, we are able to implement deep, native support for quaternions into the ALTRO solver, making it possible to solve more challenging problems with higher performance than other algorithms.
 
-    The methods presented here can easily be leveraged to adapt other classes
-    of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future
-    work may include adaptation of methods for state estimation, localization, design, or 
-    other methods for motion planning such as direct collocation.
-
-    Future work will focus on continued 
-    refinement of ALTRO and performance improvements for use in real-time model-predictive control applications, as well as extensions to multi-body robotic
-    systems, such as humanoids or quadrupeds, represented in
-    ``maximal'' coordinates that include the 3D orientation of each body \cite{brudigam2020linear}.
+    In future work, we plan to apply ALTRO to real-time model-predictive control problems for aerial vehicles like quadrotors and airships that experience large attitude changes.
+The methods we have presented can also be leveraged to adapt other classes of gradient or Newton-based algorithms to exploit the structure of 3D rotations.
+Future directions beyond trajectory optimization may include simulation and planning methods that leverage maximal-coordinate formulations of multi-body dynamics~\cite{brudigam2020linear}, system-identification for complex multi-body systems and fixed-wing aircraft in post-stall conditions, and state estimation for spacecraft with sparse measurements.
 
 \section*{Acknowledgements}
-This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503). This material is based upon work supported by the National Science Foundation Graduate
+This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503).
+This research was carried out in part at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration and funded through JPL’s Strategic University Research Partnerships (SURP) program.
+This material is based upon work supported by the National Science Foundation Graduate
 Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and

### 2020-10-14 paper/main.tex
Final changes to the flexible sat section.

@@
         A spacecraft with flexible appendages was tasked to perform a 150 degree slew maneuver while
-        ensuring that a body-mounted camera did not point within 40 degrees ``keep-out zone'' around the sun vector. The spacecraft dynamics are presented in detail in~\cite{Tracy2020}, and are based
-        on equation \eqref{eq:rigid_body_dynamics} with the addition of six states to account for three flexible modes. Control torques are generated by four reaction wheels.
-
-        \todo{Kevin: add some info on cost function, initial guess (if any), etc.}
-
+        ensuring that a body-mounted camera did not point within 40 degrees of a  ``keep-out zone'' around the sun vector.
+The spacecraft dynamics are presented in detail in~\cite{Tracy2020}, and are based
+        on equation \eqref{eq:rigid_body_dynamics} with the addition of six states to account for three flexible modes.
+Control torques are generated by four reaction wheels.
+The slew is motivated by a quadratic cost function that penalizes error from the final attitude, as well as excessive flexible displacement.
+        % \todo{Kevin: add some info on cost function, initial guess (if any), etc.}
         We enforce the camera keep-out zone can with the following constraint,
@@
         \end{figure}
-        % \todo{Talk a little about the results in the plot here.}
 
-        \todo{Kevin: looks like both versions of ALTRO worked fine on this problem, so we should probably not say all this group stuff.
-Maybe just comment on the timing results and try to offer an explanation.}
-
-        ALTRO is able to reason about this nonlinear constraint, and account for the group
-        structure present in the quaternion. From Figure \ref{fig:keepout} we can see that
-        without the camera constraint present, the trajectory will pass through the keep-out
+        ALTRO is able to converge on a feasible trajectory for this problem without any initial guess, with the resulting attitude history represented in Fig.
+\ref{fig:keepout}.
+Without the camera constraint present, the trajectory will pass through the keep-out
         zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
-        is able to converge on a trajectory that performs the slew in an efficient manner, while
+        is able to generate a trajctory that performs the slew in an efficient manner, while
         keeping the camera sufficiently far from the sun.

### 2020-10-14 paper/main.tex
typo fix

@@
         zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
-        is able to generate a trajctory that performs the slew in an efficient manner, while
+        is able to generate a trajectory that performs the slew in an efficient manner, while
         keeping the camera sufficiently far from the sun.

### 2020-10-14 paper/main.tex
Update timing results

--- 2020-10-14 paper/main.tex
@@
     The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
-    \begin{align} \label{eq:newton_foc}
-            \nabla r(q) &= \pdv{r}{q} G(q) \\
-              &= -2 H^T R(q)^T \left( \sum_i R({}^B \hat{v}_i) \right) G(q) .
-    \end{align}
+    \begin{equation} \label{eq:newton_foc}
+            \nabla r(q) = \pdv{r}{q} G(q)
+    \end{equation}
+    \begin{equation}
+            \nabla r_i(q) = -2 H^T R(q)^T R(\inframe{B}{\hat{v}_i}) G(q)
+    \end{equation}
     Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
@@
 
-    Timing results are summarized in Table \ref{tab:timing_results}. All experiments were
-    solved to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 3rd
-    order Runge-Kutta integrator. All results were run on a desktop computer with an AMD
+    Timing results are summarized in Table \ref{tab:timing_results}. All experiments are 
+    solved to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 4th 
+    order Runge-Kutta integrator. The results were run on a desktop computer with an AMD
     Ryzen 2950x processor with 40 GB of RAM. Code for all experiments is available on
@@
         fit from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do
-        a barrel roll by constraining the terminal state to upside-down (see Fig.
-        \ref{fig:barrellroll}). To mitigate issues with integration error and drift in the magnitude of the quaternion,
-        the following constraint function was used to enforce a terminal orientation of $\bar{q}$:
+        a barrel roll by constraining the terminal state to be upside-down (see Fig.
+        \ref{fig:barrellroll}). To avoid issues with the double-cover and integration error,
+        the following was used to constrain a terminal orientation to $\bar{q}$:
         \begin{equation}
@@
         relationship between the attitude state and the goal and constraints. For this 
-        relatively simple problem, we gained a modest 7\% improvement in runtime, despite
+        relatively simple problem, we gained a modest 11\% improvement in runtime, despite
         the additional matrix multiplications when calculating the cost and constraint
@@
     systems with arbitrary attitude using standard linear algebra and vector calculus.
-    The application of this methodology is straightforward and
-    yields substantial improvements in the convergence of Newton and DDP-based methods, while also offering improvements for nonlinear constrained
-    trajectory optimization for floating-base systems.
+    We have demonstrated that the application of this methodology is straightforward and
+    yields substantial improvements in the convergence of Newton-based methods (see Fig. 
+    \ref{fig:wahba_convergence}), while also offering improvements for nonlinear constrained
+    trajectory optimization for floating-base systems (see Table \ref{tab:timing_results}).
 
-    Many state-of-the-art trajectory optimization methods, including direct collocation and sequential convex programming, rely on general-purpose optimization solvers whose internal numerical methods are not exposed to the user.
-Therefore, these methods are unable to exploit the full structure of both the trajectory optimization problem and the rotation group at a low level.
-In contrast, we are able to implement deep, native support for quaternions into the ALTRO solver, making it possible to solve more challenging problems with higher performance than other algorithms.
+    \todo{This is a really nice pitch. Can we say something like this in the intro?}
+    With the modifications presented, ALTRO can solve problems few other methods for
+    trajectory optimization can. Many state-of-the-art methods, such as direct
+    collocation or sequential convex programming, rely on commercial, proprietary, or
+    general-purpose solvers whose internal numerics often are abstracted away from the
+    user. By exploiting the unique structure of both the trajectory optimization problem
+    and the rotation group, ALTRO will likely be able to solve more challenging problems
+    with performance.
 
-    In future work, we plan to apply ALTRO to real-time model-predictive control problems for aerial vehicles like quadrotors and airships that experience large attitude changes.
-The methods we have presented can also be leveraged to adapt other classes of gradient or Newton-based algorithms to exploit the structure of 3D rotations.
-Future directions beyond trajectory optimization may include simulation and planning methods that leverage maximal-coordinate formulations of multi-body dynamics~\cite{brudigam2020linear}, system-identification for complex multi-body systems and fixed-wing aircraft in post-stall conditions, and state estimation for spacecraft with sparse measurements.
+    The methods presented here can easily be leveraged to adapt other classes
+    of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future
+    work may include adaptation of methods for state estimation, localization, design, or 
+    other methods for motion planning such as direct collocation.
+
+    Future work will focus on continued 
+    refinement of ALTRO and performance improvements for use in real-time model-predictive control applications, as well as extensions to multi-body robotic
+    systems, such as humanoids or quadrupeds, represented in
+    ``maximal'' coordinates that include the 3D orientation of each body \cite{brudigam2020linear}.
 

### 2020-10-14 examples/Flexible_Spacecraft/solve_con.jl

@@ -13,14 +13,18 @@ using Altro
 
 const TO = TrajectoryOptimization
 
+using PlanningWithAttitude: post_process, dcm_from_q
+
 ##
 prob,opts = SatelliteKeepOutProblem(vecstate=true, costfun=LQRCost)
 prob,opts = SatelliteKeepOutProblem(vecstate=false, costfun=QuatLQRCost)
-solver = ALTROSolver(prob,opts)
+solver = ALTROSolver(prob, opts)
 solve!(solver)
 
 
 ##
+bodyvec = prob.constraints[3].bodyvec
+keepoutdir = prob.constraints[3].keepoutdir
 ω_hist, q_hist, u_hist,η_hist, ηd_hist, r_hist,θ_hist = post_process(solver)
 
 keepout_truth = zeros(size(q_hist,2))
@@ -31,6 +35,10 @@ for i = 1:size(q_hist,2)
 	q = q_hist[:,i]
 	g_hist[:,i] = q[2:4]/q[1]
 end
+
+using Plots
+plot(keepout_truth)
+plot(controls(solver))
 
 mat"
 figure

### 2020-10-14 examples/quadrotor_flip.jl

@@ -16,9 +16,6 @@ using Rotations
 using Colors
 
 ## Try ALTRO
-# prob = gen_quad_flip(UnitQuaternion, slack=false, vecmodel=false, renorm=true,
-#     costfun=QuatLQRCost
-# )
 prob,opts = QuadFlipProblem(UnitQuaternion, slack=false, vecmodel=false, renorm=true,
     costfun=QuatLQRCost
 )

### 2020-10-14 examples/trajopt_examples.jl

@@ -42,7 +42,7 @@ results_path = joinpath(@__DIR__, "timing_results.jld2")
 Barrell Roll
 """
 args = (
-    integration = RK3,
+    integration = RK4,
     termcon = :quatvec,
     # projected_newton = false,
     show_summary=true
@@ -51,12 +51,11 @@ args = (
 ## Original Method
 prob,opts = YakProblems(vecstate=true, costfun=:Quadratic; args...) 
 solver = ALTROSolver(prob, opts)
-solve!(solver)
 add_result!(:barrellroll, false, solver)
 
 ## Modified Method
 prob,opts = YakProblems(vecstate=false, costfun=:QuatLQR; args...) 
-solver = ALTROSolver(prob, opts)
+solver = ALTROSolver(prob, opts, show_summary=true)
 add_result!(:barrellroll, true, solver)
 
 @save results_path results 
@@ -70,7 +69,7 @@ solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true)
 add_result!(:quadflip, false, solver)
 
 ## Modified Method
-prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost)
+prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost; args...)
 solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true)
 add_result!(:quadflip, true, solver)
 
@@ -85,7 +84,7 @@ solver = ALTROSolver(prob, opts)
 add_result!(:satellite, false, solver)
 
 ## Modified Method
-prob,opts = SatelliteKeepOutProblem(vecstate=false, costfun=QuatLQRCost)
+prob,opts = SatelliteKeepOutProblem(vecstate=false, costfun=QuatLQRCost; args...)
 solver = ALTROSolver(prob, opts)
 add_result!(:satellite, true, solver)
 

### 2020-10-14 src/flexible_spacecraft_dynamics.jl

@@ -161,10 +161,10 @@ function dcm_from_q(q)
 	return I + 2*skew(v)*(s*I + skew(v))
 end
 
-function SatelliteKeepOutProblem(constrained::Bool=true; vecstate::Bool=false, costfun=LQRCost)
+function SatelliteKeepOutProblem(constrained::Bool=true; vecstate::Bool=false, 
+        costfun=LQRCost, N = 401, termcon = :quatvec, integration=RD.RK4, kwargs...)
 	# Discretization
 	tf = 100.0
-	N = 401
 	dt = tf/(N-1)
 
 	# Model
@@ -209,7 +209,7 @@ function SatelliteKeepOutProblem(constrained::Bool=true; vecstate::Bool=false, c
 	end
 
 	## Solve
-	prob = Problem(model, obj, xf, tf, x0=x0,constraints = cons,N=N;integration=RD.RK4)
+	prob = Problem(model, obj, xf, tf, x0=x0,constraints = cons,N=N;integration=integration)
 	# U0 = [@SVector randn(m) for k = 1:N-1] .* 1e-1
 	# initial_controls!(prob, U0)
 
@@ -223,7 +223,8 @@ function SatelliteKeepOutProblem(constrained::Bool=true; vecstate::Bool=false, c
 		projected_newton_tolerance = 1e-4,
 		show_summary = true,
 		projected_newton = false,
-		verbose_pn = 1
+        verbose_pn = 1;
+        kwargs...
 	)
 	return prob, opts
 end

### 2020-10-14 src/quadflip_problem.jl

@@ -1,7 +1,7 @@
 using RobotZoo: Quadrotor
 
 function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=false, 
-        renorm::Bool=false, costfun=LQRCost, integration=RD.RK4)
+        renorm::Bool=false, costfun=LQRCost, integration=RD.RK3, termcon=:none, kwargs...)
     model = Quadrotor{Rot}()
     if renorm
         model = QuatRenorm(model)
@@ -122,7 +122,7 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
     end
     initial_states!(prob, X_guess)
 
-    opts = SolverOptions(
+    opts = SolverOptions(;
         cost_tolerance=1e-5,
         cost_tolerance_intermediate=1e-5,
         constraint_tolerance=1e-5,
@@ -133,7 +133,8 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
         show_summary=false,
         verbose_pn=false,
         verbose=0,
-        projected_newton=true
+        projected_newton=true, 
+        kwargs...
     )
     return prob, opts
 end

### 2020-10-14 paper/main.tex
small tweaks

--- 2020-10-14 paper/main.tex
@@
 Control torques are generated by four reaction wheels.
-The slew is motivated by a quadratic cost function that penalizes error from the final attitude, as well as excessive flexible displacement.
+A quadratic cost function penalizes error from the desired final attitude as well as displacement of the flexible modes.
         % \todo{Kevin: add some info on cost function, initial guess (if any), etc.}
@@
         where $^Br_{cam}$ is the camera line-of-sight unit vector in the body frame and $^Wr_{sun}$ is the unit vector pointing to the sun in the world frame.
-The attitudes that satisfy this constraint comprise a non-convex set, with the
-        constraint itself being nonlinear in $q$.
+The attitudes that satisfy this constraint comprise a non-convex set, with the constraint itself being nonlinear in $q$.
         \begin{figure}[ht]
@@
 
-        ALTRO is able to converge on a feasible trajectory for this problem without any initial guess, with the resulting attitude history represented in Fig.
+        ALTRO is able to converge to a locally optimal trajectory for this problem without an initial guess (all controls were initialized to zero).
+The resulting attitude trajectory is depicted in Fig.
 \ref{fig:keepout}.
-Without the camera constraint present, the trajectory will pass through the keep-out
-        zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
-        is able to generate a trajectory that performs the slew in an efficient manner, while
-        keeping the camera sufficiently far from the sun.
-
-        The modifications didn't have a significant impact on runtime performance for this
-        example. In our experience, the computational benefits of the proposed methodology are
-        problem-dependent, especially for highly nonlinear systems. The following examples result
-        in more dynamic and aerobatic behaviors, where the computational benefit of the proposed
-        methodology is significant, including an example where the naive method failed to find a
-        good solution.
+Without enforcing the camera constraint, the trajectory passes through the keep-out zone.
+As noted in Table \ref{tab:timing_results}, the quaternion modifications did not result a significant improvement over the naive implementation of ALTRO for this problem, indicating that the computational benefits are problem dependent.
+We hypothesize that the more dynamic behaviors in the other examples benefit more from the quaternion modifications than the relatively slow-moving spacecraft.
 

### 2020-10-14 paper/main.tex
Small tweaks

--- 2020-10-14 paper/main.tex
@@
                 & -I_3 
-            \end{bmatrix} q
+            \end{bmatrix} q.
         \end{equation}
@@
     The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
-    \begin{equation} \label{eq:newton_foc}
-            \nabla r(q) = \pdv{r}{q} G(q)
-    \end{equation}
-    \begin{equation}
-            \nabla r_i(q) = -2 H^T R(q)^T R(\inframe{B}{\hat{v}_i}) G(q)
-    \end{equation}
+    \begin{align} \label{eq:newton_foc}
+            \nabla r(q) &= \pdv{r}{q} G(q) \\
+              &= -2 H^T R(q)^T \left( \sum_i R({}^B \hat{v}_i) \right) G(q) .
+    \end{align} 
     Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:
@@
         % \todo{Kevin: add some info on cost function, initial guess (if any), etc.}
-        We enforce the camera keep-out zone can with the following constraint,
+        We enforce the camera keep-out zone with the following constraint,
         \begin{equation}
@@
         example. In our experience, the computational benefits of the proposed methodology are
-        problem-dependent, especially for highly nonlinear systems. The following examples result
-        in more dynamic and aerobatic behaviors, where the computational benefit of the proposed
-        methodology is significant, including an example where the naive method failed to find a
-        good solution.
+        problem-dependent, especially for highly nonlinear systems.
 

### 2020-10-14 paper/main.tex

@@
 
-        ALTRO is able to converge on a feasible trajectory for this problem without any initial guess, with the resulting attitude history represented in Fig.
+        ALTRO is able to converge to a locally optimal trajectory for this problem without an initial guess (all controls were initialized to zero).
+The resulting attitude trajectory is depicted in Fig.
 \ref{fig:keepout}.
-Without the camera constraint present, the trajectory will pass through the keep-out
-        zone, making the trajectory infeasible. With the keep-out zone constraint present, ALTRO
-        is able to generate a trajectory that performs the slew in an efficient manner, while
-        keeping the camera sufficiently far from the sun.
-
-        The modifications didn't have a significant impact on runtime performance for this
-        example. In our experience, the computational benefits of the proposed methodology are
-        problem-dependent, especially for highly nonlinear systems.
+Without enforcing the camera constraint, the trajectory passes through the keep-out zone.
+As noted in Table \ref{tab:timing_results}, the quaternion modifications did not result in a significant improvement over the naive implementation of ALTRO for this problem, indicating that the computational benefits are problem dependent.
+We hypothesize that the more dynamic behaviors in the other examples benefit more from the quaternion modifications than the relatively slow-moving spacecraft.
 

### 2020-10-15 SE3_tuning.jl
Clean out old files

@@ -1,35 +1,0 @@
-# Test SE(3) controller
-Rot = UnitQuaternion{Float64,CayleyMap}
-# lee = Dynamics.LeeQuad(Rot)
-model = Dynamics.Quadrotor2{Rot}()
-# eigen(lee.J)
-# r = maximum(eval(model.J)) / maximum(eval(lee.J))
-# kx = 59.02 * r + 1
-# kv = 24.3  * r + 0.3
-# kR = 8.8   * r + 2
-# kO = 1.54  * r + 0.05
-kx = 2.71 + 1.0
-kv = 1.01 + 0.5
-kR = 2.26*Diagonal(@SVector [1,1,1.])
-kO = 0.3*Diagonal(@SVector [1,1,1.])
-x0 = zeros(model)[1]
-u0 = Dynamics.trim_controls(model)
-
-tf = 10.0
-dt = 1e-4
-times = range(0,tf,step=dt)
-Xref = [copy(x0) for k = 1:length(times)]
-Xdref = [copy(x0) for k = 1:length(times)]
-bref = [@SVector [1,0,0.] for k = 1:length(times)]
-
-se3 = SE3Tracking(model, Xref, Xdref, bref, collect(times), kx=kx, kv=kv, kR=kR[1], kO=kO[1])
-cntrl = PlanningWithAttitude.HFCA(model, Xref, Xdref, bref, collect(times), kx=kx, kv=kv, kR=kR, kO=kO)
-get_control(cntrl, x0, 0.0)
-get_control(se3, x0, 0.0)
-
-u = normalize(@SVector randn(3))
-# u = @SVector [1,0,0.]
-qinit = expm(u * deg2rad(170))
-xinit = Dynamics.build_state(model, [0,0.2,0], qinit, randn(3)*0.1, randn(3)*0.2)
-X = simulate(model, cntrl, xinit, tf)
-visualize!(vis, model, X[1:100:length(X)], tf)

### 2020-10-15 basic_benchmarks.jl

@@ -1,301 +1,0 @@
-using TrajectoryOptimization
-using StaticArrays
-using BenchmarkTools, Statistics
-using LinearAlgebra
-using RExLabUtils
-import RExLabUtils: boxplot
-
-import TrajectoryOptimization: ∇rotate, ∇composition1, ∇composition2
-
-# Generate Random Rotations
-N = 1000  # number of rotations
-qs = [rand(UnitQuaternion{Float64}) for i = 1:N]
-ps = [MRP(q) for q in qs]
-es = [RPY(q) for q in qs]
-
-# Generate random vector
-r = @SVector rand(3)
-
-# Make sure multiplication and rotation is working the same for each representation
-@assert qs[2]*qs[1]*r ≈ ps[2]*ps[1]*r ≈ es[2]*es[1]*r
-
-# Statistical method for comparison
-method = median
-
-
-# Benchmark Single Composition
-b1_q = @benchmark $qs[2]*$qs[1]
-b1_p = @benchmark $ps[2]*$ps[1]
-b1_e = @benchmark $es[2]*$es[1]
-judge(method(b1_q), method(b1_p))
-judge(method(b1_q), method(b1_e))
-method(b1_e).time/method(b1_q).time
-method(b1_p).time/method(b1_q).time
-
-# Benchmark Many Compositions
-b2_q = @benchmark foldr(*, $qs)
-b2_p = @benchmark foldr(*, $ps)
-b2_e = @benchmark foldr(*, $es)
-judge(method(b2_q), method(b2_p))
-judge(method(b2_q), method(b2_e))
-method(b2_e).time/method(b2_q).time
-method(b2_p).time/method(b2_q).time
-
-# Benchmark Single Rotation
-b3_q = @benchmark $qs[1]*$r
-b3_p = @benchmark $ps[1]*$r
-b3_e = @benchmark $es[1]*$r
-judge(method(b3_q), method(b3_p))
-judge(method(b3_q), method(b3_e))
-method(b3_e).time/method(b3_q).time
-method(b3_p).time/method(b3_q).time
-
-# Benchmark Many Rotations
-b4_q = @benchmark foldr(*, $qs, init=r)
-b4_p = @benchmark foldr(*, $ps, init=r)
-b4_e = @benchmark foldr(*, $es, init=r)
-judge(method(b4_q), method(b4_p))
-judge(method(b4_q), method(b4_e))
-method(b4_e).time/method(b4_q).time
-method(b4_p).time/method(b4_q).time
-
-
[…]

### 2020-10-15 examples/ipopt_test.jl

@@ -1,203 +1,0 @@
-using TrajectoryOptimization
-using RobotDynamics, Rotations
-using Ipopt
-using LinearAlgebra, StaticArrays
-using ForwardDiff
-using SparseArrays
-using BenchmarkTools
-import RobotZoo.Quadrotor
-const TO = TrajectoryOptimization
-
-## Define the problem
-model = Quadrotor{UnitQuaternion}()
-n,m = size(model)
-N,tf = 51,5.0
-dt = tf/(N-1)
-
-# Objective
-Q = Diagonal(RobotDynamics.fill_state(model, 1,1,10,10.)*1e-3)
-R = Diagonal(SA[1,1,1,1.])
-Qf = (N-1)*Q*1e3
-x0 = zeros(model)[1] 
-xf = RobotDynamics.build_state(model, 
-    RBState([2,1,0.5], expm(SA[0,0,1]*pi/4), zeros(3), zeros(3))
-)
-utrim = zeros(model)[2]
-obj = LQRObjective(Q, R, Qf, xf, N, uf=utrim)
-prob = Problem(model, obj, xf, tf, x0=x0)
-initial_controls!(prob, utrim)
-rollout!(prob)
-initial_controls!(prob, utrim .+ 1)
-J0 = cost(prob)
-
-# Trajectory
-X = states(prob)
-U = controls(prob)
-Z = Float64[]
-for k = 1:N-1
-    append!(Z, X[k])
-    append!(Z, U[k])
-end
-append!(Z, X[N])
-
-## Solve with Altro
-using Altro
-solver = ALTROSolver(prob, show_summary=true)
-solve!(solver)
-
-## Solve with Ipopt
-function dynamics_constraint_structure(n,m,N)
-    P = (N-1)*n
-    NN = N*n + (N-1)*m
-    Gstruct = spzeros(Int,P,NN) 
-    ind1 = 1:n
-    ind2 = 1:n+m
-    blk = collect(reshape(1:(n+m)*n, n, n+m))
-    for k = 1:N-1
-        Gstruct[ind1,ind2] .= blk
-        blk .+= n*(n+m)
-        ind1 = ind1 .+ n
-        ind2 = ind2 .+ (n+m)
-    end
-    ind1 = 1:n
-    ind2 = ind1 .+ (n+m) 
-    idx = (N-1)*n*(n+m)
-    for k = 1:N-1
-        for (i,j) in zip(ind1,ind2)
-            idx += 1
[…]

### 2020-10-15 examples/mlqr.jl

@@ -1,141 +1,0 @@
-
-using TrajectoryOptimization
-import TrajectoryOptimization.Dynamics.trim_controls
-const TO = TrajectoryOptimization
-using PlanningWithAttitude
-using Parameters
-using TrajOptPlots
-using StaticArrays
-using LinearAlgebra
-using MeshCat
-using BenchmarkTools
-using JLD2
-using DataFrames
-import LinearAlgebra: normalize
-import TrajectoryOptimization: state_diff
-import PlanningWithAttitude: generate_ICs, run_MC
-using Plots
-
-model = Dynamics.Quadrotor2{MRP{Float64}}()
-if !isdefined(Main,:vis)
-    vis = Visualizer()
-    open(vis)
-    set_mesh!(vis, model)
-end
-
-percentage(x) = count(x) / length(x)
-
-#---------- Run Monte Carlo Analysis of Local Controllers ----------#
-xmin = @SVector [-1,-1,-1, 1,0,0,0, -5,-5,-5, -5,-5,-5]
-xmax = @SVector [ 1, 1, 1, 1,0,0,0,  5, 5, 5,  5, 5, 5]
-xmin = RBState(xmin)
-xmax = RBState(xmax)
-ICs = generate_ICs(model,xmin,xmax,2000)
-@time data2 = run_MC(ICs, dt_cntrl=0.01, dt=1e-3,
-    types=[RPY{Float64}, CayleyMap, IdentityMap, SE3Tracking])
-
-df = DataFrame(data2)
-df.success = df.term_err .< 0.1
-by(df, :name, :success=>percentage)
-dfs = df[df.success, :]
-by(dfs, :name, (:max_err=>mean, :max_err=>maximum, :avg_err=>mean))
-df[.!df.success,:]
-
-# x0 = zero(RBState)
-# X = test_controller(ExponentialMap, ICs[1], dt_cntrl=0.01, tf=10.0)
-# visualize!(vis, X[1:100:length(X)], tf)
-
-
-# Calculate controller speed
-function get_controls(X, cntrl)
-    for x in X
-        PlanningWithAttitude.get_control(cntrl, x, 0.0)
-    end
-end
-dt = 0.01
-tf = 10
-
-model = Dynamics.Quadrotor2()
-Q = Diagonal(Dynamics.fill_error_state(model, 200,200,50,50.))
-R = Diagonal(@SVector fill(1.0, 4))
[…]

### 2020-10-15 examples/tvlqr.jl

@@ -1,218 +1,0 @@
-using TrajectoryOptimization
-using PlanningWithAttitude
-using StaticArrays
-using LinearAlgebra
-using TrajOptPlots
-using MeshCat
-using DataFrames
-using Random
-using Distributions
-using Plots
-using PGFPlotsX
-using BenchmarkTools
-import TrajectoryOptimization: ExponentialMap
-include("../problems/quadrotor_problems.jl" )
-
-model = Dynamics.Quadrotor2()
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
-end
-delete!(vis)
-set_mesh!(vis, model)
-
-# Solve Quadrotor zig-zag problem
-Rot = MRP{Float64}
-solver = gen_quad_zigzag(Rot, costfun=:Quadratic, use_rot=false)
-solve!(solver)
-visualize!(vis, solver)
-
-# Convert trajectory to general RBState
-Xref = RBState(solver.model, solver.Z)
-Uref = controls(solver)
-push!(Uref, Dynamics.trim_controls(model))
-dt_ref = solver.Z[1].dt
-
-# Track with different model
-Rot = UnitQuaternion{Float64,CayleyMap}
-model = Dynamics.Quadrotor2{Rot}(use_rot=true, mass=.5)
-dt = 1e-4
-Nsim = Int(solver.tf/dt) + 1
-inds = Int.(range(1,Nsim,length=solver.N))
-Q = Diagonal(Dynamics.fill_error_state(model, 200, 200, 50, 50))
-R = Diagonal(@SVector fill(1.0, 4))
-mlqr = TVLQR(model, Q, R, Xref, Uref, dt_ref)
-
-noise = MvNormal(Diagonal(fill(5.,6)))
-model_ = Dynamics.NoisyRB(model, noise)
-X = simulate(model_, mlqr, Xref[1], solver.tf, w=0.0)
-res_lqr = [RBState(model, x) for x in X[inds]]
-norm(res_lqr .⊖ Xref)
-
-visualize!(vis, model, X[1:100:length(X)], solver.tf)
-
-# Track with SE3 controller
-
-# get reference trajectories
-b_ = map(Xref) do x
-    # prefer pointing torwards the goal
-    x.r - Xref[end].r + 0.01*@SVector rand(3)
-end
-X_ = map(Xref) do x
-    Dynamics.build_state(model, x)
-end
-Xd_ = map(1:solver.N) do k
[…]

### 2020-10-15 examples/zig_zag.jl

@@ -1,387 +1,0 @@
-
-using TrajectoryOptimization
-using PlanningWithAttitude
-using Statistics
-using Random
-using StaticArrays
-using LinearAlgebra
-using MeshCat
-using TrajOptPlots
-using DataFrames
-using JLD2
-using PGFPlotsX
-const TO = TrajectoryOptimization
-include("../problems/quadrotor_problems.jl")
-
-model = Dynamics.Quadrotor2{UnitQuaternion{Float64,CayleyMap}}()
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
-    set_mesh!(vis, model)
-end
-
-solver = gen_quad_zigzag(RPY{Float64}, use_rot=false,
-    costfun=:Quadratic, constrained=false, normcon=false)
-solver = gen_quad_zigzag(UnitQuaternion{Float64,IdentityMap}, use_rot=false,
-    costfun=:Quadratic, constrained=false, normcon=false)
-solver = gen_quad_zigzag(UnitQuaternion{Float64,CayleyMap}, use_rot=true,
-    costfun=:ErrorQuad, constrained=false, normcon=false)
-solver.opts.verbose = true
-solve!(solver)
-iterations(solver)
-visualize!(vis, solver)
-
-plot(controls(solver))
-waypoints!(vis, solver.model, solver.Z, length=31)
-
-function log_solve!(solver, data; name=rot_type(solver), kwargs...)
-    b = benchmark_solve!(solver, data; kwargs...)
-    push!(data[:costfun], cost_type(solver))
-    push!(data[:rotation], rot_type(solver))
-    push!(data[:name], name)
-end
-
-function run_all(gen_prob;kwargs...)
-    # Initial data structure
-    data = Dict{Symbol,Vector}(:name=>Symbol[], :costfun=>Symbol[], :iterations=>Int[], :rotation=>Symbol[],
-        :time=>Float64[], :cost=>Float64[])
-
-    # Baseline methods
-    println("Baseline methods")
-    log_solve!(gen_prob(UnitQuaternion{Float64, IdentityMap}, use_rot=false), data; kwargs...)
-    # log_solve!(gen_prob(RodriguesParam{Float64}, use_rot=false), data; kwargs...)
-    # log_solve!(gen_prob(MRP{Float64}, use_rot=false), data; kwargs...)
-    log_solve!(gen_prob(RPY{Float64}, use_rot=false), data; kwargs...)
-
[…]

### 2020-10-15 figs/sphere_plane.jl

@@ -1,19 +1,0 @@
-using MeshCat
-using GeometryTypes
-using Plots
-using CoordinateTransformations
-
-# vis = Visualizer()
-# open(vis)
-
-rad = 0.5f0
-w,h = 1.5,1.5
-point = HyperSphere(Point3f0(0,0,0), 0.01f0)
-ball = HyperSphere(Point3f0(0,0,0), rad)
-plane = HyperRectangle(Vec(-w/2,-h/2,rad), Vec(w,h,0.01))
-s = 0.8
-setobject!(vis["ball"], ball, MeshPhongMaterial(color=RGBA(0.5*s,0.5*s,1*s, 1.0)))
-setobject!(vis["ball"]["plane"], plane, MeshBasicMaterial(color=RGBA(0.7,0,0,1)))
-setobject!(vis["ball"]["point"], point, MeshBasicMaterial(color=RGBA(0,0,0,1)))
-settransform!(vis["ball"], LinearMap(RotX(deg2rad(-30))))
-settransform!(vis["ball"]["point"], Translation(0,0,rad))

### 2020-10-15 problems/quadrotor_problems.jl

@@ -1,197 +1,0 @@
-
-
-function gen_quad_line(Rot; use_rot=Rot<:UnitQuaternion, costfun=:Quadratic)
-    max_con_viol = 1.0e-8
-    T = Float64
-    verbose = true
-
-    opts_ilqr = iLQRSolverOptions{T}(verbose=verbose,
-    cost_tolerance=1e-4,
-    iterations=300)
-
-    opts_al = AugmentedLagrangianSolverOptions{T}(verbose=verbose,
-    opts_uncon=opts_ilqr,
-    iterations=40,
-    cost_tolerance=1.0e-5,
-    cost_tolerance_intermediate=1.0e-4,
-    constraint_tolerance=max_con_viol,
-    penalty_scaling=10.,
-    penalty_initial=1.)
-
-    model = Dynamics.Quadrotor2{Rot}(use_rot=use_rot)
-    # model = Dynamics.LeeQuad(Rot, use_rot=use_rot)
-    # model = Dynamics.BodyFrameLeeQuad(Rot, use_rot=use_rot)
-    n,m = size(model)
-
-    # discretization
-    N = 101 # number of knot points
-    tf = 5.0
-    dt = tf/(N-1) # total time
-
-    # Initial condition
-    x0_pos = @SVector [-10., -10., 1.]
-    x0 = Dynamics.build_state(model, x0_pos, I(UnitQuaternion), zeros(3), zeros(3))
-    utrim = Dynamics.trim_controls(model)
-
-    # Final state
-    xf_pos = @SVector [ 10.,  10., 1.]
-    xf = Dynamics.build_state(model, xf_pos, I(UnitQuaternion), zeros(3), zeros(3))
-
-    # cost
-    costfun == :QuatLQR ? sq = 0 : sq = 1
-    if n == 13
-        rm_quat = @SVector [1,2,3,4,5,6,8,9,10,11,12,13]
-    else
-        rm_quat = @SVector [1,2,3,4,5,6,7,8,9,10,11,12]
-    end
-    Q_diag = Dynamics.fill_state(model, 1e-3, 1e-3*sq, 1e-3, 1e-3)
-    Q = Diagonal(Q_diag)
-    R = Diagonal(@SVector fill(1e-3,m))
-    s_term = 100 # terminal scaling
-
-    if costfun == :QuatLQR
-        cost_nom = QuatLQRCost(Q, R, xf, utrim, w=1e-2)
[…]

### 2020-10-15 quadrotor_barrell.jl

@@ -1,154 +1,0 @@
-
-using TrajectoryOptimization
-using ForwardDiff
-using Random
-using StaticArrays
-using LinearAlgebra
-using MeshCat
-using TrajOptPlots
-using Rotations
-using Plots
-using Quaternions
-import TrajectoryOptimization: Rotation
-const TO = TrajectoryOptimization
-
-max_con_viol = 1.0e-8
-T = Float64
-verbose = true
-
-opts_ilqr = iLQRSolverOptions{T}(verbose=verbose,
-    cost_tolerance=1e-4,
-    iterations=300)
-
-opts_al = AugmentedLagrangianSolverOptions{T}(verbose=verbose,
-    opts_uncon=opts_ilqr,
-    iterations=40,
-    cost_tolerance=1.0e-5,
-    cost_tolerance_intermediate=1.0e-4,
-    constraint_tolerance=max_con_viol,
-    penalty_scaling=10.,
-    penalty_initial=1.)
-
-model = Dynamics.Quadrotor2{UnitQuaternion{Float64,CayleyMap}}()
-# model = Dynamics.Quadrotor2{UnitQuaternion{Float64,MRPMap}}()
-# model = Dynamics.Quadrotor2{UnitQuaternion{Float64,IdentityMap}}()
-n,m = 13,4
-
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
-    set_mesh!(vis, model)
-end
-
-# discretization
-N = 101 # number of knot points
-tf = 5.0
-dt = tf/(N-1) # total time
-
-# Initial condition
-x0 = Dynamics.build_state(model, [0,-1,1], I(UnitQuaternion), zeros(3), zeros(3))
-
-# Final condition
-xf = Dynamics.build_state(model, [0,1,1], -I(UnitQuaternion), zeros(3), zeros(3))
-
-
-# Intermediate frames
-qs = [UnitQuaternion(Quat(RotZ(ang))) for ang in range(0,2pi,length=N)]
-qs = [ExponentialMap(ang*@SVector [0,0,1.]) for ang in range(0,pi,length=N)]
-ys = range(x0[2],xf[2],length=N)
[…]

### 2020-10-15 satellite.jl

@@ -1,61 +1,0 @@
-
-using TrajectoryOptimization
-import TrajectoryOptimization: stage_cost, cost_expansion
-using TrajOptPlots
-const TO = TrajectoryOptimization
-
-# using PlanningWithAttitude
-# import PlanningWithAttitude: cost_gradient, stage_cost, cost_hessian
-# include("visualization.jl")
-using StaticArrays
-using LinearAlgebra
-using BenchmarkTools
-using Plots
-using ForwardDiff
-using Random
-using MeshCat
-
-qref = rand(UnitQuaternion)
-q = rand(UnitQuaternion)
-Q = Diagonal(@SVector fill(1.,3))
-qval = SVector(q)
-
-function mycost(q::SVector{4})
-    dq = CayleyMap(qref\UnitQuaternion(q))
-    0.5*dq'Q*dq
-end
-
-function mycost(g::SVector{3})
-    mycost(SVector(q*CayleyMap(g)))
-end
-
-mycost(qval)
-g = @SVector zeros(3)
-mycost(g)
-
-grad = ForwardDiff.gradient(mycost,g)
-dq = qref\q
-err = CayleyMap(dq)
-err'Q*TO.jacobian(CayleyMap,qref\q)*Lmult(qref)'Lmult(q)*Vmat()' ≈ grad'
-Vmat()*Lmult(dq)'TO.jacobian(CayleyMap,dq)'Q*err ≈ grad
-
-
-hess = ForwardDiff.hessian(mycost,g)
-hess1 = Vmat()*Lmult(dq)'TO.jacobian(CayleyMap,dq)'Q*TO.jacobian(CayleyMap,dq)*Lmult(dq)*Vmat()'
-hess2 = Vmat()*Lmult(dq)'TO.∇jacobian(CayleyMap,dq,Q*err)*Lmult(dq)*Vmat()'
-hess1+hess2 ≈ hess
-
-Qω = Diagonal(@SVector fill(3.,3))
-R = Diagonal(@SVector fill(2.,3))
-ωref = @SVector rand(3)
-cost1 = SatDiffCost(model, Qω, Q, R, SVector(qref), ωref)
-
-ω = @SVector rand(3)
-x = [ωref; qval]
-u = @SVector zeros(3)
-z = KnotPoint(x,u,0.1)
-G = TO.state_diff_jacobian(model,x)
-TO.stage_cost(cost1, x, u)
[…]

### 2020-10-15 satellite_360.jl

@@ -1,231 +1,0 @@
-using TrajectoryOptimization
-import TrajectoryOptimization: stage_cost, cost_expansion
-using TrajOptPlots
-const TO = TrajectoryOptimization
-
-# using PlanningWithAttitude
-# import PlanningWithAttitude: cost_gradient, stage_cost, cost_hessian
-# include("visualization.jl")
-using StaticArrays
-using LinearAlgebra
-using BenchmarkTools
-using Plots
-using ForwardDiff
-using Random
-using MeshCat
-
-
-
-# Dynamics
-model = Dynamics.Satellite2()
-N = 101
-tf = 5.0
-
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
-    set_mesh!(vis, model)
-end
-
-# Objective
-struct SatCost <: CostFunction
-    Q::Diagonal{Float64,SVector{3,Float64}}
-    R::Diagonal{Float64,SVector{3,Float64}}
-    q::SVector{3,Float64}
-    q0::SVector{4,Float64}
-    Qq::Float64
-    c::Float64
-end
-
-function SatCost(Q::Diagonal,R::Diagonal,q0::SVector, Qq=0.0, ω0=@SVector zeros(3))
-    q = -Q*ω0
-    c = 0.5*ω0'Q*ω0
-    SatCost(Q,R,q,q0,Qq,c)
-end
-
-function stage_cost(cost::SatCost, x::SVector, u::SVector=(@SVector zeros(3)))
-    ω = @SVector [x[1], x[2], x[3]]
-    q = @SVector [x[4], x[5], x[6], x[7]]
-    J = 0.5*(ω'cost.Q*ω + u'cost.R*u) + cost.q'ω + cost.c
-    d = cost.q0'q
-    if d ≥ 0
-        J += cost.Qq*(1-d)
-    else
-        J += cost.Qq*(1+d)
-    end
-    return J
-end
-
-
-function cost_expansion(cost::SatCost, model::Dynamics.Satellite, z::KnotPoint{T,N,M}, G) where {T,N,M}
-    Q = cost.Q
-    q0 = cost.q0
-    u = control(z)
-    ω = @SVector [z.z[1], z.z[2], z.z[3]]
-    q = @SVector [z.z[4], z.z[5], z.z[6], z.z[7]]
[…]

### 2020-10-15 satellite_tests.jl

@@ -1,406 +1,0 @@
-using TrajectoryOptimization
-const TO = TrajectoryOptimization
-using TrajOptPlots
-
-using JLD
-using DataFrames
-using StaticArrays
-using LinearAlgebra
-using BenchmarkTools
-using Plots
-using ForwardDiff
-using Random
-using MeshCat
-using Statistics
-import LinearAlgebra: normalize
-
-# Dynamics
-model = Dynamics.Satellite2()
-size(model)
-x,u = rand(model)
-dynamics(model,x,u)
-N = 101
-tf = 5.0
-
-if !isdefined(Main,:vis)
-    vis = Visualizer(); open(vis);
-    set_mesh!(vis, model)
-end
-
-angwrap(x) = min(x,360-x)
-
-function final_angle_error(solver)
-    q = Dynamics.orientation(solver.model, state(solver.Z[end]))
-    qf = Dynamics.orientation(solver.model, solver.xf)
-    rad2deg(angle(qf\q))
-end
-
-function total_angle(solver)
-    θ = 0.0
-    X = states(solver)
-    model = solver.model
-    for k = 1:solver.N-1
-        q1 = Dynamics.orientation(model, X[k])
-        q2 = Dynamics.orientation(model, X[k+1])
-        θ += angle(q2\q1)
-    end
-    return rad2deg(θ)
-end
-
-
-function gen_sat_prob(Rot=UnitQuaternion{Float64,CayleyMap};
-        ang=deg2rad(90), cost=:QuatLQR, use_rot=true)
-    # Model
-    model = Dynamics.Satellite2{Rot}(use_rot=use_rot)
-    rot_dim = Rot <: UnitQuaternion ? 4 : 3
-    N = 101
-    tf = 5.0
-
-    # Objective
-    Q_diag = @SVector [1e-2,1e-2,1e-2, 0e-2,0e-2,0e-2,0e-2]
-    R_diag = @SVector fill(1e-3,3)
-    Q_rot = @SVector fill(1e-1, rot_dim)
-    θ = ang
-    u = normalize(-1 .+ 2*@SVector rand(3))
-    # u = @SVector [1,0,0.]
-    iω = @SVector [1,2,3]
-    iq = @SVector [4,5,6,7]
-    ωf = @SVector zeros(3)
-    qf = expm(θ*u)
-    xf = Dynamics.build_state(model, ωf, qf)
-
-    costs = map(1:N) do k
[…]

### 2020-10-15 src/ilqr/attitude_costs.jl

@@ -1,120 +1,0 @@
-export
-    SatCost
-
-struct SatCost <: CostFunction
-    Q::Diagonal{Float64,SVector{3,Float64}}
-    R::Diagonal{Float64,SVector{3,Float64}}
-    q::SVector{3,Float64}
-    q0::SVector{4,Float64}
-    Qq::Float64
-    c::Float64
-end
-
-function SatCost(Q::Diagonal,R::Diagonal,q0::SVector, Qq=0.0, ω0=@SVector zeros(3))
-    q = -Q*ω0
-    c = 0.5*ω0'Q*ω0
-    SatCost(Q,R,q,q0,Qq,c)
-end
-
-stage_cost(cost::SatCost, z::KnotPoint) = stage_cost(cost, state(z), control(z))
-function stage_cost(cost::SatCost, x::SVector, u::SVector)
-    ω = @SVector [x[1], x[2], x[3]]
-    q = @SVector [x[4], x[5], x[6], x[7]]
-    J = 0.5*(ω'cost.Q*ω + u'cost.R*u) + cost.q'ω + cost.c
-    d = cost.q0'q
-    if d ≥ 0
-        J += cost.Qq*(1-d)
-    else
-        J += cost.Qq*(1+d)
-    end
-    return J
-end
-
-function cost_gradient(solver::iLQRSolver, cost::SatCost, z::KnotPoint)
-    if is_terminal(z)
-        dt_x = 1.0
-        dt_u = 0.0
-    else
-        dt_x = z.dt
-        dt_u = z.dt
-    end
-    Q = cost.Q
-    u = control(z)
-    ω = @SVector [z.z[1], z.z[2], z.z[3]]
-    q = @SVector [z.z[4], z.z[5], z.z[6], z.z[7]]
-    d = cost.q0'q
-    Qω = Q*ω + cost.q
-    if d ≥ 0
-        Qq = -cost.Qq*cost.q0'Lmult(q)*Vmat()'
-    else
-        Qq =  cost.Qq*cost.q0'Lmult(q)*Vmat()'
-    end
-    Qx = [Qω; Qq']
-    Qu = cost.R*u
-    return Qx*dt_x, Qu*dt_u
-end
-
[…]

### 2020-10-15 src/ilqr/ilqr.jl

@@ -1,337 +1,0 @@
-export
-	solve!,
-	rollout!
-
-############################################################################################
-#                                   SOLVE
-############################################################################################
-function solve!(solver::iLQRSolver)
-    set_verbosity!(solver.opts)
-    clear_cache!(solver.opts)
-
-    solver.stats.iterations = 0
-    solver.ρ[1] = solver.opts.bp_reg_initial
-    solver.dρ[1] = 0.0
-    Z = solver.Z; Z̄ = solver.Z̄;
-
-    n,m,N = size(solver)
-    J = Inf
-    _J = get_J(solver.obj)
-
-    # Initial rollout
-    rollout!(solver)
-
-    cost!(solver.obj, Z)
-    J_prev = sum(_J)
-
-
-    for i = 1:solver.opts.iterations
-		ΔV = backwardpass!(solver)
-		J = forwardpass!(solver, ΔV, J_prev)
-
-        # check for cost blow up
-        if J > solver.opts.max_cost_value
-            # @warn "Cost exceeded maximum cost"
-            return solver
-        end
-
-        for k = 1:N
-            Z[k].z = Z̄[k].z
-        end
-
-        dJ = abs(J - J_prev)
-        J_prev = copy(J)
-        # gradient_todorov!(solver)
-
-        record_iteration!(solver, J, dJ)
-        evaluate_convergence(solver) ? break : nothing
-    end
-    return solver
-end
-
-"""
-Stash iteration statistics
-"""
-function record_iteration!(solver::iLQRSolver, J, dJ)
-    solver.stats.iterations += 1
-    i = solver.stats.iterations::Int
-    solver.stats.cost[i] = J
-    solver.stats.dJ[i] = dJ
-    # solver.stats.gradient[i] = mean(solver.grad)
-
-    @logmsg InnerLoop :iter value=i
-    @logmsg InnerLoop :cost value=J
-    @logmsg InnerLoop :dJ   value=dJ
-    @logmsg InnerLoop :grad value=solver.stats.gradient[i]
-    # @logmsg InnerLoop :zero_count value=solver.stats[:dJ_zero_counter][end]
-    if solver.opts.verbose
-        print_level(InnerLoop)
-    end
-    return nothing
-end
-
-"""
-    Calculate the problem gradient using heuristic from iLQG (Todorov) solver
-"""
[…]

### 2020-10-15 src/ilqr/ilqr_solver.jl

@@ -1,155 +1,0 @@
-export
-    iLQRSolverOptions,
-    iLQRSolver,
-    initial_controls!,
-    states,
-    controls
-
-@with_kw mutable struct iLQRStats{T}
-    iterations::Int = 0
-    cost::Vector{T} = [0.]
-    dJ::Vector{T} = [0.]
-    gradient::Vector{T} = [0.]
-    dJ_zero_counter::Int = 0
-end
-
-function reset!(stats::iLQRStats, N=0)
-    stats.iterations = 0
-    stats.cost = zeros(N)
-    stats.dJ = zeros(N)
-    stats.gradient = zeros(N)
-    stats.dJ_zero_counter = 0
-end
-
-
-@with_kw mutable struct iLQRSolverOptions{T}
-    # Options
-
-    "Print summary at each iteration."
-    verbose::Bool=false
-
-    "Live plotting."
-    live_plotting::Symbol=:off # :state, :control
-
-    "dJ < ϵ, cost convergence criteria for unconstrained solve or to enter outerloop for constrained solve."
-    cost_tolerance::T = 1.0e-4
-
-    "gradient type: :todorov, :feedforward."
-    gradient_type::Symbol = :todorov
-
-    "gradient_norm < ϵ, gradient norm convergence criteria."
-    gradient_norm_tolerance::T = 1.0e-5
-
-    "iLQR iterations."
-    iterations::Int = 300
-
-    "restricts the total number of times a forward pass fails, resulting in regularization, before exiting."
-    dJ_counter_limit::Int = 10
-
-    "use square root method backward pass for numerical conditioning."
-    square_root::Bool = false
-
-    "forward pass approximate line search lower bound, 0 < line_search_lower_bound < line_search_upper_bound."
-    line_search_lower_bound::T = 1.0e-8
-
-    "forward pass approximate line search upper bound, 0 < line_search_lower_bound < line_search_upper_bound < ∞."
-    line_search_upper_bound::T = 10.0
-
-    "maximum number of backtracking steps during forward pass line search."
-    iterations_linesearch::Int = 20
-
-    # Regularization
-    "initial regularization."
-    bp_reg_initial::T = 0.0
-
-    "regularization scaling factor."
-    bp_reg_increase_factor::T = 1.6
-
[…]

### 2020-10-15 src/ilqr/knotpoint.jl

@@ -1,122 +1,0 @@
-export
-    KnotPoint,
-    Traj,
-    state,
-    control,
-    set_states!,
-    set_controls!
-
-"""
-Stores critical information corresponding to each knot point in the trajectory optimization
-problem, including the state and control values, as well as the time and time step length.
-
-# Getters
-Use the following methods to access values from a `KnotPoint`:
-```julia
-x  = state(z::KnotPoint)    # returns the n-dimensional state as a SVector
-u  = control(z::KnotPoint)  # returns the m-dimensional control vector as a SVector
-t  = z.t                    # current time
-dt = z.dt                   # time step length
-```
-
-# Setters
-Use the following methods to set values in a `KnotPoint`:
-```julia
-set_state!(z::KnotPoint, x)
-set_control!(z::KnotPoint, u)
-z.t = t
-z.dt = dt
-```
-
-# Constructors
-```julia
-KnotPoint(x, u, dt, t=0.0)
-KnotPoint(x, m, t=0.0)  # for terminal knot point
-```
-
-Use `is_terminal(z::KnotPoint)` to determine if a `KnotPoint` is a terminal knot point (e.g.
-has no time step length and z.t == tf).
-"""
-mutable struct KnotPoint{T,N,M,NM}
-    z::SVector{NM,T}
-    _x::SVector{N,Int}
-    _u::SVector{M,Int}
-    dt::T # time step
-    t::T  # total time
-end
-
-function KnotPoint(x::AbstractVector, u::AbstractVector, dt::Float64, t=0.0)
-    n = length(x)
-    m = length(u)
-    xinds = ones(Bool, n+m)
-    xinds[n+1:end] .= 0
-    _x = SVector{n}(1:n)
-    _u = SVector{m}(n .+ (1:m))
-    z = SVector{n+m}([x;u])
-    KnotPoint(z, _x, _u, dt, t)
-end
-
-# Constructor for terminal time step
-function KnotPoint(x::AbstractVector, m::Int, t=0.0)
-    u = zeros(m)
-    KnotPoint(x, u, 0., t)
-end
-
-@inline state(z::KnotPoint) = z.z[z._x]
-@inline control(z::KnotPoint) = z.z[z._u]
[…]

### 2020-10-15 src/ilqr/logger.jl

@@ -1,321 +1,0 @@
-using Logging
-
-"Log Level for outer loop iterations (e.g. Augmented Lagrangian). LogLevel(-100)"
-const OuterLoop = LogLevel(-100)
-"Log Level for inner loop iterations (e.g. iLQR). LogLevel(-200)"
-const InnerLoop = LogLevel(-200)
-"Log Level for internal solve methods (e.g. forward pass for iLQR). LogLevel(-500)"
-const InnerIters = LogLevel(-500)
-
-function default_logger(verbose::Bool)
-    verbose == false ? min_level = Logging.Warn : min_level = InnerLoop
-
-    logger = SolverLogger(min_level)
-    inner_cols = [:iter, :cost, :expected, :z, :α, :info]
-    inner_widths = [5,     14,      12,    10, 10,    50]
-    outer_cols = [:iter, :total, :c_max, :info]
-    outer_widths = [6,          7,        12,        50]
-    add_level!(logger, InnerLoop, inner_cols, inner_widths, print_color=:green,indent=4)
-    add_level!(logger, OuterLoop, outer_cols, outer_widths, print_color=:yellow,indent=0)
-    return logger
-end
-
-
-"""
-Holds logging information about a particular print level, meant to assemble
-a table of output where `cols` gives the order and names of columns, `widths`
-are the column widths, and `print` will turn on/off printing the column.
-
-All values can be cached at any moment in time to accumulate a history of the
-data.
-
-# Constructors
-```julia
-LogData(metadata::NamedTuple=(color=:default, header_frequency=10, indent=0))
-LogData(cols, widths; do_print=trues(length(cols)), vartypes=fill(Any, length(cols)),
-    color=:default, header_frequency=10, indent=0)
-```
-"""
-mutable struct LogData
-    cols::Vector{Symbol}
-    widths::Vector{Int}
-    print::BitArray{1}
-    data::Dict{Symbol,Any}
-    cache::Dict{Symbol,Vector}
-    freq::Int
-    color::Symbol
-    indent::Int
-    function LogData(cols,widths,print::BitArray{1},data::Dict{Symbol,Any},
-            cache::Dict{Symbol,Vector}, freq::Int,color::Symbol,indent::Int)
-        new(cols,widths,print,data,cache,freq,color,indent)
-    end
-end
-
-Base.getindex(ldata::LogData,index::Symbol) = ldata.data[index]
-
[…]

### 2020-10-15 src/ilqr/models.jl

@@ -1,84 +1,0 @@
-export
-	AbstractModel,
-    Satellite,
-	dynamics,
-	discrete_dynamics,
-	discrete_jacobian,
-	dynamics_jacobian,
-	orientation
-
-abstract type AbstractModel end
-
-@inline Base.position(::AbstractModel, x::SVector) = @SVector [x[1], x[2], x[3]]
-@inline orientation(::AbstractModel, x::SVector) = @SVector [x[4], x[5], x[6], x[7]]
-
-
-function discrete_dynamics(model::AbstractModel, x::SVector{N,T}, u::SVector{M,T},
-		dt::T) where {N,M,T}
-    k1 = dynamics(model, x, u)*dt;
-    k2 = dynamics(model, x + k1/2, u)*dt;
-    k3 = dynamics(model, x - k1 + 2*k2, u)*dt;
-    x + (k1 + 4*k2 + k3)/6
-end
-@inline discrete_dynamics(model::AbstractModel, z::KnotPoint) =
-	discrete_dynamics(model, state(z), control(z), z.dt)
-
-function discrete_jacobian(model::AbstractModel,
-        z::KnotPoint{T,N,M,NM}) where {T,N,M,NM}
-    ix,iu,idt = z._x, z._u, NM+1
-    fd_aug(s) = discrete_dynamics(model, s[ix], s[iu], s[idt])
-    s = [z.z; @SVector [z.dt]]
-    ForwardDiff.jacobian(fd_aug, s)
-end
-
-
-struct Satellite <: AbstractModel
-    J::Diagonal{Float64,SVector{3,Float64}}
-end
-
-Satellite() = Satellite(Diagonal(@SVector ones(3)))
-
-Base.size(::Satellite) = 7,3
-Base.position(::Satellite, x::SVector) = @SVector zeros(3)
-
-function dynamics(model::Satellite, x::SVector, u::SVector)
-    ω = @SVector [x[1], x[2], x[3]]
-    q = normalize(@SVector [x[4], x[5], x[6], x[7]])
-    J = model.J
-
-    ωdot = J\(u - ω × (J*ω))
-    qdot = 0.5*Lmult(q)*Vmat()'ω
-    return [ωdot; qdot]
-end
-
-function state_diff(model::Satellite, x::SVector, x0::SVector)::SVector{6}
-    ω = @SVector [x[1], x[2], x[3]]
-    q = @SVector [x[4], x[5], x[6], x[7]]
[…]

### 2020-10-15 src/ilqr/objective.jl

@@ -1,84 +1,0 @@
-export
-    Objective,
-    QuadraticCost,
-    LQRCost,
-    cost
-
-abstract type CostFunction end
-
-struct Objective{C<:CostFunction}
-    costs::Vector{C}
-    J::Vector{Float64}
-end
-
-function Objective(costs::Vector{<:CostFunction})
-    N = length(costs)
-    return Objective(costs, zeros(N))
-end
-
-function Objective(cost::CostFunction, N::Int)
-    costs = [cost for k = 1:N]
-    return Objective(costs, zeros(N))
-end
-
-function Objective(cost::CostFunction, costterm::CostFunction, N::Int)
-    costs = [cost for k = 1:N-1]
-    push!(costs, costterm)
-    return Objective(costs, zeros(N))
-end
-
-@inline get_J(obj::Objective) = obj.J
-
-
-function cost!(obj::Objective, Z::Traj)
-    N = length(Z)
-    J = obj.J
-    for k in eachindex(obj.costs)
-        k == N ? dt = 1.0 : dt = Z[k].dt
-        J[k] = stage_cost(obj.costs[k], Z[k])*dt
-    end
-end
-
-function cost(obj::Objective, Z::Traj)
-    cost!(obj, Z)
-    return sum(obj.J)
-end
-
-
-struct QuadraticCost{N,M} <: CostFunction
-    Q::Diagonal{Float64,SVector{N,Float64}}
-    R::Diagonal{Float64,SVector{M,Float64}}
-    q::SVector{N,Float64}
-    r::SVector{M,Float64}
-    c::Float64
-end
-
-function stage_cost(cost::QuadraticCost, z::KnotPoint)
-    x,u = state(z), control(z)
-    return 0.5*(x'cost.Q*x + u'cost.R*u) + cost.q'x + cost.r'u + cost.c
-end
-
-function gradient(cost::QuadraticCost, z::KnotPoint)
-    x,u = state(z), control(z)
-    Qx = cost.Q*x + cost.q
-    Qu = cost.R*u + cost.r
-    return Qx, Qu
-end
-
-function hessian(cost::QuadraticCost, z::KnotPoint{T,N,M}) where {T,N,M}
-    x,u = state(z), control(z)
-    Qxx = cost.Q
-    Quu = cost.R
-    Qux = @SMatrix zeros(M,N)
-    return Qxx, Quu, Qux
-end
-
[…]

### 2020-10-15 src/ilqr/quaternions.jl

@@ -1,49 +1,0 @@
-export
-    cayley_map,
-    cayley_jacobian,
-    Lmult,
-    Rmult,
-    Vmat,
-    Tmat
-
-function Lmult(q::SVector{4})
-    @SMatrix [q[1] -q[2] -q[3] -q[4];
-              q[2]  q[1] -q[4]  q[3];
-              q[3]  q[4]  q[1] -q[2];
-              q[4] -q[3]  q[2]  q[1]]
-end
-
-function Rmult(q::SVector{4})
-    @SMatrix [q[1] -q[2] -q[3] -q[4];
-              q[2]  q[1]  q[4] -q[3];
-              q[3] -q[4]  q[1]  q[2];
-              q[4]  q[3] -q[2]  q[1]]
-end
-
-function Vmat()
-    @SMatrix [
-        0 1 0 0;
-        0 0 1 0;
-        0 0 0 1.
-    ]
-end
-
-function Tmat()
-    @SMatrix [
-        1  0  0  0;
-        0 -1  0  0;
-        0  0 -1  0;
-        0  0  0 -1;
-    ]
-end
-
-function cayley_map(g::SVector{3})
-    M = 1/sqrt(1+g'g)
-    @SVector [M, M*g[1], M*g[2], M*g[3]]
-end
-
-function cayley_jacobian(g::SVector{3})
-    n = 1+g'g
-    ni = 1/n
-    [-g'; -g*g' + I*n]*ni*sqrt(ni)
-end

### 2020-10-15 src/lqr_monte_carlo.jl

@@ -1,94 +1,0 @@
-
-function test_controller(Rot, x0::RBState; dt=1e-4, tf=10.0, dt_cntrl=dt)
-    model = Dynamics.Quadrotor2{UnitQuaternion{Float64,CayleyMap}}()
-    Q_ = (200.,200,50,50)
-    R = Diagonal(@SVector fill(1.0,4))
-    xref = zeros(model)[1]
-    uref = trim_controls(model)
-    if Rot <: TrajectoryOptimization.DifferentialRotation
-        model = Dynamics.Quadrotor2{UnitQuaternion{Float64,Rot}}()
-        Q = Diagonal(Dynamics.fill_error_state(model, Q_...))
-        cntrl = MLQR(model, dt_cntrl, Q, R, xref, uref)
-        xinit = Dynamics.build_state(model, x0)
-        res = simulate(model, cntrl, xinit, tf, dt=dt, w=0.0)
-    elseif Rot <: Rotation
-        model = Dynamics.Quadrotor2{Rot}(use_rot=false)
-        Q = Diagonal(Dynamics.fill_error_state(model, Q_...))
-        xref = zeros(model)[1]
-        cntrl = LQR(model, dt_cntrl, Q, R, xref, uref)
-        xinit = Dynamics.build_state(model, x0)
-        try
-            res = simulate(model, cntrl, xinit, tf, dt=dt, w=0.0)
-        catch
-            N = tf/dt + 1
-            res = [zero(xref)*NaN for i = 1:N]
-        end
-    elseif Rot == SE3Tracking
-        Rot = UnitQuaternion{Float64,IdentityMap}
-        model = Dynamics.Quadrotor2{Rot}(use_rot=false)
-        xref = zeros(model)[1]
-
-        # Build reference trajectory
-        times = range(0,tf,step=dt)
-        Xref = [copy(xref) for k = 1:length(times)]
-        bref = [@SVector [1,0,0.] for k = 1:length(times)]
-        Xdref = [@SVector zeros(13) for k = 1:length(times)]
-
-        # cntrl = SE3Tracking(model, Xref, Xdref, bref, collect(times))
-        cntrl = HFCA(model, Xref, Xdref, bref, collect(times))
-        xinit = Dynamics.build_state(model, x0)
-        res = simulate(model, cntrl, xinit, tf, dt=dt, w=0.0)
-    end
-    res = [RBState(model,x) for x in res]
-    return res
[…]

### 2020-10-15 src/plotting.jl

@@ -1,110 +1,0 @@
-export
-    rot_type,
-    cost_type,
-    short_names,
-    bar_comparison
-
-using PGFPlotsX
-using Plots
-using TrajOptPlots
-using MeshCat
-using CoordinateTransformations
-
-function TrajOptPlots.visualize!(vis, X::Vector{<:RBState}, tf)
-    fps = Int(floor(length(X)/tf))
-    anim = MeshCat.Animation(fps)
-    for k in eachindex(X)
-        atframe(anim, k) do
-            x = X[k].r
-            q = X[k].q
-            settransform!(vis["robot"], compose(Translation(x), LinearMap(Quat(q))))
-        end
-    end
-    setanimation!(vis, anim)
-    return anim
-end
-
-function rot_type(solver::TrajectoryOptimization.AbstractSolver)
-    rot_type(Dynamics.rotation_type(get_model(solver)))
-end
-
-function rot_type(Rot)
-    if Rot <: UnitQuaternion
-        return Symbol(retraction_map(Rot))
-    elseif Rot <: TrajectoryOptimization.DifferentialRotation
-        return Symbol(Rot)
-    elseif Rot <: MRP
-        return :MRP
-    elseif Rot <: RodriguesParam
-        return :RP
-    elseif Rot <: RPY
-        return :RPY
-    elseif Rot == SE3Tracking
-        return :SE3
-    end
-end
-
-function cost_type(solver)
-    obj = TO.get_objective(solver)
-    if obj isa TrajectoryOptimization.ALObjective
-        costfun = typeof(obj.obj[1])
-    else
-        costfun = typeof(obj[1])
-    end
-    if costfun <: QuadraticCost
-        return :Quadratic
-    elseif costfun <: SatDiffCost
-        return :SatDiff
-    elseif costfun <: QuadraticQuatCost
-        return :QuatLQR
-    elseif costfun <: ErrorQuadratic
-        return :ErrorQuadratic
-    end
-end
-
-function short_names(s::Symbol)
-    if s == :ExponentialMap
-        :Exp
-    elseif s == :CayleyMap
-        :Cay
-    elseif s == :MRPMap
-        :dMRP
-    elseif s == :VectorPart
-        :Vec
-    elseif s == :IdentityMap
-        :Quat
-    elseif s == :ReNorm
-        :Norm
-    elseif s == :NormCon
-        :Con
-    elseif s == :QuatSlack
-        :Slack
-    else
-        s
-    end
-end
[…]

### 2020-10-15 src/rbstate.jl

@@ -1,101 +1,0 @@
-export
-    RBState,
-    randbetween
-
-struct RBState{T}
-    r::SVector{3,T}
-    q::UnitQuaternion{T,TO.DEFAULT_QUATDIFF}
-    v::SVector{3,T}
-    ω::SVector{3,T}
-end
-
-function RBState(r::AbstractVector, q::Rotation, v::AbstractVector, ω::AbstractVector)
-    r_ = @SVector [r[1],r[2],r[3]]
-    q_ = UnitQuaternion{TO.DEFAULT_QUATDIFF}(q)
-    v_ = @SVector [v[1],v[2],v[3]]
-    ω_ = @SVector [ω[1],ω[2],ω[3]]
-    RBState(r_, q_, v_, ω_)
-end
-
-function RBState(x::SVector{13})
-    r_ = @SVector [x[1],x[2],x[3]]
-    q_ = UnitQuaternion(x[4], x[5], x[6], x[7])
-    v_ = @SVector [x[8],x[9],x[10]]
-    ω_ = @SVector [x[11],x[12],x[13]]
-    RBState(r_, q_, v_, ω_)
-end
-
-function RBState(model::RigidBody, x::SVector)
-    r,q,v,ω = Dynamics.parse_state(model, x)
-    RBState(r,UnitQuaternion(q),v,ω)
-end
-
-function RBState(model::RigidBody, Z::Traj)
-    [RBState(model, state(z)) for z in Z]
-end
-
-function Dynamics.build_state(model::RigidBody{R}, rbs::RBState) where R
-    Dynamics.build_state(model, rbs.r, rbs.q, rbs.v, rbs.ω)
-end
-
-function Base.:+(s1::RBState, s2::RBState)
-    RBState(s1.r+s2.r, s1.q*s2.q, s1.v+s2.v, s1.ω+s2.ω)
-end
-
-function Base.:-(s1::RBState, s2::RBState)
-    RBState(s1.r-s2.r, s2.q\s1.q, s1.v-s2.v, s1.ω-s2.ω)
-end
-
-function TrajectoryOptimization.:⊖(s1::RBState, s2::RBState, rmap=CayleyMap)
-    dx = s1.r-s2.r
-    dq = rmap(s2.q\s1.q)
-    dv = s1.v-s2.v
-    dw = s1.ω-s2.ω
-    @SVector [dx[1], dx[2], dx[3], dq[1], dq[2], dq[3],
[…]

### 2020-10-15 src/tracking_control.jl

@@ -1,385 +1,0 @@
-export
-    get_control,
-    TVLQR,
-    LQR,
-    MLQR,
-    simulate,
-    SE3Tracking,
-    HFCA,
-    test_ICs
-
-import TrajectoryOptimization.Dynamics: trim_controls, build_state
-abstract type AbstractController end
-abstract type LQRController <: AbstractController end
-abstract type TimeVaryingController <: AbstractController end
-abstract type TrackingController <: TimeVaryingController end
-
-import TrajectoryOptimization: state_diff, state_diff_jacobian
-
-function get_k(cntrl::TimeVaryingController, t)
-    times = get_times(cntrl)
-    searchsortedlast(times,t)
-end
-
-
-""" TVLQR
-"""
-struct TVLQR{L,TK} <: TimeVaryingController
-    model::L
-    Z::Traj
-    K::Vector{TK}
-    t::Vector{Float64}
-end
-
-""" Assumes the states of the trajectory Z match the states of the model
-"""
-function TVLQR(model::AbstractModel, Q, R, Z::Traj)
-    A,B = linearize(model, Z)
-    K = tvlqr(A,B,Q,R)
-    t = [z.t for z in Z]
-    TVLQR(model,Z,K,t)
-end
-
-function TVLQR(model::RigidBody, Q, R, X::Vector{<:RBState}, U::Vector{<:AbstractVector}, dt)
-    # Convert the trajectory to the orientation representation of the model
-    Z = Traj(model, X, U, dt)
-
-    # Call as normal
-    TVLQR(model, Q, R, Z::Traj)
-end
-
-@inline get_times(cntrl::TVLQR) = cntrl.t
-
-function get_control(cntrl::TVLQR, x, t)
-    k = get_k(cntrl, t)
-    zref = cntrl.Z[k]
-    dx = state_diff(cntrl.model, x, state(zref))
-    return cntrl.K[k]*dx + control(zref)
-end
-
-function tvlqr(A, B, Q, R)
-    n,m = size(B[1])
-    K = [@SMatrix zeros(m,n) for k = 1:length(A)]
-    tvlqr!(K, A, B, Q, R)
-end
-
-function tvlqr!(K, A, B, Q, R)
-    N = length(A)
-
[…]

### 2020-10-15 src/vecquat.jl

@@ -1,335 +1,0 @@
-struct VecQuat{T} <: Rotation{3,T}
-    w::T
-    x::T
-    y::T
-    z::T
-
-    @inline function VecQuat{T}(w, x, y, z, normalize::Bool = true) where T
-        if normalize
-            inorm = inv(sqrt(w*w + x*x + y*y + z*z))
-            new{T}(w*inorm, x*inorm, y*inorm, z*inorm)
-        else
-            new{T}(w, x, y, z)
-        end
-    end
-
-    VecQuat{T}(q::VecQuat) where T = new{T}(q.w, q.x, q.y, q.z)
-end
-
-# ~~~~~~~~~~~~~~~ Constructors ~~~~~~~~~~~~~~~ #
-# Use default map
-function VecQuat(w,x,y,z, normalize::Bool = true)
-    types = promote(w,x,y,z)
-    VecQuat{eltype(types)}(w,x,y,z, normalize)
-end
-
-# Pass in Vectors
-@inline function (::Type{Q})(q::AbstractVector, normalize::Bool = true) where Q <: VecQuat
-    check_length(q, 4)
-    Q(q[1], q[2], q[3], q[4], normalize)
-end
-@inline (::Type{Q})(q::StaticVector{4}, normalize::Bool = true) where Q <: VecQuat =
-    Q(q[1], q[2], q[3], q[4], normalize)
-
-# Copy constructors
-VecQuat(q::VecQuat) = q
-
-# Convert from UnitQuaternion
-(::Type{Q})(q::UnitQuaternion) where Q <: VecQuat = Q(q.w, q.x, q.y, q.z)
-(::Type{Q})(q::Vecquat) where Q <: UnitQuatenrion = Q(q.w, q.x, q.y, q.z)
-
-# ~~~~~~~~~~~~~~~ StaticArrays Interface ~~~~~~~~~~~~~~~ #
-function (::Type{Q})(t::NTuple{9}) where Q<:VecQuat
-    #=
-    This function solves the system of equations in Section 3.1
-    of https://arxiv.org/pdf/math/0701759.pdf. This cheap method
-    only works for matrices that are already orthonormal (orthogonal
-    and unit length columns). The nearest orthonormal matrix can
-    be found by solving Wahba's problem:
[…]

### 2020-10-15 test_controllers.jl

@@ -1,131 +1,0 @@
-using TrajectoryOptimization
-import TrajectoryOptimization.Dynamics.trim_controls
-const TO = TrajectoryOptimization
-using Parameters
-using TrajOptPlots
-using StaticArrays
-using LinearAlgebra
-using MeshCat
-using BenchmarkTools
-using JLD2
-using DataFrames
-using PlanningWithAttitude
-import LinearAlgebra: normalize
-
-
-
-############################################################################################
-#                             SE3 QUADROTOR CONTROLLER
-############################################################################################
-Rot = UnitQuaternion{Float64,CayleyMap}
-model = Dynamics.LeeQuad()
-kx = 59.02
-kv = 24.3
-kR = 8.8
-kO = 1.54
-x0 = zeros(model)[1]
-u0 = Dynamics.trim_controls(model)
-
-tf = 10.0
-dt = 1e-4
-times = range(0,tf,step=dt)
-Xref = [copy(x0) for k = 1:length(times)]
-bref = [@SVector [1,0,0.] for k = 1:length(times)]
-
-cntrl = SE3Tracking(model, Xref, bref, collect(times), kx=kx, kv=kv, kR=kR, kO=kO)
-r = @SVector [1,0,0.]
-qinit = expm(r * deg2rad(10))
-qinit = TO.rotmat_to_quat(R0)
-xinit = Dynamics.build_state(model, zeros(3), qinit, zeros(3), zeros(3))
-X = simulate(model, cntrl, xinit, tf)
-visualize!(vis, model, X[1:10:length(X)], tf)
-
-R0 = @SMatrix [1 0 0;
-      0 -0.9995 -0.0314;
-      0 0.0314 -0.9995]
-R0'
-inv(R0)
-model.J
-model.mass
-
-############################################################################################
-#                                 LQR vs MLQR COMPARISON
-############################################################################################
-function build_controller(model,dt,Q_,R)
-    Q = Diagonal(Dynamics.fill_error_state(model, Q_...))
-    n,m = size(model)
-    n = TO.state_diff_size(model)
-    x0 = zeros(model)[1]
-    u0 = trim_controls(model)
-    cntrl = MLQR(model, dt, Q, R, x0, u0)
[…]

### 2020-10-15 visualization.jl

@@ -1,38 +1,0 @@
-using PlanningWithAttitude
-using MeshCat
-using GeometryTypes
-using CoordinateTransformations
-using FileIO
-using Plots
-
-function Plots.plot(X::Vector{SVector{N,Float64}}, inds=1:N; kwargs...) where N
-    A = Array(hcat(X...))[inds,:]
-    plot(A'; kwargs...)
-end
-
-function Plots.plot!(X::Vector{SVector{N,Float64}}, inds=1:N; kwargs...) where N
-    A = Array(hcat(X...))[inds,:]
-    plot!(A'; kwargs...)
-end
-
-function visualize!(vis, model::AbstractModel, Z::Traj)
-    X = states(Z)
-    fps = Int(round(length(Z)/Z[end].t))
-    anim = MeshCat.Animation(fps)
-    for k in eachindex(Z)
-        atframe(anim, k) do
-            x = position(model, X[k])
-            q = orientation(model, X[k])
-            q = Quat(q...)
-            settransform!(vis["robot"], compose(Translation(x), LinearMap(q)))
-        end
-    end
-    setanimation!(vis, anim)
-    return anim
-end
-
-function set_mesh!(vis, model::AbstractModel; kwargs...)
-    setobject!(vis["robot"], get_mesh!(model; kwargs...))
-end
-
-get_mesh!(::Satellite; dims=[1,1,2]*0.5) = HyperRectangle(Vec((-dims/2)...), Vec(dims...))

### 2020-12-28 paper/main.tex
Clean out old files
Update Wahba problem

@@
 \usepackage{tikzscale}
+\usepgfplotslibrary{fillbetween}
 % \pgfplotsset{every axis plot post/.style={
@@
 \newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
+\newcommand{\added}[1]{\textcolor{blue}{#1}}
 
@@
 \begin{abstract}
-Planning trajectories for floating-base robotic systems that experience
-large attitude changes is challenging due to the nontrivial group structure of 3D
+Planning trajectories for floating-base robotic systems that experience large
+attitude changes is challenging due to the nontrivial group structure of 3D
 rotations. This paper introduces a powerful and accessible approach for
-optimization-based planning on the space of rotations using only standard linear algebra and vector calculus. We demonstrate the effectiveness of the approach by adapting Newton's
-method to solve the canonical Wahba's problem, and modify the trajectory optimization solver ALTRO to plan directly on the space of unit 
-quaternions, achieving superior convergence on problems involving significant changes 
-in attitude.
+optimization-based planning on the space of rotations using only standard
+linear algebra and vector calculus. We demonstrate the effectiveness of the
+approach by adapting Newton's method to solve the canonical Wahba's problem,
+and modify the trajectory optimization solver ALTRO to plan directly on the
+space of unit quaternions, achieving superior convergence on problems
+involving significant changes in attitude.
 %\keywords{trajectory optimization, motion planning, control, quaternions}
@@
         respectively. The space of unit quaternions, $\Q = \{q : \norm{q}_2 = 1\}$, is a 
-        a double-cover of the rotation group $SO(3)$, since $q$ and $-q$ represent the same 
+        double-cover of the rotation group $SO(3)$, since $q$ and $-q$ represent the same 
         rotation \cite{markley2014fundamentals}.
@@
     multiplicative method is able to maintain the fast convergence rates typical of
-    Newton's method.
+    Newton's method. \added{By comparing our method with the global solution obtained
+    from an SVD decomposition, we see that our method recovers the globally optimal
+    solution within a small number of iterations.}
     \begin{figure}
@@
         \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
-        \caption{Convergence comparison for Wahba's problem. By modfying Newton's method
-        with the methods of section \ref{sec:Quaternion_Calculus}, quadratic convergence
-        rates are achieved, while a na\"ive approach stalls after only a few
-        iterations.}
+        \caption{Convergence comparison for Wahba's problem. \added{The error
+        is the angle between the current solution and the globally optimal
+        solution computed using SVD decomposition \todo{Kevin: add citation}.
+        The thick line is the average result of 100 trials with randomized
+        orientations and measurements. The thin lines are the maximum and
+        minimum over all 100 trials.} By modfying Newton's method with the
+        methods of section \ref{sec:Quaternion_Calculus}, quadratic
+        convergence rates are achieved, while a na\"ive approach stalls after
+        only a few iterations.}
         \label{fig:wahba_convergence}

### 2020-12-28 examples/Wahbas_problem/NLLS_Wahba.jl

@@ -1,6 +1,6 @@
 # activate the virtual environment
 # cd(dirname(@__FILE__))
-import Pkg; Pkg.activate(@__DIR__)
+import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
 
 using ForwardDiff, LinearAlgebra, Statistics, JLD2
 # using MATLAB
@@ -350,7 +350,7 @@ function projected_newton_wahba()
 end
 
 ## run trials
-trials = 25
+trials = 100 
 max_iters = 21
 mn = zeros(trials,max_iters)
 pn = zeros(trials,max_iters)
@@ -431,6 +431,33 @@ p = @pgf Axis(
     ),
     Legend("naive", "modified")
 )
+# pgfsave("paper/figures/wahba_convergence.tikz", p, include_preamble=false)
+m_max = maximum(mn, dims=1)
+m_min = minimum(mn, dims=1)
+p_max = maximum(pn, dims=1)
+p_min = minimum(pn, dims=1)
+
+p = @pgf Axis(
+    {
+        xlabel="iterations",
+        ylabel="angle error (degrees)",
+        "ymode=log",
+        xmajorgrids,
+        ymajorgrids,
+        "no markers",
+        "legend style={at={(0.1,0.1)},anchor=south west}"
+    },
+    PlotInc({"orange, ultra thick"}, Coordinates(1:5, p_average)),
+    PlotInc({"orange, name path=A, line width=0.1pt, forget plot"}, Coordinates(1:5, p_max[1:5])),
+    PlotInc({"orange, name path=B, line width=0.1pt, forget plot"}, Coordinates(1:5, p_min[1:5])),
+    PlotInc({"orange!10, forget_plot"}, "fill between [of=A and B];"),
+    PlotInc({"cyan, ultra thick"}, Coordinates(1:5, m_average)),
+    PlotInc({"cyan, solid, name path=C, line width=0.1pt, forget plot"}, Coordinates(1:5, m_max[1:5])),
+    PlotInc({"cyan, solid, name path=D, line width=0.1pt, forget plot"}, Coordinates(1:5, m_min[1:5])),
+    PlotInc({"cyan!10, forget plot"}, "fill between [of=C and D];"),
+    Legend("naive", "modified")
+)
+print_tex(p)
 pgfsave("paper/figures/wahba_convergence.tikz", p, include_preamble=false)
 
 mat"

### 2020-12-29 examples/quadflip_mc.jl
Flip montecarlo

@@ -1,0 +1,84 @@
+import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+using RobotDynamics
+import RobotZoo.Quadrotor
+using TrajectoryOptimization
+using Altro
+using BenchmarkTools
+const TO = TrajectoryOptimization
+const RD = RobotDynamics
+
+using Random
+using StaticArrays
+using LinearAlgebra
+using Rotations
+
+using PlanningWithAttitude
+using JLD2
+using DataFrames
+
+args = (
+    integration = RK4,
+    termcon = :quatvec,
+    projected_newton = true,
+    show_summary=true,
+    verbose = 0,
+    iterations = 100,
+    static_bp = false 
+)
+
+## Solve problem
+prob,opts = QuadFlipProblem(vecmodel=true, renorm=true; args...)
+solver_vec = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, static_bp=false)
+prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost; args...)
+solver_err = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, static_bp=false)
+n,m = size(prob)
+solve!(solver_vec)
+solve!(solver_err)
+visualize!(vis, Quadrotor(), get_trajectory(solver_vec))
+
+## Get reference trajectories
+Xvec = states(solver_vec)
+Uvec = [u[SOneTo(4)] for u in controls(solver_vec)]
+Xerr = states(solver_err)
+Uerr = [u[SOneTo(4)] for u in controls(solver_err)]
+
+## Solve with perturbed initial guess
+prob,opts = QuadFlipProblem(vecmodel=true, renorm=true; args...)
+solver_vec = ALTROSolver(prob, opts, R_inf=1e3, infeasible=true, projected_newton=false, iterations=400)
+prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost; args...)
+solver_err = ALTROSolver(prob, opts, R_inf=1e3, infeasible=true, projected_newton=false, iterations=400)
+
+## Run Monte-Carlo Analysis
+T = 100 
+iters = zeros(T,2)
+status = fill(Altro.UNSOLVED,T,2) 
+dt = [z.dt for z in prob.Z]
+using Random
+Random.seed!(1)
+
+for i = 1:T
+    println("Iteration $i")
[…]

### 2020-12-29 src/airplane_problem.jl

@@ -1,4 +1,4 @@
-using TrajectoryOptimization: LieLQRCost
+# using TrajectoryOptimization: LieLQRCost
 using ForwardDiff
 using RobotDynamics
 
@@ -10,6 +10,7 @@ function YakProblems(;
         costfun=:Quadratic, 
         termcon=:goal,
         quatnorm=:none,
+        heading=0.0,  # deg
         kwargs...
     )
     model = RobotZoo.YakPlane(UnitQuaternion)
@@ -36,10 +37,12 @@ function YakProblems(;
         # Initial and final condition
         p0 = MRP(0.997156, 0., 0.075366) # initial orientation
         pf = MRP(0., -0.0366076, 0.) # final orientation (upside down)
+        dq = expm(SA[0,0,1]*deg2rad(heading))
+        pf = pf * dq
 
         x0 = RD.build_state(model, [-3,0,1.5], p0, [5,0,0], [0,0,0])
         utrim  = @SVector  [41.6666, 106, 74.6519, 106]
-        xf = RD.build_state(model, [3,0,1.5], pf, [5,0,0], [0,0,0])
+        xf = RD.build_state(model, [3,0,1.5], pf, dq * [5,0,0.], [0,0,0])
 
         # Xref trajectory
         x̄0 = RBState(model, x0)
@@ -70,7 +73,6 @@ function YakProblems(;
             costfuns = map(Xref) do xref
                 QuatLQRCost(Q, R, xf, utrim, w=0.1)
             end
-            costfun = QuatLQRCost(Q, R, xf, utrim; w=0.1)
             costterm = QuatLQRCost(Qf, R, xf, utrim; w=200.0)
             costfuns[end] = costterm
         elseif costfun == :LieLQR

### 2020-12-29 src/quadflip_problem.jl

@@ -20,9 +20,10 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
     x0 = RobotDynamics.build_state(model, x0_pos, UnitQuaternion(I), zeros(3), zeros(3))
 
     # cost
+    s = costfun == LQRCost
     Q_diag = RobotDynamics.build_state(model, 
-        [1e-2,1e-2,5e-2], 
-        fill(1e-5,rsize), 
+        [1e-6,1e-6,1e-6], 
+        fill(1e-5,rsize)*s, 
         fill(1e-3,3), 
         fill(1e-2,3)
     )
@@ -34,17 +35,18 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
     # waypoints
     ex = @SVector [1.0, 0, 0]
     wpts = [
-        ((@SVector [0, 0.5, 1.5,]),   expm(ex*deg2rad(90))),
+        ((@SVector [0, 1.0, 1.0,]),   expm(ex*deg2rad(00))),
+        ((@SVector [0, 1.5, 1.5,]),   expm(ex*deg2rad(90))),
         # ((@SVector [0, 0.2, 1.5]),    expm(ex*deg2rad(90))),
         # ((@SVector [0, 0.0, 2.0]),    expm(ex*deg2rad(135))),
-        ((@SVector [0, 0.0, 2.5]),    expm(ex*deg2rad(180))),
+        ((@SVector [0, 1.0, 2.5]),    expm(ex*deg2rad(180))),
         # ((@SVector [0, 0.0, 2.0]),    expm(ex*deg2rad(225))),
-        ((@SVector [0,-0.5, 1.5]),    expm(ex*deg2rad(-90))),
+        ((@SVector [0, 0.5, 1.5]),    expm(ex*deg2rad(270))),
         ((@SVector [0, 0.65, 1.0]),    expm(ex*deg2rad(360))),
         ((@SVector [0, 1.0, 1.0]),    expm(ex*deg2rad(360))),
     ]
     # times = [35, 41, 47, 51, 55, 61, 70, 101]
-    times = [45, 51, 55, 75, 101]
+    times = [20, 45, 51, 55, 75, 101]
 
     """
     Costs
@@ -52,12 +54,12 @@ function QuadFlipProblem(Rot=UnitQuaternion; slack::Bool=false, vecmodel::Bool=f
[…]

### 2020-12-29 src/vecmodel.jl

@@ -5,4 +5,8 @@ end
 
 RobotDynamics.state_dim(model::VecModel) = RD.state_dim(model.model)
 RobotDynamics.control_dim(model::VecModel) = RD.control_dim(model.model)
-RobotDynamics.dynamics(model::VecModel, args...) = RD.dynamics(model.model, args...) 
+RobotDynamics.discrete_dynamics(::Type{Q}, model::VecModel, z::RobotDynamics.AbstractKnotPoint) where Q <: RD.Explicit = 
+    discrete_dynamics(Q, model.model, z) 
+RobotDynamics.discrete_dynamics(::Type{Q}, model::VecModel, args...) where Q <: RD.Explicit = 
+    discrete_dynamics(Q, model.model, args...) 
+# RobotDynamics.dynamics(model::VecModel, args...) = RD.dynamics(model.model, args...) 

### 2020-12-29 paper/main.tex
Update with revision

@@
         \begin{equation} \label{eq:rigid_body_dynamics}
-            x = \begin{bmatrix} r \\ R \\ v \\ \omega \end{bmatrix}, \quad 
+            x = \begin{bmatrix} r \\ q \\ v \\ \omega \end{bmatrix}, \quad 
             \dot{x} = \begin{bmatrix} 
@@
 
-        Let $\phi \in \R^3$ be a differential rotation applied to a function with
-        quaternion inputs $y = h(q): \Q \to \R^p$, such that
+        Let $\phi \in \R^3$ be a differential rotation \added{from the group identity to the 
+        input of a vector-valued function of a quaternion} 
+        $y = h(q): \Q \to \R^p$, such that
         \begin{equation} \label{eq:vector_function}
@@
 
-    \subsection{Results}
-    Figure \ref{fig:wahba_convergence} compares the multiplicative Gauss-Newton method
-    with a na\"ive Newton's method in which the quaternion is simply projected back onto
-    the unit sphere at every iteration. The na\"ive method makes progress initially, but
-    quickly stalls. By correctly handling the group structure of unit quaternions, the
-    multiplicative method is able to maintain the fast convergence rates typical of
-    Newton's method. \added{By comparing our method with the global solution obtained
-    from an SVD decomposition, we see that our method recovers the globally optimal
-    solution within a small number of iterations.}
+    \subsection{Results} Figure \ref{fig:wahba_convergence} compares the
+    multiplicative Gauss-Newton method with a na\"ive Newton's method in
+    which the quaternion is simply projected back onto the unit sphere at
+    every iteration. The na\"ive method makes progress initially, but quickly
+    stalls. By correctly handling the group structure of unit quaternions,
+    the multiplicative method is able to maintain the fast convergence rates
+    typical of Newton's method. \added{By comparing our method with the
+    global solution obtained from an SVD decomposition
+    \cite{markley1999estimate}, we see that our method recovers the globally
+    optimal solution within a small number of iterations.}
+
     \begin{figure}
-
         \centering
@@
         is the angle between the current solution and the globally optimal
-        solution computed using SVD decomposition \todo{Kevin: add citation}.
+        solution computed using SVD decomposition.
         The thick line is the average result of 100 trials with randomized
@@
 
-\section{Trajectory Optimization on $\R^n \times SO(3)$} \label{sec:trajopt}
+\section{Trajectory Optimization \added{for Rigid Bodies}} \label{sec:trajopt}
     Here we outline the modifications to the ALTRO solver \cite{howell2019altro} to
     solve trajectory optimization problems for rigid bodies, which extends easily to
-    arbitrary systems whose state is in $\R^n \times SO(3)$.
+    arbitrary systems whose state is in $\R^n \times \added{\Q}$.
 %ALTRO is an efficient solver for constrained nonlinear optimization problems that uses iterative LQR (iLQR) with an augmented Lagrangian framework.
@@
     \begin{mini}[2]
-        {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell_k(x_k,u_k) }{}{}
+        {x_{1:N},u_{1:N-1}}{\ell_f(x_N) + \sum_{k=1}^{N-1} \ell_k(x_k,u_k) }{}{}
         \addConstraint{x_{k+1} = f(x_k,u_k)}
@@
             r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
-        \end{bmatrix} .
+        \end{bmatrix}.
     \end{equation}
@@
     where $\bar{u}_k$ is the control value from the previous iteration, and $\delta x$ is
-    computed using \eqref{eq:state_error}.%, with $x_k$ being the current state estimate and $\bar{x}_k$ the state from the previous iteration. 
+    computed using \eqref{eq:state_error}. %, with $x_k$ being the current state estimate and $\bar{x}_k$ the state from the previous iteration. 
     For more details on the ALTRO algorithm, we refer the reader to~\cite{howell2019altro}.
@@
         \end{align}
-        where $\sign$ denotes the signum function.
+        where $\sign$ denotes the signum function. \added{This cost function is particularly
+        useful for rotations since it eliminates the ambiguity arising from the quaternion 
+        double-cover of $SO(3)$.}
 
@@
     solved to a constraint satisfaction tolerance of $10^{-5}$ and discretized with a 4th 
-    order Runge-Kutta integrator. The results were run on a desktop computer with an AMD
-    Ryzen 2950x processor with 40 GB of RAM. Code for all experiments is available on
+    order Runge-Kutta integrator. The results were run on a laptop computer with a 2.8 GHz 
+    i7-1165G7 processor with 16 GB of RAM. Code for all experiments is available on
     \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
@@
         \centering
+        \caption{Trajectory Optimization Timing Results (naive/modified)}
         \input{figures/timing_results.tex}
         \label{tab:timing_results}
-        \caption{Trajectory Optimization Timing Results (naive/modified)}
     \end{table}
@@
 
-        A 180 degree barrel roll trajectory for a fixed-wing airplane was optimized. The
-        airplane's dynamics model consists of the a simple rigid body as defined in
-        Section \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag
-        fit from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do
-        a barrel roll by constraining the terminal state to upside-down (see Fig.
-        \ref{fig:barrellroll}). To mitigate issues with integration error and drift in the magnitude of the quaternion,
-        the following constraint function was used to enforce a terminal orientation of $\bar{q}$:
+        A 180 degree barrel roll trajectory for a fixed-wing airplane was
+        optimized. The airplane's dynamics model consists of the a simple
+        rigid body as defined in Section \ref{sec:rigidbody_dynamics} with
+        forces and torques due to lift and drag fit from wind tunnel data
+        \cite{manchester2016udp}. The airplane was tasked to do a barrel roll
+        by constraining the terminal state to upside-down (see Fig.
+        \ref{fig:barrellroll}). To mitigate issues with integration error and
+        drift in the magnitude of the quaternion, the following constraint
+        function was used to enforce a terminal orientation of $\bar{q}$:
         \begin{equation}
@@
         relationship between the attitude state and the goal and constraints. For this 
-        relatively simple problem, we gained a modest 11\% improvement in runtime, despite
+        relatively simple problem, we gained a modest 31\% improvement in runtime, despite
         the additional matrix multiplications when calculating the cost and constraint
@@
         the weight matrix for the waypoints, and
-        $\mathcal{W} = \{45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
+        $\mathcal{W} = \{20,45,51,55,75,101\}$, $\mathcal{N} = \{1:101\} \setminus
         \mathcal{W}$.
-        Four intermediary ``waypoints'' were used to encourage the
-        quadrotor to reach angles of \ang{90}, \ang{180}, \ang{270}, and \ang{360} around
-        an approximately circular arc. The last waypoint was used to encourage the
-        quadrotor to move towards the final goal. The quadrotor was constrained to stay
-        above the floor and move to a goal state two meters away in the $+y$ direction. The
-        solver was provided a dynamically infeasible initial trajectory that linearly
-        interpolates between the initial and final states, rotating the quadrotor around
-        the x-axis a full \ang{360}.
+        Four intermediary ``waypoints'' were used to encourage the quadrotor
+        to reach angles of \ang{90}, \ang{180}, \ang{270}, and \ang{360}
+        around an approximately circular arc. The last waypoint was used to
+        encourage the quadrotor to move towards the final goal, and the first
+        kept it above the floor before starting the loop. The solver was
+        provided a dynamically infeasible initial trajectory that linearly
+        interpolates between the initial and final states, rotating the
+        quadrotor around the x-axis a full \ang{360}.
 
-	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
-	    ALTRO. The original version of ALTRO, even after significant tuning efforts, was
-        not able to converge to the desired solution. 
-	    This behavior is common and expected when attempting optimization that does not
-	    properly account for the group structure of rotations. It is also worth noting
-	    that this problem could not be solved using any three-parameter attitude
-	    representation, since it passes through the singularities at $90\degree,
-	    180\degree$, and $360\degree$ associated with Euler angles, Rodrigues parameters,
-	    and Modified Rodrigues Parameters, respectively.
+	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as
+	    generated using ALTRO. \added{To compare the convergence properties of the
+	    two methods, the optimal state and control trajectories were
+	    perturbed with random Gaussian white noise with a mean of 1 for
+	    position, linear velocity, and angular velocity, 0.1 for the
+        controls, and 145 degrees for the orientation (about a random axis). 
+        As shown in Fig. \ref{fig:flip_success}, the modified method converges
+        more reliably than the na\"ive method.}
+	    It is also worth noting that this problem could not be solved using
+	    any three-parameter attitude representation, since it passes through
+	    the singularities at $90\degree, 180\degree$, and $360\degree$
+	    associated with Euler angles, Rodrigues parameters, and Modified
+	    Rodrigues Parameters, respectively.
 
@@
             \label{fig:quad_flip}
+        \end{figure}
+
+        \begin{figure}
+            \centering    
+            \begin{tikzpicture}
+                \begin{axis}[
+                    ybar,
+                    width=\columnwidth, 
+                    height=5cm, 
+                    enlarge x limits=0.5,
+                    enlarge y limits=0.2,
+                    ylabel={Percent Success},
+                    xtick={1,2},
+                    xticklabels={Na\"ive, Modified},
+                    bar width=0.6,
+                    nodes near coords,
+                    bar shift=0pt,
+                    axis on top
+                ]
+                \addplot [orange, fill] coordinates {(1,86)};
+                \addplot [cyan, fill]   coordinates {(2,97)};
+                \end{axis}
+            \end{tikzpicture}
+            \caption{Convergence comparison for quadrotor flip. Percent of 100 trials 
+            that successfully converged, where each trial is initialized with locally-optimal
+            trajectories perturbed with significant Gaussian white noise.}
+            \label{fig:flip_success}
         \end{figure}

### 2020-12-29 examples/trajopt_examples.jl

@@ -64,13 +64,13 @@ add_result!(:barrellroll, true, solver)
 Quadrotor Flip
 """
 ## Original Method
-prob,opts = QuadFlipProblem(vecmodel=true)
-solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true)
+prob,opts = QuadFlipProblem(vecmodel=true, renorm=true; args...)
+solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, static_bp=false)
 add_result!(:quadflip, false, solver)
 
 ## Modified Method
 prob,opts = QuadFlipProblem(vecmodel=false, renorm=true, costfun=QuatLQRCost; args...)
-solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true)
+solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, static_bp=false)
 add_result!(:quadflip, true, solver)
 
 @save results_path results
@@ -101,15 +101,14 @@ df = DataFrame(results)
 df.time .= round.(df.time, digits=2)
 df.prob = String.(df.problem)
 
-
 mod = df[df.errstate,:]
 org = df[.!df.errstate,:]
 tab = vcat(map(1:3) do i
     [mod.prob[i] @sprintf("%i / %i", org.iters[i], mod.iters[i]) @sprintf("%.2f / %.2f", org.time[i], mod.time[i])]
 end...)
 
-tab[2,2] = "— / $(mod.iters[2])"  # remove unsucessful quadrotor results
-tab[2,3] = "— / $(mod.time[2])"
+# tab[2,2] = "— / $(mod.iters[2])"  # remove unsucessful quadrotor results
+# tab[2,3] = "— / $(mod.time[2])"
 latex_tabular(joinpath(@__DIR__,"..","paper/figures/timing_results.tex"),
     Tabular("lll"),
     [
@@ -154,3 +153,30 @@ p = @pgf Axis(
     Legend("naive","modified")
 )
 pgfsave(joinpath(@__DIR__, "..", "paper/figures/c_max_convergence.tikz"), p, include_preamble=false)
+
+
+## Generate visualizations
+using TrajOptPlots, MeshCat, Colors
+vis = Visualizer()
+open(vis)
+
+# Barrell Roll
+prob,opts = YakProblems(vecstate=false, costfun=:QuatLQR; args...) 
[…]

### 2020-12-29 examples/yak_barrellroll.jl

@@ -1,10 +1,11 @@
 import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
 using RobotDynamics
-import RobotZoo.Quadrotor
+import RobotZoo.YakPlane
 using TrajectoryOptimization
 using Altro
 using BenchmarkTools
 using LaTeXTabulars
+using LaTeXStrings
 const TO = TrajectoryOptimization
 const RD = RobotDynamics
 
@@ -15,15 +16,17 @@ using Rotations
 
 using PlanningWithAttitude
 using JLD2
-# using TrajOptPlots
-# using MeshCat
+using TrajOptPlots
+using MeshCat
+vis = Visualizer()
+open(vis)
 
 args = (
     constraint_tolerance=1e-5,
     integration = RK4,
     termcon = :quatvec,
-    projected_newton = false,
-    show_summary=true
+    projected_newton = true,
+    show_summary=true,
 )
 
 ##
@@ -31,22 +34,29 @@ args = (
 Compare Cost Functions
 """
 
-prob,opts = YakProblems(vecstate=false, costfun=:ErrorQuadratic; args...) 
+delete!(vis)
+TrajOptPlots.set_mesh!(vis, YakPlane(UnitQuaternion))
+prob,opts = YakProblems(vecstate=false, costfun=:QuatLQR, heading=0; args...) 
+prob,opts = YakProblems(vecstate=true, costfun=:Quadratic, heading=0; args...) 
 solver = ALTROSolver(prob, opts)
 solve!(solver)
-ilqr = Altro.get_ilqr(solver)
+visualize!(vis, YakPlane(UnitQuaternion), get_trajectory(solver))
 
-function run_example(;vecstate=false, costfun=:Quadratic, args...)
-    prob,opts = YakProblems(vecstate=vecstate, costfun=costfun; args...) 
+function run_example(;vecstate=false, costfun=:Quadratic, heading=0.0, args...)
+    prob,opts = YakProblems(vecstate=vecstate, costfun=costfun, heading=heading; args...) 
     solver = ALTROSolver(prob, opts)
-    b = benchmark_solve!(solver)
+    # b = benchmark_solve!(solver)
+    solve!(solver)
     push!(cost_comparison[:costfun], costfun)
     push!(cost_comparison[:errstate], !vecstate)
-    push!(cost_comparison[:time], median(b).time)
+    # push!(cost_comparison[:time], median(b).time)
[…]

### 2020-12-29 examples/quadrotor_flip.jl
delete old example files

@@ -1,48 +1,0 @@
-import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
-using RobotDynamics
-import RobotZoo.Quadrotor
-using TrajectoryOptimization
-using PlanningWithAttitude
-using Altro
-using TrajOptPlots
-using BenchmarkTools
-const TO = TrajectoryOptimization
-
-using Random
-using StaticArrays
-using LinearAlgebra
-using MeshCat
-using Rotations
-using Colors
-
-## Try ALTRO
-prob,opts = QuadFlipProblem(UnitQuaternion, slack=false, vecmodel=false, renorm=true,
-    costfun=QuatLQRCost
-)
-solver = ALTROSolver(prob, opts, R_inf=1e-4, infeasible=true, show_summary=true, 
-    projected_newton=true)
-solve!(solver)
-
-## Original Method
-prob2,opts = QuadFlipProblem(UnitQuaternion, slack=false, vecmodel=true)
-solver2 = ALTROSolver(prob2, opts, R_inf=1e-4, infeasible=true, show_summary=true,
-    projected_newton=true)
-solve!(solver2)
-
-## Visualize
-using TrajOptPlots, Blink
-if !isdefined(Main, :vis)
-    vis = Visualizer()
-    open(vis, Blink.Window())
-end
-delete!(vis)
-quad = Quadrotor()
-TrajOptPlots.set_mesh!(vis, quad)
-visualize!(vis, quad, get_trajectory(solver))
-visualize!(vis, quad, get_trajectory(solver2))
-
-## Waypoints
-delete!(vis)
-TrajOptPlots.waypoints!(vis, quad, get_trajectory(solver), 
-    inds=[1,15,20,25,28,30,32,35,40,42,44,46,48,50,52,54,56,58,60,62,65,68,70,75,90,100],
-    color=HSL(colorant"green"), color_end=HSL(colorant"red"))

### 2020-12-29 examples/terminal_quats.jl

@@ -1,70 +1,0 @@
-import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
-using RobotDynamics
-import RobotZoo.Quadrotor
-using TrajectoryOptimization
-# using PlanningWithAttitude
-using Altro
-# using TrajOptPlots
-using BenchmarkTools
-const TO = TrajectoryOptimization
-const RD = RobotDynamics
-
-using Random
-using StaticArrays
-using LinearAlgebra
-# using MeshCat
-using Rotations
-
-
-
-## Test constraints
-model = Quadrotor()
-qf = expm(SA[0,0,1]*deg2rad(135))
-x, = rand(model)
-xf = RD.build_state(model, [2,3,1], qf, zeros(3), zeros(3))
-con1 = QuatGeoCon(13,qf,SA[4,5,6,7])
-con2 = QuatGeoCon(13,-qf,SA[4,5,6,7])
-TO.evaluate(con1, x)[1] ≈ TO.evaluate(con2, x)[1]
-TO.evaluate(con1, xf)[1] ≈ 0
-jac = TO.gen_jacobian(con1)
-TO.jacobian!(jac, con1, x)
-
-con3 = QuatErr(13, qf, SA[4,5,6,7])
-con4 = QuatErr(13,-qf, SA[4,5,6,7])
-TO.evaluate(con3, x) ≈ TO.evaluate(con4, x)
-TO.evaluate(con3, xf) ≈ zeros(3)
-jac = TO.gen_jacobian(con3)
-TO.jacobian!(jac, con3, x)
-
-con5 = QuatVecEq(13, qf, SA[4,5,6,7])
-con6 = QuatVecEq(13,-qf, SA[4,5,6,7])
-TO.evaluate(con5, x) ≈ TO.evaluate(con6, x)
-TO.evaluate(con5, xf) ≈ zeros(3) 
-jac = TO.gen_jacobian(con5)
-TO.jacobian!(jac, con5, x)
-
-##
-model = Quadrotor()
-N = 51
-tf = 5.0
-dt = tf / (N-1)
-
-# Objective
-x0,u0 = zeros(model) 
-Q = [fill(1,3), fill(0.1, 6)]
-R = fill(1e-2,4)
-costfun = TO.LieLQRCost(RD.LieState(model), Q, R, xf)
-costfun_term = TO.LieLQRCost(RD.LieState(model), Q .* 100, R, xf)
-obj = Objective(costfun, costfun_term, N)
-
-# Constraints
[…]

### 2020-12-29 examples/yak_barrellroll.jl

@@ -1,279 +1,0 @@
-import Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
-using RobotDynamics
-import RobotZoo.YakPlane
-using TrajectoryOptimization
-using Altro
-using BenchmarkTools
-using LaTeXTabulars
-using LaTeXStrings
-const TO = TrajectoryOptimization
-const RD = RobotDynamics
-
-using Random
-using StaticArrays
-using LinearAlgebra
-using Rotations
-
-using PlanningWithAttitude
-using JLD2
-using TrajOptPlots
-using MeshCat
-vis = Visualizer()
-open(vis)
-
-args = (
-    constraint_tolerance=1e-5,
-    integration = RK4,
-    termcon = :quatvec,
-    projected_newton = true,
-    show_summary=true,
-)
-
-##
-"""
-Compare Cost Functions
-"""
-
-delete!(vis)
-TrajOptPlots.set_mesh!(vis, YakPlane(UnitQuaternion))
-prob,opts = YakProblems(vecstate=false, costfun=:QuatLQR, heading=0; args...) 
-prob,opts = YakProblems(vecstate=true, costfun=:Quadratic, heading=0; args...) 
-solver = ALTROSolver(prob, opts)
-solve!(solver)
-visualize!(vis, YakPlane(UnitQuaternion), get_trajectory(solver))
-
-function run_example(;vecstate=false, costfun=:Quadratic, heading=0.0, args...)
-    prob,opts = YakProblems(vecstate=vecstate, costfun=costfun, heading=heading; args...) 
-    solver = ALTROSolver(prob, opts)
-    # b = benchmark_solve!(solver)
-    solve!(solver)
-    push!(cost_comparison[:costfun], costfun)
-    push!(cost_comparison[:errstate], !vecstate)
-    # push!(cost_comparison[:time], median(b).time)
-    push!(cost_comparison[:time], solver.stats.tsolve*1e6) 
-    push!(cost_comparison[:iters], iterations(solver))
-    push!(cost_comparison[:cost], solver.stats.cost)
-    push!(cost_comparison[:c_max], solver.stats.c_max)
-    push!(cost_comparison[:is_outer], solver.stats.iteration_outer)
-    push!(cost_comparison[:heading], heading)
-    push!(cost_comparison[:status], status(solver))
-end
-
-cost_comparison = Dict(
-    :costfun => Symbol[], 
[…]

### 2020-12-29 paper/main.tex
tiny edits

@@
 
-        Let $\phi \in \R^3$ be a differential rotation \added{from the group identity to the 
+        Let $\phi \in \R^3$ be a differential rotation \added{applied to the 
         input of a vector-valued function of a quaternion} 
@@
     typical of Newton's method. \added{By comparing our method with the
-    global solution obtained from an SVD decomposition
+    global solution obtained from a singular-value decomposition
     \cite{markley1999estimate}, we see that our method recovers the globally
@@
         is the angle between the current solution and the globally optimal
-        solution computed using SVD decomposition.
+        solution computed using a singular-value decomposition.
         The thick line is the average result of 100 trials with randomized

### 2021-01-01 paper/main.tex
minor tweak (back to original)
added some notes about frame conventions

@@
         forces in the world frame, and $\inframe{B}{\tau(x,u)}$ are the moments in the
-        body frame.
+        body frame. \added{We use the convention that ``attitude'' always refers to a rotation from body to world frame.}
 
@@
 
-        Let $\phi \in \R^3$ be a differential rotation \added{applied to the 
-        input of a vector-valued function of a quaternion} 
+        Let $\phi \in \R^3$ be a differential rotation applied to a function with
+        quaternion inputs 
         $y = h(q): \Q \to \R^p$, such that
@@
         \end{equation}
-        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by
+        \added{Note that we chose to represent $\phi$ in the body frame, consistent with the standard definition of angular velocity, and therefore it is applied to $q$ through right (rather than left) multiplication.} We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by
         differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at

### 2021-01-05 paper/main.tex
removed unnecessary frame comment
Update formatting

@@
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
-\IEEEoverridecommandlockouts                              % This command is only needed if 
+% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
+\documentclass[letterpaper, 10 pt, journal, twoside]{ieeetran}
+% \IEEEoverridecommandlockouts                              % This command is only needed if 
                                                           % you want to use the \thanks command
-\overrideIEEEmargins                                      % Needed to meet printer requirements.
-\usepackage[top=60pt, left=48pt, right=48pt, bottom=43pt]{geometry}
+% \overrideIEEEmargins                                      % Needed to meet printer requirements.
+% \usepackage[top=60pt, left=48pt, right=48pt, bottom=43pt]{geometry}
 
@@
 \newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
-\newcommand{\added}[1]{\textcolor{blue}{#1}}
+\newcommand{\added}[1]{\textcolor{black}{#1}}
+
+% Header, Title, etc.
+\markboth{IEEE Robotics and Automation Letters. Preprint Version. Accepted Dec, 2020}
+{Jackson \MakeLowercase{et al.}: Planning with Attitude}
 
@@
 
-\author{Brian E. Jackson$^1$, Kevin Tracy$^1$, and Zachary Manchester$^1$%
-    \thanks{
-        $^1$Robotics Institute, 
-        Carnegie Mellon University, 
-        5000 Forbes Ave, Pittsburgh, PA, USA
-    }
+% Make room for more info lines in the \author command 
+\author{Brian E Jackson$^{1}$, Kevin Tracy$^{1}$, and Zachary Manchester$^{1}$%
+\thanks{Manuscript received: October, 15, 2020; Revised January 2, 2021; Accepted January 4, 2021.}%Use only for final RAL version
+\thanks{This paper was recommended for publication by Editor Nancy Amato upon evaluation of the Associate Editor and Reviewers' comments.
+This work was supported by a NASA Early Career Faculty Award, NASA JPL, and NSF GFRP Grant No. DGE-1656518.} %Use only for final RAL version
+\thanks{$^{1}$All authors are with the Robotics Institute, Carnegie Mellon University, Pittsburgh, PA, USA 
+        {\tt\footnotesize brian.jackson@cmu.edu}}%
+\thanks{Digital Object Identifier (DOI): see top of this page.}
 }
@@
 
+\begin{IEEEkeywords}
+    Motion and Path Planning, Optimization and Optimal Control, Computational Geometry, Underactuated Robots, Motion Control, 
+\end{IEEEkeywords}
+
 \section{INTRODUCTION}
 
-    Many robotic systems---including quadrotors, airplanes, satellites, autonomous
+    \IEEEPARstart{M}{any} robotic systems---including quadrotors, airplanes, satellites, autonomous
     underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
@@
         forces in the world frame, and $\inframe{B}{\tau(x,u)}$ are the moments in the
-        body frame. \added{We use the convention that ``attitude'' always refers to a rotation from body to world frame.}
+        body frame.
 

### 2021-05-05 paper/main.tex
ICRA version
Minor tweaks
Small updates

@@
 % Bibliography
-\usepackage{biblatex}
+\usepackage[style=ieee]{biblatex}
 \addbibresource{references.bib}
@@
 \thanks{$^{1}$All authors are with the Robotics Institute, Carnegie Mellon University, Pittsburgh, PA, USA 
-        {\tt\footnotesize brian.jackson@cmu.edu}}%
+        {\tt\footnotesize brianjackson@cmu.edu}}%
 \thanks{Digital Object Identifier (DOI): see top of this page.}
@@
 	    \begin{equation} \label{eq:quat_hessian}
-            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q}q,
+            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) - I_3 \pdv{h}{q}q,
 	    \end{equation}
@@
     \begin{equation}
-    	\phi_k = ({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
+    	\phi_k = -({\nabla r}^T {\nabla r})^{-1} {\nabla r}^T r(q_k).
     \end{equation}
@@
     \end{equation}
-    and $E(x_k) \in \R^{12 \times 13}$ is the error-state Jacobian:
+    and $E(x_k) \in \R^{13 \times 12}$ is the error-state Jacobian:
     \begin{equation} \label{eq:state_error_jacobian}
@@
         \delta \ell(x,u) \approx  \half \delta x^T \ell_{xx} \delta x
-            + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta u \\
+            + \half \delta u^T \ell_{uu} \delta u + \delta_u^T \ell_{ux} \delta x \\
             + \ell_x^T \delta x^T + \ell_u^T \delta u.
@@
         \begin{align}
-            \nabla J_\text{geo} &= \sign(q_d^T q) q_g^T G(q) \\
-            \nabla^2 J_\text{geo} &= \sign(q_d^T q) I_3 q_g^T q ,
+            \nabla J_\text{geo} &= -\sign(q_g^T q) q_g^T G(q) \\
+            \nabla^2 J_\text{geo} &= \sign(q_g^T q) I_3 q_g^T q ,
         \end{align}
